### 2022-05-17 README.Rmd
Initial commit

---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, eval = FALSE,
  fig.width = 9, fig.height = 9,
  dpi = 320,
  fig.path = "figures/"
)
```

# Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany

This repository contains the documentation, results, and code of a project evaluating the use of a semi-parametric nowcasting approach for COVID-19 hospitalisations in Germany. See the documentation for further details. This project is part of a [wider collaboration](https://covid19nowcasthub.de) assessing a range of nowcasting methods whilst providing an ensemble nowcast of COVID-19 Hospital admissions in Germany by date of positive test. This ensemble should be used for any policy related work rather than the nowcasts provided in this repository. See [here](https://covid19nowcasthub.de) for more on this nowcasting collaboration.

## Citation

If making use of the results of this analysis or reusing the analysis pipeline please cite:

If making using of the models evaluated in this analysis please also cite [`epinowcast`](https://epiforecasts.io/epinowcast):

> Sam Abbott (2021). epinowcast: Hierarchical nowcasting of right censored epidemological counts, DOI: 10.5281/zenodo.5637165

A BibTeX entry for LaTeX users is also available:

  @Article{,
    title = {epinowcast: Hierarchical nowcasting of right censored epidemological counts},
    author = {Sam Abbott},
    journal = {Zenodo},
    year = {2021},
    doi = {10.5281/zenodo.5637165},
  }

## Documentation

```{r, child = "docs/chunks/documentation.Rmd", eval = TRUE}
```

## Key files and folders

Folder/File | Purpose
---|---
[`writeup`](writeup/) | Summary paper and additional supplementary information as `Rmarkdown` documents.
[`_targets.Rmd`](_targets.Rmd) | Analysis workflow for interactive use.
[`R`](R/) | R functions used in the analysis and for evaluation.
[`data`](data/) | Input data and summarised output generated by steps in the analysis.
[`analyses`](analyses/) | Ad-hoc analyses not part of the overarching workflow. This includes a synthetic case study and a simplified example using Germany hospitalisation data.
[`.devcontainer`](.devcontainer/) | Resources for reproducibility using `vscode` and `docker`.

## Dependencies

All dependencies can be installed using the following,

```{r}
remotes::install_dev_deps()
```

Alternatively a docker [container](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/.devcontainer/Dockerfile) and [image](https://github.com/epiforecasts/eval-germany-sp-nowcasting/pkgs/container/eval-germany-sp-nowcasting) is provided. An easy way to make use of this is using the Remote development extension of `vscode`.

## Analyses

This analysis in this repository has been implemented using the [`targets`](https://docs.ropensci.org/targets/) package and associated packages. The workflow is defined in [`_targets.md`](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/_targets.md) and can be explored interactively using [`_targets.Rmd`](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/_targets.Rmd) `Rmarkdown` document. The workflow can be visualised as the following graph.

![](figures/targets-graph.png)

This complete analysis can be recreated using the following (note this may take quite some time even with a fairly large amount of available compute),

```{bash}
bash bin/update-targets.sh
```

Alternative the following `targets` functions may be used to interactively explore the workflow:

- Run the workflow sequentially.

```{r}
targets::tar_make()
```

- Run the workflow using all available workers.

```{r}
targets::tar_make_future(workers = future::availableCores())
```

- Explore a graph of the workflow.

```{r}
targets::tar_visnetwork(targets_only = TRUE)
```

Watch the workflow as it runs in a `shiny` app.

```{r}
targets::tar_watch(targets_only = TRUE)
```

To use our archived version of the interim results (and so avoid long run times) use the following to download it. Note that this process has not been rigorously tested across environments and so may not work seamlessly).

```{r}
source(here::here("R", "targetss-archive.R"))
get_targets_archive()
```

### 2022-05-17 _targets.Rmd
Initial commit

---
title: "Analysis pipeline: Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany"
author: "Sam Abbott"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Pipeline

The analysis pipeline for this work can be regenerated by rendering this file,

```{r, eval = FALSE}
rmarkdown::render("_targets.Rmd")
```

The pipeline can then be run using,

```{r, eval = FALSE}
tar_make()
```

The complete pipeline can be visualised using,

```{r, eval = FALSE}
tar_visnetwork()
```

Alternatively the pipeline can be explored interactively using this notebook or updated programmatically using the scripts in `bin`. We also provide an archived version of our `targets` workflow if only wanting to reproduce sections of our analysis. This can be downloaded using the following,

```{r, eval = FALSE}
source(here::here("R", "targets-archive.R"))
get_targets_archive()
```

# Setup

Set up the workflow pipeline and options. We first load the `targets` package and remove the potentially outdated workflow.

```{r}
library(targets)
library(tarchetypes)
library(data.table)
library(epinowcast)
library(ggplot2)
library(purrr, quietly = TRUE)
library(here)
library(future)
library(future.callr)
tar_unscript()
```

We now define shared global options across our workflow and load R functions from the `R` folder.

```{targets globals, tar_globals = TRUE}
library(targets)
library(tarchetypes)
library(cmdstanr)
library(data.table)
library(epinowcast)
library(ggplot2)
library(purrr, quietly = TRUE)
library(here)
library(future)
library(future.callr)
plan(callr)
functions <- list.files(here("R"), full.names = TRUE)
walk(functions, source)
rm("functions")
set_cmdstan_path()  
tar_option_set(
  packages = c("data.table", "epinowcast", "scoringutils", "ggplot2", "purrr",
               "cmdstanr", "here"),
  deployment = "worker",
  memory = "transient",
  workspace_on_error = TRUE,
  error = "continue",
  garbage_collection = TRUE
)
```

# Ingest data and stratify by location and report date

* Watch the URL at which hospitalisation data is hosted in order to trigger an update to the workflow when new data is uploaded

```{targets hospitalisation_url}
tar_url(
  hospitalisation_url,
  "https://raw.githubusercontent.com/KITmetricslab/hospitalization-nowcast-hub/main/data-truth/COVID-19/COVID-19_hospitalizations_preprocessed.csv" # nolint
)
```

* Download and process hospitalisation data. Also download and combined data on public holidays in Germany.

```{targets hospitalisations, tar_simple = TRUE}
get_germany_hospitalisations(url = hospitalisation_url)
```

* Define date to start nowcasting

```{targets start_date, tar_simple = TRUE}
as.Date("2021-10-01")
```

* Define dates to nowcast.

```{targets nowcast_dates, tar_simple = TRUE}
unique(
  hospitalisations[reference_date >= start_date]$reference_date
)
```

* Define age groups

```{targets age_groups, tar_simple = TRUE}
unique(hospitalisations$age_group)
```

* Define locations to nowcast using age group data (currently limited to just national level).

```{targets locations, tar_simple = TRUE}
"DE"
```

* Define locations in which to nowcast only for the overall count and not for each age group.

```{targets other_locations, tar_simple = TRUE}
setdiff(unique(hospitalisations$location), locations)
```

* Define maximum allowed reporting delay (in days).

```{targets max_report_delay, tar_simple = TRUE}
40
```

* Define hospitalisations by date of report. Note that here the target is defined to never update if data exists for that nowcast date. This is because the truth date for previous days can change slightly over time as negative cases are adjusted for leading to spurious refits.

```{targets hospitalisations_by_date_report}
tar_target(
  hospitalisations_by_date_report,
  enw_retrospective_data(
    hospitalisations,
    rep_date = nowcast_dates,
    ref_days = max_report_delay
  )[, nowcast_date := nowcast_dates],
  map(nowcast_dates),
  iteration = "list",
  cue = tar_cue(mode = "never")
)
```

* Define latest available data.

```{targets latest_hospitalisations, tar_simple = TRUE}
 enw_latest_data(hospitalisations)
```

* Make latest observations into 7 day incidence.

```{targets, latest_7day_hospitalisations, tar_simple = TRUE}
copy(latest_hospitalisations)[,
  confirm := frollsum(confirm, n =  7), by = c("age_group", "location")
][!is.na(confirm)]
```

* Define "complete" data (i.e more than 28 days have passed since first reported)

```{targets, complete_hospitalisations, tar_simple = TRUE}
latest_hospitalisations[reference_date < (max(reference_date) - 28)]
```

* Similarly, define "complete" 7 day hospitalisation incidence.

```{targets, complete_7day_hospitalisations, tar_simple = TRUE}
latest_7day_hospitalisations[reference_date < (max(reference_date) - 28)]
```

* Plot reporting delay percentage by date, location, and age group.

# Fit models and produce nowcasts

## Model and model settings

* Compile the model for multithread use.

```{targets, epinowcast_model}
tar_file(
  epinowcast_model,
  compile_model(),
  deployment = "main"
)
```
* Define stan settings shared across models and used for fitting

```{targets epinowcast_settings, tar_simple = TRUE}
list(
  save_warmup = FALSE,
  output_loglik = FALSE,
  pp = FALSE,
  iter_warmup = 1000,
  iter_sampling = 1000,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 1,
  adapt_delta = 0.95,
  show_messages = FALSE,
  refresh = 0
)
```

## Priors

* Define a set of uninformed priors using the defaults from `epinowcast`.

```{targets uninformed_priors, tar_simple = TRUE}
enw_priors()
```

* Define a set of observations to use as a source for informed priors. Here we use overall national hospitalisations and data from the 1st of July 2021.

```{targets prior_obs}
tar_target(
  prior_obs,
  enw_retrospective_data(
    hospitalisations[location == "DE"][age_group == "00+"],
    rep_date = as.Date("2021-07-01"), ref_days = max_report_delay
  ),
  cue = tar_cue(mode = "never")
)
```

* Fit a model to national overall hospitalisations only and extract posterior estimates for the delay distribution and overdispersion to use as informed priors for other models by assuming a normal distribution with their posterior standard deviations scaled by 10 times.

```{targets priors, tar_simple = TRUE} 
do.call(prior_epinowcast, c(
  list(
    prior_obs, max_delay = max_report_delay, scale = 10,
    priors = uninformed_priors
  ),
  epinowcast_settings
))
```

## Models

* Intercept only model.

```{targets fixed_nowcast}
tar_target(
  fixed_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report,
    tar_loc = locations,
    model = fixed_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations),
  cue = tar_cue(mode = "never")
)
```

* Intercept model with day of the week reporting effects.

```{targets dow_nowcast}
tar_target(
  dow_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report,
    tar_loc = locations,
    model = dow_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations),
  cue = tar_cue(mode = "never")
)
```

* Age group random effect model with day of the week effect reporting effects.

```{targets age_nowcast}
tar_target(
  age_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report,
    tar_loc = locations,
    model = age_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations),
  cue = tar_cue(mode = "never")
)
```

* Age group random effect model with a weekly random walk and a day of the week reporting effect.

```{targets week_nowcast}
tar_target(
  week_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report,
    tar_loc = locations,
    model = week_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations),
  cue = tar_cue(mode = "never")
)
```

* Age group random effect model with an age group specific weekly random walk,  and a day of the week reporting effect.

```{targets age_week_nowcast}
tar_target(
  age_week_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report,
    tar_loc = locations,
    model = age_week_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations),
  cue = tar_cue(mode = "never")
)
```

* Independent age group model with a weekly random walk and a day of the week reporting model.

```{targets independent_nowcast}
tar_target(
  independent_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report[age_group == age_groups],
    tar_loc = locations,
    model = independent_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations, age_groups),
  cue = tar_cue(mode = "never")
)
```

* Independent age group model with a weekly random walk and a day of the week reference and reporting model.

```{targets independent_ref_dow_nowcast}
tar_target(
  independent_ref_dow_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report[age_group == age_groups],
    tar_loc = locations,
    model = independent_ref_dow_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations, age_groups),
  cue = tar_cue(mode = "never")
)
```

* Model for locations without age groups. As for the previous model this has a weekly random walk and a day of the week reporting and reference model. This was updated on the 6th of December 2021 from the independent model without a day of the week effect for reference date (to enable this note that we have overridden the default `cue` settings here so that once a model has been fit it is never run again for that combination of locations and dates regardless of upstream changes).

```{targets overall_only_nowcast}
tar_target(
  overall_only_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report[age_group == "00+"],
    tar_loc = other_locations,
    model = independent_ref_dow_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, other_locations),
  cue = tar_cue(mode = "never")
)
```

# Postprocess

* Combine nowcasts from each model

```{targets combined_nowcasts, tar_simple = TRUE}
rbindlist(list(
  fixed_nowcast,
  dow_nowcast,
  age_nowcast,
  week_nowcast,
  age_week_nowcast,
  independent_nowcast,
  overall_only_nowcast,
  independent_ref_dow_nowcast
), use.names = TRUE, fill = TRUE)[,
   model := factor(
    model,
    levels = c("Reference: Fixed, Report: Fixed",
               "Reference: Fixed, Report: Day of week",
               "Reference: Age, Report: Day of week",
               "Reference: Age and week, Report: Day of week",
               "Reference: Age and week by age, Report: Day of week",
               "Independent by age, Reference: Week, Report: Day of week",
               "Independent by age, Reference: Week and day of week, Report: Day of week")
   )
  ][, id := 1:.N]
```

* Extract summarised daily nowcast. As a temporary measure here we adjust quantiles that are more than 25% of the median when there is evidence of a fitting issue (based on divergent transistions and R hat values).

```{targets summarised_nowcast, tar_simple = TRUE}
combined_nowcasts |> 
  adjust_posteriors(
    target = "daily", 
    max_ratio = 0.25, 
    rhat_bound = 1.1,
    per_dt_bound = 0.2
) |> 
  unnest_nowcasts(target = "daily") 
```

* Extract summarised 7 day nowcast. As a temporary measure here we adjust quantiles that are more than 25% of the median when there is evidence of a fitting issue (based on divergent transistions and R hat values).

```{targets summarised_7day_nowcast, tar_simple = TRUE}
combined_nowcasts |> 
  adjust_posteriors(
    target = "seven_day", 
    max_ratio = 0.25, 
    rhat_bound = 1.1,
    per_dt_bound = 0.2
) |> 
  unnest_nowcasts(target = "seven_day") 
```

* Save daily nowcasts stratified by nowcasting date.

```{targets save_daily_nowcasts}
tar_file(
  save_daily_nowcasts,
  summarised_nowcast[nowcast_date == nowcast_dates] |>
    save_csv(
      filename = paste0(nowcast_dates, ".csv"),
      path = here("data/nowcasts/daily"),
      allow_empty = FALSE
    ),
  map(nowcast_dates),
  cue = tar_cue(mode = "never")
)
```

* Save 7 day nowcasts stratified by nowcasting date

```{targets save_7day_nowcasts}
tar_file(
  save_7day_nowcasts,
  summarised_7day_nowcast[nowcast_date == nowcast_dates] |>
    save_csv(
      filename = paste0(nowcast_dates, ".csv"),
      path = here("data/nowcasts/seven_day"),
      allow_empty = FALSE
    ),
  map(nowcast_dates),
  cue = tar_cue(mode = "never")
)
```

* Define and format the hierarchical nowcast for submission. As a temporary measure here we adjust quantiles that are more than 25% of the median when there is evidence of a fitting issue (based on divergent transistions and R hat values).

```{targets hierarchical_submission_nowcast}
tar_target(
  hierarchical_submission_nowcast,
  age_week_nowcast[nowcast_date == nowcast_dates] |>
    adjust_posteriors(
      target = "seven_day", 
      max_ratio = 0.25, 
      rhat_bound = 1.1,
      per_dt_bound = 0.2
    ) |>
    select_var("seven_day") |>
    rbindlist() |>
    format_for_submission(),
  map(nowcast_dates),
  iteration = "list",
  cue = tar_cue(mode = "never")
)
```

* Format and save 7 day hierarchical nowcasts for hub submission.

```{targets save_hierarchical_submission}
tar_file(
  save_hierarchical_submission,
  save_csv(
    hierarchical_submission_nowcast,
      filename = paste0(nowcast_dates, ".csv"),
      path = here("data/nowcasts/submission/hierarchical")
  ),
  map(hierarchical_submission_nowcast, nowcast_dates)
)
```

* Define and format the independent nowcast for submission. As a temporary measure here we adjust quantiles that are more than 25% of the median when there is evidence of a fitting issue (based on divergent transistions and R hat values). On the 6th of December the submission model was updated to include a reference day of the week effect.

```{targets independent_submission_nowcast}
tar_target(
  independent_submission_nowcast,
  rbind(
    independent_ref_dow_nowcast[nowcast_date == nowcast_dates],
    overall_only_nowcast[nowcast_date == nowcast_dates]
  ) |> 
   adjust_posteriors(
      target = "seven_day", 
      max_ratio = 0.25, 
      rhat_bound = 1.1,
      per_dt_bound = 0.2
    ) |>
    select_var("seven_day") |>
    rbindlist() |>
    format_for_submission(),
  map(nowcast_dates),
  iteration = "list",
  cue = tar_cue(mode = "never")
)
```

* Format and save 7 day independent nowcasts for hub submission.

```{targets save_independent_submission}
tar_file(
  save_independent_submission,
  save_csv(
    independent_submission_nowcast,
    filename = paste0(nowcast_dates, ".csv"),
    path = here("data/nowcasts/submission/independent")
  ),
  map(independent_submission_nowcast, nowcast_dates)
)
```

* Save latest hospitalisation data

```{targets save_latest_daily_hospitalisations}
tar_file(
  save_latest_daily_hospitalisations,
  save_csv(
    latest_hospitalisations,
    filename = paste0("daily.csv"),
    path = here("data/observations")
  )
)
```

* Save 7 day hospitalisation data

```{targets save_latest_7day_hospitalisations}
tar_file(
  save_latest_7day_hospitalisations,
  save_csv(
    latest_7day_hospitalisations,
    filename = paste0("seven_day.csv"),
    path = here("data/observations")
  )
)
```

# Evaluation

* Extract and save model fitting diagnostics.

```{targets diagnostics, tar_simple = TRUE}
combined_nowcasts[, c("daily", "seven_day") := NULL]
```

```{targets save_diagnostics}
list(
  tar_file(
    save_all_diagnostics,
      save_csv(
        diagnostics,
        filename = "all.csv",
        path = here("data/diagnostics")
      )
  ),
  tar_file(
    save_high_rhat_diagnostics,
      save_csv(
        diagnostics[max_rhat > 1.05],
        filename = "high-rhat.csv",
        path = here("data/diagnostics")
      )
  ),
  tar_file(
    save_high_divergent_transitions,
      save_csv(
        diagnostics[per_divergent_transitions > 0.1],
        filename = "high-divergent-transitions.csv",
        path = here("data/diagnostics")
      )
  ),
  tar_file(
    save_failures,
      save_csv(
        diagnostics[failure == TRUE],
        filename = "fitting-failed.csv",
        path = here("data/diagnostics")
      )
  )
)
```

* Extract and save model run-time at the national level aggregated across age-group.

```{targets save-run-time}
tar_file(
  save_run_time,
  combined_nowcasts |>
    summarise_runtimes() |>
    save_csv(
      filename = "run-times.csv",
      path = here("data/diagnostics")
    )
)
```

* Filter nowcasts to only include those with "complete" (more than 28 days of reports) data for all age groupsand with horizons between 0 days and -7 days from the nowcast date.

```{targets scored_nowcasts, tar_simple = TRUE}
summarised_nowcast[location == locations][
  reference_date < (max(nowcast_date) - 28)][,
  holiday := NULL][,
  horizon := as.numeric(as.Date(reference_date) - nowcast_date)][
  horizon >= -7
]
```

* Score daily nowcasts overall, by age group, by horizon, by nowcast date, and by reference date on both the natural and log scales (corresponding to absolute and relative scoring). These summarised scores are then saved to `data/scores`.

```{targets, score}
tar_map(
  list(score_by = list(
    "overall", "age_group", "horizon", "reference_date", "nowcast_date"
  )),
  tar_target(
    scores,
    enw_score_nowcast(
      scored_nowcasts, complete_hospitalisations, 
      summarise_by = drop_string(c(score_by, "model"), "overall"),
      log = FALSE
    )
  ),
  tar_target(
    log_scores,
    enw_score_nowcast(
      scored_nowcasts, complete_hospitalisations, 
      summarise_by = drop_string(c(score_by, "model"), "overall"),
      log = TRUE
    )
  ),
  tar_file(
    save_scores,
    save_csv(
      rbind(scores[, scale := "natural"], log_scores[, scale := "log"]),
      filename = paste0(paste(score_by, sep = "-"), ".csv"),
      path = here("data/scores")
    )
  )
)
```

# Visualise

**Currently these are not produced. See the real time report instead**

* Plot most recent daily nowcast by location, age group, and model.

```{r plot-latest-nowcast, eval = FALSE}
tar_target(
  plot_latest_nowcast,
  enw_plot_nowcast_quantiles(
    summarised_nowcast[nowcast_date == max(nowcast_date)][
                       location == locations][
                       reference_date >= (nowcast_date - 28)]
  ) +
  facet_grid(vars(age_group), vars(model), scales = "free_y"),
  map(locations),
  iteration = "list"
)
```

* Plot most recent seven day nowcast by location, age group, and model.

```{r plot-latest-7day-nowcast, eval = FALSE}
tar_target(
  plot_latest_7day_nowcast,
  enw_plot_nowcast_quantiles(
    summarised_7day_nowcast[nowcast_date == max(nowcast_date)][
                            location == locations][
                            reference_date >= (nowcast_date - 28)]
  ) +
  facet_grid(vars(age_group), vars(model), scales = "free_y"),
  map(locations),
  iteration = "list"
)
```

* Plot daily nowcasts at each horizon from the date of the nowcast to 7 days previously by location, age group and model.

```{r plot_nowcast_horizon, eval = FALSE}
tar_map(
  list(horizons = 0:7),
  tar_target(
    plot_nowcast_horizon,
    enw_plot_nowcast_quantiles(
      scored_nowcasts[location == locations][horizon == -horizons][,
                      confirm := NA],
      latest_obs = latest_hospitalisations[
        location == locations][
        reference_date >= min(scored_nowcasts$reference_date)][
        reference_date <= max(scored_nowcasts$reference_date)
      ],
      log = TRUE
    ) +
    facet_grid(vars(age_group), vars(model), scales = "free_y"),
    map(locations),
    iteration = "list"
  )
)
```

* Plot 7 day nowcasts at each horizon from the date of the nowcast to 7 days previously by location, age group and model.

```{r plot_7day_nowcast_horizon, eval = FALSE}
tar_map(
  list(horizons = 0:7),
  tar_target(
    plot_7day_nowcast_horizon,
    enw_plot_nowcast_quantiles(
      summarised_7day_nowcast[
        reference_date < (max(nowcast_date) - 28)][,
        holiday := NULL][,
        horizon := as.numeric(as.Date(reference_date) - nowcast_date)][
        location == locations][
        horizon == -horizons][,
        confirm := NA],
      latest_obs = latest_7day_hospitalisations[
        location == locations][
        reference_date >= min(scored_nowcasts$reference_date)][
        reference_date <= max(scored_nowcasts$reference_date)
      ],
      log = TRUE
    ) +
    facet_grid(vars(age_group), vars(model), scales = "free_y"),
    map(locations),
    iteration = "list"
  )
)
```

* Plot daily nowcasts for locations without age groups stratified nowcasts.

* Plot 7 day nowcasts for locations without age group stratified nowcasts.

* Plot scores relative to the baseline by location and age group on both the natural and log scale.

### 2022-05-17 docs/chunks/documentation.Rmd
Initial commit

Document | Purpose
---|---
[Summary](https://epiforecasts.io/eval-germany-sp-nowcasting/) | A summary of this work.
[Paper](https://epiforecasts.io/eval-germany-sp-nowcasting/paper.pdf) | The academic paper write up of this work.
[Supplementary information](https://epiforecasts.io/eval-germany-sp-nowcasting/si.html) | The supplementary information for the write up of this work.
[Real-time model evaluation](https://epiforecasts.io/eval-germany-sp-nowcasting/real-time/) | A report visualising and evaluating nowcasts from the various model configurations considered here in real-time.
[Real-time method evaluation](https://epiforecasts.io/eval-germany-sp-nowcasting/real-time-method-comparison/) | A report visualising and evaluating nowcasts from the various methods (from this project and other groups) submitted to the [Germany nowcasting hub](https://covid19nowcasthub.de) in real-time.
[Project README](https://github.com/epiforecasts/eval-germany-sp-nowcasting) | Overarching project README. Includes links to resources, a summary of key files, and reproducibility information.
[Analysis pipeline](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/_targets.md) | The `targets` based analysis pipeline.
[Analysis archive](https://github.com/epiforecasts/eval-germany-sp-nowcasting/releases/tag/latest) | An archived version of the `_targets` directory. Download using `get_targets_archive()`.
[Data](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/data/README.md) | Documentation for input data and summarised output from the analysis.
[bin](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/bin/README.md) | Documentation for orchestration of nowcast estimation, publishing, and archiving.
[News](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/NEWS.md) | Dated development notes.
[epinowcast](https://epiforecasts.io/epinowcast/index.html) | The documentation for `epinowcast` the R package used to implement the models evaluated here. See this [case study](https://epiforecasts.io/epinowcast/articles/germany-age-stratified-nowcasting.html) for a simplified version of this analysis.
[Germany nowcasting hub](https://covid19nowcasthub.de)  | The homepage (containing a dashboard and information) for the Germany nowcasting hub project to which nowcasts from this evaluation are submitted along with others produced by independent groups.

### 2022-05-17 docs/chunks/introduction.Rmd
Initial commit

COVID-19 hospitalisations in Germany are released by date of positive test rather than by date of admission. This has some advantages when they are used as a tool for surveillance as these data are closer to the date of infection and so easier to link to underlying transmission dynamics and public health interventions. Unfortunately, however, when released in this way the latest data are right-censored meaning that final hospitalisations for a given day are initially underreported. This issue is often found in data sets used for the surveillance of infectious diseases and can lead to delayed or biased decision making. Fortunately, when data from a series of days is available we can estimate the level of censoring and provide estimates for the truncated hospitalisations adjusted for truncation with appropriate uncertainty. This is usually known as a nowcast.

### 2022-05-17 docs/real-time-method-comparison/index.Rmd
Initial commit

---
title: "Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany"
subtitle: "Real-time comparison to other nowcasting methods"
author: Sam Abbott 
bibliography: ../../writeup/library.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 4
    includes:
      before_body: ../header.html
      after_body: ../footer.html
---

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, dpi = 300, warning = FALSE)
library(knitr)
library(here)
library(data.table)
library(DT)
library(janitor)
library(purrr)
library(ggplot2)

source(here("R", "utils.R"))
source(here("R", "plot.R"))
source(here("R", "get-hub-forecasts.R"))
source(here("R", "evaluation.R"))
```

**This is a preliminary real-time comparison between the semi-parametric nowcasting method evaluated in this project and the other methods submitted to the [German nowcasting hub](https://covid19nowcasthub.de). It is not the official prespecified evaluation the protocol for which can be found [here](https://osf.io/mru75/). See our [news](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/NEWS.md) for details of any changes to this evaluation.**

# Introduction

In this real time evaluation report we provide preliminary visualisations, evaluation, and exploration of our nowcasting methodology [@epinowcast] for COVID-19 hospitalisations in Germany by date of postive test compared to the other nowcasting methods submitted to the [Germany nowcasting hub](https://covid19nowcasthub.de). For more details of our methodology and the specifics of the models shown in this report please see our [project summary](https://epiforecasts.io/eval-germany-sp-nowcasting/), our accompanying [paper](https://epiforecasts.io/eval-germany-sp-nowcasting/paper.pdf), and the [German nowcasting hub](https://covid19nowcasthub.de). This report complements our [real-time evaluation](https://epiforecasts.io/eval-germany-sp-nowcasting/real-time/) of models implemented using our methodology. It is updated each day (at roughly 8:00 GMT) as new data and nowcasts become available and may evolve over time. See our [news section](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/NEWS.md) for a list of dated changes.

We first visualise current nowcasts across methods and age groups at the national level as well as overall at the subnational level. We also plot nowcasts at the date of estimation across sequential nowcasts allowing us to summarise the performance of multiple nowcasts on a single plot, again for age groups at the national level and overall at the subnational level.

To quantify comparative performance we make use of proper scoring rules [@scoringutils] on both the natural and log scales (corresponding to absolute and relative performance)  to observed data reported at least 28 days ago aggregating scores first across all targets and then stratifying in turn by age group, nowcast horizon, date of postive test, date of report, and location. See Bosse et al. [@scoringutils] for information on how to interpret these scores. Unlike our [real-time evaluation of our models](https://epiforecasts.io/eval-germany-sp-nowcasting/real-time/) here we evaluate performance using 7 day incidence (as this is the target of interest for the [German nowcasting hub](https://covid19nowcasthub.de)) across the complete 28 days of the nowcast and for all locations and age groups available.

The code for this report can be found [here](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/docs/real-time-method-comparison/index.Rmd) and the data that it uses can be found [here](https://github.com/KITmetricslab/hospitalization-nowcast-hub/tree/main/data-processed) (though you may find it easier to access using the [data accessing code](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/R/get-hub-forecasts.R) that supports this report.

# Visualisation

In the following sections nowcasts are visualised for the latest estimation date and then by estimation date for all models considered. Visualising a nowcast from a single estimation date corresponds to plotting a real-time nowcast whilst plotting across estimation dates for that date is useful for understanding the performance of nowcasting models across a number of nowcasts in a concise way.

## Nowcasts

Nowcasts based on the latest available data by age group on the national level and overall on the subnational level. Due to a technical issue resolved on the 5th of January 2022 the 10% quantile is excluded for all nowcasts so that they are evaluated on an even footing.

```{r load-nowcasts}
start_using_memoise(".cache")
nowcast_dates <- seq(as.Date("2021-11-22"), Sys.Date(), by = "day")
hub_nowcasts <- get_hub_forecasts(
  "KITmetricslab/hospitalization-nowcast-hub",
  dates = nowcast_dates
)
hub_nowcasts <- hub_nowcasts[quantile != 0.1]
hub_nowcasts <- format_hub_nowcasts(hub_nowcasts)

latest_seven_day_hosp <- load_obs(here("data", "observations", "seven_day.csv"))
complete_seven_day_hosp <- latest_seven_day_hosp[
  reference_date < (max(reference_date) - 28)
][]
```

### National

```{r latest-national-nowcast, fig.width = 24, fig.height = 16}
plot_nowcast(
  hub_nowcasts[location == "DE"][
               nowcast_date == max(nowcast_date)] |>
  map_to_dummy_quantiles(),
  latest_seven_day_hosp[location == "DE"],
  max_delay = 28
) +
  facet_grid(vars(age_group), vars(model), scales = "free_y")
```

### Sub-national

```{r latest-subnational-nowcast, fig.width = 24, fig.height = 24}
plot_nowcast(
  hub_nowcasts[!(location == "DE")][
                 nowcast_date == max(nowcast_date)] |>
  map_to_dummy_quantiles(),
  latest_seven_day_hosp[!(location == "DE")][age_group == "00+"],
  max_delay = 28
) +
  facet_grid(vars(location), vars(model), scales = "free_y")
```

## Nowcasts at estimation date

Nowcast estimates at the date of estimation for sequential nowcasts by age group at the national level and overall at the subnational level.

### National

```{r horizon-national-nowcast, fig.width = 24, fig.height = 16}
plot_nowcast(
  hub_nowcasts[location == "DE"][
               horizon == 0] |>
  map_to_dummy_quantiles(),
  latest_seven_day_hosp[location == "DE"]
) +
  facet_grid(vars(age_group), vars(model), scales = "free_y")
```

### Subnational

```{r horizon-subnational-nowcast, fig.width = 24, fig.height = 24}
plot_nowcast(
  hub_nowcasts[!(location == "DE")][
               horizon == 0] |>
  map_to_dummy_quantiles(),
  latest_seven_day_hosp[!(location == "DE")][age_group == "00+"]
) +
  facet_grid(vars(location), vars(model), scales = "free_y")
```

# Evaluation

In this section we evaluate and compare the performance of each nowcasting method using proper scoring rules [@scoringutils] on both the natural and the log scale to observed data reported at least 28 days ago. This corresponds to evaluating absolute and relative error. Unlike our [real-time evaluation of our models](https://epiforecasts.io/eval-germany-sp-nowcasting/real-time/) here we evaluate performance using 7 day incidence (as this is the target of interest for the [German nowcasting hub](https://covid19nowcasthub.de)) across the complete 28 days of the nowcast and for all locations and age groups available. We explore overall scores as well as scores stratified by age group, by nowcast horizon, and by date of postive test, by report date, and by location. We drop the SZ-hosp_nowcast from all stratified evaluation due to its poor overall performance compared to other approaches and because this masks other meaningful between model differences. We also drop the first nowcast (from the 24th of November 2021) from the RKI-weekly_report model from all stratified evaluatioon as this was the only nowcast for which estimates were made for the day of nowcast and the subsequent two days from this model, performance was poor, and this masked other meaningful between model differences.

## Overall nowcast model scores

```{r calc-overall-scores}
scores <- score_nowcast(hub_nowcasts, complete_seven_day_hosp)
hub_nowcasts <- hub_nowcasts[!model %in% "SZ-hosp_nowcast"]
hub_nowcasts <- hub_nowcasts[
  !(model %in% "RKI-weekly_report" & nowcast_date == "2021-11-24")
]
```

```{r plot-overall-scores, fig.height = 4, fig.width = 8}
plot_scores(
  scores, y = interval_score, x = model, col = model,
  fill = model, group = model
) +
  facet_wrap(vars(scale), scales = "free_x") +
  labs(x = "Model") +
  theme(legend.position = "none") +
  coord_flip()
```

### Natural scale scores (absolute)

```{r overall-scores}
fancy_datatable(scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-overall-scores}
fancy_datatable(scores[scale == "log"][, scale := NULL])
```

## Nowcast model scores by horizon

```{r calc-horizon-scores}
horizon_scores <- score_nowcast(
  hub_nowcasts, complete_seven_day_hosp, by = c("horizon", "model")
)
```

```{r plot-horizon, fig.height = 4, fig.width = 8}
plot_scores(
  horizon_scores, y = interval_score, x = horizon, col = model, fill = model,
  group = model
) +
  facet_wrap(vars(scale), scales = "free_y") +
  labs(x = "Nowcast horizon (with 0 being the date of nowcast)")
```

### Natural scale scores (absolute)

```{r horizon-scores}
fancy_datatable(horizon_scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-horizon-scores}
fancy_datatable(horizon_scores[scale == "log"][, scale := NULL])
```

## Nowcast model scores by age group

```{r calc-age-group-scores}
age_group_scores <- score_nowcast(
  hub_nowcasts, complete_seven_day_hosp, by = c("age_group", "model")
)
```

```{r plot-age-group, fig.height = 4, fig.width = 8}
plot_scores(
  age_group_scores, y = interval_score, x = age_group, col = model,
  fill = model, group = model
) +
  facet_wrap(vars(scale), scales = "free_y") +
  labs(x = "Age group")
```

### Natural scale scores (absolute)

```{r age-group-scores}
fancy_datatable(age_group_scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-age-group-scores}
fancy_datatable(age_group_scores[scale == "log"][, scale := NULL])
```

## Nowcast model scores by date of positive test

```{r calc-ref-date-scores}
reference_date_scores <- score_nowcast(
  hub_nowcasts, complete_seven_day_hosp, by = c("reference_date", "model")
)
```

```{r plot-ref-date, fig.height = 4, fig.width = 8}
plot_scores(
  reference_date_scores, y = interval_score, x = reference_date, col = model,
  fill = model, group = model
) +
  facet_wrap(vars(scale), scales = "free_y") +
  labs(x = "Reference date")
```

### Natural scale scores (absolute)

```{r ref-date-scores}
fancy_datatable(reference_date_scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-ref-date-scores}
fancy_datatable(reference_date_scores[scale == "log"][, scale := NULL])
```

## Nowcast model scores by report date

```{r calc-rep-date-scores}
report_date_scores <- score_nowcast(
  hub_nowcasts, complete_seven_day_hosp, by = c("nowcast_date", "model")
)
```

```{r plot-rep-date, fig.height = 4, fig.width = 8}
plot_scores(
  report_date_scores, y = interval_score, x = nowcast_date, col = model,
  fill = model, group = model
) +
  facet_wrap(vars(scale), scales = "free_y") +
  labs(x = "Report date")
```

### Natural scale scores (absolute)

```{r rep-date-scores}
fancy_datatable(report_date_scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-rep-date-scores}
fancy_datatable(report_date_scores[scale == "log"][, scale := NULL])
```

## Nowcast model scores by location

```{r calc-location-scores}
location_scores <- score_nowcast(
  hub_nowcasts, complete_seven_day_hosp, by = c("location", "model")
)
```

```{r plot-location-scores, fig.height = 8, fig.width = 8}
plot_scores(
  location_scores, y = interval_score, x = location, col = model,
  fill = model, group = model
) +
  facet_wrap(vars(scale), scales = "free_x") +
  labs(x = "Location") +
  coord_flip()
```

### Natural scale scores (absolute)

```{r location-scores}
fancy_datatable(location_scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-location-scores}
fancy_datatable(location_scores[scale == "log"][, scale := NULL])
```

# References

### 2022-05-17 docs/real-time/index.Rmd
Initial commit

---
title: "Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany"
subtitle: "Real-time evaluation and current nowcasts"
author: Sam Abbott 
bibliography: ../../writeup/library.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 4
    includes:
      before_body: ../header.html
      after_body: ../footer.html
---

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, dpi = 300, warning = FALSE)
library(knitr)
library(here)
library(data.table)
library(DT)
library(janitor)
library(purrr)
library(ggplot2)

source(here("R", "utils.R"))
source(here("R", "plot.R"))
```

# Introduction

In this real time evaluation report we provide preliminary visualisations, evaluation, and exploration of our nowcasting methodology [@epinowcast] for COVID-19 hospitalisations in Germany by date of postive test. For more details of our methodology and the specifics of the models shown in this report please see our [project summary](https://epiforecasts.io/eval-germany-sp-nowcasting/) and our accompanying [paper](https://epiforecasts.io/eval-germany-sp-nowcasting/paper.pdf) for full method details. This report is updated each day (at roughly 9:00 GMT) as new data and nowcasts become available and may evolve over time. See our [news section](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/NEWS.md) for a list of dated changes.

We first visualise current nowcasts across models and age groups at the national level as well as overall at the subnational level. We also plot nowcasts at the date of estimation across sequential nowcasts allowing us to summarise the performance of multiple nowcasts on a single plot, again for age groups at the national level and overall at the subnational level.

To quantify comparative performance we make use of proper scoring rules [@scoringutils] on both the natural and log scales (corresponding to absolute and relative performance) to observed data reported at least 28 days ago aggregating scores first across all targets and then stratifying in turn by age group, nowcast horizon, date of postive test, and date of report. See Bosse et al. [@scoringutils] for information on how to interpret these scores.

To explore other aspects of our models performance we highlight models that have problematic fitting diagnostics and summarise the estimation time for each model. To learn more about the model fitting diagnostics we use here see [@stan] and [@cmdstanr].

The code for this report can be found [here](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/docs/real-time/index.Rmd) and the data that it uses can be found [here](https://github.com/epiforecasts/eval-germany-sp-nowcasting/tree/main/data).

# Visualisation

In the following sections nowcasts are visualised for the latest estimation date and then by estimation date for all models considered. Visualising a nowcast from a single estimation date corresponds to plotting a real-time nowcast whilst plotting across estimation dates for that date is useful for understanding the performance of nowcasting models across a number of nowcasts in a concise way.

## Nowcasts

Nowcasts based on the latest available data by age group on the national level and overall on the subnational level.

```{r load-nowcasts}
daily_nowcasts <- load_nowcasts(here("data", "nowcasts", "daily"))
daily_nowcasts[, mean := NA][, median := NA]

seven_day_nowcasts <- load_nowcasts(here("data", "nowcasts", "seven_day"))

latest_hosp <- load_obs(here("data", "observations", "daily.csv"))
latest_seven_day_hosp <- load_obs(here("data", "observations", "seven_day.csv"))
```

### National

```{r latest-national-nowcast, fig.width = 32, fig.height = 16}
plot_nowcast(
  daily_nowcasts[location == "DE"][
                 nowcast_date == max(nowcast_date)][,
                 confirm := NA],
  latest_hosp[location == "DE"],
  max_delay = 28
) +
  facet_grid(vars(age_group), vars(model), scales = "free_y")
```

### Sub-national

```{r latest-subnational-nowcast, fig.width = 16, fig.height = 16}
plot_nowcast(
  daily_nowcasts[!(location == "DE")][
                 nowcast_date == max(nowcast_date)][,
                 confirm := NA],
  latest_hosp[!(location == "DE")][age_group == "00+"],
  max_delay = 28
) +
  facet_wrap(vars(location), scales = "free_y")
```

## Nowcasts at estimation date

Nowcast estimates at the date of estimation for sequential nowcasts by age group at the national level and overall at the subnational level.

### National

```{r horizon-national-nowcast, fig.width = 30, fig.height = 16}
plot_nowcast(
  daily_nowcasts[location == "DE"][
                 horizon == 0][,
                 confirm := NA],
  latest_hosp[location == "DE"]
) +
  facet_grid(vars(age_group), vars(model), scales = "free_y")
```

### Subnational

```{r horizon-subnational-nowcast, fig.width = 16, fig.height = 16}
plot_nowcast(
  daily_nowcasts[!(location == "DE")][
                 horizon == 0][,
                 confirm := NA],
  latest_hosp[!(location == "DE")][age_group == "00+"]
) +
  facet_wrap(vars(location), scales = "free_y")
```

# Evaluation

In this section we evaluate and compare the performance of each nowcasting model using proper scoring rules [@scoringutils] on both the natural and the log scale to observed data reported at least 28 days ago. This corresponds to evaluating absolute and relative error. Only nowcasts from the national level are considered as sub-national nowcasts are not available for all models. For evaluation we consider only forecast targets for which the data is minimally informative which we defined as the 7 days prior to the date of the nowcast. We explore overall scores as well as scores stratified by age group, by nowcast horizon, and by date of postive test, and by report date.

## Overall nowcast model scores

```{r load-overall-scores}
scores <- load_diagnostics(here("data/scores/overall.csv"))
```

```{r plot-overall-scores, fig.height = 4, fig.width = 8}
plot_scores(
  scores, y = interval_score, x = model, col = model,
  fill = model, group = model
) +
  facet_wrap(vars(scale), scales = "free_x") +
  labs(x = "Model") +
  theme(legend.position = "none") +
  coord_flip()
```

### Natural scale scores (absolute)

```{r overall-scores}
fancy_datatable(scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-overall-scores}
fancy_datatable(scores[scale == "log"][, scale := NULL])
```

## Nowcast model scores by horizon

```{r load-horizon-scores}
horizon_scores <- load_diagnostics(here("data/scores/horizon.csv"))
```

```{r plot-horizon, fig.height = 4, fig.width = 8}
plot_scores(
  horizon_scores, y = interval_score, x = horizon, col = model, fill = model,
  group = model
) +
  facet_wrap(vars(scale), scales = "free_y") +
  labs(x = "Nowcast horizon (with 0 being the date of nowcast)")
```

### Natural scale scores (absolute)

```{r horizon-scores}
fancy_datatable(horizon_scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-horizon-scores}
fancy_datatable(horizon_scores[scale == "log"][, scale := NULL])
```

## Nowcast model scores by age group

```{r load-age-group-scores}
age_group_scores <- load_diagnostics(here("data/scores/age_group.csv"))
```

```{r plot-age-group, fig.height = 4, fig.width = 8}
plot_scores(
  age_group_scores, y = interval_score, x = age_group, col = model,
  fill = model, group = model
) +
  facet_wrap(vars(scale), scales = "free_y") +
  labs(x = "Age group")
```

### Natural scale scores (absolute)

```{r age-group-scores}
fancy_datatable(age_group_scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-age-group-scores}
fancy_datatable(age_group_scores[scale == "log"][, scale := NULL])
```

## Nowcast model scores by date of positive test

```{r load-ref-date-scores}
reference_date_scores <- load_diagnostics(
  here("data/scores/reference_date.csv")
)
```

```{r plot-ref-date, fig.height = 8, fig.width = 8}
plot_scores(
  reference_date_scores, y = interval_score, x = reference_date, col = model,
  fill = model, group = model
) +
  facet_wrap(vars(scale), scales = "free_y") +
  labs(x = "Reference date")
```

### Natural scale scores (absolute)

```{r ref-date-scores}
fancy_datatable(reference_date_scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-ref-date-scores}
fancy_datatable(reference_date_scores[scale == "log"][, scale := NULL])
```

## Nowcast model scores by report date

```{r load-rep-date-scores}
report_date_scores <- load_diagnostics(here("data/scores/nowcast_date.csv"))
```

```{r plot-rep-date, fig.height = 8, fig.width = 8}
plot_scores(
  report_date_scores, y = interval_score, x = nowcast_date, col = model,
  fill = model, group = model
) +
  facet_wrap(vars(scale), scales = "free_y") +
  labs(x = "Report date")
```

### Natural scale scores (absolute)

```{r rep-date-scores}
fancy_datatable(report_date_scores[scale == "natural"][, scale := NULL])
```

### Log scale scores (relative)

```{r relative-rep-date-scores}
fancy_datatable(report_date_scores[scale == "log"][, scale := NULL])
```

# Diagnostics

This section summarises model fitting diagnostics. Whilst most model fits successfully a small subset have issues for as yet unidentified reasons. These issues can be split into issues with convergence and issues with exploring the posterior distribution (often leading to unreliable posterior samples). Resolving these issues is an area of [active research](https://github.com/epiforecasts/epinowcast/issues/22). In a smaller subset of cases model fitting may faily entirely. In this instance a nowcast using the default model from `epinowcast` (a fixed lognormal delay) is used and a flag is raised indicating this has happened. These instances are also summarised below. Finally, model run-time is also summarised for national level nowcasts.

## Convergence issues

```{r rhat}
rhat <- load_diagnostics(here("data/diagnostics/high-rhat.csv"))
fancy_datatable(rhat)
```

## Divergent transitions

```{r dts}
dts <- load_diagnostics(here("data/diagnostics/high-divergent-transitions.csv"))
fancy_datatable(dts)
```

## Fitting failures

```{r failures}
failures <- load_diagnostics(here("data/diagnostics/fitting-failed.csv"))
fancy_datatable(failures)
```

## Run time

In this section, we summarise the run time of each nowcasting model using 2 cores each on a standard Azure virtual machine. These run-times are indicative only but the relative difference between models should hold across machines. To make the independent model run times comparable with other models we have combined the run time across all age groups. However, this model is naively parallel and so could in theory be run on separate compute nodes per age group (so as we have 7 age groups this would result in a 7 times speed up). As all models make use of the support in `stan` for within chain parallisation all models could also be estimated with significantly reduced run time with the allocation of increased compute resources. Run times are shown in minutes. Note that there is sometimes competition for resources on the server where these estimates are run and this can lead to extended run times for all models.

```{r run-time}
run_times <- load_diagnostics(here("data/diagnostics/run-times.csv"))
```

### Summary table

```{r run-time-summary}
summary_run_times <- run_times[,
  .(mean = mean(run_time_mins),
    median = median(run_time_mins),
    sd = sd(run_time_mins),
    max = max(run_time_mins),
    min = min(run_time_mins)
  ), by = c("model")
]
cols <- c("mean", "median", "sd", "max", "min")
summary_run_times <- summary_run_times[,
 (cols) := lapply(.SD, round, digits = 1), .SDcols = cols
]
fancy_datatable(summary_run_times)
```

### By model and nowcast date

```{r}
ggplot(run_times) +
  aes(x = model, y = run_time_mins, col = model, fill = model) + 
  geom_violin(alpha = 0.4) +
  geom_jitter(alpha = 0.4) +
  scale_y_log10() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "Estimation time (minutes)", x = "") +
  coord_flip()
```

# References

### 2022-05-17 writeup/paper.Rmd
Initial commit

---
title: "Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany"
output: 
  pdf_document: default
  html_document:
    keep_md: true
    toc: true
    theme: united
  word_document: default
bibliography: library.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
header-includes:
   - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, eval = TRUE,
  fig.width = 9, fig.height = 9,
  dpi = 320,
  fig.path = "figures/"
)
```

# Authors

S. Abbott (1), S. Funk (1)

Correspondence to: sam.abbott@lshtm.ac.uk

# Affiliations

1. Center for the Mathematical Modelling of Infectious Diseases, London School of Hygiene & Tropical Medicine, London WC1E 7HT, United Kingdom

# Abstract

**TODO: Summarise the project in a structured abstract (background, methods, results and conclusions)**

# Introduction

```{r, child = "../docs/chunks/introduction.Rmd"}
```

**TODO: What else has been done for nowcasting**

In this work, we aim to evaluate a series of novel semi-parametric nowcasting model formulations in real-time and provide an example workflow to allow others to do similarly using German COVID-19 hospitalisations by date of positive test at the national level both overall and by age group, and at the state level. This project is part of a [wider collaboration](https://covid19nowcasthub.de) assessing a range of nowcasting methods whilst providing an ensemble nowcast of COVID-19 Hospital admissions in Germany by date of positive test. This ensemble should be used for any policy-related work rather than the nowcasts provided in this repository. See [here](https://covid19nowcasthub.de) for more on this nowcasting collaboration.

# Methods

## Data

**TODO: Data source and processing details**

## Models

In the following sections we provide methodological and implementation details for the nowcasting framework implemented in `epinowcast` and applied here. Our approach is an extension of that proposed by Günther et al.[@gunther2021] which was itself an extension of the model proposed by Höhle and Heiden[@hohle] and implemented in the `surveillance` R package[@surveillance]. Compared to the model proposed in Günther et al.[@gunther2021], `epinowcast` adds an optional parametric assumption for the underlying delay from occurrence to report, support for jointly nowcasting multiple related datasets, a flexible formula interface allowing for the specification of a large range of models, and an efficient implementation in `stan` which makes use of sparse design matrices and within chain parallelisation to reduce runtimes [@stan; @cmdstanr].

### Decomposition into expected final notifications and report delay components

We follow the approach of Höhle and Heiden[@hohle] and consider the distribution of notifications ($n_{gtd}$) by time of occurrence ($t$) and reporting delay ($d$) conditional on the final observed count $N_{gt}$ for each dataset ($g$) such that,

\begin{equation}
  N_{gt} = \sum_{d=0}^{D} n_{gtd}
\end{equation}

where $D$ represents the maximum delay between time of occurrence and time of report which in theory could be infinite but in practice we set to a value in order to make the model identifiable and computationally feasible. This formulation means that $n_{gtd} \mid N_{gt}$ is multinomial with a probability vector ($p_{gtd}$) of length $D$ for each $t$ and $g$ that needs to be estimated at the same time as estimating the expected number of final notifications $\mathbb{E}[N_{gt}] = \lambda_{gt}$.

An alternative approach, not explored here, is to consider each $n_{gtd}$ independently at which point the model can be defined as a standard regression with the appropriate observation model and adjustment for reporting delay (i.e it becomes a Poisson or Negative Binomial regression)[@bastos]. An implementation of this approach is available in Bastos et al.[@bastos]. More work needs to be done to evaluate which of these approaches produces more accurate nowcasts for epidemiological count data.

### Expected final notifications

Here we follow the approach of Günther et al.[@gunther2021] and specify the model for expected final notifications as a first order random walk. This model can in principle be any model such as a more complex time-series approach, a gaussian process, or a mechanistic or semi-mechanistic compartmental model. Extending the flexibility of this model is an area of further work as is evaluating the benefits and tradeoffs of more complex approaches.

\begin{align}
  \log (\lambda_{gt}) &\sim \text{Normal}\left(\log (\lambda_{gt-1}) , \sigma^{\lambda}_{g} \right) \\
  \log (\lambda_{g0}) &\sim \text{Normal}\left(\log (N_{g0}), 1 \right) \\ \sigma^{\lambda}_{g} &\sim \text{Half-Normal}\left(0, 1\right)
\end{align}

### Delay distribution

Again following the approach of Günther et al.[@gunther2021] we define the delay distribution ($p_{gtd}$) as a discrete time hazard model ($h_{gtd} =\text{P} \left(\text{delay}=d|\text{delay} \geq d, W_{gtd}\right)$) but we extend this model to decompose $W_{gtd}$ into 3 components: hazard derived from a parametric delay distribution ($\gamma_{gtd}$) dependent on covariates at the date of occurrence, hazard not derived from a parametric distribution ($\delta_{gtd}$) dependent on covariates at the date of occurrence, and hazard dependent on covariates referenced to the date of report ($\epsilon_{gtd}$).

For first component ($\gamma_{gtd}$) we assume that the probability of reporting $p^{\prime}_{gtd}$ on a given date given follow a parametric distribution (in the baseline case a discretised log normal distribution) with the log mean and log standard deviation being defined using an intercept and arbitrary shared, reference date indexed, covariates with fixed ($\alpha_{i}$) and random ($\beta_{i}$) coefficients,

\begin{align}
  p^{\prime}_{gtd} &\sim \text{LogNormal} \left(\mu_{gt}, \upsilon_{gt} \right) \\
  \mu_{gt} &= \mu_0 + \alpha_{\mu} X_{\gamma} + \beta_{\mu} Z_{\gamma} \\
  \upsilon_{gt} &= \text{exp} \left( \upsilon_0 + \alpha_{\upsilon} X_{\gamma} + \beta_{\upsilon} Z_{\gamma} \right)
\end{align}

The parametric logit hazard (probability of report on a given date conditional on not already having reported) for this component of the model is then,

\begin{equation}
  \gamma_{gtd} = \text{logit} \left(\frac{p^{\prime}_{gtd}}{\left(1 -\sum^{d-1}_{d^{\prime}=0} p^{\prime}_{gtd^{\prime}} \right)} \right)
\end{equation}

The non-distributional logit hazarad components for the date of occurrence and report are then again defined using an intercept and arbitrary shared covariates with fixed ($\alpha_{i}$) and random ($\beta_{i}$) coefficients.

\begin{align}
  \delta_{gtd} &= \mu_0 + \alpha_{\delta} X_{\delta} + \beta_{\delta} Z_{\delta} \\
  \epsilon_{gtd} &= \epsilon_0 + \alpha_{\epsilon} X_{\epsilon} + \beta_{\epsilon} Z_{\epsilon}
\end{align}

The overall hazard for each group, occurrence time, and delay is then,

\begin{equation}
  \text{logit} (h_{gtd}) = \gamma_{gtd} + \delta_{gtd} +  \upsilon_{gtd},\ h_{gtD} = 1
\end{equation}

where the hazard on the final day has been assumed to be 1 in order to enforce the constraint that all reported observations are reported within the specified maximum delay. The probability of report for a given delay, occurrence date, and group is then as follows,

\begin{equation}
  p_{gt0} = h_{gt0},\ p_{gtd} = \left(1 -\sum^{d-1}_{d^{\prime}=0} p_{gtd^{\prime}} \right) \times h_{gtd}
\end{equation}

All ($\alpha_{i}$) and random ($\beta_{i}$) coefficients have standard normal priors by default with standard half-normal priors for pooled standard deviations.

### Observation model and nowcast

Expected notifications by time of occurrence ($t$) and reporting delay can now be found by multiplying expected final notifications for each $t$ with the probability of reporting for each day of delay ($p_{gtd}$). We then assume a negative binomial observation model with a joint overdispersion parameter (with a 1 over square root standard half normal prior) and produce a nowcast of final observed notifications at each occurrence time by summing posterior estimates for each observed notification for that occurrence time.

\begin{align}
  n_{gtd} \mid \lambda_{gt},p_{gtd}  &\sim \text{NB}(\lambda_{gt} \times p_{gtd}, \phi),\ t=1,...,T. \\
  \frac{1}{\phi^2} &\sim \text{Half-Normal}(0, 1) \\
  N_{gt} &= \sum_{d=0}^{D} n_{gtd}
\end{align}

### Specific model formulations

**TODO: Detail specific model formulations**

We explore two primary models and submit nowcasts from these models to the nowcasting hub. The first of these is fit independently to each data set by age and location. Hospitalisations are modelled using a random walk on the log scale. Reporting delays are then modelled parametrically using a lognormal distribution with the log mean and log standard deviation each modelled using a weekly random walk with a pooled standard deviation and a random effect for the day of the week (introduced on the 6th of December 2021) with public holidays assumed to be reported like Sundays. Report date effects are again modelled using a random effect for day of the week with public holidays assumed to be reported like Sundays. The second model is fit jointly to age groups but is otherwise structured in the same way as the unpooled model except that report day of the week effects and the observation overdispersion are assumed to be joint across age groups, age groups are assumed to have a random intercept for both the log mean and the log standard deviation of the reporting delay distribution, and there is no random effect for reference day of the week. We also consider a series of pooled models which sequentially include the features of our most complex model. These are: age groups are fit jointly, day of the week reporting effects, a random intercept for each age group, and a random walk by positive test week shared across age groups.

Model | Formula
Reference: Fixed, Report: Fixed | 
Reference: Fixed, Report: Day of week |
Reference: Age, Report: Day of week |
Reference: Age and week, Report: Day of week |
Reference: Age and week by age, Report: Day of week | 
Independent by age, Reference: Week, Report: Day of week |
Independent by age, Reference: Week and day of week, Report: Day of week |

## Implementation

The model is implemented as part of the `epinowcast` R package [@epinowcast] using `stan` and `cmdstanr`[@stan; @cmdstanr]. Sparse design matrices have been used for all covariates to limit the number of probability mass functions that need to be calculated. `epinowcast` incorporates additional functionality written in R[@R] to enable plotting nowcasts and posterior predictions, summarising nowcasts, and scoring them using `scoringutils`[@scoringutils]. All functionality is modular allowing users to extend and alter the underlying model whilst continuing to use the package framework.

## Evaluation

**TODO: Model evaluation. What targets. What scoring measures. What aggregation**
**TODO: Evaluation of performance for submitted models vs others submitted to forecasting hub.**
**TODO: Add links to nowcasting hub and evaluation against those models.**

We evaluate these models first visually across a range of nowcasting dates and then quantitatively using proper scoring rules [@scoringutils] on both the natural and log scales (corresponding to absolute and relative performance) aggregating scores first across all targets and then stratifying in turn by age group, nowcast horizon, date of postive test, and date of report. We also explore other aspects of our models performance by highlighting models that have problematic fitting diagnostics and summarising the estimation time for each model. We provide a [report](https://epiforecasts.io/eval-germany-sp-nowcasting/real-time/) of this evaluation that is updated in real-time as new data and nowcasts become available.

## Reproducibility

All analysis was carried out using R version 4.1.2 [@R]. The analysis pipeline described here is available as a `targets` workflow [@targets] along with an archive of all interim results generated using `gittargets` [@gittargets] and stored using `piggyback` [@piggyback]. A Dockerfile, and prebuilt archived image, has been made available with the code to enhance reproducibility [@Boettiger:2015dw].

# Resullts

*This project is still underway and so results are not finalised. Please see our [real-time evaluation report](https://epiforecasts.io/eval-germany-sp-nowcasting/real-time/#Summary) for our initial findings.*

# Dicussion

**TODO: Summarise project**
**TODO: Discussion strengths and limitations of the evaluation**
**TODO: Compare approach evaluated here to the literature**
**TODO: Discuss further work**
**TODO Summarise conclusions.**

**Software availability**

Source code is available from: https://github.com/epiforecasts/eval-germany-sp-nowcasting/

License: MIT

Models are implemented in the `epinowcast` R package. This is available from: https://github.com/epiforecasts/epinowcast/

**Data availability**

Input data and summarised results are available from: https://github.com/epiforecasts/eval-germany-sp-nowcasting/tree/main/data

Data from all interim stages of the analysis is from: https://github.com/epiforecasts/eval-germany-sp-nowcasting/releases/tag/latest

License: MIT

**Grant information**

This work was supported by the Wellcome Trust through a Wellcome Senior Research Fellowship to SF [210758].

**Competing interests**

There are no competing interests.

## References

<div id = 'refs'></div>

### 2022-05-17 writeup/si.Rmd
Initial commit

---
title: "Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany"
subtitle: "Supplementary Information"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 4
    includes:
      before_body: ../docs/header.html
      after_body: ../docs/footer.html
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
bibliography: library.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
header-includes:
   - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, eval = TRUE,
  fig.width = 9, fig.height = 9,
  dpi = 320,
  fig.path = "figures/"
)
```

## Authors

S. Abbott (1), S. Funk (1)

Correspondence to: sam.abbott@lshtm.ac.uk

## Affiliations

1. Center for the Mathematical Modelling of Infectious Diseases, London School of Hygiene & Tropical Medicine, London WC1E 7HT, United Kingdom

## Overview

This analysis uses the `targets` package for reproducibility. The full analysis structure can be seen below:

```{r, message = FALSE, eval = FALSE}
library(targets)
tar_glimpse()
```

![](../figures/targets-graph.png)

## References

<div id = 'refs'></div>

### 2022-05-17 R/get-hub-forecasts.R

@@ -1,0 +1,175 @@
+start_using_memoise <- function(path = tempdir()) {
+  message("Using a cache at: ", path)
+  options("useMemoise" = TRUE, cache_path = path)
+}
+
+stop_using_memoise <- function() {
+  if (!is.null(options("useMemoise"))) {
+    options("useMemoise" = NULL)
+  }
+}
+
+reset_cache <- function() {
+  unlink(getOption("cache_path"), recursive = TRUE)
+  memoise::cache_filesystem(getOption("cache_path"))
+  return(invisible(NULL))
+}
+
+# Get all the paths for the repository
+get_github_paths <- function(repo, branch = "main", depth = 0) {
+  return <- gh::gh(
+    "/repos/{repo}/git/trees/{branch}?recursive={depth}",
+    method = "GET",
+    repo = repo,
+    branch = branch,
+    depth = depth
+  )
+
+  dt <- purrr::map(
+    return$tree,
+    ~ data.table::data.table(
+      path = .$path
+    )
+  )
+  dt <- data.table::rbindlist(dt)
+  return(dt[])
+}
+
+# Filter paths for targets of interest
+filter_paths <- function(paths, string) {
+  paths <- paths[grepl(string, path)]
+  return(paths)
+}
+
+# Filter paths for models and dates we want from a target folder
+get_hub_forecast_paths <- function(repo, branch = "main",
+                                   folder = "data-processed", models, dates) {
+
+  paths <- get_github_paths(repo = repo, branch = branch, depth = 15)
+
+  paths <- paths |>
+    filter_paths(paste0(folder, "/")) |>
+    filter_paths(".csv")
+
+  if (!missing(models)) {
+    paths <- purrr::map(models, ~ filter_paths(paths, .))
+    paths <- data.table::rbindlist(paths)
+  }
+
+  if (!missing(dates)) {
+    paths <- purrr::map(dates, ~ filter_paths(paths, .))
+    paths <- data.table::rbindlist(paths)
+  }
+  return(paths[])
+}
+
+# Download a forecast with error protection
+# use of fread means that data can be filtered prior to loading into R
+download_forecast <- function(repo, path, branch, ...) {
+  sfread <- data.table::fread
+  if (!is.null(getOption("useMemoise"))) {
+    if (getOption("useMemoise")) {
+      ch <- memoise::cache_filesystem(getOption("cache_path"))
+      sfread <- memoise::memoise(sfread, cache = ch)
+    }
+  }
[…]

### 2022-05-17 R/model-tools.R

@@ -1,0 +1,118 @@
+compile_model <- function(...) {
+  model <- epinowcast::enw_model(
+    thread = TRUE, dir = here::here("data"),
+    ...
+  )
+  return(model$stan_file())
+}
+
+load_model <- function() {
+  model <- epinowcast::enw_model(
+    dir = here::here("data"), thread = TRUE
+  )
+  return(model)
+}
+
+prior_epinowcast <- function(obs, priors, max_delay = 40, scale = 5, ...) {
+  pobs <- enw_preprocess_data(obs,
+    max_delay = max_delay,
+    rep_holidays = "holiday"
+  )
+
+  model <- load_model()
+
+  nowcast <- epinowcast(
+    pobs,
+    model = model,
+    ...
+  )
+
+  priors <- enw_posterior_as_prior(
+    nowcast,
+    priors = priors,
+    variables = c("logmean_int", "logsd_int", "sqrt_phi"),
+    scale = scale
+  )
+
+  return(priors)
+}
+
+summarise_nowcast <- function(nowcast, model,
+                              probs = c(
+                                0.025, 0.05, 0.1, 0.2, 0.25, 0.5, 0.75,
+                                0.8, 0.9, 0.95, 0.975
+                              )) {
+  samples <- summary(nowcast, type = "nowcast_samples")
+
+  samples[sample < confirm, sample := confirm]
+  samples[is.na(sample), sample := confirm]
+
+  daily <- enw_summarise_samples(samples, probs)
+
+  cols <- c("confirm", "sample")
+  samples[, (cols) := lapply(.SD, data.table::frollsum, n = 7),
+    .SDcols = cols, by = c(".draw", "age_group", "location")
+  ]
+  samples <- samples[!is.na(sample)]
+
+  # Summarise 7 day nowcast
+  seven_day <- enw_summarise_samples(samples, probs = probs)
+
+  out <- data.table::data.table(
+    model = model,
+    nowcast_date = max(daily$reference_date),
+    daily = list(daily),
+    seven_day = list(seven_day)
+  )
+  out <- cbind(
+    out,
+    nowcast[
+      ,
+      .(
+        max_rhat, divergent_transitions, per_divergent_transitions,
+        max_treedepth, no_at_max_treedepth, per_at_max_treedepth, run_time,
+        failure
+      )
+    ]
+  )
+  return(out[])
+}
+
+default_nowcast_on_error <- function(nowcast, pobs, model,
[…]

### 2022-05-17 R/models.R

@@ -1,0 +1,259 @@
+fixed_epinowcast <- function(obs, max_delay = 40, ...) {
+  pobs <- enw_preprocess_data(
+    obs,
+    rep_holidays = "holiday",
+    max_delay = max_delay, by = "age_group"
+  )
+
+  reference_effects <- enw_formula(pobs$metareference)
+  report_effects <- enw_formula(pobs$metareport)
+
+  model <- load_model()
+
+  nowcast <- epinowcast(
+    pobs,
+    model = model,
+    reference_effects = reference_effects,
+    report_effects = report_effects,
+    ...
+  )
+
+  nowcast <- default_nowcast_on_error(nowcast, pobs, model, ...)
+
+  out <- summarise_nowcast(nowcast, model = "Reference: Fixed, Report: Fixed")
+  return(out[])
+}
+
+dow_epinowcast <- function(obs, max_delay = 40, ...) {
+  pobs <- enw_preprocess_data(
+    obs,
+    rep_holidays = "holiday",
+    max_delay = max_delay, by = "age_group"
+  )
+
+  reference_effects <- enw_formula(pobs$metareference)
+  report_effects <- enw_formula(pobs$metareport, random = "day_of_week")
+
+  model <- load_model()
+
+  nowcast <- epinowcast(
+    pobs,
+    model = model,
+    reference_effects = reference_effects,
+    report_effects = report_effects,
+    ...
+  )
+
+  nowcast <- default_nowcast_on_error(nowcast, pobs, model, ...)
+
+  out <- summarise_nowcast(
+    nowcast,
+    model = "Reference: Fixed, Report: Day of week"
+  )
+  return(out[])
+}
+
+age_epinowcast <- function(obs, max_delay = 40, ...) {
+  pobs <- enw_preprocess_data(
+    obs,
+    rep_holidays = "holiday",
+    max_delay = max_delay, by = "age_group"
+  )
+
+  reference_effects <- enw_formula(pobs$metareference, random = "age_group")
+  report_effects <- enw_formula(pobs$metareport, random = "day_of_week")
+
+  model <- load_model()
+
+  nowcast <- epinowcast(
+    pobs,
+    model = model,
+    reference_effects = reference_effects,
+    report_effects = report_effects,
+    ...
+  )
+
+  nowcast <- default_nowcast_on_error(nowcast, pobs, model, ...)
+
+  out <- summarise_nowcast(
+    nowcast,
+    model = "Reference: Age, Report: Day of week"
+  )
[…]

### 2022-05-17 R/plot.R

@@ -1,0 +1,59 @@
+plot_nowcast <- function(summarised_nowcast, latest_obs,
+                                     max_delay = Inf, ...) {
+   summarised_nowcast <- summarised_nowcast[
+        reference_date >= (max(as.Date(nowcast_date)) - max_delay)
+   ]
+
+   suppressWarnings(summarised_nowcast[, holiday := NULL])
+
+   plot <- epinowcast::enw_plot_nowcast_quantiles(
+      summarised_nowcast,
+      latest_obs = latest_obs[
+        reference_date >= min(summarised_nowcast$reference_date)][
+        reference_date <= max(summarised_nowcast$reference_date)
+      ], ...
+  )
+  return(plot)
+}
+
+plot_scores <- function(scores, ...) {
+  ggplot2::ggplot(scores) +
+    ggplot2::aes(...) +
+    ggplot2::geom_point(size = 1.2) +
+    ggplot2::geom_line(size = 1.1, alpha = 0.4) +
+    ggplot2::scale_fill_brewer(palette = "Paired") +
+    ggplot2::scale_color_brewer(palette = "Paired") +
+    labs(y = "Weighted interval score") +
+    theme_bw() +
+    theme(legend.position = "bottom") +
+    guides(fill = guide_legend(title = "Model", nrow = 4),
+           col = guide_legend(title = "Model", nrow = 4))
+}
+
+plot_relative_scores <- function(score, baseline) {
+  score <- data.table::as.data.table(score)
+  fixed_score <- score[
+    model %in% baseline,
+    .(reference_date, age_group, fixed_is = interval_score)
+  ]
+  score <- merge(score, fixed_score, by = c("reference_date", "age_group"))
+
+  score <- score[, interval_score := interval_score / fixed_is]
+  score <- score[!model %in% baseline]
+  plot <- ggplot(score) +
+    aes(x = reference_date, y = interval_score, col = model) +
+    geom_hline(yintercept = 1, linetype = 2, size = 1.2, alpha = 0.5) +
+    geom_line(size = 1.1, alpha = 0.6) +
+    geom_point(size = 1.2) +
+    facet_wrap(vars(age_group)) +
+    scale_color_brewer(palette = "Paired") +
+    scale_y_log10(labels = scales::percent)
+
+  plot <- enw_plot_theme(plot) +
+    labs(
+      x = "Reference date",
+      y = paste0("Weighted interval score (", baseline, ")")
+    ) + # nolint
[…]

### 2022-05-17 R/postprocess.R

@@ -1,0 +1,124 @@
+adjust_quantile <- function(quantile, median, max_ratio = 0.5) {
+  q_ratio <- quantile / median
+
+  if (max(q_ratio, na.rm = TRUE) > 1) {
+    quantile <- ifelse(
+      q_ratio >= max_ratio + 1,
+      (max_ratio + 1) * median,
+      quantile
+    )
+  }else{
+    quantile <- ifelse(
+      q_ratio <= (1 - max_ratio),
+      max_ratio * median,
+      quantile
+    )
+  }
+  return(quantile)
+}
+
+adjust_posteriors <- function(nowcasts, target, max_ratio = 0.25,
+                              rhat_bound = 1.1, per_dt_bound = 0.2) {
+  unadjusted_nowcasts <- nowcasts[
+    !(max_rhat > rhat_bound | per_divergent_transitions > per_dt_bound)
+  ]
+  adjusted_nowcasts <- nowcasts[
+    max_rhat > rhat_bound | per_divergent_transitions > per_dt_bound
+  ]
+  if (nrow(adjusted_nowcasts) > 0) {
+    adjusted_nowcasts <- adjusted_nowcasts[,
+      (target) := purrr::map(
+        get(target),
+        function(dt) {
+          cols <- grep("^q[0-9]", colnames(dt), value = TRUE)
+          dt[,
+            (cols) := purrr::map(.SD, adjust_quantile, median = median,
+                                  max_ratio = max_ratio),
+            .SDcols = cols
+          ]
+        })
+    ]
+  nowcasts <- rbind(unadjusted_nowcasts, adjusted_nowcasts)
+  }
+  return(nowcasts)
+}
+
+unnest_nowcasts <- function(nowcasts, target) {
+
+  nowcasts <- nowcasts[,
+    (target) := pmap(
+      list(get(target), model, nowcast_date),
+        function(df, m, d) {
+          df[, model := m][, nowcast_date := d]
+        })][,
+    rbindlist(get(target), use.names = TRUE, fill = TRUE)][
+    order(location, age_group, nowcast_date, reference_date, model)
+    ]
+
+  data.table::setcolorder(nowcasts, neworder = c("model", "nowcast_date"))
+  nowcasts[, nowcast_date := as.Date(nowcast_date)]
+  nowcasts[, report_date := nowcast_date]
+  return(nowcasts[])
+}
+
+format_for_submission <- function(nowcast, horizon = -28,
[…]

### 2022-05-17 R/utils.R

@@ -1,0 +1,98 @@
+#' Save a plot and return the path for targets
+#' @importFrom ggplot2 ggsave
+save_plot <- function(plot, filename, path, ...) {
+  dir.create(path, recursive = TRUE, showWarnings = FALSE)
+  path <- file.path(path, file)
+  ggplot2::ggsave(path, plot, ...)
+  return(filename)
+}
+
+#' Save a dataframe to a csv and return the path for targets
+save_csv <- function(dt, filename, path, allow_empty = TRUE) {
+  dir.create(path, recursive = TRUE, showWarnings = FALSE)
+  path <- file.path(path, filename)
+
+  if (allow_empty | nrow(dt) > 0) {
+    data.table::fwrite(dt, path)
+  }
+  return(path)
+}
+
+drop_string <- function(var, string) {
+  var[!grepl(string, var)]
+}
+
+select_var <- function(dt, var) {
+  dt[[var]]
+}
+
+format_from_csv <- function(dt) {
+  if (!is.null(dt$age_group)) {
+      dt[
+      ,
+        age_group := factor(
+          age_group,
+          levels = c("00+", "00-04", "05-14", "15-34", "35-59", "60-79", "80+")
+        )
+    ]
+    }
+    dt[,
+    model := factor(
+      model,
+      levels = c("Reference: Fixed, Report: Fixed",
+                "Reference: Fixed, Report: Day of week",
+                "Reference: Age, Report: Day of week",
+                "Reference: Age and week, Report: Day of week",
+                "Reference: Age and week by age, Report: Day of week",
+                "Independent by age, Reference: Week, Report: Day of week",
+                "Independent by age, Reference: Week and day of week, Report: Day of week" # nolint
+                )
+    )
+    ]
+  return(dt[])
+}
+
+load_nowcasts <- function(path) {
+  nowcasts <- fs::dir_ls(
+    path,
+    glob = "*.csv"
+  ) |>
+    purrr::map(data.table::fread) |>
+    data.table::rbindlist(use.names = TRUE, fill = TRUE)
+  nowcasts[, horizon := as.numeric(
+    as.Date(reference_date) - as.Date(nowcast_date)
+  )]
+  nowcasts <- format_from_csv(nowcasts)
+  return(nowcasts[])
+}
+
+load_obs <- function(path) {
+  obs <- fread(path)
+  obs[
+    ,
+    age_group := factor(
[…]

### 2022-05-17 _targets_r/globals/globals.R

@@ -1,0 +1,24 @@
+library(targets)
+library(tarchetypes)
+library(cmdstanr)
+library(data.table)
+library(epinowcast)
+library(ggplot2)
+library(purrr, quietly = TRUE)
+library(here)
+library(future)
+library(future.callr)
+plan(callr)
+functions <- list.files(here("R"), full.names = TRUE)
+walk(functions, source)
+rm("functions")
+set_cmdstan_path()  
+tar_option_set(
+  packages = c("data.table", "epinowcast", "scoringutils", "ggplot2", "purrr",
+               "cmdstanr", "here"),
+  deployment = "worker",
+  memory = "transient",
+  workspace_on_error = TRUE,
+  error = "continue",
+  garbage_collection = TRUE
+)

### 2022-05-17 _targets_r/targets/age_groups.R

@@ -1,0 +1,3 @@
+tar_target(age_groups, {
+  unique(hospitalisations$age_group)
+})

### 2022-05-17 _targets_r/targets/combined_nowcasts.R

@@ -1,0 +1,23 @@
+tar_target(combined_nowcasts, {
+  rbindlist(list(
+    fixed_nowcast,
+    dow_nowcast,
+    age_nowcast,
+    week_nowcast,
+    age_week_nowcast,
+    independent_nowcast,
+    overall_only_nowcast,
+    independent_ref_dow_nowcast
+  ), use.names = TRUE, fill = TRUE)[,
+     model := factor(
+      model,
+      levels = c("Reference: Fixed, Report: Fixed",
+                 "Reference: Fixed, Report: Day of week",
+                 "Reference: Age, Report: Day of week",
+                 "Reference: Age and week, Report: Day of week",
+                 "Reference: Age and week by age, Report: Day of week",
+                 "Independent by age, Reference: Week, Report: Day of week",
+                 "Independent by age, Reference: Week and day of week, Report: Day of week")
+     )
+    ][, id := 1:.N]
+})

### 2022-05-17 _targets_r/targets/complete_7day_hospitalisations.R

@@ -1,0 +1,3 @@
+tar_target(complete_7day_hospitalisations, {
+  latest_7day_hospitalisations[reference_date < (max(reference_date) - 28)]
+})

### 2022-05-17 _targets_r/targets/complete_hospitalisations.R

@@ -1,0 +1,3 @@
+tar_target(complete_hospitalisations, {
+  latest_hospitalisations[reference_date < (max(reference_date) - 28)]
+})

### 2022-05-17 _targets_r/targets/dow_nowcast.R

@@ -1,0 +1,13 @@
+tar_target(
+  dow_nowcast,
+  nowcast(
+    obs = hospitalisations_by_date_report,
+    tar_loc = locations,
+    model = dow_epinowcast,
+    priors = priors,
+    max_delay = max_report_delay,
+    settings = epinowcast_settings
+  ),
+  cross(hospitalisations_by_date_report, locations),
+  cue = tar_cue(mode = "never")
+)

### 2022-05-17 _targets_r/targets/epinowcast_model.R

@@ -1,0 +1,5 @@
+tar_file(
+  epinowcast_model,
+  compile_model(),
+  deployment = "main"
+)

### 2022-05-17 _targets_r/targets/epinowcast_settings.R

@@ -1,0 +1,15 @@
+tar_target(epinowcast_settings, {
+  list(
+    save_warmup = FALSE,
+    output_loglik = FALSE,
+    pp = FALSE,
+    iter_warmup = 1000,
+    iter_sampling = 1000,
+    chains = 2,
+    parallel_chains = 2,
+    threads_per_chain = 1,
+    adapt_delta = 0.95,
+    show_messages = FALSE,
+    refresh = 0
+  )
+})

### 2022-05-17 _targets_r/targets/hierarchical_submission_nowcast.R

@@ -1,0 +1,16 @@
+tar_target(
+  hierarchical_submission_nowcast,
+  age_week_nowcast[nowcast_date == nowcast_dates] |>
+    adjust_posteriors(
+      target = "seven_day", 
+      max_ratio = 0.25, 
+      rhat_bound = 1.1,
+      per_dt_bound = 0.2
+    ) |>
+    select_var("seven_day") |>
+    rbindlist() |>
+    format_for_submission(),
+  map(nowcast_dates),
+  iteration = "list",
+  cue = tar_cue(mode = "never")
+)

### 2022-05-17 _targets_r/targets/hospitalisation_url.R

@@ -1,0 +1,4 @@
+tar_url(
+  hospitalisation_url,
+  "https://raw.githubusercontent.com/KITmetricslab/hospitalization-nowcast-hub/main/data-truth/COVID-19/COVID-19_hospitalizations_preprocessed.csv" # nolint
+)

### 2022-05-17 _targets_r/targets/hospitalisations.R

@@ -1,0 +1,3 @@
+tar_target(hospitalisations, {
+  get_germany_hospitalisations(url = hospitalisation_url)
+})

### 2022-05-17 _targets_r/targets/hospitalisations_by_date_report.R

@@ -1,0 +1,11 @@
+tar_target(
+  hospitalisations_by_date_report,
+  enw_retrospective_data(
+    hospitalisations,
+    rep_date = nowcast_dates,
+    ref_days = max_report_delay
+  )[, nowcast_date := nowcast_dates],
+  map(nowcast_dates),
+  iteration = "list",
+  cue = tar_cue(mode = "never")
+)

### 2022-05-17 _targets_r/targets/independent_nowcast.R

@@ -1,0 +1,13 @@
+tar_target(
+  independent_nowcast,
+  nowcast(
+    obs = hospitalisations_by_date_report[age_group == age_groups],
+    tar_loc = locations,
+    model = independent_epinowcast,
+    priors = priors,
+    max_delay = max_report_delay,
+    settings = epinowcast_settings
+  ),
+  cross(hospitalisations_by_date_report, locations, age_groups),
+  cue = tar_cue(mode = "never")
+)

### 2022-05-17 _targets_r/targets/latest_7day_hospitalisations.R

@@ -1,0 +1,5 @@
+tar_target(latest_7day_hospitalisations, {
+  copy(latest_hospitalisations)[,
+    confirm := frollsum(confirm, n =  7), by = c("age_group", "location")
+  ][!is.na(confirm)]
+})

### 2022-05-17 _targets_r/targets/locations.R

@@ -1,0 +1,3 @@
+tar_target(locations, {
+  "DE"
+})

### 2022-05-17 _targets_r/targets/max_report_delay.R

@@ -1,0 +1,3 @@
+tar_target(max_report_delay, {
+  40
+})

### 2022-05-17 _targets_r/targets/nowcast_dates.R

@@ -1,0 +1,5 @@
+tar_target(nowcast_dates, {
+  unique(
+    hospitalisations[reference_date >= start_date]$reference_date
+  )
+})

### 2022-05-17 _targets_r/targets/priors.R

@@ -1,0 +1,9 @@
+tar_target(priors, {
+  do.call(prior_epinowcast, c(
+    list(
+      prior_obs, max_delay = max_report_delay, scale = 10,
+      priors = uninformed_priors
+    ),
+    epinowcast_settings
+  ))
+})

### 2022-05-17 _targets_r/targets/save_7day_nowcasts.R

@@ -1,0 +1,11 @@
+tar_file(
+  save_7day_nowcasts,
+  summarised_7day_nowcast[nowcast_date == nowcast_dates] |>
+    save_csv(
+      filename = paste0(nowcast_dates, ".csv"),
+      path = here("data/nowcasts/seven_day"),
+      allow_empty = FALSE
+    ),
+  map(nowcast_dates),
+  cue = tar_cue(mode = "never")
+)

### 2022-05-17 _targets_r/targets/save_daily_nowcasts.R

@@ -1,0 +1,11 @@
+tar_file(
+  save_daily_nowcasts,
+  summarised_nowcast[nowcast_date == nowcast_dates] |>
+    save_csv(
+      filename = paste0(nowcast_dates, ".csv"),
+      path = here("data/nowcasts/daily"),
+      allow_empty = FALSE
+    ),
+  map(nowcast_dates),
+  cue = tar_cue(mode = "never")
+)

### 2022-05-17 _targets_r/targets/save_diagnostics.R

@@ -1,0 +1,34 @@
+list(
+  tar_file(
+    save_all_diagnostics,
+      save_csv(
+        diagnostics,
+        filename = "all.csv",
+        path = here("data/diagnostics")
+      )
+  ),
+  tar_file(
+    save_high_rhat_diagnostics,
+      save_csv(
+        diagnostics[max_rhat > 1.05],
+        filename = "high-rhat.csv",
+        path = here("data/diagnostics")
+      )
+  ),
+  tar_file(
+    save_high_divergent_transitions,
+      save_csv(
+        diagnostics[per_divergent_transitions > 0.1],
+        filename = "high-divergent-transitions.csv",
+        path = here("data/diagnostics")
+      )
+  ),
+  tar_file(
+    save_failures,
+      save_csv(
+        diagnostics[failure == TRUE],
+        filename = "fitting-failed.csv",
+        path = here("data/diagnostics")
+      )
+  )
+)

### 2022-05-17 _targets_r/targets/save_hierarchical_submission.R

@@ -1,0 +1,9 @@
+tar_file(
+  save_hierarchical_submission,
+  save_csv(
+    hierarchical_submission_nowcast,
+      filename = paste0(nowcast_dates, ".csv"),
+      path = here("data/nowcasts/submission/hierarchical")
+  ),
+  map(hierarchical_submission_nowcast, nowcast_dates)
+)

### 2022-05-17 _targets_r/targets/save_latest_7day_hospitalisations.R

@@ -1,0 +1,8 @@
+tar_file(
+  save_latest_7day_hospitalisations,
+  save_csv(
+    latest_7day_hospitalisations,
+    filename = paste0("seven_day.csv"),
+    path = here("data/observations")
+  )
+)

### 2022-05-17 _targets_r/targets/save_latest_daily_hospitalisations.R

@@ -1,0 +1,8 @@
+tar_file(
+  save_latest_daily_hospitalisations,
+  save_csv(
+    latest_hospitalisations,
+    filename = paste0("daily.csv"),
+    path = here("data/observations")
+  )
+)

### 2022-05-17 _targets_r/targets/score.R

@@ -1,0 +1,29 @@
+tar_map(
+  list(score_by = list(
+    "overall", "age_group", "horizon", "reference_date", "nowcast_date"
+  )),
+  tar_target(
+    scores,
+    enw_score_nowcast(
+      scored_nowcasts, complete_hospitalisations, 
+      summarise_by = drop_string(c(score_by, "model"), "overall"),
+      log = FALSE
+    )
+  ),
+  tar_target(
+    log_scores,
+    enw_score_nowcast(
+      scored_nowcasts, complete_hospitalisations, 
+      summarise_by = drop_string(c(score_by, "model"), "overall"),
+      log = TRUE
+    )
+  ),
+  tar_file(
+    save_scores,
+    save_csv(
+      rbind(scores[, scale := "natural"], log_scores[, scale := "log"]),
+      filename = paste0(paste(score_by, sep = "-"), ".csv"),
+      path = here("data/scores")
+    )
+  )
+)

### 2022-05-17 _targets_r/targets/scored_nowcasts.R

@@ -1,0 +1,8 @@
+tar_target(scored_nowcasts, {
+  summarised_nowcast[location == locations][
+    reference_date < (max(nowcast_date) - 28)][,
+    holiday := NULL][,
+    horizon := as.numeric(as.Date(reference_date) - nowcast_date)][
+    horizon >= -7
+  ]
+})

### 2022-05-17 _targets_r/targets/start_date.R

@@ -1,0 +1,3 @@
+tar_target(start_date, {
+  as.Date("2021-10-01")
+})

### 2022-05-17 _targets_r/targets/summarised_7day_nowcast.R

@@ -1,0 +1,10 @@
+tar_target(summarised_7day_nowcast, {
+  combined_nowcasts |> 
+    adjust_posteriors(
+      target = "seven_day", 
+      max_ratio = 0.25, 
+      rhat_bound = 1.1,
+      per_dt_bound = 0.2
+  ) |> 
+    unnest_nowcasts(target = "seven_day") 
+})

### 2022-05-17 _targets_r/targets/summarised_nowcast.R

@@ -1,0 +1,10 @@
+tar_target(summarised_nowcast, {
+  combined_nowcasts |> 
+    adjust_posteriors(
+      target = "daily", 
+      max_ratio = 0.25, 
+      rhat_bound = 1.1,
+      per_dt_bound = 0.2
+  ) |> 
+    unnest_nowcasts(target = "daily") 
+})

### 2022-05-17 _targets_r/targets/uninformed_priors.R

@@ -1,0 +1,3 @@
+tar_target(uninformed_priors, {
+  enw_priors()
+})

### 2022-05-17 analyses/germany-example.R

@@ -1,0 +1,44 @@
+library(epinowcast)
+suppressMessages(library(data.table, quietly = TRUE))
+library(purrr)
+library(here)
+
+# load dev code
+functions <- list.files(here("R"), full.names = TRUE)
+walk(functions, source)
+
+# set number of cores to use
+options(mc.cores = 4)
+
+# get processed germany data and reprocess
+germany_hosp <- get_germany_hospitalisations()
+
+# national only and only for the last 28 days of data
+germany_nat_hosp <- germany_hosp[location == "DE"]
+germany_nat_hosp <-
+  germany_nat_hosp[reference_date >= (max(reference_date) - 27)]
+
+# Preprocess data
+pobs <- enw_preprocess_data(germany_nat_hosp, by = "age_group", max_delay = 30)
+
+# Construct design matrices for the desired reference date effects
+reference_effects <- enw_formula(pobs$metareference[[1]])
+
+# Construct design matrices for the desired report day effects
+report_effects <- enw_formula(pobs$metareport[[1]], random = "day_of_week")
+
+# fit model to example data and produce a nowcast
+est <- epinowcast(pobs,
+  report_effects = report_effects,
+  reference_effects = reference_effects,
+  debug = FALSE, pp = TRUE, save_warmup = FALSE,
+)
+
+# observations linked to truncation adjusted estimates
+summary(nowcast)
+
+# Plot nowcast vs latest observations
+plot(est, obs = latest_cases)
+
+# Plot posterior prediction for observed cases at date of report
+plot(est, obs = latest_cases, type = "pos", log = TRUE)

### 2022-05-17 analyses/synthetic-example.R

@@ -1,0 +1,63 @@
+library(epinowcast)
+library(covidregionaldata)
+suppressMessages(library(data.table, quietly = TRUE))
+suppressMessages(library(rstan, quietly = TRUE))
+library(purrr)
+library(here)
+
+# load dev code
+functions <- list.files(here("R"), full.names = TRUE)
+walk(functions, source)
+
+# set number of cores to use
+options(mc.cores = 4)
+# get example case counts
+start_using_memoise()
+latest_cases <- get_national_data("UK", verbose = FALSE)
+latest_cases <- setDT(latest_cases)[date >= as.Date("2021-07-01")]
+latest_cases <- latest_cases[, .(date, confirm = cases_new)]
+
+# get a range of dates to generate synthetic data for
+scenarios <- enw_random_intercept_scenario(
+  obs = latest_cases,
+  snapshots = seq(6 * 7, 0, by = -1),
+  logmean = 1.9, logmean_sd = 0.1,
+  logsd = 1, logsd_sd = 0.1
+)
+
+# simulate observations
+scenarios <- enw_simulate_lnorm_trunc_obs(scenarios, latest_cases)
+sim_reported_cases <- rbindlist(scenarios$reported_cases)
+setnames(sim_reported_cases, "date", "reference_date")
+sim_reported_cases <- rbind(
+  copy(sim_reported_cases)[, age := 1],
+  copy(sim_reported_cases)[, age := 2]
+)
+
+# Preprocess data
+pobs <- enw_preprocess_data(sim_reported_cases, by = "age")
+
+# Construct design matrices for the desired reference date effects
+reference_effects <- enw_formula(pobs$metareference[[1]])
+
+# Construct design matrices for the desired report day effects
+report_effects <- enw_formula(pobs$metareport[[1]], random = "day_of_week")
+
+# fit model to example data and produce nowcast
+est <- epinowcast(pobs,
+  report_effects = report_effects,
+  reference_effects = reference_effects,
+  debug = FALSE, pp = TRUE, save_warmup = FALSE
+)
+
+# summarise nowcast
+summary(est)
+
+# Plot nowcast vs latest observations
+plot(est, latest_obs = copy(latest_cases)[, reference_date := date])
+
+# Plot posterior prediction for observed cases at date of report
+plot(est,
+  latest_obs = copy(latest_cases)[, reference_date := date],
+  type = "pos", log = TRUE
+)

### 2022-05-17 bin/update-if-new-data.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+URL=https://raw.githubusercontent.com/KITmetricslab/hospitalization-nowcast-hub/main/data-truth/COVID-19/COVID-19_hospitalizations_preprocessed.csv
+CSV=data/COVID-19_hospitalizations_preprocessed.csv 
+
+if [ ! -f $CSV ] 
+  then
+  echo No cached data found - updating estimates
+  wget -nc $URL -P data/
+  bash bin/update-targets-and-publish.sh
+  rm $CSV
+  else
+  echo Cache found - pipeline running - no action taken
+fi

### 2022-05-17 bin/update-targets-and-publish.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+git pull -X theirs
+
+bash bin/render-targets.sh > render.log 2>&1
+
+# Run the update twice just in case a random error occurs the first time
+bash bin/make-parallel-targets.sh > targets.log 2>&1
+bash bin/make-parallel-targets.sh >> targets.log 2>&1
+
+bash bin/update-git-remote.sh > git-remote.log 2>&1
+
+bash bin/update-targets-remote.sh > targets-remote.log 2>&1

### 2022-05-17 summary/index.Rmd
Initialize gh-pages

--- 2022-05-17 writeup/paper.Rmd
@@
 title: "Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany"
+subtitle: "Summary"
+author: Sam Abbott, Sebastian Funk
+bibliography: ../../writeup/library.bib
+csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
+date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
 output:
-  pdf_document: default
   html_document:
-    keep_md: true
+    theme: cosmo
     toc: true
-    theme: united
-  word_document: default
-bibliography: library.bib
-csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
-header-includes:
-   - \usepackage{float}
+    toc_float: true
+    toc_depth: 4
+    includes:
+      before_body: ../header.html
+      after_body: ../footer.html
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE, eval = TRUE,
-  fig.width = 9, fig.height = 9,
-  dpi = 320,
-  fig.path = "figures/"
-)
+```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, dpi = 330, warning = FALSE)
 ```
 
-# Authors
+**This project is under active development. Current results are preliminary and may change as more data is collected or as the methodology is improved.**
 
-S. Abbott (1), S. Funk (1)
+# Background
 
-Correspondence to: sam.abbott@lshtm.ac.uk
-
-# Affiliations
-
-1. Center for the Mathematical Modelling of Infectious Diseases, London School of Hygiene & Tropical Medicine, London WC1E 7HT, United Kingdom
-
-# Abstract
-
-**TODO: Summarise the project in a structured abstract (background, methods, results and conclusions)**
-
-# Introduction
-
-```{r, child = "../docs/chunks/introduction.Rmd"}
+```{r, child = "../chunks/introduction.Rmd"}
 ```
-
-**TODO: What else has been done for nowcasting**
 
@@
 
-## Data
-
-**TODO: Data source and processing details**
-
-## Models
-
-In the following sections we provide methodological and implementation details for the nowcasting framework implemented in `epinowcast` and applied here.
-Our approach is an extension of that proposed by Günther et al.[@gunther2021] which was itself an extension of the model proposed by Höhle and Heiden[@hohle] and implemented in the `surveillance` R package[@surveillance].
-Compared to the model proposed in Günther et al.[@gunther2021], `epinowcast` adds an optional parametric assumption for the underlying delay from occurrence to report, support for jointly nowcasting multiple related datasets, a flexible formula interface allowing for the specification of a large range of models, and an efficient implementation in `stan` which makes use of sparse design matrices and within chain parallelisation to reduce runtimes [@stan; @cmdstanr].
-
-### Decomposition into expected final notifications and report delay components
-
-We follow the approach of Höhle and Heiden[@hohle] and consider the distribution of notifications ($n_{gtd}$) by time of occurrence ($t$) and reporting delay ($d$) conditional on the final observed count $N_{gt}$ for each dataset ($g$) such that,
-
-\begin{equation}
-  N_{gt} = \sum_{d=0}^{D} n_{gtd}
-\end{equation}
-
-where $D$ represents the maximum delay between time of occurrence and time of report which in theory could be infinite but in practice we set to a value in order to make the model identifiable and computationally feasible.
-This formulation means that $n_{gtd} \mid N_{gt}$ is multinomial with a probability vector ($p_{gtd}$) of length $D$ for each $t$ and $g$ that needs to be estimated at the same time as estimating the expected number of final notifications $\mathbb{E}[N_{gt}] = \lambda_{gt}$.
-
-An alternative approach, not explored here, is to consider each $n_{gtd}$ independently at which point the model can be defined as a standard regression with the appropriate observation model and adjustment for reporting delay (i.e it becomes a Poisson or Negative Binomial regression)[@bastos].
-An implementation of this approach is available in Bastos et al.[@bastos].
-More work needs to be done to evaluate which of these approaches produces more accurate nowcasts for epidemiological count data.
-
-### Expected final notifications
-
-Here we follow the approach of Günther et al.[@gunther2021] and specify the model for expected final notifications as a first order random walk.
-This model can in principle be any model such as a more complex time-series approach, a gaussian process, or a mechanistic or semi-mechanistic compartmental model.
-Extending the flexibility of this model is an area of further work as is evaluating the benefits and tradeoffs of more complex approaches.
-
-\begin{align}
-  \log (\lambda_{gt}) &\sim \text{Normal}\left(\log (\lambda_{gt-1}) , \sigma^{\lambda}_{g} \right) \\
-  \log (\lambda_{g0}) &\sim \text{Normal}\left(\log (N_{g0}), 1 \right) \\ \sigma^{\lambda}_{g} &\sim \text{Half-Normal}\left(0, 1\right)
-\end{align}
-
-### Delay distribution
-
-Again following the approach of Günther et al.[@gunther2021] we define the delay distribution ($p_{gtd}$) as a discrete time hazard model ($h_{gtd} =\text{P} \left(\text{delay}=d|\text{delay} \geq d, W_{gtd}\right)$) but we extend this model to decompose $W_{gtd}$ into 3 components: hazard derived from a parametric delay distribution ($\gamma_{gtd}$) dependent on covariates at the date of occurrence, hazard not derived from a parametric distribution ($\delta_{gtd}$) dependent on covariates at the date of occurrence, and hazard dependent on covariates referenced to the date of report ($\epsilon_{gtd}$).
-
-For first component ($\gamma_{gtd}$) we assume that the probability of reporting $p^{\prime}_{gtd}$ on a given date given follow a parametric distribution (in the baseline case a discretised log normal distribution) with the log mean and log standard deviation being defined using an intercept and arbitrary shared, reference date indexed, covariates with fixed ($\alpha_{i}$) and random ($\beta_{i}$) coefficients,
-
-\begin{align}
-  p^{\prime}_{gtd} &\sim \text{LogNormal} \left(\mu_{gt}, \upsilon_{gt} \right) \\
-  \mu_{gt} &= \mu_0 + \alpha_{\mu} X_{\gamma} + \beta_{\mu} Z_{\gamma} \\
-  \upsilon_{gt} &= \text{exp} \left( \upsilon_0 + \alpha_{\upsilon} X_{\gamma} + \beta_{\upsilon} Z_{\gamma} \right)
-\end{align}
-
-The parametric logit hazard (probability of report on a given date conditional on not already having reported) for this component of the model is then,
-
-\begin{equation}
-  \gamma_{gtd} = \text{logit} \left(\frac{p^{\prime}_{gtd}}{\left(1 -\sum^{d-1}_{d^{\prime}=0} p^{\prime}_{gtd^{\prime}} \right)} \right)
-\end{equation}
-
-The non-distributional logit hazarad components for the date of occurrence and report are then again defined using an intercept and arbitrary shared covariates with fixed ($\alpha_{i}$) and random ($\beta_{i}$) coefficients.
-
-\begin{align}
-  \delta_{gtd} &= \mu_0 + \alpha_{\delta} X_{\delta} + \beta_{\delta} Z_{\delta} \\
-  \epsilon_{gtd} &= \epsilon_0 + \alpha_{\epsilon} X_{\epsilon} + \beta_{\epsilon} Z_{\epsilon}
-\end{align}
-
-The overall hazard for each group, occurrence time, and delay is then,
-
-\begin{equation}
-  \text{logit} (h_{gtd}) = \gamma_{gtd} + \delta_{gtd} +  \upsilon_{gtd},\ h_{gtD} = 1
-\end{equation}
-
-where the hazard on the final day has been assumed to be 1 in order to enforce the constraint that all reported observations are reported within the specified maximum delay.
-The probability of report for a given delay, occurrence date, and group is then as follows,
-
-\begin{equation}
-  p_{gt0} = h_{gt0},\ p_{gtd} = \left(1 -\sum^{d-1}_{d^{\prime}=0} p_{gtd^{\prime}} \right) \times h_{gtd}
-\end{equation}
-
-All ($\alpha_{i}$) and random ($\beta_{i}$) coefficients have standard normal priors by default with standard half-normal priors for pooled standard deviations.
-
-### Observation model and nowcast
-
-Expected notifications by time of occurrence ($t$) and reporting delay can now be found by multiplying expected final notifications for each $t$ with the probability of reporting for each day of delay ($p_{gtd}$).
-We then assume a negative binomial observation model with a joint overdispersion parameter (with a 1 over square root standard half normal prior) and produce a nowcast of final observed notifications at each occurrence time by summing posterior estimates for each observed notification for that occurrence time.
-
-\begin{align}
-  n_{gtd} \mid \lambda_{gt},p_{gtd}  &\sim \text{NB}(\lambda_{gt} \times p_{gtd}, \phi),\ t=1,...,T. \\
-  \frac{1}{\phi^2} &\sim \text{Half-Normal}(0, 1) \\
-  N_{gt} &= \sum_{d=0}^{D} n_{gtd}
-\end{align}
-
-### Specific model formulations
-
-**TODO: Detail specific model formulations**
+We follow the approach of Höhle and Heiden[@hohle] and consider the distribution of notifications by date of positive test and reporting delay conditional on the final observed count for each target dataset.
+This results in an estimation process where expected hospitalisations by date of positive test are estimated jointly with the delay distribution for each date of positive test.
+When combined this gives the expected final hospitalisations, and once an observation model has been assumed (in our case a negative binomial model) the estimated hospitalisations by date of report can be recovered.
+Aggregating these estimates then gives estimated hospitalisations by date of positive test adjusted for right censoring.
 
@@
 Hospitalisations are modelled using a random walk on the log scale.
-Reporting delays are then modelled parametrically using a lognormal distribution with the log mean and log standard deviation each modelled using a weekly random walk with a pooled standard deviation and a random effect for the day of the week (introduced on the 6th of December 2021) with public holidays assumed to be reported like Sundays.
+Reporting delays are then modelled parametrically using a lognormal distribution with the log mean and log standard deviation each modelled using a weekly random walk with a pooled standard deviation, and a random effect for the day of the week (introduced on the 6th of December 2021) with public holidays assumed to be reported like Sundays.
 Report date effects are again modelled using a random effect for day of the week with public holidays assumed to be reported like Sundays.
@@
 
-Model | Formula
-Reference: Fixed, Report: Fixed | 
-Reference: Fixed, Report: Day of week |
-Reference: Age, Report: Day of week |
-Reference: Age and week, Report: Day of week |
-Reference: Age and week by age, Report: Day of week | 
-Independent by age, Reference: Week, Report: Day of week |
-Independent by age, Reference: Week and day of week, Report: Day of week |
-
-## Implementation
-
-The model is implemented as part of the `epinowcast` R package [@epinowcast] using `stan` and `cmdstanr`[@stan; @cmdstanr].
-Sparse design matrices have been used for all covariates to limit the number of probability mass functions that need to be calculated.
-`epinowcast` incorporates additional functionality written in R[@R] to enable plotting nowcasts and posterior predictions, summarising nowcasts, and scoring them using `scoringutils`[@scoringutils].
-All functionality is modular allowing users to extend and alter the underlying model whilst continuing to use the package framework.
-
-## Evaluation
-
-**TODO: Model evaluation. What targets. What scoring measures. What aggregation**
-**TODO: Evaluation of performance for submitted models vs others submitted to forecasting hub.**
-**TODO: Add links to nowcasting hub and evaluation against those models.**
-
 We evaluate these models first visually across a range of nowcasting dates and then quantitatively using proper scoring rules [@scoringutils] on both the natural and log scales (corresponding to absolute and relative performance) aggregating scores first across all targets and then stratifying in turn by age group, nowcast horizon, date of postive test, and date of report.
@@
 
-## Reproducibility
+All models are implemented using the `epinowcast` R package [@epinowcast].
+The nowcasting and evaluation pipeline is implemented using the `targets` R package[@targets].
+All input data, interim data, and output data are available and should also be fully reproducible from the provided code.
+Please see the resources section for details.
+Further details on our methodology are included in our [paper](https://epiforecasts.io/eval-germany-sp-nowcasting/paper.pdf).
 
-All analysis was carried out using R version 4.1.2 [@R].
-The analysis pipeline described here is available as a `targets` workflow [@targets] along with an archive of all interim results generated using `gittargets` [@gittargets] and stored using `piggyback` [@piggyback].
-A Dockerfile, and prebuilt archived image, has been made available with the code to enhance reproducibility [@Boettiger:2015dw].
+# Resources
 
-# Resullts
+```{r, child = "../chunks/documentation.Rmd"}
+```
 
-*This project is still underway and so results are not finalised.
-Please see our [real-time evaluation report](https://epiforecasts.io/eval-germany-sp-nowcasting/real-time/#Summary) for our initial findings.*
-
-# Dicussion
-
-**TODO: Summarise project**
-**TODO: Discussion strengths and limitations of the evaluation**
-**TODO: Compare approach evaluated here to the literature**
-**TODO: Discuss further work**
-**TODO Summarise conclusions.**
-
-**Software availability**
-
-Source code is available from: https://github.com/epiforecasts/eval-germany-sp-nowcasting/
-
-License: MIT
-
-Models are implemented in the `epinowcast` R package. This is available from: https://github.com/epiforecasts/epinowcast/
-
-**Data availability**
-
-Input data and summarised results are available from: https://github.com/epiforecasts/eval-germany-sp-nowcasting/tree/main/data
-
-Data from all interim stages of the analysis is from: https://github.com/epiforecasts/eval-germany-sp-nowcasting/releases/tag/latest
-
-License: MIT
-
-**Grant information**
-
-This work was supported by the Wellcome Trust through a Wellcome Senior Research Fellowship to SF [210758].
-
-**Competing interests**
-
-There are no competing interests.
-
-## References
-
-<div id = 'refs'></div>
+# References

### 2022-05-17 README.Rmd
simplify and add requirments

@@
 knitr::opts_chunk$set(
-  echo = TRUE, eval = FALSE,
-  fig.width = 9, fig.height = 9,
-  dpi = 320,
-  fig.path = "figures/"
+  echo = TRUE, eval = FALSE
 )
@@
 
-This repository contains the documentation, results, and code of a project evaluating the use of a semi-parametric nowcasting approach for COVID-19 hospitalisations in Germany.
-See the documentation for further details.
-This project is part of a [wider collaboration](https://covid19nowcasthub.de) assessing a range of nowcasting methods whilst providing an ensemble nowcast of COVID-19 Hospital admissions in Germany by date of positive test.
-This ensemble should be used for any policy related work rather than the nowcasts provided in this repository.
-See [here](https://covid19nowcasthub.de) for more on this nowcasting collaboration.
+This repository contains the documentation, results, and code of a project evaluating ... See the documentation for further details.
 
 ## Citation
-
-If making use of the results of this analysis or reusing the analysis pipeline please cite:
-
-If making using of the models evaluated in this analysis please also cite [`epinowcast`](https://epiforecasts.io/epinowcast):
-
-> Sam Abbott (2021). epinowcast: Hierarchical nowcasting of right censored epidemological counts, DOI: 10.5281/zenodo.5637165
-
-A BibTeX entry for LaTeX users is also available:
-
-  @Article{,
-    title = {epinowcast: Hierarchical nowcasting of right censored epidemological counts},
-    author = {Sam Abbott},
-    journal = {Zenodo},
-    year = {2021},
-    doi = {10.5281/zenodo.5637165},
-  }
 
@@
 ---|---
-[`writeup`](writeup/) | Summary paper and additional supplementary information as `Rmarkdown` documents.
+[`writeup`](docs/) | Summary paper and additional supplementary information as `Rmarkdown` documents.
 [`_targets.Rmd`](_targets.Rmd) | Analysis workflow for interactive use.
@@
 [`data`](data/) | Input data and summarised output generated by steps in the analysis.
-[`analyses`](analyses/) | Ad-hoc analyses not part of the overarching workflow. This includes a synthetic case study and a simplified example using Germany hospitalisation data.
 [`.devcontainer`](.devcontainer/) | Resources for reproducibility using `vscode` and `docker`.
@@
 
-All dependencies can be installed using the following,
+Dependencies are managed using [`renv`].
 
-```{r}
-remotes::install_dev_deps()
-```
-
-Alternatively a docker [container](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/.devcontainer/Dockerfile) and [image](https://github.com/epiforecasts/eval-germany-sp-nowcasting/pkgs/container/eval-germany-sp-nowcasting) is provided.
+Alternatively a docker [container](https://github.com/epiforecasts/simplfied-forecaster-evaluation/blob/main/.devcontainer/Dockerfile) and [image](https://github.com/epiforecasts/simplfied-forecaster-evaluation/pkgs/container/simplfied-forecaster-evaluation) is provided.
 An easy way to make use of this is using the Remote development extension of `vscode`.
@@
 ## Analyses
-
-This analysis in this repository has been implemented using the [`targets`](https://docs.ropensci.org/targets/) package and associated packages.
-The workflow is defined in [`_targets.md`](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/_targets.md) and can be explored interactively using [`_targets.Rmd`](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/_targets.Rmd) `Rmarkdown` document.
-The workflow can be visualised as the following graph.
-
-![](figures/targets-graph.png)
-
-This complete analysis can be recreated using the following (note this may take quite some time even with a fairly large amount of available compute),
-
-```{bash}
-bash bin/update-targets.sh
-```
-
-Alternative the following `targets` functions may be used to interactively explore the workflow:
-
-- Run the workflow sequentially.
-
-```{r}
-targets::tar_make()
-```
-
-- Run the workflow using all available workers.
-
-```{r}
-targets::tar_make_future(workers = future::availableCores())
-```
-
-- Explore a graph of the workflow.
-
-```{r}
-targets::tar_visnetwork(targets_only = TRUE)
-```
-
-Watch the workflow as it runs in a `shiny` app.
-
-```{r}
-targets::tar_watch(targets_only = TRUE)
-```
-
-To use our archived version of the interim results (and so avoid long run times) use the following to download it.
-Note that this process has not been rigorously tested across environments and so may not work seamlessly).
-
-```{r}
-source(here::here("R", "targetss-archive.R"))
-get_targets_archive()
-```

### 2022-05-17 docs/index.Rmd
simplify and add requirments

---
title: "Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany"
subtitle: "Summary"
author: Sam Abbott, Sebastian Funk
bibliography: ../../writeup/library.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 4
    includes:
      before_body: header.html
      after_body: footer.html
---

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, dpi = 330, warning = FALSE)
```

# Background

```{r, child = "../chunks/introduction.Rmd"}
```

# Resources

```{r, child = "../chunks/documentation.Rmd"}
```

# References

### 2022-05-17 R/evaluation.R

@@ -1,12 +1,0 @@
-score_nowcast <- function(nowcast, obs, by = "model") {
-  scores <- epinowcast::enw_score_nowcast(
-      nowcast, obs, log = FALSE,
-      summarise_by = by
-  )
-  log_scores <- epinowcast::enw_score_nowcast(
-      nowcast, obs, log = TRUE,
-      summarise_by = by
-  )
-  scores <- rbind(scores[, scale := "natural"], log_scores[, scale := "log"])
-  return(scores[])
-}

### 2022-05-17 R/models.R

@@ -1,259 +1,0 @@
-fixed_epinowcast <- function(obs, max_delay = 40, ...) {
-  pobs <- enw_preprocess_data(
-    obs,
-    rep_holidays = "holiday",
-    max_delay = max_delay, by = "age_group"
-  )
-
-  reference_effects <- enw_formula(pobs$metareference)
-  report_effects <- enw_formula(pobs$metareport)
-
-  model <- load_model()
-
-  nowcast <- epinowcast(
-    pobs,
-    model = model,
-    reference_effects = reference_effects,
-    report_effects = report_effects,
-    ...
-  )
-
-  nowcast <- default_nowcast_on_error(nowcast, pobs, model, ...)
-
-  out <- summarise_nowcast(nowcast, model = "Reference: Fixed, Report: Fixed")
-  return(out[])
-}
-
-dow_epinowcast <- function(obs, max_delay = 40, ...) {
-  pobs <- enw_preprocess_data(
-    obs,
-    rep_holidays = "holiday",
-    max_delay = max_delay, by = "age_group"
-  )
-
-  reference_effects <- enw_formula(pobs$metareference)
-  report_effects <- enw_formula(pobs$metareport, random = "day_of_week")
-
-  model <- load_model()
-
-  nowcast <- epinowcast(
-    pobs,
-    model = model,
-    reference_effects = reference_effects,
-    report_effects = report_effects,
-    ...
-  )
-
-  nowcast <- default_nowcast_on_error(nowcast, pobs, model, ...)
-
-  out <- summarise_nowcast(
-    nowcast,
-    model = "Reference: Fixed, Report: Day of week"
-  )
-  return(out[])
-}
-
-age_epinowcast <- function(obs, max_delay = 40, ...) {
-  pobs <- enw_preprocess_data(
-    obs,
-    rep_holidays = "holiday",
-    max_delay = max_delay, by = "age_group"
-  )
-
-  reference_effects <- enw_formula(pobs$metareference, random = "age_group")
-  report_effects <- enw_formula(pobs$metareport, random = "day_of_week")
-
-  model <- load_model()
-
-  nowcast <- epinowcast(
-    pobs,
-    model = model,
-    reference_effects = reference_effects,
-    report_effects = report_effects,
-    ...
-  )
-
-  nowcast <- default_nowcast_on_error(nowcast, pobs, model, ...)
-
-  out <- summarise_nowcast(
-    nowcast,
-    model = "Reference: Age, Report: Day of week"
-  )
[…]

### 2022-05-17 R/plot.R

@@ -1,59 +1,0 @@
-plot_nowcast <- function(summarised_nowcast, latest_obs,
-                                     max_delay = Inf, ...) {
-   summarised_nowcast <- summarised_nowcast[
-        reference_date >= (max(as.Date(nowcast_date)) - max_delay)
-   ]
-
-   suppressWarnings(summarised_nowcast[, holiday := NULL])
-
-   plot <- epinowcast::enw_plot_nowcast_quantiles(
-      summarised_nowcast,
-      latest_obs = latest_obs[
-        reference_date >= min(summarised_nowcast$reference_date)][
-        reference_date <= max(summarised_nowcast$reference_date)
-      ], ...
-  )
-  return(plot)
-}
-
-plot_scores <- function(scores, ...) {
-  ggplot2::ggplot(scores) +
-    ggplot2::aes(...) +
-    ggplot2::geom_point(size = 1.2) +
-    ggplot2::geom_line(size = 1.1, alpha = 0.4) +
-    ggplot2::scale_fill_brewer(palette = "Paired") +
-    ggplot2::scale_color_brewer(palette = "Paired") +
-    labs(y = "Weighted interval score") +
-    theme_bw() +
-    theme(legend.position = "bottom") +
-    guides(fill = guide_legend(title = "Model", nrow = 4),
-           col = guide_legend(title = "Model", nrow = 4))
-}
-
-plot_relative_scores <- function(score, baseline) {
-  score <- data.table::as.data.table(score)
-  fixed_score <- score[
-    model %in% baseline,
-    .(reference_date, age_group, fixed_is = interval_score)
-  ]
-  score <- merge(score, fixed_score, by = c("reference_date", "age_group"))
-
-  score <- score[, interval_score := interval_score / fixed_is]
-  score <- score[!model %in% baseline]
-  plot <- ggplot(score) +
-    aes(x = reference_date, y = interval_score, col = model) +
-    geom_hline(yintercept = 1, linetype = 2, size = 1.2, alpha = 0.5) +
-    geom_line(size = 1.1, alpha = 0.6) +
-    geom_point(size = 1.2) +
-    facet_wrap(vars(age_group)) +
-    scale_color_brewer(palette = "Paired") +
-    scale_y_log10(labels = scales::percent)
-
-  plot <- enw_plot_theme(plot) +
-    labs(
-      x = "Reference date",
-      y = paste0("Weighted interval score (", baseline, ")")
-    ) + # nolint
[…]

### 2022-05-17 R/postprocess.R

@@ -1,124 +1,0 @@
-adjust_quantile <- function(quantile, median, max_ratio = 0.5) {
-  q_ratio <- quantile / median
-
-  if (max(q_ratio, na.rm = TRUE) > 1) {
-    quantile <- ifelse(
-      q_ratio >= max_ratio + 1,
-      (max_ratio + 1) * median,
-      quantile
-    )
-  }else{
-    quantile <- ifelse(
-      q_ratio <= (1 - max_ratio),
-      max_ratio * median,
-      quantile
-    )
-  }
-  return(quantile)
-}
-
-adjust_posteriors <- function(nowcasts, target, max_ratio = 0.25,
-                              rhat_bound = 1.1, per_dt_bound = 0.2) {
-  unadjusted_nowcasts <- nowcasts[
-    !(max_rhat > rhat_bound | per_divergent_transitions > per_dt_bound)
-  ]
-  adjusted_nowcasts <- nowcasts[
-    max_rhat > rhat_bound | per_divergent_transitions > per_dt_bound
-  ]
-  if (nrow(adjusted_nowcasts) > 0) {
-    adjusted_nowcasts <- adjusted_nowcasts[,
-      (target) := purrr::map(
-        get(target),
-        function(dt) {
-          cols <- grep("^q[0-9]", colnames(dt), value = TRUE)
-          dt[,
-            (cols) := purrr::map(.SD, adjust_quantile, median = median,
-                                  max_ratio = max_ratio),
-            .SDcols = cols
-          ]
-        })
-    ]
-  nowcasts <- rbind(unadjusted_nowcasts, adjusted_nowcasts)
-  }
-  return(nowcasts)
-}
-
-unnest_nowcasts <- function(nowcasts, target) {
-
-  nowcasts <- nowcasts[,
-    (target) := pmap(
-      list(get(target), model, nowcast_date),
-        function(df, m, d) {
-          df[, model := m][, nowcast_date := d]
-        })][,
-    rbindlist(get(target), use.names = TRUE, fill = TRUE)][
-    order(location, age_group, nowcast_date, reference_date, model)
-    ]
-
-  data.table::setcolorder(nowcasts, neworder = c("model", "nowcast_date"))
-  nowcasts[, nowcast_date := as.Date(nowcast_date)]
-  nowcasts[, report_date := nowcast_date]
-  return(nowcasts[])
-}
-
-format_for_submission <- function(nowcast, horizon = -28,
[…]

### 2022-05-17 R/utils.R

@@ -1,98 +1,0 @@
-#' Save a plot and return the path for targets
-#' @importFrom ggplot2 ggsave
-save_plot <- function(plot, filename, path, ...) {
-  dir.create(path, recursive = TRUE, showWarnings = FALSE)
-  path <- file.path(path, file)
-  ggplot2::ggsave(path, plot, ...)
-  return(filename)
-}
-
-#' Save a dataframe to a csv and return the path for targets
-save_csv <- function(dt, filename, path, allow_empty = TRUE) {
-  dir.create(path, recursive = TRUE, showWarnings = FALSE)
-  path <- file.path(path, filename)
-
-  if (allow_empty | nrow(dt) > 0) {
-    data.table::fwrite(dt, path)
-  }
-  return(path)
-}
-
-drop_string <- function(var, string) {
-  var[!grepl(string, var)]
-}
-
-select_var <- function(dt, var) {
-  dt[[var]]
-}
-
-format_from_csv <- function(dt) {
-  if (!is.null(dt$age_group)) {
-      dt[
-      ,
-        age_group := factor(
-          age_group,
-          levels = c("00+", "00-04", "05-14", "15-34", "35-59", "60-79", "80+")
-        )
-    ]
-    }
-    dt[,
-    model := factor(
-      model,
-      levels = c("Reference: Fixed, Report: Fixed",
-                "Reference: Fixed, Report: Day of week",
-                "Reference: Age, Report: Day of week",
-                "Reference: Age and week, Report: Day of week",
-                "Reference: Age and week by age, Report: Day of week",
-                "Independent by age, Reference: Week, Report: Day of week",
-                "Independent by age, Reference: Week and day of week, Report: Day of week" # nolint
-                )
-    )
-    ]
-  return(dt[])
-}
-
-load_nowcasts <- function(path) {
-  nowcasts <- fs::dir_ls(
-    path,
-    glob = "*.csv"
-  ) |>
-    purrr::map(data.table::fread) |>
-    data.table::rbindlist(use.names = TRUE, fill = TRUE)
-  nowcasts[, horizon := as.numeric(
-    as.Date(reference_date) - as.Date(nowcast_date)
-  )]
-  nowcasts <- format_from_csv(nowcasts)
-  return(nowcasts[])
-}
-
-load_obs <- function(path) {
-  obs <- fread(path)
-  obs[
-    ,
-    age_group := factor(
[…]

### 2022-05-17 analyses/synthetic-example.R

@@ -1,63 +1,0 @@
-library(epinowcast)
-library(covidregionaldata)
-suppressMessages(library(data.table, quietly = TRUE))
-suppressMessages(library(rstan, quietly = TRUE))
-library(purrr)
-library(here)
-
-# load dev code
-functions <- list.files(here("R"), full.names = TRUE)
-walk(functions, source)
-
-# set number of cores to use
-options(mc.cores = 4)
-# get example case counts
-start_using_memoise()
-latest_cases <- get_national_data("UK", verbose = FALSE)
-latest_cases <- setDT(latest_cases)[date >= as.Date("2021-07-01")]
-latest_cases <- latest_cases[, .(date, confirm = cases_new)]
-
-# get a range of dates to generate synthetic data for
-scenarios <- enw_random_intercept_scenario(
-  obs = latest_cases,
-  snapshots = seq(6 * 7, 0, by = -1),
-  logmean = 1.9, logmean_sd = 0.1,
-  logsd = 1, logsd_sd = 0.1
-)
-
-# simulate observations
-scenarios <- enw_simulate_lnorm_trunc_obs(scenarios, latest_cases)
-sim_reported_cases <- rbindlist(scenarios$reported_cases)
-setnames(sim_reported_cases, "date", "reference_date")
-sim_reported_cases <- rbind(
-  copy(sim_reported_cases)[, age := 1],
-  copy(sim_reported_cases)[, age := 2]
-)
-
-# Preprocess data
-pobs <- enw_preprocess_data(sim_reported_cases, by = "age")
-
-# Construct design matrices for the desired reference date effects
-reference_effects <- enw_formula(pobs$metareference[[1]])
-
-# Construct design matrices for the desired report day effects
-report_effects <- enw_formula(pobs$metareport[[1]], random = "day_of_week")
-
-# fit model to example data and produce nowcast
-est <- epinowcast(pobs,
-  report_effects = report_effects,
-  reference_effects = reference_effects,
-  debug = FALSE, pp = TRUE, save_warmup = FALSE
-)
-
-# summarise nowcast
-summary(est)
-
-# Plot nowcast vs latest observations
-plot(est, latest_obs = copy(latest_cases)[, reference_date := date])
-
-# Plot posterior prediction for observed cases at date of report
-plot(est,
-  latest_obs = copy(latest_cases)[, reference_date := date],
-  type = "pos", log = TRUE
-)

### 2022-05-17 bin/update-git-remote.sh

@@ -1,7 +1,0 @@
-#!/bin/bash
-
-git add data _targets_r _targets.md _targets.Rmd
-date=$(date +'%Y-%m-%d')
-git commit -m "$date - nowcast update"
-git pull -Xours
-git push

### 2022-05-17 bin/update-targets.sh

@@ -1,5 +1,0 @@
-#!/bin/bash
-
-bash bin/render-targets.sh > render.log 2>&1
-
-nohup bash bin/make-parallel-targets.sh > targets.log 2>&1 &

### 2022-05-17 R/get-data.R
further pruning and add renv support

@@ -1,52 +1,0 @@
-get_public_holidays <- function(today = Sys.Date(),
-                                url = "https://raw.githubusercontent.com/KITmetricslab/hospitalization-nowcast-hub/main/other_data/public_holidays.csv") { # nolint
-  message("Downloading public holiday data available on the: ", today)
-
-  holidays <- data.table::fread(url)
-  holidays <- data.table::melt(holidays, id.vars = "date")
-  holidays <- holidays[value == TRUE]
-  holidays <- holidays[
-    ,
-    .(report_date = date, location = variable, holiday = value)
-  ]
-  return(holidays[])
-}
-
-get_germany_hospitalisations <- function(today = Sys.Date(),
-                                         url = "https://raw.githubusercontent.com/KITmetricslab/hospitalization-nowcast-hub/main/data-truth/COVID-19/COVID-19_hospitalizations_preprocessed.csv") { # nolint
-  message("Downloading hospitalisation data available on the: ", today)
-
-  germany_hosp <- data.table::fread(url)
-
-  germany_hosp <- data.table::melt(
-    germany_hosp,
-    variable.name = "delay",
-    value.name = "confirm",
-    id.vars = c("date", "location", "age_group")
-  )
-  data.table::setnames(germany_hosp, "date", "reference_date")
-
-  germany_hosp[, report_date := as.Date(reference_date) + 0:(.N - 1),
-    by = c("reference_date", "location", "age_group")
-  ]
-  germany_hosp <- germany_hosp[report_date <= max(reference_date)]
-  germany_hosp[, delay := NULL]
-  germany_hosp[is.na(confirm), confirm := 0]
-  germany_hosp[, confirm := cumsum(confirm),
-    by = c("reference_date", "location", "age_group")
-  ]
-  germany_hosp[
-    ,
-    age_group := factor(
-      age_group,
-      levels = c("00+", "00-04", "05-14", "15-34", "35-59", "60-79", "80+")
-    )
-  ]
-  holidays <- get_public_holidays(today)
-  germany_hosp <- merge(
-    germany_hosp, holidays,
-    by = c("report_date", "location"), all.x = TRUE, all.y = FALSE
-  )
-  germany_hosp[is.na(holiday), holiday := FALSE]
-  return(germany_hosp[])
-}

### 2022-05-17 _targets_r/targets/age_groups.R

@@ -1,3 +1,0 @@
-tar_target(age_groups, {
-  unique(hospitalisations$age_group)
-})

### 2022-05-17 _targets_r/targets/age_nowcast.R

@@ -1,13 +1,0 @@
-tar_target(
-  age_nowcast,
-  nowcast(
-    obs = hospitalisations_by_date_report,
-    tar_loc = locations,
-    model = age_epinowcast,
-    priors = priors,
-    max_delay = max_report_delay,
-    settings = epinowcast_settings
-  ),
-  cross(hospitalisations_by_date_report, locations),
-  cue = tar_cue(mode = "never")
-)

### 2022-05-17 _targets_r/targets/age_week_nowcast.R

@@ -1,13 +1,0 @@
-tar_target(
-  age_week_nowcast,
-  nowcast(
-    obs = hospitalisations_by_date_report,
-    tar_loc = locations,
-    model = age_week_epinowcast,
-    priors = priors,
-    max_delay = max_report_delay,
-    settings = epinowcast_settings
-  ),
-  cross(hospitalisations_by_date_report, locations),
-  cue = tar_cue(mode = "never")
-)

### 2022-05-17 _targets_r/targets/combined_nowcasts.R

@@ -1,23 +1,0 @@
-tar_target(combined_nowcasts, {
-  rbindlist(list(
-    fixed_nowcast,
-    dow_nowcast,
-    age_nowcast,
-    week_nowcast,
-    age_week_nowcast,
-    independent_nowcast,
-    overall_only_nowcast,
-    independent_ref_dow_nowcast
-  ), use.names = TRUE, fill = TRUE)[,
-     model := factor(
-      model,
-      levels = c("Reference: Fixed, Report: Fixed",
-                 "Reference: Fixed, Report: Day of week",
-                 "Reference: Age, Report: Day of week",
-                 "Reference: Age and week, Report: Day of week",
-                 "Reference: Age and week by age, Report: Day of week",
-                 "Independent by age, Reference: Week, Report: Day of week",
-                 "Independent by age, Reference: Week and day of week, Report: Day of week")
-     )
-    ][, id := 1:.N]
-})

### 2022-05-17 _targets_r/targets/complete_7day_hospitalisations.R

@@ -1,3 +1,0 @@
-tar_target(complete_7day_hospitalisations, {
-  latest_7day_hospitalisations[reference_date < (max(reference_date) - 28)]
-})

### 2022-05-17 _targets_r/targets/epinowcast_model.R

@@ -1,5 +1,0 @@
-tar_file(
-  epinowcast_model,
-  compile_model(),
-  deployment = "main"
-)

### 2022-05-17 _targets_r/targets/fixed_nowcast.R

@@ -1,13 +1,0 @@
-tar_target(
-  fixed_nowcast,
-  nowcast(
-    obs = hospitalisations_by_date_report,
-    tar_loc = locations,
-    model = fixed_epinowcast,
-    priors = priors,
-    max_delay = max_report_delay,
-    settings = epinowcast_settings
-  ),
-  cross(hospitalisations_by_date_report, locations),
-  cue = tar_cue(mode = "never")
-)

### 2022-05-17 _targets_r/targets/hospitalisation_url.R

@@ -1,4 +1,0 @@
-tar_url(
-  hospitalisation_url,
-  "https://raw.githubusercontent.com/KITmetricslab/hospitalization-nowcast-hub/main/data-truth/COVID-19/COVID-19_hospitalizations_preprocessed.csv" # nolint
-)

### 2022-05-17 _targets_r/targets/hospitalisations.R

@@ -1,3 +1,0 @@
-tar_target(hospitalisations, {
-  get_germany_hospitalisations(url = hospitalisation_url)
-})

### 2022-05-17 _targets_r/targets/hospitalisations_by_date_report.R

@@ -1,11 +1,0 @@
-tar_target(
-  hospitalisations_by_date_report,
-  enw_retrospective_data(
-    hospitalisations,
-    rep_date = nowcast_dates,
-    ref_days = max_report_delay
-  )[, nowcast_date := nowcast_dates],
-  map(nowcast_dates),
-  iteration = "list",
-  cue = tar_cue(mode = "never")
-)

### 2022-05-17 _targets_r/targets/independent_submission_nowcast.R

@@ -1,19 +1,0 @@
-tar_target(
-  independent_submission_nowcast,
-  rbind(
-    independent_ref_dow_nowcast[nowcast_date == nowcast_dates],
-    overall_only_nowcast[nowcast_date == nowcast_dates]
-  ) |> 
-   adjust_posteriors(
-      target = "seven_day", 
-      max_ratio = 0.25, 
-      rhat_bound = 1.1,
-      per_dt_bound = 0.2
-    ) |>
-    select_var("seven_day") |>
-    rbindlist() |>
-    format_for_submission(),
-  map(nowcast_dates),
-  iteration = "list",
-  cue = tar_cue(mode = "never")
-)

### 2022-05-17 _targets_r/targets/latest_7day_hospitalisations.R

@@ -1,5 +1,0 @@
-tar_target(latest_7day_hospitalisations, {
-  copy(latest_hospitalisations)[,
-    confirm := frollsum(confirm, n =  7), by = c("age_group", "location")
-  ][!is.na(confirm)]
-})

### 2022-05-17 _targets_r/targets/locations.R

@@ -1,3 +1,0 @@
-tar_target(locations, {
-  "DE"
-})

### 2022-05-17 _targets_r/targets/max_report_delay.R

@@ -1,3 +1,0 @@
-tar_target(max_report_delay, {
-  40
-})

### 2022-05-17 _targets_r/targets/nowcast_dates.R

@@ -1,5 +1,0 @@
-tar_target(nowcast_dates, {
-  unique(
-    hospitalisations[reference_date >= start_date]$reference_date
-  )
-})

### 2022-05-17 _targets_r/targets/prior_obs.R

@@ -1,8 +1,0 @@
-tar_target(
-  prior_obs,
-  enw_retrospective_data(
-    hospitalisations[location == "DE"][age_group == "00+"],
-    rep_date = as.Date("2021-07-01"), ref_days = max_report_delay
-  ),
-  cue = tar_cue(mode = "never")
-)

### 2022-05-17 _targets_r/targets/priors.R

@@ -1,9 +1,0 @@
-tar_target(priors, {
-  do.call(prior_epinowcast, c(
-    list(
-      prior_obs, max_delay = max_report_delay, scale = 10,
-      priors = uninformed_priors
-    ),
-    epinowcast_settings
-  ))
-})

### 2022-05-17 _targets_r/targets/save-run-time.R

@@ -1,9 +1,0 @@
-tar_file(
-  save_run_time,
-  combined_nowcasts |>
-    summarise_runtimes() |>
-    save_csv(
-      filename = "run-times.csv",
-      path = here("data/diagnostics")
-    )
-)

### 2022-05-17 _targets_r/targets/save_7day_nowcasts.R

@@ -1,11 +1,0 @@
-tar_file(
-  save_7day_nowcasts,
-  summarised_7day_nowcast[nowcast_date == nowcast_dates] |>
-    save_csv(
-      filename = paste0(nowcast_dates, ".csv"),
-      path = here("data/nowcasts/seven_day"),
-      allow_empty = FALSE
-    ),
-  map(nowcast_dates),
-  cue = tar_cue(mode = "never")
-)

### 2022-05-17 _targets_r/targets/save_daily_nowcasts.R

@@ -1,11 +1,0 @@
-tar_file(
-  save_daily_nowcasts,
-  summarised_nowcast[nowcast_date == nowcast_dates] |>
-    save_csv(
-      filename = paste0(nowcast_dates, ".csv"),
-      path = here("data/nowcasts/daily"),
-      allow_empty = FALSE
-    ),
-  map(nowcast_dates),
-  cue = tar_cue(mode = "never")
-)

### 2022-05-17 _targets_r/targets/save_latest_7day_hospitalisations.R

@@ -1,8 +1,0 @@
-tar_file(
-  save_latest_7day_hospitalisations,
-  save_csv(
-    latest_7day_hospitalisations,
-    filename = paste0("seven_day.csv"),
-    path = here("data/observations")
-  )
-)

### 2022-05-17 _targets_r/targets/score.R

@@ -1,29 +1,0 @@
-tar_map(
-  list(score_by = list(
-    "overall", "age_group", "horizon", "reference_date", "nowcast_date"
-  )),
-  tar_target(
-    scores,
-    enw_score_nowcast(
-      scored_nowcasts, complete_hospitalisations, 
-      summarise_by = drop_string(c(score_by, "model"), "overall"),
-      log = FALSE
-    )
-  ),
-  tar_target(
-    log_scores,
-    enw_score_nowcast(
-      scored_nowcasts, complete_hospitalisations, 
-      summarise_by = drop_string(c(score_by, "model"), "overall"),
-      log = TRUE
-    )
-  ),
-  tar_file(
-    save_scores,
-    save_csv(
-      rbind(scores[, scale := "natural"], log_scores[, scale := "log"]),
-      filename = paste0(paste(score_by, sep = "-"), ".csv"),
-      path = here("data/scores")
-    )
-  )
-)

### 2022-05-17 _targets_r/targets/scored_nowcasts.R

@@ -1,8 +1,0 @@
-tar_target(scored_nowcasts, {
-  summarised_nowcast[location == locations][
-    reference_date < (max(nowcast_date) - 28)][,
-    holiday := NULL][,
-    horizon := as.numeric(as.Date(reference_date) - nowcast_date)][
-    horizon >= -7
-  ]
-})

### 2022-05-17 _targets_r/targets/summarised_nowcast.R

@@ -1,10 +1,0 @@
-tar_target(summarised_nowcast, {
-  combined_nowcasts |> 
-    adjust_posteriors(
-      target = "daily", 
-      max_ratio = 0.25, 
-      rhat_bound = 1.1,
-      per_dt_bound = 0.2
-  ) |> 
-    unnest_nowcasts(target = "daily") 
-})

### 2022-05-17 _targets_r/targets/uninformed_priors.R

@@ -1,3 +1,0 @@
-tar_target(uninformed_priors, {
-  enw_priors()
-})

### 2022-05-17 docs/paper/paper.Rmd
add some structure to the proposed paper

---
title: Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble
preprint: false
author: 
  - name: Sam Abbott
    affiliation: 1
    corresponding: true
    email: sam.abbott@lshtm.ac.uk
affiliation:
  - code: 1
    address: Cold Spring Harbor Laboratory, One Bungtown Road Cold Spring Harbor, NY 11724
abstract: > 
  **Background**

  **Methods**

  **Results**

  **Conclusions**
bibliography: library.bib
output:
  rticles::peerj_article: default
  bookdown::pdf_book:
    base_format: rticles::peerj_article # for using bookdown features like \@ref()
---

# Introduction {-}

- Why

- Extension of why

- What

- Aim:

# Methods {-}

## Data {-}

## Model definition {-}

## Forecast evaluation {-}

## Implementation {-}

# Resuls {-}

## Data summary {-}

- Date range of forecasts
- Forecasts per country
- Summarises general trends in incidence in the data

## Visualisation of forecasts by horizon {-}

- Figure 1:
  - X: date
  - Y: incidence rate
  - Colour: Model
  - X facet: Country
  - Y facet: Horizon

## Forecast evaluation {-}

- Figure 2:
  Sub-figure 1:
  Sub-figure 2:
  Sub-figure 3:

## Forecast calibration {-}

- Figure 3:
  Sub-figure 1:
  Sub-figure 2:
  Sub-figure 3:

# Discussion {-}

## Summary {-}

## Stengths and weaknesses {-}

## Literature context {-}

## Further work {-}

## Conclusions {-}

# Acknowledgments {-}

# References

### 2022-05-20 docs/paper/paper.Rmd
add methods outline

@@
 
-## Model definition {-}
+## Model {-}
+
+### Motivation
+
+The package currently supports two models both with various options.
+A baseline single strain model and a two strain model with flexible variable dynamics.
+Both share the same underlying structure with expected notifications being assumed to be a product of past expected notifications, and an exponentiated time-varying term.
+This term is then modelled in both instances as a differenced autoregressive process with a signal lag term.
+In the two strain model this process can either be modelled jointly for variants, as a correlated process, or independently.
+By default it is assumed to vary inline with the timestep of the data but this can be altered to speed up computation or to improve out of sample performance.
+Observed cases are estimated from expected cases by assuming either a negative binomial observation model.
+Detailed model definitions are given below.
+Prior values when given represent the package defaults and can, and generally should, altered by the user based on their domain specific knowledge.
+
+### Definition
+
+We model the mean ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(P)) process by time unit ($t$).
+The model is initialised by assuming that the initial reported cases are representative with a small amount of error (2.5%) for each $t \lt P$.
+
+\begin{align}
+  \lambda_t &\sim \text{LogNormal}\left(\log C_t , 0.025 \times \log C_t \right),\ t \leq P \\
+  \lambda_t &= \text{exp}\left(r_t\right) \sum_{p = 1}^{P} \alpha_p  \lambda_{t-p},\ t \gt P  
+\end{align}
+
+Where $r_t$ can be interpreted as the growth rate and the exponential of $r_t$ as the effective reproduction number ($R_t$) assuming a mean generation time equal to the scaling rate used (see the documentation for `scale_r` in `?forecast()`).
+$r_t$ is then itself modelled as a piecewise constant differenced AR(1) process,
+
+\begin{align}
+  r_1 &\sim \text{Normal}\left(0, 0.25 \right) \\
+  r_t &= r_{t-1} + \epsilon_t  \\
+  \epsilon_1 &= \eta_1 \\
+  \epsilon_t &= \beta \epsilon_{t-1} + \eta_t
+\end{align}
+
+Where,
+
+\begin{align}
+  \eta_t &\sim \text{Normal}\left(0, \sigma \right) \\
+  \sigma &\sim \text{Normal}\left(0, 0.2 \right) \\
+  \beta &\sim \text{Normal}\left(0, 0.5 \right)
+\end{align}
+
+We then assume a negative binomial observation model with overdispersion $\phi_c$ for reported cases ($C_t$),
+
+\begin{align}
+  C_{t} &\sim \text{NegBinomial}\left(\lambda_t, \phi_c\right) \\
+  \frac{1}{\sqrt{\phi_c}} &\sim \text{Normal}(0, 0.5) 
+\end{align}
+
+Where $\sigma$, and $\frac{1}{\sqrt{phi_c}}$ are truncated to be greater than 0 and $\beta$ is truncated to be between -1 and 1.
+Optionally a Poisson observation model may instead be used (see the documentation for `overdispersion` in `?forecast()`).
+
+### Summary statistics
+
+As well as posterior predictions and forecasts for notifcations  the model also return several epidemiological summary statistics which may be useful for drawing inferences about underlying transmission dynamics.
+These are the log scale growth rate ($g^{o, \delta}_t$), and the instantaneous effective reproduction number ($R^{o, \delta}_t$).
+These are calculated as follows:
+
+\begin{align}
+  g^{o, \delta}_t &= T_s r^{o, \delta}_t \\
+  R^{o, \delta}_t &= \text{exp}\left(T_s r^{o, \delta}_t\right) \\
+\end{align}
+
+$T_s$ is a user set scaling parameter that defines the timespan over which the summary metrics apply dependent on the time step of the data.
+It can be set using the `scale_r` and defaults to 1 which returns summary statistics scaled to the timestep of the data.
+Depending on the setup of the model used these summary measures will be more or less related to their epidemiological definitions.
+In particular, adding a weighting to past expected cases that is more complex than a simple lag may cause interpretation issues.
 
@@
 
-# Resuls {-}
+The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`) [@forecast.vocs].
+The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
+`cmstanr` surfaces several settings which trade-off between sampling speed and the robustness of approach.
+Here we take a conservative approach, as model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum treedepth setting to 15.
+For real-time usage convergance was not assessed but during model development the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergant transitions and exceedance of the maximum tree depth and posterior predictions visually compared to observed data [@cmdstanr].
+
+To download and manipulate forecasts from the ECDC forecasting hub [@EuroHub] we use  the `data.table` (`1.14.2`) [@data.table] and `gh` (`1.3.0`) [@gh] R packages.
+We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, to visualise forecasts and summary measures, and to summarise forecasts.
+Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils].
+
+To ensure reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with built Docker image [@Boettiger:2015dw] (via GitHub Actions) is provided.
+Weekly forecasts also made use of `renv` and Docker to ensure reproducibility with GitHub actions used to produce the forecasts ensuring we meet our limited compute goal and that our implementation is independent from available compute resources democratising  access to our approach.
+The code for this analysis can be found here: https://github.com/epiforecasts/simplified-forecaster-evaluation
+The code for the forecating model defined above along with the infrastructure required to forecast using GitHub Actions can be found here: https://github.com/seabbs/ecdc-weekly-growth-forecasts
+
+# Results {-}
 

### 2022-05-20 data-raw/get-hub-forecasts.R

@@ -23,7 +23,7 @@ models <- c("epiforecasts-weeklygrowth", "EuroCOVIDhub-ensemble")
 up_to <- Sys.Date()
 forecast_dates <- seq(as.Date("2022-01-15"), up_to, by = "day")
 
-# Check
+# Get forecasts for models of interest
 hub_forecasts <- get_hub_forecasts(
   "covid19-forecast-hub-europe/covid19-forecast-hub-europe",
   dates = forecast_dates,

### 2022-05-20 data-raw/update.sh

@@ -1,0 +1,11 @@
+
+# Download metadata on forecasts  in the window of interest.
+
+# Download hub forecasts of interest
+Rscript data-raw/get-hub-forecasts.R
+
+# Download truth data
+Rscript data-raw/get-truth.R
+
+# Calculate per forecast and quantile scores
+Rscript data-raw/score-forecasts.R

### 2022-05-20 docs/paper/paper.Rmd
some more planning detail

@@
 ## Data {-}
+
+- Data from JHU as used as truth data by the ECDC forecasting hub.
+  - When accessed
+  - Aggregated to epiweek as in the ECDC forecasting GitHub
+  - Countries included
+  - Dates of data included
+- Forecasts from the ECDC forecast hub
+  - Forecasts dowloaded for the model proposed here and the ECDC ensemble
+  - Forecast format
+  - Forecases from when and horizon
+  - Ensemble construction and reference
 
@@
 
-As well as posterior predictions and forecasts for notifcations  the model also return several epidemiological summary statistics which may be useful for drawing inferences about underlying transmission dynamics.
+As well as posterior predictions and forecasts for notifcations the model also returns several epidemiological summary statistics which may be useful for drawing inferences about underlying transmission dynamics.
 These are the log scale growth rate ($g^{o, \delta}_t$), and the instantaneous effective reproduction number ($R^{o, \delta}_t$).
@@
 ## Forecast evaluation {-}
+
+- Visualise forecasts across time horizons in examplar countries.
+- Weighted interval score.
+- Calibration and coverage.
+- Relative interval score.
+- Relative interval score overall, distribution of scores, by horizon both overall and distribution, by location overall and distribution.
+- Relative interval score by interval range by horizon
+- Calibration at 50 and 90 credible intervals
+- Coverage across intervals.
 
@@
 - Forecasts per country
+- Total number of forecast models and average per location and per forecast date
 - Summarises general trends in incidence in the data

### 2022-05-26 docs/paper/paper.Rmd
draft figure 1

@@
 abstract: > 
-  **Background**
-
-  **Methods**
-
-  **Results**
-
-  **Conclusions**
+  We made the best model. It is both simple and amazing.
 bibliography: library.bib
@@
 
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  eval = TRUE,
+  echo = FALSE,
+  message = FALSE
+)
+```
+
 # Introduction {-}
@@
 - Why
+  - ECDC and CDC forecasting hubs
+  - Ensemble shown to outperform individual models in many cases
+  - Thought of as being more desirable due to reliability etc vs multiple models
+  - Many models are complex and require large amounts of compute time.
+  - As the ensemble is made up of these models it means it is extremely resource demandning to construct requiring significant compute as well as multiple teams.
+  - As the ensemble is constructed from multiple models it functions as effectively a black box. This means it may be hard for policy-makers to understand its underyling assumptions
+  - Here we seek to refine forecasting ideas from our previous work alongside our experiances forecasting COVID-19 as contributors to the ECDC and CDC forecasting hubs to develop a simplified forecasting model that has characteristics fo the ensemble forecast whilst maintianing a interpretability and low resource demands.
 
 - Extension of why
+   - Brief dicussion of implications of modelling NPIs/future changes in behaviour
+   - Brief discussion of reporting patterns in data. Daily/Weekly.
+   - Brief discussion of EpiNow2 and related models
+   - Brief dicussion of statistical modelling approaches (use MUNI-ARIMA as an example).
+   - Brief discussion of scoring of hub models
+   - Summary of observed performance of the hub ensembles
+    - Auto-correlated
+    - Typically relatively trend following though shows some ability to forecast future changes.
 
 - What
+  - Simplified auto-regressive growth rate model assuming a non-stationary trend
+  -  Explicit modelling of the assumption that rapid growth/decay of cases will typically result in future growth tending towards zero.
+This seeks to capture potential implementation and relaxation of NPIs as well as individuals attitudes to risk (assuming this is based on current growth rates and not case incidence).
+  - Uses weekly vs daily data to side step issues of reporting periodicity and to replicate the observed auto-corelation of the hub ensembles.
+  - Simple extension of an already implemented model and resuse of its modular components.
+  - Implemented with a focus on reproducibility, simplicity, and low resource use. 
+  - The model runs using GitHub actions in a framework supplied by the ECDC forecast hub. We provide our own tools so that others can run the model using GitHub actions democratising our approach.
 
-- Aim:
+- Aim: In this work we define a simplified forecasting model designed to replicate some of the observed characteristics of successful infectious disease ensemble models without sacrificing interpretability and with low resource requirements.
+We evaluate the performance of this model against the ECDC forecasting hub ensemble in real-tim, and discuss implications of our findings along with areas for future research.
 
@@
 ## Data {-}
+
+```{r packages}
+library(data.table)
+library(ggplot2)
+library(scales)
+library(lemon)
+library(here)
+```
+
+```{r load-functions}
+source(here("R", "utils.R"))
+```
+
+```{r data}
+# Load JHU frozen truth data. See data-raw/get-truth.R for munging
+truth <- fread(here("data", "truth.csv"))
+
+# Load population data
+population <- fread(here("data", "population.csv"))
+
+# Load ECDC forecast metadata. See data-raw/get-hub-metadata.R for munging
+#metadata <- fread(here("data", "forecast-metadata.csv"))
+
+# Load ECDC forecasts. See data-raw/get-hub-forecasts.R for munging
+# Merge with truth and rescale to incidence rate per 100,000
+forecasts <- fread(here("data", "forecasts.csv")) |>
+  merge_forecasts_with_truth(truth) |>
+  rescale_to_incidence_rate(population, scale = 1000)
+```
 
@@
 
-The package currently supports two models both with various options.
-A baseline single strain model and a two strain model with flexible variable dynamics.
-Both share the same underlying structure with expected notifications being assumed to be a product of past expected notifications, and an exponentiated time-varying term.
-This term is then modelled in both instances as a differenced autoregressive process with a signal lag term.
-In the two strain model this process can either be modelled jointly for variants, as a correlated process, or independently.
+The same underlying structure with expected notifications being assumed to be a product of past expected notifications, and an exponentiated time-varying term.
+This term is then modelled in both instances as a differenced autoregressive process with a single lag term.
 By default it is assumed to vary inline with the timestep of the data but this can be altered to speed up computation or to improve out of sample performance.
-Observed cases are estimated from expected cases by assuming either a negative binomial observation model.
-Detailed model definitions are given below.
-Prior values when given represent the package defaults and can, and generally should, altered by the user based on their domain specific knowledge.
+Observed cases are estimated from expected cases by assuming a negative binomial observation model.
+A detailed model definition is given below.
 
@@
 
-We model the mean ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(P)) process by time unit ($t$).
-The model is initialised by assuming that the initial reported cases are representative with a small amount of error (2.5%) for each $t \lt P$.
+We model the expectation ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(P)) process by epidemiological week ($t$).
+The model is initialised by assuming that the initial reported cases are representative with a small amount of error (2.5%).
+We assume a negative binomial observation model with overdispersion $\phi$ for reported cases ($C_t$).
 
-\begin{align}
-  \lambda_t &\sim \text{LogNormal}\left(\log C_t , 0.025 \times \log C_t \right),\ t \leq P \\
-  \lambda_t &= \text{exp}\left(r_t\right) \sum_{p = 1}^{P} \alpha_p  \lambda_{t-p},\ t \gt P  
-\end{align}
+\begin{align*}
+  \lambda_0 &\sim \text{LogNormal}\left(\log C_0 , 0.025 \times \log C_0 \right)\\
+  \lambda_t &= C_{t-1} e^{r_t},\ t > 0  \\
+  C_{t} \mid \lambda_{t} &\sim \text{NB}\left(\lambda_t, \phi\right)
+\end{align*}
 
-Where $r_t$ can be interpreted as the growth rate and the exponential of $r_t$ as the effective reproduction number ($R_t$) assuming a mean generation time equal to the scaling rate used (see the documentation for `scale_r` in `?forecast()`).
-$r_t$ is then itself modelled as a piecewise constant differenced AR(1) process,
+Where $r_t$ can be interpreted as the growth rate.
+$r_t$ is then modelled as a piecewise constant differenced AR(1) process modified such that the dependence of $r_{t-1}$ is multiplied by a decay factor ($\xi_{+,-}$) that varies dynamically according to the sign of $r_{t-1}$.
+The assumptions of this modelling approach are that the growth rate is non-stationary with a trend that is independent of the current growth rate (the differenced AR(1) process), the additional decay factor encodes the belief that larger absolute growth rates will tend more quickly towards no growth and that this process may work differently for positive or negative growth rates.
+This process can be defined as follows,
 
-\begin{align}
-  r_1 &\sim \text{Normal}\left(0, 0.25 \right) \\
-  r_t &= r_{t-1} + \epsilon_t  \\
-  \epsilon_1 &= \eta_1 \\
-  \epsilon_t &= \beta \epsilon_{t-1} + \eta_t
-\end{align}
+\begin{align*}
+  r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
+  r_t &= \left(\mathcal{H}(r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
+  \epsilon_t &= \mathcal{H}(t > 0) \beta \epsilon_{t-1} + \eta_t
+\end{align*}
 
-Where,
+Where $\mathcal{H}$ is the Heaviside step function and is defined such that it attains the value of 1 if the argument is true and the value of 0 otherwise.
+The following priors are used,
 
-\begin{align}
-  \eta_t &\sim \text{Normal}\left(0, \sigma \right) \\
+\begin{align*}
+  \xi_{+} &\sim \text{Beta}\left(3, 1 \right) \\
+  \xi_{-} &\sim \text{Beta}\left(3, 1 \right) \\
+  \beta &\sim \text{Normal}\left(0, 0.25 \right) \\
+  \eta_t &\sim \text{Half-Normal}\left(0, \sigma \right) \\
   \sigma &\sim \text{Normal}\left(0, 0.2 \right) \\
-  \beta &\sim \text{Normal}\left(0, 0.5 \right)
-\end{align}
+  \frac{1}{\sqrt{\phi}} &\sim \text{Half-Normal}(0, 1) 
+\end{align*}
 
-We then assume a negative binomial observation model with overdispersion $\phi_c$ for reported cases ($C_t$),
-
-\begin{align}
-  C_{t} &\sim \text{NegBinomial}\left(\lambda_t, \phi_c\right) \\
-  \frac{1}{\sqrt{\phi_c}} &\sim \text{Normal}(0, 0.5) 
-\end{align}
-
-Where $\sigma$, and $\frac{1}{\sqrt{phi_c}}$ are truncated to be greater than 0 and $\beta$ is truncated to be between -1 and 1.
-Optionally a Poisson observation model may instead be used (see the documentation for `overdispersion` in `?forecast()`).
+Where $\sigma$, and $\frac{1}{\sqrt{phi}}$ are truncated to be greater than 0 and $\beta$ is truncated to be between -1 and 1.
+The Beta priors for $\xi{+,-}$ have been chosen to be weakly informative that the reduction towards 0 growth is relatively slow.
+Similarly the prior for $\beta$ has been chosen to be weakly informative that there is weak auto-correlation in differenced growth rates.
+$\sigma$ has also been made weakly informative under the assumption that the potential change in growth rates in a single timestep should be relatively small.
 
@@
 
-\begin{align}
+\begin{align*}
   g^{o, \delta}_t &= T_s r^{o, \delta}_t \\
-  R^{o, \delta}_t &= \text{exp}\left(T_s r^{o, \delta}_t\right) \\
-\end{align}
+  R^{o, \delta}_t &= e^{\left(T_s r^{o, \delta}_t\right)} \\
+\end{align*}
 
@@
 The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
-`cmstanr` surfaces several settings which trade-off between sampling speed and the robustness of approach.
+`cmstanr` surfaces several settings which trade-off between sampling speed and the robustness of approach
 Here we take a conservative approach, as model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum treedepth setting to 15.
@@
 
+```{r vis-forecasts}
+locs <- c("United Kingdom", "Germany", "Slovakia", "Italy")
+
+forecasts[location_name %in% locs] |>
+  DT(, date := target_end_date) |>
+    plot_predictions(
+      by = c("horizon", "location_name"), range = c(30, 60, 90)
+    ) +
+    facet_rep_grid(
+      location_name ~ horizon, scales = "free_y", repeat.tick.labels = "y"
+    ) +
+    aes(fill = model, col = model) +
+    scale_y_continuous(labels = comma) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    labs(x = "Date", fill = "Model", col = "Model", ramp = "Range")
+```
+
 ## Forecast evaluation {-}
@@
   Sub-figure 1:
+    - Density/histogram of overall relative scores + mean line
   Sub-figure 2:
+    - Density/histogram of relative scores + mean line by horizon
   Sub-figure 3:
+     - Density/histogram of relative scores + mean line by location
 
@@
   Sub-figure 1:
+    - Calibraton of forecasts by horizon and model
   Sub-figure 2:
+    - Coverage of forecasts by quantile
   Sub-figure 3:
+    - Relative interval score by quantile by horizon.
 
@@
 
+  - Emulators in weather forecasting and elsewhere
+  - ECDC and CDC ensemble papers
+  - Other forecasting models contributed to the ensembles
+  - Baseline models
+
 ## Further work {-}
+
+  - Other similar model formulations. More lag terms. Reduced run-times using approximate Bayesian approaches.
 

### 2022-05-26 R/utils.R

@@ -1,0 +1,29 @@
+merge_forecasts_with_truth <- function(forecasts, truth) {
+  forecasts <- data.table::copy(forecasts)
+  truth <- data.table::copy(truth)
+
+  data.table::setnames(truth, "date", "target_end_date")
+
+  forecasts_with_truth <- truth[
+    forecasts, on = c("location", "target_end_date")
+  ]
+
+  # Remove point forecasts
+  forecasts_with_truth <- forecasts_with_truth[!is.na(quantile)]
+
+  # Remove forecasts with no truth data
+  forecasts_with_truth <- forecasts_with_truth[!is.na(true_value)]
+
+  return(forecasts_with_truth)
+}
+
+rescale_to_incidence_rate <- function(forecasts, population, scale = 1e5) {
+  forecasts <- data.table::copy(forecasts)
+  population <- data.table::copy(population)
+
+  forecasts_with_population <- forecasts[population, on = c("location")] |>
+    data.table::DT(, true_value := true_value / population * scale) |>
+    data.table::DT(, prediction := prediction / population * scale)
+
+  return(forecasts_with_population)
+}

### 2022-05-26 data-raw/score-forecasts.R

@@ -3,6 +3,9 @@ library(data.table)
 library(scoringutils)
 library(here)
 
+# Load functions
+source(here("R", "utils.R"))
+
 # Load truth data
 truth <- fread(here("data", "truth.csv"))
 
@@ -10,14 +13,7 @@ truth <- fread(here("data", "truth.csv"))
 forecasts <- fread(here("data", "forecasts.csv"))
 
 # Merge forecasts and truth
-setnames(truth, "date", "target_end_date")
-forecasts_with_truth <- truth[forecasts, on = c("location", "target_end_date")]
-
-# Remove point forecasts
-forecasts_with_truth <- forecasts_with_truth[!is.na(quantile)]
-
-# Remove forecasts with no truth data
-forecasts_with_truth <- forecasts_with_truth[!is.na(true_value)]
+forecasts_with_truth <- merge_forecasts_with_truth(forecasts, truth)
 
 # Score forecasts
 scores <- score(forecasts_with_truth)

### 2022-05-26 data-raw/update.sh

@@ -7,5 +7,8 @@ Rscript data-raw/get-hub-forecasts.R
 # Download truth data
 Rscript data-raw/get-truth.R
 
+# Get population data
+Rscript data-raw/get-population.R
+
 # Calculate per forecast and quantile scores
 Rscript data-raw/score-forecasts.R

### 2022-05-26 docs/paper/paper.Rmd
figure tuning

@@
   merge_forecasts_with_truth(truth) |>
-  rescale_to_incidence_rate(population, scale = 1000)
+  rescale_to_incidence_rate(population, scale = 1e4)
 ```
@@
 ```{r vis-forecasts}
-locs <- c("United Kingdom", "Germany", "Slovakia", "Italy")
+locs <- c("United Kingdom", "Germany", "Slovakia", "Italy", "Poland", "Greece")
 
@@
     ) +
-    facet_rep_grid(
-      location_name ~ horizon, scales = "free_y", repeat.tick.labels = "y"
-    ) +
+    facet_grid(
+      location_name ~ horizon) +
     aes(fill = model, col = model) +
-    scale_y_continuous(labels = comma) +
+    scale_y_log10(labels = comma, oob = squish, limits = c(0.1, 1000)) +
     scale_fill_brewer(palette = "Dark2") +

### 2022-05-26 docs/paper/paper.Rmd
wrap up figure 1

@@
   merge_forecasts_with_truth(truth) |>
-  rescale_to_incidence_rate(population, scale = 1e4)
+  rescale_to_incidence_rate(population, scale = 1e4) |>
+  rename_models()
+
+# Load forecast scores
+scores <- fread(here("data", "scores.csv")) |>
+  rename_models()
 ```
@@
     scale_color_brewer(palette = "Dark2") +
-    labs(x = "Date", fill = "Model", col = "Model", ramp = "Range")
+    labs(x = "Date", fill = "Model", col = "Model", ramp = "Range",
+         y = "Notified test positive cases per 10,000 population")
 ```

### 2022-05-30 docs/paper/paper.Rmd
finalise figure 1

@@
 library(scales)
-library(lemon)
 library(here)
@@
 
+```{r empirical-coverage, eval = FALSE}
+empirical_coverage <- scores |>
+  DT(, horizon := as.character(horizon)) |>
+  add_coverage(ranges = c(50, 90), by = c("model", "horizon")) |>
+  summarise_scores(fun = signif, digits = 2, by = c("model", "horizon"))
+```
 ## Implementation {-}
@@
 
-- Figure 1:
-  - X: date
-  - Y: incidence rate
-  - Colour: Model
-  - X facet: Country
-  - Y facet: Horizon
-
-```{r vis-forecasts}
+```{r vis-forecasts, fig.height = 9, fig.width = 9, fig.cap = "Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in the United Kingdom, Germany, Slovakia, Italy, Poland, and Greece by forecast horizon (one to four weeks).
+Incidence rates are shown on a log scale.
+30%, 60%, and 90% credible intervals are shown.
+The black line and points are the notified cases as of the date of data extraction rather than those available at the time."}
 locs <- c("United Kingdom", "Germany", "Slovakia", "Italy", "Poland", "Greece")

### 2022-05-30 R/utils.R

@@ -33,7 +33,7 @@ rescale_to_incidence_rate <- function(forecasts, population, scale = 1e5) {
 rename_models <- function(forecasts) {
   forecasts <- data.table::copy(forecasts) |>
     data.table::DT(, model := data.table::fcase(
-      model %in% "EuroCOVIDhub-ensemble", "ECDC ensemble",
+      model %in% "EuroCOVIDhub-ensemble", "Ensemble",
       model %in% "epiforecasts-weeklygrowth", "Surrogate"
     ))
   return(forecasts[])

### 2022-05-30 docs/paper/paper.Rmd
add relative scoring

--- 2022-05-26 docs/paper/paper.Rmd
@@
   echo = FALSE,
-  message = FALSE
+  message = FALSE,
+  cache = TRUE
 )
@@
 library(data.table)
+library(scoringutils)
 library(ggplot2)
 library(scales)
-library(lemon)
 library(here)
@@
 
+```{r coverage}
+ranges <- c(30, 60, 90)
+
+coverage <- scores |>
+  copy() |>
+  DT(, horizon := as.character(horizon)) |>
+  DT(, coverage_deviation := NULL) |>
+  add_coverage(ranges = ranges, by = c("model", "horizon")) |>
+  summarise_scores(fun = signif, digits = 2, by = c("model", "horizon")) |>
+  melt(
+    measure.vars = patterns("coverage_"), variable.name = "range",
+    value.name = "coverage"
+  )|>
+  DT(, range := gsub("coverage_", "", range)) |>
+  DT(, range := paste0(range, "% interval"))
+```
+
+```{r relative-interval-score}
+relative_interval_score <- scores |>
+  copy() |>
+  setorderv(c(
+    "location", "target_end_date", "horizon", "quantile", "model"
+  )) |>
+  DT(,
+    relative_interval_score := interval_score / shift(interval_score),
+    by = c("location", "target_end_date", "horizon", "quantile")
+  ) |>
+  DT(model %in% "Surrogate")
+```
+
 ## Implementation {-}
@@
 
-- Figure 1:
-  - X: date
-  - Y: incidence rate
-  - Colour: Model
-  - X facet: Country
-  - Y facet: Horizon
-
-```{r vis-forecasts}
+```{r vis-forecasts, fig.height = 9, fig.width = 9, fig.cap = "Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in the Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).
+Incidence rates are shown on a log scale.
+30%, 60%, and 90% credible intervals are shown.
+The black line and points are the notified cases as of the date of data extraction rather than those available at the time."}
 locs <- c("United Kingdom", "Germany", "Slovakia", "Italy", "Poland", "Greece")
@@
   facet_grid(
-      location_name ~ horizon) +
+    location_name ~ horizon, scales = "free_y"
+  ) +
   aes(fill = model, col = model) +
-    scale_y_log10(labels = comma, oob = squish, limits = c(0.1, 1000)) +
+  scale_y_log10(labels = comma, oob = squish, limits = c(NA, NA)) +
   scale_fill_brewer(palette = "Dark2") +
@@
 
+```{r plot-coverage}
+hlines <- data.table(
+  range = paste0(ranges, "% interval"), nominal = ranges / 100
+)
+
+plot_cov_levels <- ggplot(coverage) +
+  aes(
+    x = horizon, y = coverage, colour = model, group = interaction(range, model)
+  ) +
+  geom_line() +
+  geom_point() +
+  geom_hline(
+    data = hlines, aes(yintercept = nominal, x = NULL), linetype = "dashed"
+  ) +
+  scale_colour_brewer(palette = "Dark2") +
+  scale_y_continuous(labels = percent) +
+  facet_wrap(vars(range), ncol = 1) +
+  theme_scoringutils() +
+  labs(
+     y = "Proportion of data within forecast interval",
+     x = "Forecast horizon (weeks)",
+     col = "Model"
+  )
+
+plot_cov_quantile <- scores |>
+  summarise_scores(by = c("model", "quantile", "horizon")) |>
+  plot_quantile_coverage(scores) +
+  aes(col = model) +
+  scale_colour_brewer(palette = "Dark2") +
+  facet_wrap(vars(horizon)) +
+  labs(col = "Model")
+
+plot_interval_score <- relative_interval_score |>
+  geom
+```
 # Discussion {-}

### 2022-05-30 docs/paper/paper.Rmd
figure 1 and 3 in place

@@
   bookdown::pdf_book:
-    base_format: rticles::peerj_article # for using bookdown features like \@ref()
+    base_format: rticles::peerj_article
+    extra_dependencies: ["float"]
 ---
@@
   message = FALSE,
-  cache = TRUE
+  cache = TRUE,
+  messages = FALSE,
+  warnings = FALSE
 )
@@
 ```{r relative-interval-score}
+cols <- c("location", "target_end_date", "horizon", "range", "model")
 relative_interval_score <- scores |>
   copy() |>
-  setorderv(c(
-    "location", "target_end_date", "horizon", "quantile", "model"
-  )) |>
+  setorderv(cols) |>
+  DT(, c(..cols, "interval_score")) |>
+  unique() |>
   DT(,
     relative_interval_score := interval_score / shift(interval_score),
-    by = c("location", "target_end_date", "horizon", "quantile")
+    by = c("location", "target_end_date", "horizon", "range")
   ) |>
@@
 
-```{r vis-forecasts, fig.height = 9, fig.width = 9, fig.cap = "Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in the Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).
+```{r vis-forecasts, fig.cap = "Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in the Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).
 Incidence rates are shown on a log scale.
-30%, 60%, and 90% credible intervals are shown.
-The black line and points are the notified cases as of the date of data extraction rather than those available at the time."}
+30\\%, 60\\%, and 90\\% credible intervals are shown.
+The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 9}
 locs <- c("United Kingdom", "Germany", "Slovakia", "Italy", "Poland", "Greece")
@@
 
-```{r plot-coverage}
+```{r plot-coverage, fig.height = 9, fig.width = 9, out.width = "95%"}
 hlines <- data.table(
@@
   scale_y_continuous(labels = percent) +
-  facet_wrap(vars(range), ncol = 1) +
+  facet_wrap(vars(range), ncol = 1, scales = "free_y") +
   theme_scoringutils() +
@@
   aes(col = model) +
+  scale_x_continuous(labels = percent) +
+  scale_y_continuous(labels = percent) +
   scale_colour_brewer(palette = "Dark2") +
+  guides(col = guide_none()) +
   facet_wrap(vars(horizon)) +
-  labs(col = "Model")
+  labs(col = "Model", y = "Percent of observations below quantile")
 
 plot_interval_score <- relative_interval_score |>
-  geom
+  copy() |>
+  DT(,
+    .(relative_interval_score = median(relative_interval_score)),
+    by = c("horizon", "range")
+  ) |>
+  ggplot() +
+  aes(x = range / 100, y = relative_interval_score, col = as.factor(horizon)) +
+  geom_hline(yintercept = 1, linetype = 2) +
+  geom_point(size = 1.2) +
+  geom_line(alpha = 0.8) +
+  scale_x_continuous(labels = percent) +
+  scale_y_continuous(trans = "log", breaks = seq(1, 4, by = 0.2)) +
+  scale_colour_brewer(palette = "Accent") +
+  theme_scoringutils() +
+  labs(
+    x = "Quantile", y = "Median relative interval score",
+    col = "Forecast horizon (weeks)"
+  )
+
+(((plot_cov_levels /
+  plot_interval_score)  +
+  plot_layout(heights = c(3, 1), guides = "collect")
+) |
+  plot_cov_quantile) +
+  plot_layout(guides = "collect", widths = c(1, 2)) +
+  plot_annotation(tag_levels = "a") &
+  theme(
+    legend.position = "bottom", legend.box="vertical", legend.margin = margin()
+  )
 ```
+
 # Discussion {-}

### 2022-05-30 docs/paper/paper.Rmd
update fig org

@@
 30\\%, 60\\%, and 90\\% credible intervals are shown.
-The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 9}
+The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 12}
 locs <- c("United Kingdom", "Germany", "Slovakia", "Italy", "Poland", "Greece")
@@
 
-```{r plot-coverage, fig.height = 9, fig.width = 9, out.width = "95%"}
+```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%"}
 hlines <- data.table(
@@
   scale_x_continuous(labels = percent) +
-  scale_y_continuous(trans = "log", breaks = seq(1, 4, by = 0.2)) +
+  scale_y_continuous(
+    trans = "log", breaks = c(seq(1, 2, by = 0.2), seq(2.4, 4, by = 0.4))
+  ) +
   scale_colour_brewer(palette = "Accent") +
@@
   theme(
-    legend.position = "bottom", legend.box="vertical", legend.margin = margin()
+    legend.position = "bottom", legend.margin = margin()
   )

### 2022-05-31 docs/paper/paper.Rmd
streamline coverage plotting and plan text

--- 2022-05-30 docs/paper/paper.Rmd
@@
 source(here("R", "utils.R"))
+source(here("R","coverage-plots.R"))
 ```
@@
 
+```{r eval-parameters}
+locs <- c("United Kingdom", "Germany", "Slovakia", "Italy", "Poland", "Greece")
+ranges <- c(30, 60, 90)
+```
+
 ```{r coverage}
-ranges <- c(30, 60, 90)
-
 coverage <- scores |>
@@
 
+- Forecast performance overview for the ensemble and surrogate.
+- Forecast performance overview by forecast horizon. Highlight degrading performance with this impacting the surrogate model to a greater degree. 
+- Discussion country level differences in perforamance and link to potential growth rate trends.
+
 ```{r vis-forecasts, fig.cap = "Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in the Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).
@@
 30\\%, 60\\%, and 90\\% credible intervals are shown.
-The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 9}
-locs <- c("United Kingdom", "Germany", "Slovakia", "Italy", "Poland", "Greece")
-
-forecasts[location_name %in% locs] |>
-  DT(, date := target_end_date) |>
-  plot_predictions(
-    by = c("horizon", "location_name"), range = c(30, 60, 90)
-  ) +
-  facet_grid(
-    location_name ~ horizon, scales = "free_y"
-  ) +
-  aes(fill = model, col = model) +
-  scale_y_log10(labels = comma, oob = squish, limits = c(NA, NA)) +
-  scale_fill_brewer(palette = "Dark2") +
-  scale_color_brewer(palette = "Dark2") +
-  labs(x = "Date", fill = "Model", col = "Model", ramp = "Range",
-        y = "Notified test positive cases per 10,000 population")
+The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 12}
+plot_forecasts(forecasts, locs, ranges)
 ```
@@
 
-- Figure 3:
-  Sub-figure 1:
-    - Calibraton of forecasts by horizon and model
-  Sub-figure 2:
-    - Coverage of forecasts by quantile
-  Sub-figure 3:
-    - Relative interval score by quantile by horizon.
+- Calibration of forecasts at credible ranges of interest.
+- Calibration across quantiles
+- Relative interval scores by quantile
 
-```{r plot-coverage, fig.height = 9, fig.width = 9, out.width = "95%"}
-hlines <- data.table(
-  range = paste0(ranges, "% interval"), nominal = ranges / 100
-)
-
-plot_cov_levels <- ggplot(coverage) +
-  aes(
-    x = horizon, y = coverage, colour = model, group = interaction(range, model)
+```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%"}
+(
+  (
+    (
+      plot_coverage_range(coverage, ranges) /
+        plot_rel_score_by_quantile(relative_interval_score)
     )  +
-  geom_line() +
-  geom_point() +
-  geom_hline(
-    data = hlines, aes(yintercept = nominal, x = NULL), linetype = "dashed"
-  ) +
-  scale_colour_brewer(palette = "Dark2") +
-  scale_y_continuous(labels = percent) +
-  facet_wrap(vars(range), ncol = 1, scales = "free_y") +
-  theme_scoringutils() +
-  labs(
-     y = "Proportion of data within forecast interval",
-     x = "Forecast horizon (weeks)",
-     col = "Model"
-  )
-
-plot_cov_quantile <- scores |>
-  summarise_scores(by = c("model", "quantile", "horizon")) |>
-  plot_quantile_coverage(scores) +
-  aes(col = model) +
-  scale_x_continuous(labels = percent) +
-  scale_y_continuous(labels = percent) +
-  scale_colour_brewer(palette = "Dark2") +
-  guides(col = guide_none()) +
-  facet_wrap(vars(horizon)) +
-  labs(col = "Model", y = "Percent of observations below quantile")
-
-plot_interval_score <- relative_interval_score |>
-  copy() |>
-  DT(,
-    .(relative_interval_score = median(relative_interval_score)),
-    by = c("horizon", "range")
-  ) |>
-  ggplot() +
-  aes(x = range / 100, y = relative_interval_score, col = as.factor(horizon)) +
-  geom_hline(yintercept = 1, linetype = 2) +
-  geom_point(size = 1.2) +
-  geom_line(alpha = 0.8) +
-  scale_x_continuous(labels = percent) +
-  scale_y_continuous(trans = "log", breaks = seq(1, 4, by = 0.2)) +
-  scale_colour_brewer(palette = "Accent") +
-  theme_scoringutils() +
-  labs(
-    x = "Quantile", y = "Median relative interval score",
-    col = "Forecast horizon (weeks)"
-  )
-
-(((plot_cov_levels /
-  plot_interval_score)  +
     plot_layout(heights = c(3, 1), guides = "collect")
   ) |
-  plot_cov_quantile) +
+  plot_coverage_quantiles(scores)
+) +
   plot_layout(guides = "collect", widths = c(1, 2)) +

### 2022-05-31 R/coverage-plots.R

@@ -1,0 +1,127 @@
+#' Plot forecasts by forecast horizon
+#'
+#' @param forecasts A data.tabble of forecasts
+#'
+#' @param locs A character vector of locations to plot.
+#'
+#' @param ranges A numeric vector of credible interval ranges to plot.
+#'
+#' @return A ggplot2 object.
+#' @importFrom scoringutils plot_predictions
+#' @import ggplot2
+#' @import data.table
+plot_forecasts <- function(forecasts, locs, ranges) {
+  forecasts[location_name %in% locs] |>
+    DT(, date := target_end_date) |>
+    plot_predictions(
+      by = c("horizon", "location_name"), range = ranges
+    ) +
+    facet_grid(
+      location_name ~ horizon, scales = "free_y"
+    ) +
+    aes(fill = model, col = model) +
+    scale_y_log10(labels = comma, oob = squish, limits = c(NA, NA)) +
+    scale_fill_brewer(palette = "Dark2") +
+    scale_color_brewer(palette = "Dark2") +
+    labs(x = "Date", fill = "Model", col = "Model", ramp = "Range",
+          y = "Notified test positive cases per 10,000 population")
+}
+
+#' Plot coverage at specified ranges
+#'
+#' @param coverage a  data.table of coverage
+#'
+#' @param ranges A numeric vector of credible interval ranges to plot.
+#'
+#' @return A ggplot2 object
+#'
+#' @import data.table
+#' @import ggplot2
+#' @import scales
+#' @importFrom scoringutils theme_scoringtuils
+plot_coverage_range <- function(coverage, ranges) {
+  hlines <- data.table(
+    range = paste0(ranges, "% interval"), nominal = ranges / 100
+  )
+
+  plot_cov_levels <- ggplot(coverage) +
+    aes(
+      x = horizon, y = coverage, colour = model,
+      group = interaction(range, model)
+    ) +
+    geom_line() +
+    geom_point() +
+    geom_hline(
+      data = hlines, aes(yintercept = nominal, x = NULL), linetype = "dashed"
+    ) +
+    scale_colour_brewer(palette = "Dark2") +
+    scale_y_continuous(labels = percent) +
+    facet_wrap(vars(range), ncol = 1, scales = "free_y") +
+    theme_scoringutils() +
+    labs(
+      y = "Proportion of data within forecast interval",
+      x = "Forecast horizon (weeks)",
+      col = "Model"
+    )
[…]

### 2022-05-31 docs/paper/paper.Rmd
update renv
add missing patchwork

@@
 library(ggplot2)
+library(patchwork)
 library(scales)
@@
 source(here("R", "utils.R"))
-source(here("R","coverage-plots.R"))
+source(here("R", "coverage-plots.R"))
 ```

### 2022-06-01 docs/paper/paper.Rmd
switch to incidence rate based scoring

@@
 source(here("R", "utils.R"))
+source(here("R", "evaluate.R"))
 source(here("R", "coverage-plots.R"))
@@
 - Visualise forecasts across time horizons in examplar countries.
-- Weighted interval score.
+- Weighted interval score on incidence rate per 10,000 people to account for population differences between locations.
 - Calibration and coverage.
@@
 ```{r coverage}
-coverage <- scores |>
-  copy() |>
-  DT(, horizon := as.character(horizon)) |>
-  DT(, coverage_deviation := NULL) |>
-  add_coverage(ranges = ranges, by = c("model", "horizon")) |>
-  summarise_scores(fun = signif, digits = 2, by = c("model", "horizon")) |>
-  melt(
-    measure.vars = patterns("coverage_"), variable.name = "range",
-    value.name = "coverage"
-  )|>
-  DT(, range := gsub("coverage_", "", range)) |>
-  DT(, range := paste0(range, "% interval"))
+coverage <- calc_coverage(scores)
 ```
@@
 ```{r relative-interval-score}
-cols <- c("location", "target_end_date", "horizon", "range", "model")
 relative_interval_score <- scores |>
-  copy() |>
-  setorderv(cols) |>
-  DT(, c(..cols, "interval_score")) |>
-  unique() |>
-  DT(,
-    relative_interval_score := interval_score / shift(interval_score),
-    by = c("location", "target_end_date", "horizon", "range")
+  calc_relative_score(
+    cols = c("location", "target_end_date", "horizon", "range")
+  )
+relative_wis <- scores |>
+  summarise_scores(
+    by = c("location_name", "target_end_date", "horizon", "model")
   ) |>
-  DT(model %in% "Surrogate")
+  calc_relative_score(cols = c("location_name", "target_end_date", "horizon"))
 ```
@@
 
+```{r}
+relative_wis |>
+  copy() |>
+  ggplot() +
+  aes(x = relative_interval_score) +
+  geom_vline(xintercept = 1, linetype = 2) +
+  geom_density() +
+  facet_wrap(vars(horizon)) +
+  scale_x_continuous(trans = "log")
+```
+
+```{r plot-eval, fig.height = 12, fig.width = 18, out.width = "95%"}
+ (
+  (
+    (
+      plot_rel_wis(coverage, ranges) /
+      plot_rel_wis_by_horizon(relative_wis)
+    )  +
+    plot_layout(heights = c(3, 1), guides = "collect")
+  ) |
+  plot_rel_wis_by_location(relative_wis)
+) +
+  plot_layout(guides = "collect", widths = c(2, 1)) +
+  plot_annotation(tag_levels = "a") &
+  theme(
+    legend.position = "bottom", legend.margin = margin(),
+    legend.justification = "centre"
+  )
+```
+
 ## Forecast calibration {-}
@@
   theme(
-    legend.position = "bottom", legend.margin = margin()
+    legend.position = "bottom", legend.margin = margin(),
+    legend.justification = "centre"
   )

### 2022-06-01 R/evaluate.R

@@ -1,0 +1,27 @@
+calc_coverage <- function(scores) {
+  scores |>
+    copy() |>
+    DT(, horizon := as.character(horizon)) |>
+    DT(, coverage_deviation := NULL) |>
+    add_coverage(ranges = ranges, by = c("model", "horizon")) |>
+    summarise_scores(fun = signif, digits = 2, by = c("model", "horizon")) |>
+    melt(
+      measure.vars = patterns("coverage_"), variable.name = "range",
+      value.name = "coverage"
+    )|>
+    DT(, range := gsub("coverage_", "", range)) |>
+    DT(, range := paste0(range, "% interval"))
+}
+
+calc_relative_score <- function(scores, cols) {
+  scores |>
+    copy() |>
+    setorderv(cols) |>
+    DT(, c(..cols, "model", "interval_score")) |>
+    unique() |>
+    DT(,
+      relative_interval_score := interval_score / shift(interval_score),
+      by = cols
+    ) |>
+    DT(model %in% "Surrogate")
+}

### 2022-06-07 docs/paper/paper.Rmd
add raw wis plot

@@
 library(scoringutils)
+library(lubridate)
+library(forcats)
 library(ggplot2)
+library(ggridges)
 library(patchwork)
@@
 source(here("R", "evaluate.R"))
+source(here("R", "evaluation-plots.R"))
 source(here("R", "coverage-plots.R"))
@@
   )
-relative_wis <- scores |>
+wis <- scores |>
   summarise_scores(
     by = c("location_name", "target_end_date", "horizon", "model")
-  ) |>
+  )
+relativ_wis <- wis |>
   calc_relative_score(cols = c("location_name", "target_end_date", "horizon"))
@@
 
-```{r vis-forecasts, fig.cap = "Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in the Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).
+```{r vis-forecasts, fig.cap = "a.) Weighted interval score at the 1 week and 4 week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom.
+b.)Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).
 Incidence rates are shown on a log scale.
 30\\%, 60\\%, and 90\\% credible intervals are shown.
-The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 12}
-plot_forecasts(forecasts, locs, ranges)
+The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 18}
+(
+  plot_wis(wis, locs) +
+   theme(strip.text.x = element_blank()) |
+  plot_forecasts(forecasts, locs, ranges) +
+    guides(fill = guide_none(), fill_ramp = guide_none(), col = guide_none())
+) +
+  plot_layout(widths = c(1, 4), guides = "collect") +
+  plot_annotation(tag_levels = "a") &
+  theme(
+    legend.position = "bottom", legend.margin = margin(),
+    legend.justification = "centre"
+  )
 ```
@@
 
-```{r}
-relative_wis |>
+```{r make-rwis-plots}
+plot_rwis_horizon <- relative_wis |>
   copy() |>
-  ggplot() +
-  aes(x = relative_interval_score) +
-  geom_vline(xintercept = 1, linetype = 2) +
-  geom_density() +
-  facet_wrap(vars(horizon)) +
-  scale_x_continuous(trans = "log")
+  DT(, horizon := fct_rev(as.factor(horizon))) |>
+  plot_relative_wis(
+    y = horizon, fill = horizon,
+    point_color = horizon
+  ) +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_color_brewer(palette = "Dark2") +
+  labs(y = "Forecast horizon (weeks)") +
+  guides(fill = guide_none(), point_color = guide_none())
+
+plot_rwis_location <- relative_wis |>
+  DT(horizon == 1 | horizon == 4) |>
+  DT(, horizon := fct_rev(as.factor(horizon))) |>
+  plot_relative_wis(
+    y = location_name, fill = horizon, alpha = 0.3,
+    jittered_points = FALSE, quantiles = NULL,  point_color = horizon
+  ) +
+  scale_fill_brewer(palette = "Dark2") +
+  scale_color_brewer(palette = "Dark2") +
+  labs(
+    y = "Forecast location", fill = "Forecast horizon (weeks)",
+    point_color = "Forecast horizon (weeks)"
+  )
+
+plot_rwis_by_month <- relative_wis |>
+  DT(horizon == 1 | horizon == 4) |>
+  DT(, target_month := month(target_end_date, label = TRUE)) |>
+  DT(, target_month := fct_rev(target_month)) |>
+  plot_relative_wis(
+    y = target_month, fill = as.factor(horizon), alpha = 0.3,
+    jittered_points = TRUE, quantiles = NULL
+  ) +
+  scale_fill_brewer(palette = "Dark2") +
+  labs(y = "Target forecast month", fill = "Forecast horizon (weeks)")
 ```
 
-```{r plot-eval, fig.height = 12, fig.width = 18, out.width = "95%"}
+```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%"}
  (
@@
     (
-      plot_rel_wis(coverage, ranges) /
-      plot_rel_wis_by_horizon(relative_wis)
+      plot_rwis_horizon / plot_rwis_by_month
     )  +
-    plot_layout(heights = c(3, 1), guides = "collect")
+    plot_layout(heights = c(4, 5), guides = "collect")
   ) |
-  plot_rel_wis_by_location(relative_wis)
+  plot_rwis_location + guides(fill = guide_none())
 ) +
-  plot_layout(guides = "collect", widths = c(2, 1)) +
+  plot_layout(guides = "collect", widths = c(1, 1)) +
   plot_annotation(tag_levels = "a") &

### 2022-06-07 R/evaluation-plots.R

@@ -27,6 +27,28 @@ plot_forecasts <- function(forecasts, locs, ranges) {
           y = "Notified test positive cases per 10,000 population")
 }
 
+plot_wis <- function(wis, locs) {
+  wis |>
+    DT(location_name %in% locs) |>
+    DT(horizon == 1 | horizon == 4) |>
+    ggplot() +
+    aes(
+      x = target_end_date, y = interval_score, col = model,
+      shape = as.factor(horizon)
+    ) +
+    geom_point(size = 1.2) +
+    geom_line(alpha = 0.8) +
+    scale_color_brewer(palette = "Dark2") +
+    theme_scoringutils() +
+    theme(legend.position = "bottom") +
+    scale_y_log10() +
+    facet_wrap(vars(location_name), scales = "free_y", ncol = 1) +
+    labs(
+      x = "Date", y = "Weighted interval score",
+      col = "Model", shape =  "Forecast horizon (weeks)"
+    )
+}
+
 plot_relative_wis <- function(relative_wis, alpha = 0.8,
                               jittered_points = TRUE,
                               quantiles = c(0.05, 0.35, 0.65, 0.95), ...) {

### 2022-06-07 docs/paper/paper.Rmd
more plot tuning

@@
   plot_relative_wis(
-    y = horizon, fill = horizon,
-    point_color = horizon
+    y = horizon, fill = horizon
   ) +
@@
   labs(y = "Forecast horizon (weeks)") +
-  guides(fill = guide_none(), point_color = guide_none())
+  guides(fill = guide_none(), point_color = guide_none(), point = guide_none())
 
@@
   scale_fill_brewer(palette = "Dark2") +
-  labs(y = "Target forecast month", fill = "Forecast horizon (weeks)")
+  labs(y = "Target forecast month", fill = "Forecast horizon (weeks)") +
+  guides(fill = guide_none())
 ```

### 2022-06-07 R/evaluation-plots.R

@@ -65,7 +65,7 @@ plot_relative_wis <- function(relative_wis, alpha = 0.8,
     ) +
     scale_x_continuous(
       trans = "log", limits = c(NA, 5),
-      breaks = c(seq(0, 2, by = 0.2), seq(2.5, 5, by = 0.5))
+      breaks = c(seq(0, 2, by = 0.4), 3, 5)
     ) +
     geom_vline(xintercept = 1, linetype = 2, size = 1.05, alpha = 0.8) +
     theme_scoringutils() +

### 2022-06-07 docs/paper/paper.Rmd
more plot tuning

@@
   plot_relative_wis(
-    y = horizon, fill = horizon
+    y = horizon, fill = horizon,
   ) +
@@
   DT(horizon == 1 | horizon == 4) |>
-  DT(, horizon := fct_rev(as.factor(horizon))) |>
   plot_relative_wis(
-    y = location_name, fill = horizon, alpha = 0.3,
+    y = location_name, fill = as.factor(horizon), alpha = 0.3,
     jittered_points = FALSE, quantiles = NULL,  point_color = horizon
@@
   ) |
-  plot_rwis_location + guides(fill = guide_none())
+  plot_rwis_location
 ) +
@@
 ) +
-  plot_layout(guides = "collect", widths = c(1, 2)) +
+  plot_layout(widths = c(1, 2), guides = "collect") +
   plot_annotation(tag_levels = "a") &

### 2022-06-07 docs/paper/paper.Rmd
update forecasts + truth

@@
 - Forecast performance overview by forecast horizon. Highlight degrading performance with this impacting the surrogate model to a greater degree. 
-- Discussion country level differences in perforamance and link to potential growth rate trends.
+- Discussion country level differences in performance and link to potential growth rate trends.
 
@@
 
-## Forecast evaluation {-}
+## Relative forecast evaluation {-}
 
-- Figure 2:
-  Sub-figure 1:
-    - Density/histogram of overall relative scores + mean line
-  Sub-figure 2:
-    - Density/histogram of relative scores + mean line by horizon
-  Sub-figure 3:
-     - Density/histogram of relative scores + mean line by location
+- Relative performance of surrogate model by forecast horizon.
+   - Give mean relative performance difference by horizon
+   - Discuss distribution.
+- Relative performance of surrogate model by month stratified by forecast horizon.
+  - Give mean relative performance by month for each horizon
+  - Discuss distribution
+- Relative performance of surrogate model by location sratified by forecast horizon.
+  - Discuss qualatative differences between locations
 

### 2022-06-07 docs/paper/paper.Rmd
add evaluation methods

@@
 $T_s$ is a user set scaling parameter that defines the timespan over which the summary metrics apply dependent on the time step of the data.
-It can be set using the `scale_r` and defaults to 1 which returns summary statistics scaled to the timestep of the data.
-Depending on the setup of the model used these summary measures will be more or less related to their epidemiological definitions.
-In particular, adding a weighting to past expected cases that is more complex than a simple lag may cause interpretation issues.
 
@@
 
-- Visualise forecasts across time horizons in examplar countries.
-- Weighted interval score on incidence rate per 10,000 people to account for population differences between locations.
-- Calibration and coverage.
-- Relative interval score.
-- Relative interval score overall, distribution of scores, by horizon both overall and distribution, by location overall and distribution.
-- Relative interval score by interval range by horizon
-- Calibration at 50 and 90 credible intervals
-- Coverage across intervals.
+In order to standardise forecasts across forecast locations we first normalised both weekly notified test positive cases and forecast test positive cases by the population in the forecast region scaled to be a rate per 10,000 people.
+We then evaluated a subset of forecasts from the following countries Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom visually by forecast horizon (1-4 weeks) for both the ensemble and surrogate model.
+
+We evaluate forecasts quantatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile based proper scoring rule that approximates the e continuous ranked probability score (CRPS).
+Both the WIS and CRPS are generalisations of absolute error in order to evaluate probablistic forecasts and are widely used to evaluate COVID-19 forecasts [@REF].
+We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks) as well as summarising across the study period.
+
+In order to understand the relative performance of the surrogate model compared to the ensemble model we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model.
+In order to maintain the properiety of this score we do this after first taking the means of scores for the relavant stratification.
+We explore relative performance by forecast horizon, by forecast month and forecast horizon, and by forecast location and forecast horizon.
+
+In addition to presenting the WIS for a subset of locations and the relative WIS for all locations we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% credible intervals and by quantile.
+Lastly we calculate and visualise the relative interval score by quantile, stratified by forecast horizon.
 
@@
 The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
-`cmstanr` surfaces several settings which trade-off between sampling speed and the robustness of approach
+`cmstanr` surfaces several settings which trade-off between sampling speed and the robustness of approach.
 Here we take a conservative approach, as model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum treedepth setting to 15.
-For real-time usage convergance was not assessed but during model development the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergant transitions and exceedance of the maximum tree depth and posterior predictions visually compared to observed data [@cmdstanr].
+For real-time usage convergance was not assessed but during model development the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergant transitions and exceedance of the maximum tree depth.
+During development posterior predictions were also visually compared to observed data [@cmdstanr].
 
@@
 We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, to visualise forecasts and summary measures, and to summarise forecasts.
-Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils].
+Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils], and the `scoringRules` R package (1.0.1).
 
 To ensure reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with built Docker image [@Boettiger:2015dw] (via GitHub Actions) is provided.
-Weekly forecasts also made use of `renv` and Docker to ensure reproducibility with GitHub actions used to produce the forecasts ensuring we meet our limited compute goal and that our implementation is independent from available compute resources democratising  access to our approach.
+Weekly forecasts are also made use of `renv` and Docker to ensure reproducibility with GitHub actions used to produce the forecasts ensuring we meet our goal of requring limited compute and that our implementation is independent from local resources resulting in democratised access to our approach.
 The code for this analysis can be found here: https://github.com/epiforecasts/simplified-forecaster-evaluation
@@
 
-## Visualisation of forecasts by horizon {-}
+## Forecast evaluation
+
+### Visualisation of forecasts by horizon {-}
 
@@
 
-## Relative forecast evaluation {-}
+### Relative forecast evaluation {-}
 
@@
 
-## Forecast calibration {-}
+### Forecast calibration {-}
 

### 2022-06-07 docs/paper/paper.Rmd
partial introduction

@@
 
-- Why
-  - ECDC and CDC forecasting hubs
-  - Ensemble shown to outperform individual models in many cases
-  - Thought of as being more desirable due to reliability etc vs multiple models
-  - Many models are complex and require large amounts of compute time.
-  - As the ensemble is made up of these models it means it is extremely resource demandning to construct requiring significant compute as well as multiple teams.
-  - As the ensemble is constructed from multiple models it functions as effectively a black box. This means it may be hard for policy-makers to understand its underyling assumptions
-  - Here we seek to refine forecasting ideas from our previous work alongside our experiances forecasting COVID-19 as contributors to the ECDC and CDC forecasting hubs to develop a simplified forecasting model that has characteristics fo the ensemble forecast whilst maintianing a interpretability and low resource demands.
+Multi-model and  multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade with significant success [@REF].
+These approaches have drawn heavily on the experiance of other domains, for example climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
+However, ensemble forecasts have range of potential downsides with the most obvious being their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project, and that forecasts from these ensembles can be difficult to link to the underling model assumption.
+In climate forecasting emulation approaches have been used to circumvent some of the  resource requirment issue by training a simplified model, usually a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
+However, within epidemiological modelling we often want to use our models for applications other than forecasting, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
+This means that non-parametric emulation is not an ideal solution. 
+In this work we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, limited computational resources, and an easily generalised implementation.
+
+The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the percieved low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and ECDC forecast hub [@ECDC] in 2021.
+All of these collaborations, which the authors (particularly SF and JB have been involved with managing),  ensemble contributions from multiple independent teams using a similar ensembling approach (most recently an unweighted median ensemble [@REF]) and have shown that their ensemble forecasts outperform most individually contributed forecasts and provide forecasts that are more robust to outliers [@REF].
+However, they have struggled to provide feedback to those contributing forecasts that would allow them to improve their forecast approaches, have found it difficult to improve the quality of their ensemble forecasts [@REF], and have struggled to maintain incentives for forecasters to continue to contribute which may impact the long-term quality of the  ensembles they produce.
+
+These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a 1 to 4 week time horizon.
+Generally ground truth data is available daily, and is updated in retrospect as new data becomes available.
+Typically, the ensemble forecast produced each week is made up of at least 5 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
+The majority of submitted forecasts make use of the available daily data, along with potentially other leading predictors to produce daily forecasts that are then aggregated to weekly forecasts.
+Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts.
+In general, 4 main classes of forecast models are used, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and it's generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
+Real-time evaluation has shown that each of these classes of models may perform well though in general semi-mechanistic models, and models that include human insights have outperformed other approaches during certain periods of time [@REF].
+We have contributed a range of forecasts to these forecast collobarations generally focussed on semi-mechanistic statistical methods [@REF], and human insight based forecasts [@REF] with our forecasts performing relatively well compared to the majority of those submitted but rarely outperforming significantly.
+The model based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within week, reporting periodicity [@REF].
+Our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
+Our previous methodology also requires signinficant computational resources running for an hour on a 16 core machine when producing forecasts from the ECDC. 
+In our model based forecasts we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
+We found that our human insight based forecasts significantly outperformed our model based forecasts and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@REF].
 
 - Extension of why
-   - Brief dicussion of implications of modelling NPIs/future changes in behaviour
-   - Brief discussion of reporting patterns in data. Daily/Weekly.
    - Brief discussion of EpiNow2 and related models
@@
 
+Based on these insights here we define a simplified auto-regressive weekly growth rate model that assumes a non-stationary trend in growth rate.
+This means we do not need to model daily reporting processes, which can cause severe performance issues due to reporting artefacts
 - What

### 2022-06-07 docs/paper/paper.Rmd
frist draft introduction

@@
     email: sam.abbott@lshtm.ac.uk
+  - name:  Johannes Bracher
+    affiliation: 2
+  - name: Sebastian Funk
+    affiliation: 1
 affiliation:
   - code: 1
-    address: Cold Spring Harbor Laboratory, One Bungtown Road Cold Spring Harbor, NY 11724
+    address: London School of Hygiene and Tropical Medicine, London, UK
+  - code: 2
+    address: Chair of Statistics and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
 abstract: > 
@@
 The model based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within week, reporting periodicity [@REF].
-Our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
+Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
 Our previous methodology also requires signinficant computational resources running for an hour on a 16 core machine when producing forecasts from the ECDC. 
@@
 We found that our human insight based forecasts significantly outperformed our model based forecasts and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@REF].
+Unlike our epidemiologically motivated forecast submissions the hub ensemble forecasts were typically robust to daily reporting artefacts.
+They also demonstrated some ability to forecast future changes in trend that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight or assumptions about future interventions.
+In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence.
+On the other hand this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
+Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting.
+There is some suggestion that this was partially driven by the ensemble forecasting having non-exponetial uncertainty, unlike many infectious disease forecasting models.
 
-- Extension of why
-   - Brief discussion of EpiNow2 and related models
-   - Brief dicussion of statistical modelling approaches (use MUNI-ARIMA as an example).
-   - Brief discussion of scoring of hub models
-   - Summary of observed performance of the hub ensembles
-    - Auto-correlated
-    - Typically relatively trend following though shows some ability to forecast future changes.
+Based on these insights from our own forecasting work and from observing the hub ensembles behaviour over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model based forecasting [@REF] with aim of producing ensemble like perform without sacrificing interpretability.
+The first simplification we make is to model only weekly data, rather than using daily data and then aggregating, this means that the impact daily reporting artefacts is mitigated.
+It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain signifcant weight in the model this leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
+The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
+This reduces computational requirements and removes the need for external information on the delay from infection to report which can potentially be misspecified leading to reduced forecast performance.
+However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
+As discussed in [@gostic] this leads to varying amounts of bias depending on the epidemic phase. 
+The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a more complex method as we have done in other work [@EpiNow2].
+This represents a parsiminous approach in that we encode our expectation that the growth rate should vary over time and allow this to influance the forecast but we include only a single lag term and hence reduce the computational overhead of the model.
+In order to model potential unobserved interventions and more generally changes in transmission we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
+This reflects a simplified interpretation for how the ensemble appears to react to potential future changes.
+The only observed for which we do not make an adaption is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
+This choice is motivated by our belief that the underlying process is an exponential one and therefore an exponential error model is a sensible choice though this is an area of current research.
+We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the current usage of a generalsied form of the absolute error for the majority of forecast evaluation.
 
-Based on these insights here we define a simplified auto-regressive weekly growth rate model that assumes a non-stationary trend in growth rate.
-This means we do not need to model daily reporting processes, which can cause severe performance issues due to reporting artefacts
-- What
-  - Simplified auto-regressive growth rate model assuming a non-stationary trend
-  -  Explicit modelling of the assumption that rapid growth/decay of cases will typically result in future growth tending towards zero.
-This seeks to capture potential implementation and relaxation of NPIs as well as individuals attitudes to risk (assuming this is based on current growth rates and not case incidence).
-  - Uses weekly vs daily data to side step issues of reporting periodicity and to replicate the observed auto-corelation of the hub ensembles.
-  - Simple extension of an already implemented model and resuse of its modular components.
-  - Implemented with a focus on reproducibility, simplicity, and low resource use. 
-  - The model runs using GitHub actions in a framework supplied by the ECDC forecast hub. We provide our own tools so that others can run the model using GitHub actions democratising our approach.
-
-- Aim: In this work we define a simplified forecasting model designed to replicate some of the observed characteristics of successful infectious disease ensemble models without sacrificing interpretability and with low resource requirements.
-We evaluate the performance of this model against the ECDC forecasting hub ensemble in real-tim, and discuss implications of our findings along with areas for future research.
+In this study, we define the model outlined above and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action. 
+We briefly outline some of the summary statistics it produces and their epidemiological relevance. 
+We then evaluate it's real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying its empirical coverage.
+We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed. 
+Finally we summarise our findings, discuss their implications, and highlight areas for potential study.
+We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low resource, surrogate model in settings where large-scale colloborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models
 
@@
 ## Model {-}
-
-### Motivation
-
-The same underlying structure with expected notifications being assumed to be a product of past expected notifications, and an exponentiated time-varying term.
-This term is then modelled in both instances as a differenced autoregressive process with a single lag term.
-By default it is assumed to vary inline with the timestep of the data but this can be altered to speed up computation or to improve out of sample performance.
-Observed cases are estimated from expected cases by assuming a negative binomial observation model.
-A detailed model definition is given below.
 

### 2022-06-07 docs/paper/paper.Rmd
more intro

@@
 
-In this study, we define the model outlined above and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action. 
+In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptionse and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action.
 We briefly outline some of the summary statistics it produces and their epidemiological relevance. 

### 2022-07-19 docs/paper/paper.Rmd
add data text

@@
 We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed. 
-Finally we summarise our findings, discuss their implications, and highlight areas for potential study.
+Finally, we summarise our findings, discuss their implications, and highlight areas for potential study.
 We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low resource, surrogate model in settings where large-scale colloborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models
@@
 
-- Data from JHU as used as truth data by the ECDC forecasting hub.
-  - When accessed
-  - Aggregated to epiweek as in the ECDC forecasting GitHub
-  - Countries included
-  - Dates of data included
-- Forecasts from the ECDC forecast hub
-  - Forecasts dowloaded for the model proposed here and the ECDC ensemble
-  - Forecast format
-  - Forecases from when and horizon
-  - Ensemble construction and reference
+  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from XX to the XX for the ensemble model (referenced as `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
+  We included all locations covered by the ECDC forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
+  Data on notified weekly cases was originally sourced from the Johns Hopkins University curated data repository (JHU, [@REF]).
+  We used the latest available observed data. As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
+  The follows the common practice of the forecasting hub projects in treating the latest available data as the forecast target regardless of the number of data revisions.
+  Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.
+
+  The ECDC forecast hub requests forecasts for 1 to 4 week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
+  No restrictions were placed on who could submit forecasts and the hub team actively invited particupation from research groups known to be involved with COVID-19 forecasting projects.
+  Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent based model, semi-mechanistic models, expert elicitation models, and ensembles of mutliple approaches (described at https://covid19forecasthub.eu/community.html).
+  Teams submitted forecasts at the latest two days after the complete dataset for the latest forecasting week became available and were allowed to use all data available at the time of submission. 
+  The ensemble forecast was constructed by taking the median of all predictive quantiles without the exlusion of any valid submitted forecast (where validity was defined as passing minimal formatting checks by the hub team).
+  Submitted forecasts and target observations are freely available with a permissive open-source license from the ECDC forecast hub GitHub repository.
+We provide code in the repository of this study to streamline access.
 
@@
 
+Whilst we found that our surrogate model performed compab
   - Other similar model formulations. More lag terms. Reduced run-times using approximate Bayesian approaches.

### 2022-07-19 docs/paper/paper.Rmd
add metadata extraction and summary statistic generation

@@
 
-# Load ECDC forecast metadata. See data-raw/get-hub-metadata.R for munging
-#metadata <- fread(here("data", "forecast-metadata.csv"))
+# Load forecast metedata. See data-raw/get-hub-metadata.R for munging
+metadata <- fread(here("data", "metadata.R"))
 
@@
 
-  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from XX to the XX for the ensemble model (referenced as `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
+  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from 15th of January 2022 to the 19th of July 2022 for the ensemble model (referenced as `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
   We included all locations covered by the ECDC forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
@@
 We provide code in the repository of this study to streamline access.
+
+```{r metadata}
+# Number of locations
+locations <- metadata |>
+  DT(, location) |>
+  unique() |>
+  length()
+
+# Number of models overall
+models <- metadata |>
+  DT(, model) |>
+  unique() |>
+  length()
+
+# Models per forecast date
+models_per_forecast <- metadata |>
+  summarise_forecasts_by(by = "forecast_date")
+
+# Models per location
+models_per_location <- metadata |>
+  summarise_forecasts_by(by = "location")
+
+# Models with a single location target
+locations_per_model <- metadata |>
+  summarise_forecasts_by(var = "location", by = "model")
+
+# Single target models per location
+single_locs_models_per_loc <- metadata |>
+  DT(locations_per_model[n == 1], on = "model") |>
+  summarise_forecasts_by(by = "location")
+
+# Global models
+global_models <- locations_per_model |>
+  DT(n > 30) |>
+  DT(, model)
+
+# Partial coverage models
+partial_coverage <- locations_per_model |>
+  DT(n > 1) |>
+  DT(n <= 30)
+
+# Models coverage of forecast datees
+forecast_dates_by_models <- metadata |>
+  summarise_forecasts_by(var = "forecast_date", by = "model") |>
+  DT(order(n))
+```
 
@@
 
+We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK). 
 - Date range of forecasts

### 2022-07-19 R/utils.R

@@ -38,3 +38,10 @@ rename_models <- function(forecasts) {
     ))
   return(forecasts[])
 }
+
+summarise_forecasts_by <- function(forecasts, var = "model", by) {
+  forecasts |>
+    DT(, c(..var, ..by)) |>
+    unique() |>
+    DT(, .(n = .N), by = by)
+  }

### 2022-07-19 data-raw/get-hub-forecasts.R

@@ -20,7 +20,8 @@ models <- c("epiforecasts-weeklygrowth", "EuroCOVIDhub-ensemble")
 
 # First available forecast submission was the 15th of Janurary 2022 and has
 # continued each week
-up_to <- Sys.Date()
+# We freeze data extraction on the 19th of July 2022
+up_to <- as.Date("2022-07-19")
 forecast_dates <- seq(as.Date("2022-01-15"), up_to, by = "day")
 
 # Get forecasts for models of interest

### 2022-07-19 data-raw/update.sh

@@ -4,6 +4,10 @@
 # Download hub forecasts of interest
 Rscript data-raw/get-hub-forecasts.R
 
+
+# Make hub metadata for study period
+Rscript data-raw/get-hub-metadata.R
+
 # Download truth data
 Rscript data-raw/get-truth.R
 

### 2022-07-20 docs/paper/paper.Rmd
more metadata detail

@@
 # Load forecast metedata. See data-raw/get-hub-metadata.R for munging
-metadata <- fread(here("data", "metadata.R"))
+metadata <- fread(here("data", "metadata.csv"))
 
@@
 locations <- metadata |>
-  DT(, location) |>
-  unique() |>
-  length()
+  DT(, uniqueN(location))
 
@@
 models <- metadata |>
-  DT(, model) |>
-  unique() |>
-  length()
+  DT(, uniqueN(model))
+
+# Number of forecast dates
+forecast_dates <- metadata |>
+  DT(, uniqueN(target_end_date))
+
+# Forecasts
 
@@
 models_per_forecast <- metadata |>
-  summarise_forecasts_by(by = "forecast_date")
+  summarise_forecasts_by(by = "target_end_date")
 
@@
 
+locs_per_model_single_date <- metadata |>
+    summarise_forecasts_by(
+    var = "location", by = c("model", "target_end_date")
+    ) |>
+    DT(, unique(.SD[n == max(n), .(n)]), by = "model")
+
+locs_per_model_date <- metadata |>
+  summarise_forecasts_by(
+    var = "location", by = c("model", "target_end_date")
+  ) |>
+  DT(locs_per_model_single_date[, .(model, max_n = n)], on = "model") |>
+  DT(, n_per := n / max_n) |>
+  DT(, .SD[!all(n == max_n)], by = "model")
+
+models_per_loc_date <- metadata |>
+  summarise_forecasts_by(var = "model", by = c("target_end_date", "location"))
+
 # Single target models per location
@@
   summarise_forecasts_by(by = "location")
+
+# Local models
+local_models <- locations_per_model |>
+  DT(n == 1) |>
+  DT(, model)
 
@@
 # Partial coverage models
-partial_coverage <- locations_per_model |>
+partial_models <- locations_per_model |>
   DT(n > 1) |>
@@
 forecast_dates_by_models <- metadata |>
-  summarise_forecasts_by(var = "forecast_date", by = "model") |>
+  summarise_forecasts_by(var = "target_end_date", by = "model") |>
   DT(order(n))
@@
 
-We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK). 
+We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK making `r locations` unique locations).
+In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the ECDC baseline model).
+Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as "epiforecasts-EpiNow2" by the hub), and our surrogate model (referred to as "epiforecasts-weeklygrowth" by the hub).
+Of the remainining models submitted `r length(local_models)` were submitted in only one location.
+These models were gernerally submitted to a subset of locations with particular clustering in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the ECDC forecasting hub project).
+Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
+`r length(partial_models)` models submitted for between 2 and 30 locations likely indicating manual curation of submissions.
+
+Across all forecast dates and locations the minimum number of independent forecasts was `r models_per_loc_date[, min(n)]` with the maxmimum being `r models_per_loc_date[, max(n)]`.
+The median number of independent forecasts per location and forecast date was `r models_per_loc_date[, median(n)]`.
+Across all forecast dates all locations recieved forecasts from at least `r models_per_location[, min(n)]` models with the median number of forecast models per location being `r models_per_location[, median(n)]`.
+
 - Date range of forecasts

### 2022-07-20 data-raw/get-hub-metadata.R

@@ -26,9 +26,14 @@ hub_forecasts <- get_hub_forecasts(
 # Only keep forecasts for incident cases
 hub_forecasts <- hub_forecasts[grepl("inc case", target)]
 
+# Drop the Hub ensemble
+hub_forecasts <- hub_forecasts |>
+  DT(!model %in% "EuroCOVIDhub-ensemble")
+  
 # Streamline to metadata
 metadata <- hub_forecasts |>
-  DT(, .(model, location, forecast_date)) |>
+  DT(target %in% "1 wk ahead inc case") |>
+  DT(, .(model, location, forecast_date, target_end_date)) |>
   unique()
 
 # Save metadata

### 2022-07-21 docs/paper/paper.Rmd
add more description of forecats patterns

@@
 partial_models <- locations_per_model |>
-  DT(n > 1) |>
+  DT(n > 2) |>
   DT(n <= 30)
+
+#Varying submissions
+varying_submissions <- locs_per_model_date |>
+  DT(partial_models[, .(model)], on = "model")
 
@@
   summarise_forecasts_by(var = "target_end_date", by = "model") |>
-  DT(order(n))
+  DT(order(n)) |>
+  DT(, n_per := n / forecast_dates) |>
+  DT(, type := fcase(
+      model %in% global_models, "global",
+      model %in% local_models, "local",
+      model %in% partial_models[, model], "partial",
+      default = "other"
+    )
+  )
 ```
@@
 
+- Incidence in the ECDC
+  - Omicron variant
+  - BA.1 and BA.2 variant
+  - Differences across countries and timings
+  - NPIs across countries
+  - testing across countries
+
 We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK making `r locations` unique locations).
@@
 Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as "epiforecasts-EpiNow2" by the hub), and our surrogate model (referred to as "epiforecasts-weeklygrowth" by the hub).
-Of the remainining models submitted `r length(local_models)` were submitted in only one location.
+Of the remaining models submitted `r length(local_models)` were submitted in only one location.
 These models were gernerally submitted to a subset of locations with particular clustering in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the ECDC forecasting hub project).
 Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
-`r length(partial_models)` models submitted for between 2 and 30 locations likely indicating manual curation of submissions.
+`r length(partial_models)` models submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targted at specific conditions.
 
@@
 The median number of independent forecasts per location and forecast date was `r models_per_loc_date[, median(n)]`.
-Across all forecast dates all locations recieved forecasts from at least `r models_per_location[, min(n)]` models with the median number of forecast models per location being `r models_per_location[, median(n)]`.
-
-- Date range of forecasts
-- Forecasts per country
-- Total number of forecast models and average per location and per forecast date
-- Summarises general trends in incidence in the data
+All locations recieved forecasts from at least `r models_per_location[, min(n)]` models with the median number of forecast models per location being `r models_per_location[, median(n)]`.
+Coverage of forecast dates varied across submitted models with `r forecast_dates_by_models[n_per == 1, .N]` models submitting for all dates, `r forecast_dates_by_models[n_per >= 0.9, .N]` models submitting for at least 90% of dates, and `r forecast_dates_by_models[n_per < 0.50, .N]` models submitting for fewer than 50% of forecast dates.
+In general, there was no clear difference in forecast date coverage between models that submitted for all locations vs a small subset but models with partial coverage of locations all also had partial coverage of forecast dates.
 

### 2022-07-21 docs/paper/paper.Rmd
add incidence rate discussion across ECDC/UK

@@
 
-- Incidence in the ECDC
-  - Omicron variant
-  - BA.1 and BA.2 variant
-  - Differences across countries and timings
-  - NPIs across countries
-  - testing across countries
+In our study period, incidence rates across ECDC nations and in the UK were primary driven by the spread of noval subvariants of concern related to the Omicron variant.
+Many nations, such as the UK saw large BA.1 waves in early January to late January that then resulted in declining incidence rates through Febuary.
+From late Febuary, througth to the end of May, most nations then saw another wave, typically with lower reported incidence rates, driven by BA.2.
+This wave was typically characterised by a lower peak than the BA.1 wave but a longer tail.
+The end of our study period was characterised by the gradual take-over of the BA.4/BA.5 subvariants that again had a lower peak and lower absolute growth rates.
+Unlike earlier periods in the pandemic, our study period did not see the use of new non-pharmaceutical interventions (NPIs) in response to increasing COVID-19 incidence in most locations.
+In addition, in general ascertainment rates likely reduced over time in most locations due to reductions in routine testing, and reductions in test availability.
+Whilst both the reduced use of NPIs and testing generally occured across nations our study period also marked an increase in the heterogenity of the response to COVID-19 pandemic with nations changing policy at different times and to different degrees.
+This is in contrast to the early COVID-19 pandemic response for which most nations took similar actions at similar times.
 

### 2022-07-21 docs/paper/paper.Rmd
update paper

@@
   )
-relativ_wis <- wis |>
+relative_wis <- wis |>
   calc_relative_score(cols = c("location_name", "target_end_date", "horizon"))

### 2022-07-21 docs/paper/paper.Rmd
tweak

@@
 In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the ECDC baseline model).
-Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as "epiforecasts-EpiNow2" by the hub), and our surrogate model (referred to as "epiforecasts-weeklygrowth" by the hub).
+Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
 Of the remaining models submitted `r length(local_models)` were submitted in only one location.
@@
 Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
-`r length(partial_models)` models submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targted at specific conditions.
+`r length(partial_models)` models submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.
 

### 2022-07-21 docs/paper/paper.Rmd
correct partial models usage

@@
 Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
-`r length(partial_models)` models submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.
+`r nrow(partial_models)` models submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.
 

### 2022-07-22 docs/paper/paper.Rmd
add  bullet S and T's

@@
 
-# Forecasts
-
 # Models per forecast date
@@
 
-# Models with a single location target
+# Models with a location target
 locations_per_model <- metadata |>
@@
 
-#Varying submissions
+# Varying submissions
 varying_submissions <- locs_per_model_date |>
@@
 
-# Models coverage of forecast datees
+# Models coverage of forecast dats
 forecast_dates_by_models <- metadata |>
@@
 - Discussion country level differences in performance and link to potential growth rate trends. 
+- Discuss differences between countries with different numbers and kinds of submitted forecast models.
 
@@
 - Relative performance of surrogate model by forecast horizon.
+   - Overall performance (mean)
    - Give mean relative performance difference by horizon
@@
 
+- Real-time evaluation of a model with development tracked using version control (showing no changes since January 2022).
+- Conducted using reproducibility best practices with all code available for reuse.
+THe model is implemented in GitHub actions so can be reused without any compute resources and is by definition fast to run.
+- Only use absolute scoring which has implications for what we consider a good model.
+This is the approach used by the hub and so makes sense here but area of active research to understand what is more appropriate.
+- Limited exploration of model formulations. Likely that there are alternatives that perform closer to the ensemble
+- As our target is not to produce the best COVID-19 forecast model, or the best COVID-19 forecast model as defined using the metrics used to evaluate by the ECDC but rather to develop a model that closely matches the Hub ensemble whilst maintaining interpretability and using limited resources it is likely that it will perform less  well than a forecast model designed with the first two criterion in mind.
+  - However, if we start from the review that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance) then our approach makes sense.
+This is the view often used in the literature.
+
 ## Literature context {-}
@@
   - Other similar model formulations. More lag terms. Reduced run-times using approximate Bayesian approaches.
+  - Role of simple models for adjusting for susceptible population. Highlight useful now but not in early pandemic
+  - Follow up study comparing these results to retrospective forecasts from early in the pandemic or on other infectious disease outbreaks (though limited by requirement of a multi-team ensemble to compare to).
 

### 2022-07-22 docs/paper/paper.Rmd
flesh out some summary and s and t points

@@
 
+In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test positive reported COVID-19 cases in ECDC nations and the UK.
+We first derived a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experiance submitting forecasts to various forecast Hubs and our observations of the performance of other models.
+We restricted ourselves to a model that could be easily understand, that produced meaningful summary statistics, and that could be run with low compute resources.
+We further provide a fully reproducibile workflow for running this model using GitHub actions and code to evaluate its performance.
+
+Over the 6 months of the study period, we found that our surrogate model performed adequately at reproducing the forecast hub ensemble with average performance that was in general slightly worse though this varied by location and forecast date.
+Performance was closest for short forecast horizons and degraded relative to the forecast hub ensemble as the horizon increased.
+We observed that our surrogate model displayed similar levels of robustness to the forecast ensemble with few outlier forecasts unlike our previous forecasting efforts [@refNikosinGermany] and also partially captured the hub ensembles propensity to forecast reductions in incidence after periods of increase.
+Unlike the forecast hub ensemble our surrogate model was well calibrated at all horizons and credible intervals explored though it was overly uncertain at short horizons.
+We found that the majority of the difference in performance between our surrogate model and the hub ensemble occured in the tails of the predictive distribution with similar central forecasts.
+
+In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptionse and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action.
+We briefly outline some of the summary statistics it produces and their epidemiological relevance. 
+We then evaluate it's real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying its empirical coverage.
+We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed. 
+Finally, we summarise our findings, discuss their implications, and highlight areas for potential study.
+We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low resource, surrogate model in settings where large-scale colloborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models
+
 ## Stengths and weaknesses {-}
@@
 This is the view often used in the literature.
+- As the majority of the difference in relative performance occurred in the tail of the predictive distribution and that performance in the central predictive distribution was fairly similar this indicates that fore post-processing could make performance more similar.
+However, this would like reduce the calibration of our surrogate model which is not to be desired.
+Instead we hypothesis that this feature is driven by potentially inappropriate use of absolute scoring methods for exponential processes though our choice here was motivated by the methodology adopted by the forecast hubs.
 

### 2022-08-22 docs/paper/paper.Rmd
use data up to the 15th of August

@@
 
-  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from 15th of January 2022 to the 19th of July 2022 for the ensemble model (referenced as `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
+  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from 15th of January 2022 to the 15th of August 2022 for the ensemble model (referenced as `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
   We included all locations covered by the ECDC forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
@@
 
-We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK making `r locations` unique locations).
+We extracted forecasts starting from the 15th of January until the 1st of August 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK making `r locations` unique locations).
 In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the ECDC baseline model).
@@
 In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test positive reported COVID-19 cases in ECDC nations and the UK.
-We first derived a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experiance submitting forecasts to various forecast Hubs and our observations of the performance of other models.
-We restricted ourselves to a model that could be easily understand, that produced meaningful summary statistics, and that could be run with low compute resources.
+We first defined a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experiance submitting forecasts to various forecast Hubs and our observations of the performance of other models.
+We restricted ourselves to a model that could be easily understood, that produced meaningful summary statistics, and that could be run with low compute resources.
 We further provide a fully reproducibile workflow for running this model using GitHub actions and code to evaluate its performance.
@@
 We observed that our surrogate model displayed similar levels of robustness to the forecast ensemble with few outlier forecasts unlike our previous forecasting efforts [@refNikosinGermany] and also partially captured the hub ensembles propensity to forecast reductions in incidence after periods of increase.
-Unlike the forecast hub ensemble our surrogate model was well calibrated at all horizons and credible intervals explored though it was overly uncertain at short horizons.
+Unlike the forecast hub ensemble our surrogate model was well calibrated for most horizons and credible intervals explored though it was overly uncertain at short horizons.
 We found that the majority of the difference in performance between our surrogate model and the hub ensemble occured in the tails of the predictive distribution with similar central forecasts.
-
-In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptionse and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action.
-We briefly outline some of the summary statistics it produces and their epidemiological relevance. 
-We then evaluate it's real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying its empirical coverage.
-We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed. 
-Finally, we summarise our findings, discuss their implications, and highlight areas for potential study.
-We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low resource, surrogate model in settings where large-scale colloborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models
 

### 2022-08-22 docs/paper/paper.Rmd
revert back to 19th of July data

@@
 
-  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from 15th of January 2022 to the 15th of August 2022 for the ensemble model (referenced as `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
+  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from 15th of January 2022 to the 19th of July 2022 for the ensemble model (referenced as `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
   We included all locations covered by the ECDC forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
@@
 
-We extracted forecasts starting from the 15th of January until the 1st of August 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK making `r locations` unique locations).
+We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK making `r locations` unique locations).
 In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the ECDC baseline model).

### 2022-08-22 data-raw/get-hub-forecasts.R

@@ -20,8 +20,8 @@ models <- c("epiforecasts-weeklygrowth", "EuroCOVIDhub-ensemble")
 
 # First available forecast submission was the 15th of Janurary 2022 and has
 # continued each week
-# We freeze data extraction on the 15th August 2022
-up_to <- as.Date("2022-08-15")
+# We freeze data extraction on the 19th of July 2022
+up_to <- as.Date("2022-07-19")
 forecast_dates <- seq(as.Date("2022-01-15"), up_to, by = "day")
 
 # Get forecasts for models of interest

### 2022-08-22 docs/paper/paper.Rmd
add section for visual comparison

@@
 Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting.
-There is some suggestion that this was partially driven by the ensemble forecasting having non-exponetial uncertainty, unlike many infectious disease forecasting models.
+There is some suggestion that this was partially driven by the ensemble forecasting having non-exponential uncertainty, unlike many infectious disease forecasting models.
 
@@
 
-### Definition
+### Definition {-}
 
@@
 
-### Summary statistics
+### Summary statistics {-}
 
@@
 
-## Forecast evaluation
+## Forecast evaluation {-}
 
@@
 
-- Forecast performance overview for the ensemble and surrogate.
-- Forecast performance overview by forecast horizon. Highlight degrading performance with this impacting the surrogate model to a greater degree. 
-- Discussion country level differences in performance and link to potential growth rate trends. 
-- Discuss differences between countries with different numbers and kinds of submitted forecast models.
+In our example set of locations, the absolute performance of the ensemble and the surrogate model was similar in all locations at short forecast horizons though varied by location.
+Performance was most comparable in Slovakia but in the United Kingdom and Germany in particular the surrogate model outperformed visibly worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
+For both the ensemble and the surrogate performance visibly decreased as the forecast horizon increased though the relative difference in performance between models was relatively stable across horizons.
+In general in the study period the ensemble appeared to be liable to underpredict and the surrogate model appeared more likely to overpredict. 
+Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
+In comparison to the ensemble model the surrogate model appeared less likely to place weight on infeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in Febuary in Slokavkia and Poland).
+Italy and Germany, which had the highest number of submitted models of the example locations had little clear difference in performance between the ensemble and surrogate models compared to other example locations.
 

### 2022-08-23 docs/paper/paper.Rmd
add figure 2 caption

@@
 output:
-  rticles::peerj_article: default
   bookdown::pdf_book:
@@
 library(data.table)
+library(purrr)
 library(scoringutils)
@@
   calc_relative_score(cols = c("location_name", "target_end_date", "horizon"))
+
+relative_wis_by_horizon <- scores |>
+  summarise_scores(by = c("horizon", "model")) |>
+  calc_relative_score(cols = c("horizon")) |>
+  DT(, ris := round(relative_interval_score, 2))
+
+relative_median_wis_by_horizon <- relative_wis |>
+  DT(,
+     .(relative_interval_score = median(relative_interval_score)),
+     by = horizon
+  ) |>
+  DT(, ris := round(relative_interval_score, 2))
+
+relative_wis_thresholds <- relative_wis |>
+  DT(,
+    .(
+      better_than = sum(relative_interval_score < 1) / .N,
+      fifty_percent = sum(relative_interval_score < 1.5) / .N,
+      more_than_100_percent = sum(relative_interval_score > 2) / .N
+    )
+  ) |>
+  DT(, map(.SD, ~ round(. * 100, digits = 0))) |>
+  DT(, map(.SD, ~ paste0(., "%")))
 ```
@@
 The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 18}
-(
+p <- (
   plot_wis(wis, locs) +
@@
   )
+suppressWarnings(print(p))
 ```
@@
 
-- Relative performance of surrogate model by forecast horizon.
-   - Overall performance (mean)
-   - Give mean relative performance difference by horizon
-   - Discuss distribution.
-- Relative performance of surrogate model by month stratified by forecast horizon.
-  - Give mean relative performance by month for each horizon
-  - Discuss distribution
-- Relative performance of surrogate model by location sratified by forecast horizon.
-  - Discuss qualatative differences between locations
+Evaluating the ensemble and surrogate models using the weighted interval score across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
+Indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by approximately 100% on average with the surrogate model performing worst at the one and four week horizon.
+However, much of this outperformance was driven by a small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
+If we instead consider median relative performance (note not a proper score) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
+Looking across the distribution of relative scores by horizon we see little evidence that relative forecast quality decreased with time, though the tail of the distribution appears slightly heavier at the 4 week time horizon (Figure \@ref(fig:assemble-eval) a).
+`r relative_wis_thresholds[, better_than]` of surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecat, and scored `r relative_wis_thresholds[, more_than_100_percent]` performed worse than 100% more than the comparable ensemble forecast.
+
+The surrogate models relative performance varied over time with sustantially worse performance in January to March compared to later in the year across all forecast horizons (Figure \@ref(fig:assemble-eval) b).
+The majority in the difference in peformance appeared to be driven by a thicker right tail with this being particular feature of forecasts at longer horizons.
+Notably forecast performance in March had a big-model distribution at the four week horizon with a substational fraction of surrogate forecasts outperforming the ensemble.
+This viariation in performance may have been linked to the BA.2 wave which peaked in most locations during this time period.
+
+There was also substantial variation across forecast locations with the surrogate performing well in some locations at some forecast horizons, for example the four week horizon in the United Kingdom, and badly in others, for example the four week forecast in Switzerland (Figure \@ref(fig: assemble-eval) c).
+In general, across locations, as observed overall, relative forecast performance was comparable across horizons though with a heavier right tail at longer horizons.
+Some locations showed less of this behaviour, for example Spain, and in some it was very dominant, for exampel Switzerland.
 
@@
 
-```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%"}
- (
+```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale.
+a.) Density of the relative score by horizon.
+Horizontal black lines give the 5\\%, 35\\%, 65\\%, and 95\\% quantiles.
+b.) Density of the relative score by month for a given forecast horizon stratified by the one and four week forecast horizon.
+c.) Density of the relative score by forecast location stratified by the one and four week forecast horizon.
+The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast.
+The vertical black lines on the y axis give gives individual relative scores."}
+p <- (
   (
@@
 
-```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%"}
- (
+```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.), b.), c.)"}
+p <- (
   (

### 2022-08-23 docs/paper/paper.Rmd
complete draft results

@@
   )
+
+overall_ae_median <- scores |>
+  DT(, as.list(summary(ae_median)), by = "model")
+
 wis <- scores |>
@@
 The majority in the difference in peformance appeared to be driven by a thicker right tail with this being particular feature of forecasts at longer horizons.
-Notably forecast performance in March had a big-model distribution at the four week horizon with a substational fraction of surrogate forecasts outperforming the ensemble.
-This viariation in performance may have been linked to the BA.2 wave which peaked in most locations during this time period.
+Notably forecast performance in March had a bimodal distribution at the four week horizon with a substational fraction of surrogate forecasts outperforming the ensemble.
+This variation in performance may have been linked to the BA.2 wave which peaked in most locations during this time period.
 
@@
 
-- Calibration of forecasts at credible ranges of interest.
-- Calibration across quantiles
-- Relative interval scores by quantile
+Overall the surrogate model was well calibrated at the 30%, 60% and 90% credible intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger credible intervals (at 30%, at 60%, and at 90%).
+When stratified by forecast horizon the ensemble forecast was best calibrated at the one week forecast horizon and then became progressively less well calibrated as foreast horizon increased with the worst calibration at the 90% credible interval (Figure \@ref(fig:plot-coverage) a).
+In comparison the surrogate forecast was less well calibrated at the one week forecast horizon with a tendancy to have a larger empirical coverage than required at longer horizons the surrogate forecast transioned to being well calibrated and then had a lower coverage than optimal though to a much lower degree than the ensemble forecast.
+This was not the case for the 90% credible interval where the surrogate model stably covered slightly more than the expected interval indicating forecasts were somewhat overly uncertain for this intervals.
 
-```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.), b.), c.)"}
+Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly precise at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
+In contrast the surrogate ensemble forecast was generally equally well calibrated at shorter horizons with a tendency to produce conservative forecasts.
+At longer horizons however quantiles below the median were liable to have poor coverage suggesting a bias towards over-predicting.
+
+Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by larger intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) c).
+This is true across forecast horizons but the magnitude of the different increases.
+This implies that the surogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecasts we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
+However, when we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
+`r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) as noted elsewhere.
+This indicates a right skew in performance for the surrogate model point forecasts not observed for the ensemble forecast which matches that observed when evaluating the full predictive distributions of both forecasts.
+
+```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate and ensemble forecasts at the 90\\%, 60\\%, and 30\\% credible intervals stratified by forecast horizon.
+Ideally a well calibrated forecast should have empirical coverage for a given credible interval that equals the range of that interval (i.e 90\\%).
+b.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast.
+c.) Empirical coverage by quantile for both the surrrogate and ensemble forecasts.
+A well calibrated forecast should have empirical quantiles that match the theoretical ones.
+The green area of this figure corresponds to conservative forecasts."}
 p <- (

### 2022-08-23 R/coverage-plots.R

@@ -91,7 +91,7 @@ plot_rel_score_by_quantile <- function(relative_interval_score) {
     scale_colour_brewer(palette = "Accent") +
     theme_scoringutils() +
     labs(
-      x = "Quantile", y = "Median relative interval score",
+      x = "Interval", y = "Median relative interval score",
       col = "Forecast horizon (weeks)"
     )
   return(plot_interval_score)

### 2022-08-23 README.Rmd
further pruning and add renv support
repository clean up

@@
 
-# Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany
+# Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble
 
-This repository contains the documentation, results, and code of a project evaluating ... See the documentation for further details.
+This repository contains the documentation, results, and code of a project evaluating a simplified forecasting model in comparison the ECDC forecasting hub ensemble.
+See the documentation for further details.
 
@@
 
-## Documentation
+## Project structure
 
-```{r, child = "docs/chunks/documentation.Rmd", eval = TRUE}
-```
+Folder | Purpose
+---|---
 
-## Key files and folders
-
-Folder/File | Purpose
----|---
-[`writeup`](docs/) | Summary paper and additional supplementary information as `Rmarkdown` documents.
-[`_targets.Rmd`](_targets.Rmd) | Analysis workflow for interactive use.
+[`ecdc-weekly-growth-forecasts`](ecdc-weekly-growth-forecasts/) | The code for the simplified forecasting model evaluated in this work.
+[`data-raw`](data-raw/) | Raw input data and scripts required to download and process it.
+[`data`](data/) | Processed data from `data-raw` ready to be used in the paper analysis.
 [`R`](R/) | R functions used in the analysis and for evaluation.
-[`data`](data/) | Input data and summarised output generated by steps in the analysis.
+[`paper`](paper/) | Summary paper and additional supplementary information as `Rmarkdown` documents.
 [`.devcontainer`](.devcontainer/) | Resources for reproducibility using `vscode` and `docker`.
@@
 An easy way to make use of this is using the Remote development extension of `vscode`.
-
-## Analyses

### 2022-08-23 paper/paper.Rmd
repository clean up

--- 2022-08-23 docs/paper/paper.Rmd
@@
 title: Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble
-preprint: false
+preprint: true
 author: 
@@
 
-There was also substantial variation across forecast locations with the surrogate performing well in some locations at some forecast horizons, for example the four week horizon in the United Kingdom, and badly in others, for example the four week forecast in Switzerland (Figure \@ref(fig: assemble-eval) c).
+There was also substantial variation across forecast locations with the surrogate performing well in some locations at some forecast horizons, for example the four week horizon in the United Kingdom, and badly in others, for example the four week forecast in Switzerland (Figure \@ref(fig:assemble-eval) c).
 In general, across locations, as observed overall, relative forecast performance was comparable across horizons though with a heavier right tail at longer horizons.
-Some locations showed less of this behaviour, for example Spain, and in some it was very dominant, for exampel Switzerland.
+Some locations showed less of this behaviour, for example Spain, and in some it was very dominant, for example Switzerland.
 
@@
 
-Overall the surrogate model was well calibrated at the 30%, 60% and 90% credible intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger credible intervals (at 30%, at 60%, and at 90%).
+Overall the surrogate model was well calibrated at the 30%, 60% and 90% credible intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger credible intervals.
 When stratified by forecast horizon the ensemble forecast was best calibrated at the one week forecast horizon and then became progressively less well calibrated as foreast horizon increased with the worst calibration at the 90% credible interval (Figure \@ref(fig:plot-coverage) a).

### 2022-08-24 paper/paper.Rmd
update publishing

@@
 The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`) [@forecast.vocs].
+We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`. Other model formulations without this would be more efficient.
 The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
@@
 In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test positive reported COVID-19 cases in ECDC nations and the UK.
-We first defined a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experiance submitting forecasts to various forecast Hubs and our observations of the performance of other models.
-We restricted ourselves to a model that could be easily understood, that produced meaningful summary statistics, and that could be run with low compute resources.
+We first defined a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experiance submitting forecasts to various forecast Hubs, and our observations of the performance of other models.
+We restricted ourselves to a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources.
 We further provide a fully reproducibile workflow for running this model using GitHub actions and code to evaluate its performance.

### 2022-08-24 paper/paper.Rmd
write results summary

@@
 In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test positive reported COVID-19 cases in ECDC nations and the UK.
-We first defined a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experiance submitting forecasts to various forecast Hubs, and our observations of the performance of other models.
+We first defined a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experience submitting forecasts to various forecast Hubs, and our observations of the performance of other models.
 We restricted ourselves to a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources.
-We further provide a fully reproducibile workflow for running this model using GitHub actions and code to evaluate its performance.
+We further provide a fully reproducibile workflow for running and evaluating this model using GitHub actions facilitating others to do the same.
 
-Over the 6 months of the study period, we found that our surrogate model performed adequately at reproducing the forecast hub ensemble with average performance that was in general slightly worse though this varied by location and forecast date.
-Performance was closest for short forecast horizons and degraded relative to the forecast hub ensemble as the horizon increased.
-We observed that our surrogate model displayed similar levels of robustness to the forecast ensemble with few outlier forecasts unlike our previous forecasting efforts [@refNikosinGermany] and also partially captured the hub ensembles propensity to forecast reductions in incidence after periods of increase.
-Unlike the forecast hub ensemble our surrogate model was well calibrated for most horizons and credible intervals explored though it was overly uncertain at short horizons.
-We found that the majority of the difference in performance between our surrogate model and the hub ensemble occured in the tails of the predictive distribution with similar central forecasts.
+Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty.
+In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased towards underprediction, and the surrrogate model appeared biased towards overprediction.
+Evaluating the relative performance of the surrogate model compard to ECDC forecast hub ensemble we found that the mean performance was substantially worse but that this was largely driven by a relatively small subset of forecasts.
+The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
+Performance varied by location and forecast date with the surrogate mdoel performing worse in the first part of 2021 which may have been linked to increasing indicence rates accross forecast locations linked to BA.2.
+Unlike our earlier forecasting efforts the effect of outlier forecasts was much reduced though the surrogate model was less likely to outperform the ensemble than our earlier forecasting model methodology [@refNikosinGermany].
+In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison the the hub ensemble.
+As noted elsewhere, the forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger credible intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to over cover at large intervals.
+Stratifying the relative weighted interval score by interval we found that larger intervals (i.e the tails of the forecast distribution) were responsible for the majority of difference in performance difference.
+Indeed, median point performance of the forecasts were comparable though the surrogate model performed susbtantially worse on average due to its skewed performance distribution.
 

### 2022-08-24 paper/paper.Rmd
update fluff  section

@@
 
+Our study benefits from having been conducted using forecasts conducted in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the ECDC forecast hub).
+This means that we can be confident we have captured how our forecast methodology behaves in real-world usage and that others can evaluate this for themselves. 
+The downsides of this approach is that it was not possible to update the surrogate model over time in response to initial evaluation or to explore other parameterisations that might be more successful.
+However, as our study has been conducted with a focus on reproducibility and openness our findings can be replicated or extended by others regardless of compute availability (due to our use of GitHub actions as a compute platform which is freely available to researchers).
+Importantly in this study we focus on replicating the forecast hub ensembles observed beaviour rather than attempting to define an optimal forecast for forecast consumers.
+This means that if the forecast ensemble is not optimal then we are optimising for the wrong target. 
+That being said understanding what represents a useful forecast is difficult and an area of active research with a widespread assumption that large-scale ensembles of independent forecasts is the current state of the art.
+Our focus on replicating the performance of the hub ensemble is also useful as it highlights some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, modelled decrease in growth rate as the forecast horizon increases
+
 - Real-time evaluation of a model with development tracked using version control (showing no changes since January 2022).
@@
 
+# Additional Information and Declarations {-}
+
+## Competing interests {-}
+
+JB and SF, and KS have recieved funding to run forecast hub projects.
+
+## Author contribution {-}
+
+SA conceived the study, developed the initial set of assumptions for the surrogate model, implemented the model into code, designed and conducted the forecast evaluation, and wrote the first draft of the manuscript.
+All other authors provided feedback on the manuscript and contributed to revisions.
+SF provided funding.
+
+## Data availability {-}
+
+All data and code is available here: https://github.com/epiforecasts/simplified-forecaster-evaluation
+
+## Funding {-}
+
+SA, and SF were funded by a Wellcome senior fellowship to SF (), KS was funded by, and JB was funded by.
+
 # Acknowledgments {-}
 
+We thank the ECDC for support the forecasting hub, and all forecasters who submitted forecasts for making this study possible.
+We thank the forecast hub team for publishing all data in an accessible format.
+We thank Molly for being a good Labrador.
+
 # References

### 2022-08-24 paper/paper.Rmd
first draft stengths and weaknesses

@@
 Multi-model and  multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade with significant success [@REF].
-These approaches have drawn heavily on the experiance of other domains, for example climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
+These approaches have drawn heavily on the experience of other domains, for example climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
 However, ensemble forecasts have range of potential downsides with the most obvious being their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project, and that forecasts from these ensembles can be difficult to link to the underling model assumption.
@@
 
-# Methods {-}
+# Materials and Methods {-}
 
@@
 This means that we can be confident we have captured how our forecast methodology behaves in real-world usage and that others can evaluate this for themselves. 
-The downsides of this approach is that it was not possible to update the surrogate model over time in response to initial evaluation or to explore other parameterisations that might be more successful.
+The downsides of this approach is that it was not possible to update the surrogate model over time in response to initial evaluation or to explore other parameterisations that might be more successful of which there are likely several.
 However, as our study has been conducted with a focus on reproducibility and openness our findings can be replicated or extended by others regardless of compute availability (due to our use of GitHub actions as a compute platform which is freely available to researchers).
-Importantly in this study we focus on replicating the forecast hub ensembles observed beaviour rather than attempting to define an optimal forecast for forecast consumers.
-This means that if the forecast ensemble is not optimal then we are optimising for the wrong target. 
-That being said understanding what represents a useful forecast is difficult and an area of active research with a widespread assumption that large-scale ensembles of independent forecasts is the current state of the art.
+Importantly in this study we focussed on replicating the forecast hub ensembles observed beaviour rather than attempting to define an optimal forecast for forecast consumers.
+It is likely that if we had instead aimed to develop a forecast methodology that minimised the evaluation critiera we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
+However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance) then our approach makes sense and may be optimal for producing a "good" forecast.
+This is the view often used in the literature when describing forecast ensembles.
 Our focus on replicating the performance of the hub ensemble is also useful as it highlights some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, modelled decrease in growth rate as the forecast horizon increases
-
-- Real-time evaluation of a model with development tracked using version control (showing no changes since January 2022).
-- Conducted using reproducibility best practices with all code available for reuse.
-THe model is implemented in GitHub actions so can be reused without any compute resources and is by definition fast to run.
-- Only use absolute scoring which has implications for what we consider a good model.
-This is the approach used by the hub and so makes sense here but area of active research to understand what is more appropriate.
-- Limited exploration of model formulations. Likely that there are alternatives that perform closer to the ensemble
-- As our target is not to produce the best COVID-19 forecast model, or the best COVID-19 forecast model as defined using the metrics used to evaluate by the ECDC but rather to develop a model that closely matches the Hub ensemble whilst maintaining interpretability and using limited resources it is likely that it will perform less  well than a forecast model designed with the first two criterion in mind.
-  - However, if we start from the review that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance) then our approach makes sense.
-This is the view often used in the literature.
-- As the majority of the difference in relative performance occurred in the tail of the predictive distribution and that performance in the central predictive distribution was fairly similar this indicates that fore post-processing could make performance more similar.
-However, this would like reduce the calibration of our surrogate model which is not to be desired.
-Instead we hypothesis that this feature is driven by potentially inappropriate use of absolute scoring methods for exponential processes though our choice here was motivated by the methodology adopted by the forecast hubs.
+It also highlights some of the differences between our surrogate forecast model and the ensemble that may lead to new insights, such as the generally poor coverage of the ensemble that could not be explained by the assumptions we used in developing our surrogate methodology.
+Whilst we normalised reported cases to be population adjusted incidence rates our results are conditional on the use of the weighted interval score as an evaluation metrice.
+As this proper scoring rule measures absolute performance this means that forecasts during periods of higher incidence are given more weight than forecasts from periods of low incidence.
+It also means that potentially overpredicting is penalised more than underpredicting as incidence rates are bounded at zero but relatively weakly bounded by populations at the upper bound (as incidence rates are typically only a small fraction of the overall population).
+This bias could explain the relatively poor performance of the surrogate model, compared to the ensemble, despite the surrogate model being comparably well calibrated.
+We consider alternative methods of forecast evaluation that would be robust to this potential source of bias but choice to stick closely to the methodology used by the ECDC forecast hub as these choices inform the development of submitted models and so are key to our findings.
 
@@
   - Other similar model formulations. More lag terms. Reduced run-times using approximate Bayesian approaches.
-  - Role of simple models for adjusting for susceptible population. Highlight useful now but not in early pandemic
+  - Role of simple models for adjusting for susceptible population. Highlight useful now but not in early pandemic.
+  - As the majority of the difference in relative performance occurred in the tail of the predictive distribution and that performance in the central predictive distribution was fairly similar this indicates that fore post-processing could make performance more similar.
+However, this would like reduce the calibration of our surrogate model which is not to be desired.
+Instead we hypothesis that this feature is driven by potentially inappropriate use of absolute scoring methods for exponential processes though our choice here was motivated by the methodology adopted by the forecast hubs.
   - Follow up study comparing these results to retrospective forecasts from early in the pandemic or on other infectious disease outbreaks (though limited by requirement of a multi-team ensemble to compare to).
@@
 
-All data and code is available here: https://github.com/epiforecasts/simplified-forecaster-evaluation
+All data and code is available here:
+
+https://github.com/epiforecasts/simplified-forecaster-evaluation
 

### 2022-08-24 paper/paper.Rmd
add paper abstract

@@
     email: sam.abbott@lshtm.ac.uk
+  - name: Katharine Sherratt
+    affiliation: 1
   - name:  Johannes Bracher
@@
 abstract: > 
-  We made the best model. It is both simple and amazing.
+  Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts, especially during the COVID-19 pandemic.
+However, these forecasts require a coordinated team to produce and are effectively black boxes.
+In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models.
+Here we use observations of the behaviour of the ECDC forecast ensemble to specify a series of model assumptions.
+We then specify a forecast model using these assumptions and compare its performance to the ECDC forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four week forecast horizon.
+We found that this surrogate model performed visibly similarly to a subset of locations though with increased uncertainty.
+Using the weighted interval score we concluded that average performance was substantially worse but that median performance was more comparable.
+The majority of the difference in performance was linked to a subset of forecasts from a period of high incidence.
+We note that despite performing worse than the hub ensemble the majority of the time our surrogate model was better calibrated, and less liable to underpredict.
+This finding was supported by the observation that the central part of the surrogate models forecast distribution performed more similarly to ensemble forecasts distribution and that the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecat performance.
+We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders requirements.
+We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations.
+Our work maybe be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to methodology developers thinking about how to develop forecasts that perform similary to current multi-model and multi-teamn forecast ensembles.
 bibliography: library.bib

### 2022-08-24 paper/paper.Rmd
add abstract

@@
 We note that despite performing worse than the hub ensemble the majority of the time our surrogate model was better calibrated, and less liable to underpredict.
-This finding was supported by the observation that the central part of the surrogate models forecast distribution performed more similarly to ensemble forecasts distribution and that the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecat performance.
+This finding was supported by the observation that the central part of the surrogate models forecast distribution performed more similarly to ensemble forecasts distribution and that the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecast performance.
 We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders requirements.
-We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations.
-Our work maybe be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to methodology developers thinking about how to develop forecasts that perform similary to current multi-model and multi-teamn forecast ensembles.
+We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
+Our work maybe be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to methodology developers thinking about how to develop forecasts that perform similary to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
 bibliography: library.bib

### 2022-08-25 paper/paper.Rmd
add further work and conclusions

@@
 
-## Stengths and weaknesses {-}
+## Limitations {-}
 
@@
 
-  - Emulators in weather forecasting and elsewhere
-  - ECDC and CDC ensemble papers
-  - Other forecasting models contributed to the ensembles
-  - Baseline models
+There are no other studies in the epidemiology literature which we are aware of that attempt to develop a forecasting model based on the observed behaviour of a multi-team, multi-model ensemble.
+There are also few studies that focus on delivering computationally feasible forecasting models in a reproducibile framework backed by an openly accesible compute platform. 
+However, the CDC, ECDC, and Germany/Poland forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles. 
+In general these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study).
+The CDC forecast hub identified the top 5 performing models and noted they made the following structural assumptions.
+However, they did not extensively compare and contrast these conclusions in order arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model.
+Similary, the Germany and Poland, and ECDC forecasting hubs were able to identify forecast models that performed comparably well but did not derive structural assumptions that led to this performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
+All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers.
+In general, during the study period all projects used the same median ensemble forecast of all submissions after having concluded it was a favourable choice. 
+The CDC forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensemble.
+Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases.
+That being said no forcast hub has switched to these for their operational forecast implying they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the communities suggested default ensemble option.
+Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation  metric.
+Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and also noted the bias towards underprediction observed in the ensemble forecasts and not the model based forecasts [@refNikosinGermany].
 
@@
 
-Whilst we found that our surrogate model performed compab
-  - Other similar model formulations. More lag terms. Reduced run-times using approximate Bayesian approaches.
-  - Role of simple models for adjusting for susceptible population. Highlight useful now but not in early pandemic.
-  - As the majority of the difference in relative performance occurred in the tail of the predictive distribution and that performance in the central predictive distribution was fairly similar this indicates that fore post-processing could make performance more similar.
-However, this would like reduce the calibration of our surrogate model which is not to be desired.
-Instead we hypothesis that this feature is driven by potentially inappropriate use of absolute scoring methods for exponential processes though our choice here was motivated by the methodology adopted by the forecast hubs.
-  - Follow up study comparing these results to retrospective forecasts from early in the pandemic or on other infectious disease outbreaks (though limited by requirement of a multi-team ensemble to compare to).
+Whilst we derived our surrogate model from a range of assumptions based on observing ensemble forecasts behaviour and the behaviour and structure of submitted models this was by no means as structure or complete a process as it could have been.
+In follow up work a more rigorous approach to this could be taken in order to further refine this set of assumptions. 
+The findings from our study may also be useful for informing these improved assumptions. 
+In addition, the model we derived based on our assumptions was likely not optimal both in terms of compute time and accuracy at reproducing ensemble like behaviour.
+Models with a more complex auto-correlation structure, and more refined approaches to localise trends should be explored in order to improve relative performance to ensemble forecasts.
+An example of a family of possible approaches are structural time series model which have many of the characteristcs implied by our assumptions for how forecast ensembles typically operate.
+As we identified that the tails of our predictive distributions were responsible for a large proportion of the difference in performance compared to the forecast ensemble.
+It may be the case that post processing of forecasts from our surrogate model to enhance their similarity to the forecast ensemble is a sensible strategy.
+This seems likely to improve out of sample performance but does not help with the secondary aim of understanding the implicit assumptions driving the performance of multi-model, multi-team infectious disease forecast ensembles.
+As we have hypothesised that the use of absolute scoring measures is inappropriate and leads to performance characteristics that are unlikely to be favoured by forecast stakeholders more work should be done in this area.
+If new forecast ensemble methods are adopted as best practice by forecast hubs then follow up work attempting to create surrogate forecat models must also use these appproaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example the tendency to be poorly calibrated.
+Lastly, here we have only explored a surrogate for an ensemble for a single disease and a limited set of locations.
+Follow up work should explore whether this behaviour holds across diseases and locations.
+However, this work is limited by the lack of large-scale forecat ensembling projects despite them showing obvious promise to improve the forecasts available to stakeholders.
 
 ## Conclusions {-}
+
+We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders requirements.
+We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
+Our work maybe be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to methodology developers thinking about how to develop forecasts that perform similary to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
 

### 2022-08-26 paper/paper.Rmd
add hugo for code review

@@
   - name: Katharine Sherratt
+    affiliation: 1
+  - name: Hugo Gruson
     affiliation: 1
@@
 SF provided funding.
+HG reviewed the code and reproducibility of the analysis.
 

### 2022-08-26 paper/paper.Rmd
review abstract

@@
 Here we use observations of the behaviour of the ECDC forecast ensemble to specify a series of model assumptions.
-We then specify a forecast model using these assumptions and compare its performance to the ECDC forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four week forecast horizon.
+We then develop a forecast model using these assumptions and compare its performance to the ECDC forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four week forecast horizon.
 We found that this surrogate model performed visibly similarly to a subset of locations though with increased uncertainty.
@@
 We note that despite performing worse than the hub ensemble the majority of the time our surrogate model was better calibrated, and less liable to underpredict.
-This finding was supported by the observation that the central part of the surrogate models forecast distribution performed more similarly to ensemble forecasts distribution and that the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecast performance.
-We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders requirements.
+The central part of the surrogate models forecast distribution performed more similarly to ensemble forecasts distribution whilst the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecast performance.
+We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders requirements.
 We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
-Our work maybe be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to methodology developers thinking about how to develop forecasts that perform similary to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
+Our work maybe useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers developing forecast approaches that perform similary, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
 bibliography: library.bib
@@
 
-We thank the ECDC for support the forecasting hub, and all forecasters who submitted forecasts for making this study possible.
+We thank the ECDC for supporting the forecasting hub, and all forecasters who submitted forecasts for making this study possible.
 We thank the forecast hub team for publishing all data in an accessible format.

### 2022-08-26 paper/paper.Rmd
add Nikos

@@
   - name: Katharine Sherratt
+    affiliation: 1
+  - name: Nikos Bosse
     affiliation: 1

### 2022-08-26 paper/paper.Rmd
comments from @bisaloo

@@
   The ensemble forecast was constructed by taking the median of all predictive quantiles without the exlusion of any valid submitted forecast (where validity was defined as passing minimal formatting checks by the hub team).
-  Submitted forecasts and target observations are freely available with a permissive open-source license from the ECDC forecast hub GitHub repository.
-We provide code in the repository of this study to streamline access.
+  Submitted forecasts and target observations are freely available from the ECDC forecast hub GitHub repository. We provide code in the repository of this study to streamline access.
 
@@
 We then evaluated a subset of forecasts from the following countries Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom visually by forecast horizon (1-4 weeks) for both the ensemble and surrogate model.
+These countries were selected in order to include forecasts based on different numbers and types of submitted forecast models, to be at least somewhat representative of the full sample of forecast locations, and to include nations for which the authors had a good understanding of local data and transmission dynamics in the study period.
 
@@
 Of the remaining models submitted `r length(local_models)` were submitted in only one location.
-These models were gernerally submitted to a subset of locations with particular clustering in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the ECDC forecasting hub project).
+These models were generally submitted to a subset of locations with particular clustering in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the ECDC forecasting hub project).
 Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.

### 2022-08-26 paper/paper.Rmd
don't cache chunks loading packages from @bisaloo review

@@
 
-```{r packages}
+```{r packages, cache = FALSE}
 library(data.table)

### 2022-08-30 paper/paper.Rmd
review introduction

@@
 In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models.
-Here we use observations of the behaviour of the ECDC forecast ensemble to specify a series of model assumptions.
-We then develop a forecast model using these assumptions and compare its performance to the ECDC forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four week forecast horizon.
-We found that this surrogate model performed visibly similarly to a subset of locations though with increased uncertainty.
-Using the weighted interval score we concluded that average performance was substantially worse but that median performance was more comparable.
-The majority of the difference in performance was linked to a subset of forecasts from a period of high incidence.
+Here we use observations of the behaviour of the ECDC forecast ensemble, combined with our own forecasting experience,  to specify a series of model assumptions.
+We then develop a forecast model using these assumptions and compare its performance to the ECDC forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon.
+We found that this surrogate model performed visibly similarly to a subset of example locations though with increased uncertainty.
+Unsurprisingly, using the weighted interval score, we concluded that average performance was substantially worse but that median performance was more comparable.
+The majority of the difference in performance was linked to a few forecasts from a period of high incidence.
 We note that despite performing worse than the hub ensemble the majority of the time our surrogate model was better calibrated, and less liable to underpredict.
 The central part of the surrogate models forecast distribution performed more similarly to ensemble forecasts distribution whilst the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecast performance.
-We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders requirements.
+We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
 We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
-Our work maybe useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers developing forecast approaches that perform similary, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
+Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers developing forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
 bibliography: library.bib
@@
 Multi-model and  multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade with significant success [@REF].
-These approaches have drawn heavily on the experience of other domains, for example climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
-However, ensemble forecasts have range of potential downsides with the most obvious being their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project, and that forecasts from these ensembles can be difficult to link to the underling model assumption.
-In climate forecasting emulation approaches have been used to circumvent some of the  resource requirment issue by training a simplified model, usually a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
-However, within epidemiological modelling we often want to use our models for applications other than forecasting, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
-This means that non-parametric emulation is not an ideal solution. 
-In this work we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, limited computational resources, and an easily generalised implementation.
+These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
+However, ensemble forecasts have a range of potential downsides with the most obvious being their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project, and that forecasts from these ensembles can be difficult to link to the underlying model assumption.
+In climate forecasting emulation approaches have been used to circumvent some of the  resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
+However, within epidemiological modelling, we often want to use our models for applications other than forecasting and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
+This means that non-parametric emulation is less suitable for use in epidemiological forecasting though for epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@REFLSHTMTBEMULATOR].
+In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, limited computational resources, and an easily generalised implementation.
 
-The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the percieved low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and ECDC forecast hub [@ECDC] in 2021.
-All of these collaborations, which the authors (particularly SF and JB have been involved with managing),  ensemble contributions from multiple independent teams using a similar ensembling approach (most recently an unweighted median ensemble [@REF]) and have shown that their ensemble forecasts outperform most individually contributed forecasts and provide forecasts that are more robust to outliers [@REF].
+The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and ECDC forecast hub [@ECDC] in 2021.
+All of these collaborations, which the authors (particularly SF and JB have been involved with managing),  ensemble contributions from multiple independent teams using a similar ensembling approach (most recently an unweighted median ensemble [@REF]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting [@REF].
 However, they have struggled to provide feedback to those contributing forecasts that would allow them to improve their forecast approaches, have found it difficult to improve the quality of their ensemble forecasts [@REF], and have struggled to maintain incentives for forecasters to continue to contribute which may impact the long-term quality of the ensembles they produce.
 
-These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a 1 to 4 week time horizon.
-Generally ground truth data is available daily, and is updated in retrospect as new data becomes available.
-Typically, the ensemble forecast produced each week is made up of at least 5 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
-The majority of submitted forecasts make use of the available daily data, along with potentially other leading predictors to produce daily forecasts that are then aggregated to weekly forecasts.
+These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a one to four-week time horizon.
+Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available.
+Typically, the ensemble forecast produced each week comprises at least 5 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
+Most submitted forecasts use available daily data, along with potentially other leading predictors, to produce daily forecasts that are then aggregated into weekly forecasts.
 Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts.
-In general, 4 main classes of forecast models are used, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and it's generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
+In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and it's generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
 Real-time evaluation has shown that each of these classes of models may perform well though in general semi-mechanistic models, and models that include human insights have outperformed other approaches during certain periods of time [@REF].
@@
 Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
-Our previous methodology also requires signinficant computational resources running for an hour on a 16 core machine when producing forecasts from the ECDC. 
+Our previous methodology also requires signinficant computational resources running for an hour on a 16 core machine when producing forecasts for the ECDC forecasting hub. 
 In our model based forecasts we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
@@
 Unlike our epidemiologically motivated forecast submissions the hub ensemble forecasts were typically robust to daily reporting artefacts.
-They also demonstrated some ability to forecast future changes in trend that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight or assumptions about future interventions.
+They also demonstrated some ability to forecast future changes in trend that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions.
 In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence.
@@
 Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting.
-There is some suggestion that this was partially driven by the ensemble forecasting having non-exponential uncertainty, unlike many infectious disease forecasting models.
+There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models.
 
-Based on these insights from our own forecasting work and from observing the hub ensembles behaviour over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model based forecasting [@REF] with aim of producing ensemble like perform without sacrificing interpretability.
+Based on these insights from observing the performance of our own forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model based forecasting [@REF] with aim of producing ensemble like performance without sacrificing interpretability.
 The first simplification we make is to model only weekly data, rather than using daily data and then aggregating, this means that the impact daily reporting artefacts is mitigated.
@@
 The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
-This reduces computational requirements and removes the need for external information on the delay from infection to report which can potentially be misspecified leading to reduced forecast performance.
+This reduces computational requirements and removes the need for external information on the delay from infection to report which can potentially be misspecified leading to potentially reduced forecast performance.
 However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
 As discussed in [@gostic] this leads to varying amounts of bias depending on the epidemic phase. 
-The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a more complex method as we have done in other work [@EpiNow2].
-This represents a parsiminous approach in that we encode our expectation that the growth rate should vary over time and allow this to influance the forecast but we include only a single lag term and hence reduce the computational overhead of the model.
+The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process based method as we have done in other work [@EpiNow2].
+This represents a parsiminous approach in that we encode our expectation that the growth rate should vary over time and allow this to influance the forecast but we include only a single lag term, reducing the computational overhead of the model.
 In order to model potential unobserved interventions and more generally changes in transmission we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
-This reflects a simplified interpretation for how the ensemble appears to react to potential future changes.
-The only observed for which we do not make an adaption is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
-This choice is motivated by our belief that the underlying process is an exponential one and therefore an exponential error model is a sensible choice though this is an area of current research.
-We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the current usage of a generalsied form of the absolute error for the majority of forecast evaluation.
+This reflects a simplified interpretation for how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence. 
+The only observation for which we do not make an adaption is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
+This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model (where variance is linked to the mean)  is a sensible choice though this is an area of current research.
+We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalsied form of the absolute error for the majority of forecast evaluation.
 
-In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptionse and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action.
+In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action.
 We briefly outline some of the summary statistics it produces and their epidemiological relevance. 

### 2022-08-30 paper/paper.Rmd
review methods

@@
 
-  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from 15th of January 2022 to the 19th of July 2022 for the ensemble model (referenced as `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
+  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
   We included all locations covered by the ECDC forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
-  Data on notified weekly cases was originally sourced from the Johns Hopkins University curated data repository (JHU, [@REF]).
-  We used the latest available observed data. As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
-  The follows the common practice of the forecasting hub projects in treating the latest available data as the forecast target regardless of the number of data revisions.
+  Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@REF].
+  We used the latest available observed data as of the 20th of August 2022.
+As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
+  The follows the common practice of the forecasting hub projects in treating the latest available data as the forecast target regardless of the number of potential data revisions.
   Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.
@@
   No restrictions were placed on who could submit forecasts and the hub team actively invited particupation from research groups known to be involved with COVID-19 forecasting projects.
-  Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent based model, semi-mechanistic models, expert elicitation models, and ensembles of mutliple approaches (described at https://covid19forecasthub.eu/community.html).
-  Teams submitted forecasts at the latest two days after the complete dataset for the latest forecasting week became available and were allowed to use all data available at the time of submission. 
+  Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent based model, semi-mechanistic models, expert elicitation models, and ensembles of mutliple approaches [@https://covid19forecasthub.eu/community.html].
+  Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week).
   The ensemble forecast was constructed by taking the median of all predictive quantiles without the exlusion of any valid submitted forecast (where validity was defined as passing minimal formatting checks by the hub team).
-  Submitted forecasts and target observations are freely available from the ECDC forecast hub GitHub repository. We provide code in the repository of this study to streamline access.
+A ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
+  Submitted forecasts and target observations are available from the ECDC forecast hub GitHub repository [@REF]. We provide code in the repository of this study to streamline access.
 
@@
 $r_t$ is then modelled as a piecewise constant differenced AR(1) process modified such that the dependence of $r_{t-1}$ is multiplied by a decay factor ($\xi_{+,-}$) that varies dynamically according to the sign of $r_{t-1}$.
-The assumptions of this modelling approach are that the growth rate is non-stationary with a trend that is independent of the current growth rate (the differenced AR(1) process), the additional decay factor encodes the belief that larger absolute growth rates will tend more quickly towards no growth and that this process may work differently for positive or negative growth rates.
+This assumes that the growth rate is non-stationary with a trend that is independent of the current growth rate (the differenced AR(1) process), the additional decay factor encodes the belief that larger absolute growth rates will tend more quickly towards no growth and that this process may work differently for positive or negative growth rates.
 This process can be defined as follows,
@@
 
-Where $\sigma$, and $\frac{1}{\sqrt{phi}}$ are truncated to be greater than 0 and $\beta$ is truncated to be between -1 and 1.
-The Beta priors for $\xi{+,-}$ have been chosen to be weakly informative that the reduction towards 0 growth is relatively slow.
+Where $\sigma$, and $\frac{1}{\sqrt{\phi}}$ are truncated to be greater than 0 and $\beta$ is truncated to be between -1 and 1.
+The Beta priors for $\xi_{+,-}$ have been chosen to be weakly informative that the reduction towards 0 growth is relatively slow.
 Similarly the prior for $\beta$ has been chosen to be weakly informative that there is weak auto-correlation in differenced growth rates.
@@
 As well as posterior predictions and forecasts for notifcations the model also returns several epidemiological summary statistics which may be useful for drawing inferences about underlying transmission dynamics.
-These are the log scale growth rate ($g^{o, \delta}_t$), and the instantaneous effective reproduction number ($R^{o, \delta}_t$).
+These are the log scale weekly growth rate ($g^{o, \delta}_t$), and the instantaneous effective reproduction number ($R^{o, \delta}_t$).
 These are calculated as follows:
@@
 
-$T_s$ is a user set scaling parameter that defines the timespan over which the summary metrics apply dependent on the time step of the data.
+$T_s = 7$ is a parameter that defines the weekly timespan over which the summary metrics apply dependent on the time step of the data.
 
@@
 
-In order to standardise forecasts across forecast locations we first normalised both weekly notified test positive cases and forecast test positive cases by the population in the forecast region scaled to be a rate per 10,000 people.
+In order to standardise forecasts across forecast locations we first normalised both weekly notified test positive cases and forecast test positive cases by the population in the forecast region scaled to be an incidence rate per 10,000 people.
 We then evaluated a subset of forecasts from the following countries Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom visually by forecast horizon (1-4 weeks) for both the ensemble and surrogate model.
-These countries were selected in order to include forecasts based on different numbers and types of submitted forecast models, to be at least somewhat representative of the full sample of forecast locations, and to include nations for which the authors had a good understanding of local data and transmission dynamics in the study period.
+These countries were selected in order to include forecasts based on different numbers and types of submitted forecast models, to be at least partially representative of the full sample of forecast locations, and to include nations for which the authors had a good understanding of local data and transmission dynamics in the study period.
 
-We evaluate forecasts quantatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile based proper scoring rule that approximates the e continuous ranked probability score (CRPS).
-Both the WIS and CRPS are generalisations of absolute error in order to evaluate probablistic forecasts and are widely used to evaluate COVID-19 forecasts [@REF].
-We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks) as well as summarising across the study period.
+We evaluate forecasts for all locations and horizons quantatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile based proper scoring rule that approximates the continuous ranked probability score (CRPS).
+Both the WIS and CRPS are generalisations of absolute error in order to evaluate probablistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the ECDC forecast hub  [@REF].
+We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).
 
-In order to understand the relative performance of the surrogate model compared to the ensemble model we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model.
+In order to understand the relative performance of the surrogate model compared to the ensemble model we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
 In order to maintain the properiety of this score we do this after first taking the means of scores for the relavant stratification.
-We explore relative performance by forecast horizon, by forecast month and forecast horizon, and by forecast location and forecast horizon.
+We explore relative performance by forecast horizon, by month and horizon, and by location and horizon.
 
 In addition to presenting the WIS for a subset of locations and the relative WIS for all locations we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% credible intervals and by quantile.
-Lastly we calculate and visualise the relative interval score by quantile, stratified by forecast horizon.
+Lastly we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon.
 
@@
 The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`)  [@forecast.vocs].
-We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`. Other model formulations without this would be more efficient.
+We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`.
+Other model formulations without feature this would be more efficient and robust.
 The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
@@
 Here we take a conservative approach, as model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum treedepth setting to 15.
-For real-time usage convergance was not assessed but during model development the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergant transitions and exceedance of the maximum tree depth.
-During development posterior predictions were also visually compared to observed data [@cmdstanr].
+For real-time usage convergance was not assessed but during model development the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergant transitions and exceedance of the maximum tree depth [@cmdstanr].
+During development posterior predictions were also visually compared to observed data.
 
@@
 We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, to visualise forecasts and summary measures, and to summarise forecasts.
-Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils], and the `scoringRules` R package (1.0.1).
+Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils], and the `scoringRules` R package (1.0.1) [@REF].
 
-To ensure reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with built Docker image [@Boettiger:2015dw] (via GitHub Actions) is provided.
-Weekly forecasts are also made use of `renv` and Docker to ensure reproducibility with GitHub actions used to produce the forecasts ensuring we meet our goal of requring limited compute and that our implementation is independent from local resources resulting in democratised access to our approach.
+To ensure reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with built Docker image [@Boettiger:2015dw] (via GitHub Actions) is provided in the code repository.
+Weekly forecasts are also made using of `renv` and GitHub actions to ensure they require limited compute and that our implementation is independent from local resources facilitating democratised access.
 The code for this analysis can be found here: https://github.com/epiforecasts/simplified-forecaster-evaluation
 The code for the forecating model defined above along with the infrastructure required to forecast using GitHub Actions can be found here: https://github.com/seabbs/ecdc-weekly-growth-forecasts
+Versions archived on Zenodo are available [@REF] and [@REF].
 

### 2022-08-30 paper/paper.Rmd
first section of results reviewed

@@
 Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts.
-In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and it's generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
+In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
 Real-time evaluation has shown that each of these classes of models may perform well though in general semi-mechanistic models, and models that include human insights have outperformed other approaches during certain periods of time [@REF].
-We have contributed a range of forecasts to these forecast collobarations generally focussed on semi-mechanistic statistical methods [@REF], and human insight based forecasts [@REF] with our forecasts performing relatively well compared to the majority of those submitted but rarely outperforming significantly.
-The model based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within week, reporting periodicity [@REF].
+We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@REF], and human insight-based forecasts [@REF] with our forecasts performing relatively well compared to the majority of those submitted but rarely outperforming significantly.
+The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within week, reporting periodicity [@REF].
 Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
-Our previous methodology also requires signinficant computational resources running for an hour on a 16 core machine when producing forecasts for the ECDC forecasting hub. 
-In our model based forecasts we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
-We found that our human insight based forecasts significantly outperformed our model based forecasts and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@REF].
-Unlike our epidemiologically motivated forecast submissions the hub ensemble forecasts were typically robust to daily reporting artefacts.
-They also demonstrated some ability to forecast future changes in trend that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions.
+Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the ECDC forecasting hub. 
+In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
+We found that our human insight-based forecasts significantly outperformed our model-based forecasts and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@REF].
+Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts.
+They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions.
 In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence.
-On the other hand this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
+On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
 Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting.
@@
 
-Based on these insights from observing the performance of our own forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model based forecasting [@REF] with aim of producing ensemble like performance without sacrificing interpretability.
-The first simplification we make is to model only weekly data, rather than using daily data and then aggregating, this means that the impact daily reporting artefacts is mitigated.
-It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain signifcant weight in the model this leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
+Based on these insights from observing the performance of our own forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with aim of producing ensemble like performance without sacrificing interpretability.
+The first simplification we make is to model only weekly data, rather than using daily data and then aggregating, this means that the impact of daily reporting artefacts is mitigated.
+It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model this leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
 The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
@@
 As discussed in [@gostic] this leads to varying amounts of bias depending on the epidemic phase. 
-The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process based method as we have done in other work [@EpiNow2].
-This represents a parsiminous approach in that we encode our expectation that the growth rate should vary over time and allow this to influance the forecast but we include only a single lag term, reducing the computational overhead of the model.
-In order to model potential unobserved interventions and more generally changes in transmission we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
-This reflects a simplified interpretation for how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence. 
+The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@EpiNow2].
+This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
+In order to model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
+This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence. 
 The only observation for which we do not make an adaption is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
 This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model (where variance is linked to the mean)  is a sensible choice though this is an area of current research.
-We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalsied form of the absolute error for the majority of forecast evaluation.
+We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations.
 
-In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action.
+In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions and summarising its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action.
 We briefly outline some of the summary statistics it produces and their epidemiological relevance. 
-We then evaluate it's real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying its empirical coverage.
+We then evaluate its real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying its empirical coverage.
 We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed. 
 Finally, we summarise our findings, discuss their implications, and highlight areas for potential study.
-We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low resource, surrogate model in settings where large-scale colloborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models
+We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models
 
@@
 
-  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
+  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
   We included all locations covered by the ECDC forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
@@
 As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
-  The follows the common practice of the forecasting hub projects in treating the latest available data as the forecast target regardless of the number of potential data revisions.
+  This follows the common practice of the forecasting hub projects in treating the latest available data as the forecast target regardless of the number of potential data revisions.
   Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.
 
-  The ECDC forecast hub requests forecasts for 1 to 4 week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
-  No restrictions were placed on who could submit forecasts and the hub team actively invited particupation from research groups known to be involved with COVID-19 forecasting projects.
-  Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent based model, semi-mechanistic models, expert elicitation models, and ensembles of mutliple approaches [@https://covid19forecasthub.eu/community.html].
+  The ECDC forecast hub requests forecasts for 1 to 4-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
+  No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
+  Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@https://covid19forecasthub.eu/community.html].
   Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week).
-  The ensemble forecast was constructed by taking the median of all predictive quantiles without the exlusion of any valid submitted forecast (where validity was defined as passing minimal formatting checks by the hub team).
-A ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
+  The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any valid submitted forecast (where validity was defined as passing minimal formatting checks by the hub team).
+An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
   Submitted forecasts and target observations are available from the ECDC forecast hub GitHub repository [@REF]. We provide code in the repository of this study to streamline access.
@@
 
-As well as posterior predictions and forecasts for notifcations the model also returns several epidemiological summary statistics which may be useful for drawing inferences about underlying transmission dynamics.
+As well as posterior predictions and forecasts for notifications the model also returns several epidemiological summary statistics which may be useful for drawing inferences about underlying transmission dynamics.
 These are the log scale weekly growth rate ($g^{o, \delta}_t$), and the instantaneous effective reproduction number ($R^{o, \delta}_t$).
@@
 
-We evaluate forecasts for all locations and horizons quantatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile based proper scoring rule that approximates the continuous ranked probability score (CRPS).
-Both the WIS and CRPS are generalisations of absolute error in order to evaluate probablistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the ECDC forecast hub  [@REF].
+We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
+Both the WIS and CRPS are generalisations of the absolute error in order to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the ECDC forecast hub  [@REF].
 We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).
 
-In order to understand the relative performance of the surrogate model compared to the ensemble model we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
-In order to maintain the properiety of this score we do this after first taking the means of scores for the relavant stratification.
+In order to understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
+In order to maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
 We explore relative performance by forecast horizon, by month and horizon, and by location and horizon.
 
-In addition to presenting the WIS for a subset of locations and the relative WIS for all locations we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% credible intervals and by quantile.
-Lastly we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon.
+In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% credible intervals and by quantile.
+Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon.
 
@@
 We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`.
-Other model formulations without feature this would be more efficient and robust.
+Other model formulations without this feature would be more efficient and robust.
 The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
-`cmstanr` surfaces several settings which trade-off between sampling speed and the robustness of approach.
-Here we take a conservative approach, as model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum treedepth setting to 15.
-For real-time usage convergance was not assessed but during model development the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergant transitions and exceedance of the maximum tree depth [@cmdstanr].
-During development posterior predictions were also visually compared to observed data.
+`cmstanr` surfaces several settings that trade-off between sampling speed and the robustness of the approach.
+Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15.
+For real-time usage convergence was not assessed but during model development, the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergent transitions and exceedance of the maximum tree depth [@cmdstanr].
+During development, posterior predictions were also visually compared to observed data.
 
 To download and manipulate forecasts from the ECDC forecasting hub [@EuroHub] we use  the `data.table` (`1.14.2`) [@data.table] and `gh` (`1.3.0`) [@gh] R packages.
-We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, to visualise forecasts and summary measures, and to summarise forecasts.
+We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, visualise forecasts and summary measures, and summarise forecasts.
 Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils], and the `scoringRules` R package (1.0.1) [@REF].
 
-To ensure reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with built Docker image [@Boettiger:2015dw] (via GitHub Actions) is provided in the code repository.
-Weekly forecasts are also made using of `renv` and GitHub actions to ensure they require limited compute and that our implementation is independent from local resources facilitating democratised access.
+To ensure the reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with a built Docker image [@Boettiger:2015dw] (via GitHub Actions) is provided in the code repository.
+Weekly forecasts are also made using `renv` and GitHub actions to ensure they require limited compute and that our implementation is independent of local resources facilitating democratised access.
 The code for this analysis can be found here: https://github.com/epiforecasts/simplified-forecaster-evaluation
-The code for the forecating model defined above along with the infrastructure required to forecast using GitHub Actions can be found here: https://github.com/seabbs/ecdc-weekly-growth-forecasts
+The code for the forecasting model defined above along with the infrastructure required to forecast using GitHub Actions can be found here: https://github.com/seabbs/ecdc-weekly-growth-forecasts
 Versions archived on Zenodo are available [@REF] and [@REF].
@@
 
-In our study period, incidence rates across ECDC nations and in the UK were primary driven by the spread of noval subvariants of concern related to the Omicron variant.
-Many nations, such as the UK saw large BA.1 waves in early January to late January that then resulted in declining incidence rates through Febuary.
-From late Febuary, througth to the end of May, most nations then saw another wave, typically with lower reported incidence rates, driven by BA.2.
-This wave was typically characterised by a lower peak than the BA.1 wave but a longer tail.
+In our study period, incidence rates across ECDC nations and in the UK were primarily driven by the spread of novel subvariants of concern related to the Omicron variant and changes in population susceptibility.
+Many nations, such as the UK saw large BA.1 waves in January that then resulted in declining incidence rates through February.
+From late February, through to the end of May, most nations then saw another wave, typically with lower reported incidence rates, driven by BA.2.
+This wave was typically characterised by a lower peak than the BA.1 wave with a more gradual decrease in incidence.
 The end of our study period was characterised by the gradual take-over of the BA.4/BA.5 subvariants that again had a lower peak and lower absolute growth rates.
 Unlike earlier periods in the pandemic, our study period did not see the use of new non-pharmaceutical interventions (NPIs) in response to increasing COVID-19 incidence in most locations.
-In addition, in general ascertainment rates likely reduced over time in most locations due to reductions in routine testing, and reductions in test availability.
-Whilst both the reduced use of NPIs and testing generally occured across nations our study period also marked an increase in the heterogenity of the response to COVID-19 pandemic with nations changing policy at different times and to different degrees.
+In addition, in general, ascertainment rates likely reduced over time in most locations due to reductions in routine testing, and reductions in test availability.
+Whilst both the reduced use of NPIs and testing generally occurred across nations our study period also marked an increase in the heterogeneity of the response to the COVID-19 pandemic with nations changing policy at different times and to different degrees.
 This is in contrast to the early COVID-19 pandemic response for which most nations took similar actions at similar times.
@@
 Of the remaining models submitted `r length(local_models)` were submitted in only one location.
-These models were generally submitted to a subset of locations with particular clustering in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the ECDC forecasting hub project).
+Single location models were clustered in a few locations, particularly Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the ECDC forecasting hub project).
 Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
-`r nrow(partial_models)` models submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.
+`r nrow(partial_models)` models were submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.
 
-Across all forecast dates and locations the minimum number of independent forecasts was `r models_per_loc_date[, min(n)]` with the maxmimum being `r models_per_loc_date[, max(n)]`.
+Across all forecast dates and locations the minimum number of independent forecasts was `r models_per_loc_date[, min(n)]` with the maximum being `r models_per_loc_date[, max(n)]`.
 The median number of independent forecasts per location and forecast date was `r models_per_loc_date[, median(n)]`.
-All locations recieved forecasts from at least `r models_per_location[, min(n)]` models with the median number of forecast models per location being `r models_per_location[, median(n)]`.
+All locations received forecasts from at least `r models_per_location[, min(n)]` models with the median number of forecast models per location being `r models_per_location[, median(n)]`.
 Coverage of forecast dates varied across submitted models with `r forecast_dates_by_models[n_per == 1, .N]` models submitting for all dates, `r forecast_dates_by_models[n_per >= 0.9, .N]` models submitting for at least 90% of dates, and `r forecast_dates_by_models[n_per < 0.50, .N]` models submitting for fewer than 50% of forecast dates.
@@
 
-In our example set of locations, the absolute performance of the ensemble and the surrogate model was similar in all locations at short forecast horizons though varied by location.
-Performance was most comparable in Slovakia but in the United Kingdom and Germany in particular the surrogate model outperformed visibly worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
+In our example set of locations, the absolute performance of the ensemble and the surrogate model was similar in all locations at short forecast horizons though this varied by location.
+Performance was most comparable in Slovakia but in the United Kingdom and Germany in particular the surrogate model performed visibly worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
 For both the ensemble and the surrogate performance visibly decreased as the forecast horizon increased though the relative difference in performance between models was relatively stable across horizons.
-In general in the study period the ensemble appeared to be liable to underpredict and the surrogate model appeared more likely to overpredict. 
+In general, in the study period, the ensemble appeared to be liable to underpredict and the surrogate model appeared more likely to over-predict. 
 Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
-In comparison to the ensemble model the surrogate model appeared less likely to place weight on infeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in Febuary in Slokavkia and Poland).
+In comparison to the ensemble model the surrogate model appeared less likely to place weight on infeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in February in Slovakia and Poland).
 Italy and Germany, which had the highest number of submitted models of the example locations had little clear difference in performance between the ensemble and surrogate models compared to other example locations.
 
-```{r vis-forecasts, fig.cap = "a.) Weighted interval score at the 1 week and 4 week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom.
+```{r vis-forecasts, fig.cap = "a.) Weighted interval scores at the 1-week and 4-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom.
 b.)Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).

### 2022-08-30 paper/paper.Rmd
grammar check

@@
 In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
-Real-time evaluation has shown that each of these classes of models may perform well though in general semi-mechanistic models, and models that include human insights have outperformed other approaches during certain periods of time [@REF].
+Real-time evaluation has shown that each of these classes of models may perform well though in general semi-mechanistic models, and models that include human insights have outperformed other approaches during certain periods [@REF].
 We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@REF], and human insight-based forecasts [@REF] with our forecasts performing relatively well compared to the majority of those submitted but rarely outperforming significantly.
-The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within week, reporting periodicity [@REF].
+The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@REF].
 Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
@@
 
-Based on these insights from observing the performance of our own forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with aim of producing ensemble like performance without sacrificing interpretability.
+Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with aim of producing ensemble-like performance without sacrificing interpretability.
 The first simplification we make is to model only weekly data, rather than using daily data and then aggregating, this means that the impact of daily reporting artefacts is mitigated.
@@
 This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
-In order to model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
+To model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
 This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence. 
@@
 
-  The ECDC forecast hub requests forecasts for 1 to 4-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
+  The ECDC forecast hub requests forecasts for one to four-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
   No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
@@
   Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week).
-  The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any valid submitted forecast (where validity was defined as passing minimal formatting checks by the hub team).
+  The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team).
 An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
@@
 
-# Models coverage of forecast dats
+# Models coverage of forecast dates
 forecast_dates_by_models <- metadata |>
@@
 We model the expectation ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(P)) process by epidemiological week ($t$).
-The model is initialised by assuming that the initial reported cases are representative with a small amount of error (2.5%).
+The model is initialised by assuming that the initially reported cases are representative with a small amount of error (2.5%).
 We assume a negative binomial observation model with overdispersion $\phi$ for reported cases ($C_t$).
@@
 
-In order to standardise forecasts across forecast locations we first normalised both weekly notified test positive cases and forecast test positive cases by the population in the forecast region scaled to be an incidence rate per 10,000 people.
+To standardise forecasts across forecast locations we first normalised both weekly notified test positive cases and forecast test positive cases by the population in the forecast region scaled to be an incidence rate per 10,000 people.
 We then evaluated a subset of forecasts from the following countries Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom visually by forecast horizon (1-4 weeks) for both the ensemble and surrogate model.
-These countries were selected in order to include forecasts based on different numbers and types of submitted forecast models, to be at least partially representative of the full sample of forecast locations, and to include nations for which the authors had a good understanding of local data and transmission dynamics in the study period.
+These countries were selected to include forecasts based on different numbers and types of submitted forecast models, to be at least partially representative of the full sample of forecast locations, and to include nations for which the authors had a good understanding of local data and transmission dynamics in the study period.
 
 We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
-Both the WIS and CRPS are generalisations of the absolute error in order to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the ECDC forecast hub  [@REF].
+Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the ECDC forecast hub  [@REF].
 We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).
 
-In order to understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
-In order to maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
+To understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
+To maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
 We explore relative performance by forecast horizon, by month and horizon, and by location and horizon.
@@
 The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
-`cmstanr` surfaces several settings that trade-off between sampling speed and the robustness of the approach.
+`cmstanr` surfaces several settings that trade off between sampling speed and the robustness of the approach.
 Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15.
@@
 
-```{r vis-forecasts, fig.cap = "a.) Weighted interval scores at the 1-week and 4-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom.
+```{r vis-forecasts, fig.cap = "a.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom.
 b.)Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).
@@
 Evaluating the ensemble and surrogate models using the weighted interval score across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
-Indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by approximately 100% on average with the surrogate model performing worst at the one and four week horizon.
+Indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by approximately 100% on average with the surrogate model performing worst at the one and four-week horizon.
 However, much of this outperformance was driven by a small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
 If we instead consider median relative performance (note not a proper score) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
-Looking across the distribution of relative scores by horizon we see little evidence that relative forecast quality decreased with time, though the tail of the distribution appears slightly heavier at the 4 week time horizon (Figure \@ref(fig:assemble-eval) a).
+Looking across the distribution of relative scores by horizon we see little evidence that relative forecast quality decreased with time, though the tail of the distribution appears slightly heavier at the four-week time horizon (Figure \@ref(fig:assemble-eval) a).
 `r relative_wis_thresholds[, better_than]` of surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecat, and scored `r relative_wis_thresholds[, more_than_100_percent]` performed worse than 100% more than the comparable ensemble forecast.
 
-The surrogate models relative performance varied over time with sustantially worse performance in January to March compared to later in the year across all forecast horizons (Figure \@ref(fig:assemble-eval) b).
-The majority in the difference in peformance appeared to be driven by a thicker right tail with this being particular feature of forecasts at longer horizons.
-Notably forecast performance in March had a bimodal distribution at the four week horizon with a substational fraction of surrogate forecasts outperforming the ensemble.
-This variation in performance may have been linked to the BA.2 wave which peaked in most locations during this time period.
+The surrogate model's relative performance varied over time with substantially worse performance from January to March compared to later in the year across all forecast horizons (Figure \@ref(fig:assemble-eval) b).
+The majority of the difference in performance appeared to be driven by a thicker right tail with this being a particular feature of forecasts at longer horizons.
+Notably, forecast performance in March had a bimodal distribution at the four-week horizon with a substantial fraction of surrogate forecasts outperforming the ensemble.
+This variation in performance may have been linked to the BA.2 wave which peaked in most locations during this period.
 
-There was also substantial variation across forecast locations with the surrogate performing well in some locations at some forecast horizons, for example the four week horizon in the United Kingdom, and badly in others, for example the four week forecast in Switzerland (Figure \@ref(fig:assemble-eval) c).
+There was also substantial variation across forecast locations with the surrogate performing well in some locations at some forecast horizons, for example, the four-week horizon in the United Kingdom, and badly in others, for example, the four-week forecast in Switzerland (Figure \@ref(fig:assemble-eval) c).
 In general, across locations, as observed overall, relative forecast performance was comparable across horizons though with a heavier right tail at longer horizons.
-Some locations showed less of this behaviour, for example Spain, and in some it was very dominant, for example Switzerland.
+Some locations showed less of this behaviour, for example, Spain, and in some, it was very dominant, for example, Switzerland.
 
@@
 ```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale.
-a.) Density of the relative score by horizon.
+a.) The density of the relative score by horizon.
 Horizontal black lines give the 5\\%, 35\\%, 65\\%, and 95\\% quantiles.
-b.) Density of the relative score by month for a given forecast horizon stratified by the one and four week forecast horizon.
-c.) Density of the relative score by forecast location stratified by the one and four week forecast horizon.
+b.) The density of the relative score by month for a given forecast horizon stratified by the one and four-week forecast horizon.
+c.) The density of the relative score by forecast location stratified by the one and four-week forecast horizon.
 The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast.
-The vertical black lines on the y axis give gives individual relative scores."}
+The vertical black lines on the y-axis give individual relative scores."}
 p <- (
@@
 Overall the surrogate model was well calibrated at the 30%, 60% and 90% credible intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger credible intervals.
-When stratified by forecast horizon the ensemble forecast was best calibrated at the one week forecast horizon and then became progressively less well calibrated as foreast horizon increased with the worst calibration at the 90% credible interval (Figure \@ref(fig:plot-coverage) a).
-In comparison the surrogate forecast was less well calibrated at the one week forecast horizon with a tendancy to have a larger empirical coverage than required at longer horizons the surrogate forecast transioned to being well calibrated and then had a lower coverage than optimal though to a much lower degree than the ensemble forecast.
-This was not the case for the 90% credible interval where the surrogate model stably covered slightly more than the expected interval indicating forecasts were somewhat overly uncertain for this intervals.
+When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon and then became progressively less well calibrated as the forecast horizon increased with the worst calibration at the 90% credible interval (Figure \@ref(fig:plot-coverage) a).
+In comparison, the surrogate forecast was less well calibrated at the one-week forecast horizon with a tendency to have a larger empirical coverage than required at longer horizons the surrogate forecast transitioned to being well calibrated and then had lower coverage than optimal though to a much lower degree than the ensemble forecast.
+This was not the case for the 90% credible interval where the surrogate model stably covered slightly more than the expected interval indicating forecasts were somewhat overly uncertain for this interval.
 
 Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly precise at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
-In contrast the surrogate ensemble forecast was generally equally well calibrated at shorter horizons with a tendency to produce conservative forecasts.
-At longer horizons however quantiles below the median were liable to have poor coverage suggesting a bias towards over-predicting.
+In contrast, the surrogate ensemble forecast was generally equally well calibrated at shorter horizons with a tendency to produce conservative forecasts.
+At longer horizons, however, quantiles below the median were liable to have poor coverage suggesting a bias towards over-predicting.
 
@@
 This is true across forecast horizons but the magnitude of the different increases.
-This implies that the surogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecasts we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
+This implies that the surrogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecasts we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
 However, when we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
@@
 ```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate and ensemble forecasts at the 90\\%, 60\\%, and 30\\% credible intervals stratified by forecast horizon.
-Ideally a well calibrated forecast should have empirical coverage for a given credible interval that equals the range of that interval (i.e 90\\%).
+Ideally, a well-calibrated forecast should have empirical coverage for a given credible interval that equals the range of that interval (i.e 90\\%).
 b.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast.
-c.) Empirical coverage by quantile for both the surrrogate and ensemble forecasts.
-A well calibrated forecast should have empirical quantiles that match the theoretical ones.
+c.) Empirical coverage by quantile for both the surrogate and ensemble forecasts.
+A well-calibrated forecast should have empirical quantiles that match the theoretical ones.
 The green area of this figure corresponds to conservative forecasts."}
@@
 
-In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test positive reported COVID-19 cases in ECDC nations and the UK.
-We first defined a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experience submitting forecasts to various forecast Hubs, and our observations of the performance of other models.
+In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test-positive reported COVID-19 cases in ECDC nations and the UK.
+We first defined a set of assumptions for how the surrogate model should behave based on our observations of the ECDC forecast hub ensemble, our experience submitting forecasts to various forecast Hubs, and our observations of the performance of other models.
 We restricted ourselves to a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources.
-We further provide a fully reproducibile workflow for running and evaluating this model using GitHub actions facilitating others to do the same.
+We further provide a fully reproducible workflow for running and evaluating this model using GitHub actions facilitating others to do the same.
 
 Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty.
-In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased towards underprediction, and the surrrogate model appeared biased towards overprediction.
-Evaluating the relative performance of the surrogate model compard to ECDC forecast hub ensemble we found that the mean performance was substantially worse but that this was largely driven by a relatively small subset of forecasts.
+In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared biased towards overprediction.
+Evaluating the relative performance of the surrogate model compared to the ECDC forecast hub ensemble we found that the mean performance was substantially worse but that this was largely driven by a relatively small subset of forecasts.
 The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
-Performance varied by location and forecast date with the surrogate mdoel performing worse in the first part of 2021 which may have been linked to increasing indicence rates accross forecast locations linked to BA.2.
-Unlike our earlier forecasting efforts the effect of outlier forecasts was much reduced though the surrogate model was less likely to outperform the ensemble than our earlier forecasting model methodology [@refNikosinGermany].
-In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison the the hub ensemble.
-As noted elsewhere, the forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger credible intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to over cover at large intervals.
-Stratifying the relative weighted interval score by interval we found that larger intervals (i.e the tails of the forecast distribution) were responsible for the majority of difference in performance difference.
-Indeed, median point performance of the forecasts were comparable though the surrogate model performed susbtantially worse on average due to its skewed performance distribution.
+Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2021 which may have been linked to increasing incidence rates across forecast locations linked to BA.2.
+Unlike our earlier forecasting efforts, the effect of outlier forecasts was much reduced though the surrogate model was less likely to outperform the ensemble than our earlier forecasting model methodology [@refNikosinGermany].
+In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison the hub ensemble.
+As noted elsewhere, the forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger credible intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to overcover at large intervals.
+Stratifying the relative weighted interval score by interval we found that larger intervals (i.e the tails of the forecast distribution) were responsible for the majority of the difference in performance.
+Indeed, the median point performance of the forecasts was comparable though the surrogate model performed substantially worse on average due to its skewed performance distribution.
 
@@
 This means that we can be confident we have captured how our forecast methodology behaves in real-world usage and that others can evaluate this for themselves. 
-The downsides of this approach is that it was not possible to update the surrogate model over time in response to initial evaluation or to explore other parameterisations that might be more successful of which there are likely several.
+The downside of this approach is that it was not possible to update the surrogate model over time in response to initial evaluation or to explore other parameterisations that might be more successful of which there are likely several.
 However, as our study has been conducted with a focus on reproducibility and openness our findings can be replicated or extended by others regardless of compute availability (due to our use of GitHub actions as a compute platform which is freely available to researchers).
-Importantly in this study we focussed on replicating the forecast hub ensembles observed beaviour rather than attempting to define an optimal forecast for forecast consumers.
-It is likely that if we had instead aimed to develop a forecast methodology that minimised the evaluation critiera we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
+Importantly in this study, we focussed on replicating the forecast hub ensembles' observed behaviour rather than attempting to define an optimal forecast for forecast consumers.
+It is likely that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
 However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance) then our approach makes sense and may be optimal for producing a "good" forecast.
@@
 It also highlights some of the differences between our surrogate forecast model and the ensemble that may lead to new insights, such as the generally poor coverage of the ensemble that could not be explained by the assumptions we used in developing our surrogate methodology.
-Whilst we normalised reported cases to be population adjusted incidence rates our results are conditional on the use of the weighted interval score as an evaluation metrice.
+Whilst we normalised reported cases to be population-adjusted incidence rates our results are conditional on the use of the weighted interval score as an evaluation metric.
 As this proper scoring rule measures absolute performance this means that forecasts during periods of higher incidence are given more weight than forecasts from periods of low incidence.
 It also means that potentially overpredicting is penalised more than underpredicting as incidence rates are bounded at zero but relatively weakly bounded by populations at the upper bound (as incidence rates are typically only a small fraction of the overall population).
-This bias could explain the relatively poor performance of the surrogate model, compared to the ensemble, despite the surrogate model being comparably well calibrated.
-We consider alternative methods of forecast evaluation that would be robust to this potential source of bias but choice to stick closely to the methodology used by the ECDC forecast hub as these choices inform the development of submitted models and so are key to our findings.
+This bias could explain the relatively poor performance of the surrogate model, compared to the ensemble, despite the surrogate model being comparably well-calibrated.
+We consider alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick closely to the methodology used by the ECDC forecast hub as these choices inform the development of submitted models and so are key to our findings.
 
@@
 There are no other studies in the epidemiology literature which we are aware of that attempt to develop a forecasting model based on the observed behaviour of a multi-team, multi-model ensemble.
-There are also few studies that focus on delivering computationally feasible forecasting models in a reproducibile framework backed by an openly accesible compute platform. 
+Few studies focus on delivering computationally feasible forecasting models in a reproducible framework backed by an openly accessible compute platform. 
 However, the CDC, ECDC, and Germany/Poland forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles. 
-In general these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study).
+In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study).
 The CDC forecast hub identified the top 5 performing models and noted they made the following structural assumptions.
-However, they did not extensively compare and contrast these conclusions in order arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model.
-Similary, the Germany and Poland, and ECDC forecasting hubs were able to identify forecast models that performed comparably well but did not derive structural assumptions that led to this performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
+However, they did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model.
+Similarly, the Germany and Poland, and ECDC forecasting hubs were able to identify forecast models that performed comparably well but did not derive structural assumptions that led to this performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
 All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers.
-In general, during the study period all projects used the same median ensemble forecast of all submissions after having concluded it was a favourable choice. 
+In general, during the study period, all projects used the same median ensemble forecast of all submissions after having concluded it was a favourable choice. 
 The CDC forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensemble.
 Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases.
-That being said no forcast hub has switched to these for their operational forecast implying they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the communities suggested default ensemble option.
-Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation  metric.
-Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and also noted the bias towards underprediction observed in the ensemble forecasts and not the model based forecasts [@refNikosinGermany].
+That being said no forecast hub has switched to these for their operational forecast implying they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option.
+Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
+Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and also noted the bias towards underprediction observed in the ensemble forecasts and not the model-based forecasts [@refNikosinGermany].
 
@@
 
-Whilst we derived our surrogate model from a range of assumptions based on observing ensemble forecasts behaviour and the behaviour and structure of submitted models this was by no means as structure or complete a process as it could have been.
-In follow up work a more rigorous approach to this could be taken in order to further refine this set of assumptions. 
+Whilst we derived our surrogate model from a range of assumptions based on observing ensemble forecasts behaviour and the behaviour and structure of submitted models this was by no means as structured or complete a process as it could have been.
+In follow-up work, a more rigorous approach to this could be taken to further refine this set of assumptions. 
 The findings from our study may also be useful for informing these improved assumptions. 
-In addition, the model we derived based on our assumptions was likely not optimal both in terms of compute time and accuracy at reproducing ensemble like behaviour.
-Models with a more complex auto-correlation structure, and more refined approaches to localise trends should be explored in order to improve relative performance to ensemble forecasts.
-An example of a family of possible approaches are structural time series model which have many of the characteristcs implied by our assumptions for how forecast ensembles typically operate.
+In addition, the model we derived based on our assumptions was likely not optimal both in terms of compute time and accuracy at reproducing ensemble-like behaviour.
+Models with a more complex auto-correlation structure and more refined approaches to localise trends should be explored to improve relative performance to ensemble forecasts.
+An example of a family of possible approaches are structural time series model which has many of the characteristics implied by our assumptions for how forecast ensembles typically operate.
 As we identified that the tails of our predictive distributions were responsible for a large proportion of the difference in performance compared to the forecast ensemble.
-It may be the case that post processing of forecasts from our surrogate model to enhance their similarity to the forecast ensemble is a sensible strategy.
-This seems likely to improve out of sample performance but does not help with the secondary aim of understanding the implicit assumptions driving the performance of multi-model, multi-team infectious disease forecast ensembles.
+It may be the case that post-processing of forecasts from our surrogate model to enhance their similarity to the forecast ensemble is a sensible strategy.
+This seems likely to improve out-of-sample performance but does not help with the secondary aim of understanding the implicit assumptions driving the performance of multi-model, multi-team infectious disease forecast ensembles.
 As we have hypothesised that the use of absolute scoring measures is inappropriate and leads to performance characteristics that are unlikely to be favoured by forecast stakeholders more work should be done in this area.
-If new forecast ensemble methods are adopted as best practice by forecast hubs then follow up work attempting to create surrogate forecat models must also use these appproaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example the tendency to be poorly calibrated.
+If new forecast ensemble methods are adopted as best practice by forecast hubs then follow-up work attempting to create surrogate forecast models must also use these approaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example, the tendency to be poorly calibrated.
 Lastly, here we have only explored a surrogate for an ensemble for a single disease and a limited set of locations.
-Follow up work should explore whether this behaviour holds across diseases and locations.
-However, this work is limited by the lack of large-scale forecat ensembling projects despite them showing obvious promise to improve the forecasts available to stakeholders.
+Follow-up work should explore whether this behaviour holds across diseases and locations.
+However, this work is limited by the lack of large-scale forecast ensembling projects despite them showing obvious promise to improve the forecasts available to stakeholders.
 
@@
 
-We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders requirements.
+We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
 We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
-Our work maybe be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to methodology developers thinking about how to develop forecasts that perform similary to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
+Our work may be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to methodology developers thinking about how to develop forecasts that perform similary to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
 
@@
 
-JB and SF, and KS have recieved funding to run forecast hub projects.
+JB and SF, and KS have received funding to run forecast hub projects.
 
@@
 
-All data and code is available here:
+All data and code are available here:
 
@@
 
-SA, and SF were funded by a Wellcome senior fellowship to SF (), KS was funded by, and JB was funded by.
+SA and SF were funded by a Wellcome senior fellowship to SF (), KS was funded by, and JB was funded by.
 

### 2022-08-30 paper/paper.Rmd
review results

@@
 ```{r vis-forecasts, fig.cap = "a.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom.
-b.)Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).
+b.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).
 Incidence rates are shown on a log scale.
@@
 However, much of this outperformance was driven by a small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
-If we instead consider median relative performance (note not a proper score) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
+If we instead consider median relative performance (note not a proper scoring rule) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
 Looking across the distribution of relative scores by horizon we see little evidence that relative forecast quality decreased with time, though the tail of the distribution appears slightly heavier at the four-week time horizon (Figure \@ref(fig:assemble-eval) a).
-`r relative_wis_thresholds[, better_than]` of surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecat, and scored `r relative_wis_thresholds[, more_than_100_percent]` performed worse than 100% more than the comparable ensemble forecast.
+`r relative_wis_thresholds[, better_than]` of surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecat, and `r relative_wis_thresholds[, more_than_100_percent]` scored performed worse than 100% more than the comparable ensemble forecast.
 
@@
 
-Overall the surrogate model was well calibrated at the 30%, 60% and 90% credible intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger credible intervals.
+Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% credible intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger credible intervals.
 When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon and then became progressively less well calibrated as the forecast horizon increased with the worst calibration at the 90% credible interval (Figure \@ref(fig:plot-coverage) a).
@@
 Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty.
-In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared biased towards overprediction.
+In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction.
 Evaluating the relative performance of the surrogate model compared to the ECDC forecast hub ensemble we found that the mean performance was substantially worse but that this was largely driven by a relatively small subset of forecasts.
 The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
-Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2021 which may have been linked to increasing incidence rates across forecast locations linked to BA.2.
+Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to increasing incidence rates across forecast locations linked to BA.2.
 Unlike our earlier forecasting efforts, the effect of outlier forecasts was much reduced though the surrogate model was less likely to outperform the ensemble than our earlier forecasting model methodology [@refNikosinGermany].
-In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison the hub ensemble.
-As noted elsewhere, the forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger credible intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to overcover at large intervals.
+In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
+The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger credible intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to overcover at large intervals.
 Stratifying the relative weighted interval score by interval we found that larger intervals (i.e the tails of the forecast distribution) were responsible for the majority of the difference in performance.
@@
 
-Our study benefits from having been conducted using forecasts conducted in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the ECDC forecast hub).
-This means that we can be confident we have captured how our forecast methodology behaves in real-world usage and that others can evaluate this for themselves. 
+Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the ECDC forecast hub).
+This means that we can be confident we have captured how our forecast methodology behaves in real-world usage, at least within our study period. 
 The downside of this approach is that it was not possible to update the surrogate model over time in response to initial evaluation or to explore other parameterisations that might be more successful of which there are likely several.
@@
 It is likely that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
-However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance) then our approach makes sense and may be optimal for producing a "good" forecast.
-This is the view often used in the literature when describing forecast ensembles.
-Our focus on replicating the performance of the hub ensemble is also useful as it highlights some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, modelled decrease in growth rate as the forecast horizon increases
+However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature, then our approach makes sense and may be optimal for producing a "good" forecast.
+Our focus on replicating the performance of the hub ensemble is also useful as it highlights some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
 It also highlights some of the differences between our surrogate forecast model and the ensemble that may lead to new insights, such as the generally poor coverage of the ensemble that could not be explained by the assumptions we used in developing our surrogate methodology.
@@
 As this proper scoring rule measures absolute performance this means that forecasts during periods of higher incidence are given more weight than forecasts from periods of low incidence.
-It also means that potentially overpredicting is penalised more than underpredicting as incidence rates are bounded at zero but relatively weakly bounded by populations at the upper bound (as incidence rates are typically only a small fraction of the overall population).
+It also means that overprediction is penalised more than underprediction as incidence rates are bounded at zero but relatively weakly bounded by populations at the upper bound (as incidence rates are typically only a small fraction of the overall population).
 This bias could explain the relatively poor performance of the surrogate model, compared to the ensemble, despite the surrogate model being comparably well-calibrated.
-We consider alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick closely to the methodology used by the ECDC forecast hub as these choices inform the development of submitted models and so are key to our findings.
+We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick closely to the methodology used by the ECDC forecast hub as these choices inform the development of submitted models and so are key to our findings.
 
@@
 Few studies focus on delivering computationally feasible forecasting models in a reproducible framework backed by an openly accessible compute platform. 
-However, the CDC, ECDC, and Germany/Poland forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles. 
-In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study).
-The CDC forecast hub identified the top 5 performing models and noted they made the following structural assumptions.
-However, they did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model.
-Similarly, the Germany and Poland, and ECDC forecasting hubs were able to identify forecast models that performed comparably well but did not derive structural assumptions that led to this performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
-All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers.
-In general, during the study period, all projects used the same median ensemble forecast of all submissions after having concluded it was a favourable choice. 
-The CDC forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensemble.
+However, the CDC [@REF], ECDC [@REF], and Germany/Poland [@REF] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles.
+In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study) and comparatively little attention has been paid to coverage.
+The CDC forecast hub identified the top 5 performing models and noted the structural assumptions they made [@REF].
+However, they did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
+Similarly, the Germany and Poland, and ECDC forecasting hubs were able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
+All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@REF; @REF; @REF].
+In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions after having concluded it was a favourable choice [@REF]. 
+The CDC forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling [@REF].
 Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases.
-That being said no forecast hub has switched to these for their operational forecast implying they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option.
+That being said no forecast hub has switched to these alternative ensemble designs for their operational forecast implying they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.
 Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
-Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and also noted the bias towards underprediction observed in the ensemble forecasts and not the model-based forecasts [@refNikosinGermany].
+Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not the model-based forecasts [@refNikosinGermany].
 
@@
 Whilst we derived our surrogate model from a range of assumptions based on observing ensemble forecasts behaviour and the behaviour and structure of submitted models this was by no means as structured or complete a process as it could have been.
-In follow-up work, a more rigorous approach to this could be taken to further refine this set of assumptions. 
-The findings from our study may also be useful for informing these improved assumptions. 
+In follow-up work, a more rigorous approach to this could be taken to further refine this set of assumptions, in particular using the input of a wider pool of researchers. 
+The findings from our study may also be useful for informing this improved set of assumptions. 
 In addition, the model we derived based on our assumptions was likely not optimal both in terms of compute time and accuracy at reproducing ensemble-like behaviour.
-Models with a more complex auto-correlation structure and more refined approaches to localise trends should be explored to improve relative performance to ensemble forecasts.
-An example of a family of possible approaches are structural time series model which has many of the characteristics implied by our assumptions for how forecast ensembles typically operate.
-As we identified that the tails of our predictive distributions were responsible for a large proportion of the difference in performance compared to the forecast ensemble.
-It may be the case that post-processing of forecasts from our surrogate model to enhance their similarity to the forecast ensemble is a sensible strategy.
+Models with a more complex auto-correlation structure and more refined approaches to localised trends should be explored to improve relative performance to ensemble forecasts.
+An example of a family of possible approaches are structural time series models which have many of the characteristics implied by our assumptions for how forecast ensembles typically operate.
+As we identified that the tails of our predictive distributions were responsible for a large proportion of the difference in performance compared to the forecast ensemble it may be the case that post-processing of forecasts from our surrogate model would enhance their similarity to the forecast ensemble.
 This seems likely to improve out-of-sample performance but does not help with the secondary aim of understanding the implicit assumptions driving the performance of multi-model, multi-team infectious disease forecast ensembles.
 As we have hypothesised that the use of absolute scoring measures is inappropriate and leads to performance characteristics that are unlikely to be favoured by forecast stakeholders more work should be done in this area.
-If new forecast ensemble methods are adopted as best practice by forecast hubs then follow-up work attempting to create surrogate forecast models must also use these approaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example, the tendency to be poorly calibrated.
-Lastly, here we have only explored a surrogate for an ensemble for a single disease and a limited set of locations.
+If new forecast ensemble methods are adopted as best practice by forecast hubs then follow-up work attempting to create surrogate forecast models should also use these approaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example, the tendency to be poorly calibrated.
+Lastly, here we have only explored a surrogate for an ensemble for a single disease and a limited set of locations meaning our findings are difficult to generalise.
 Follow-up work should explore whether this behaviour holds across diseases and locations.
@@
 We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
-Our work may be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to methodology developers thinking about how to develop forecasts that perform similary to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
+Our work may be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to researchers thinking about how to develop forecasts that perform similary to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
 

### 2022-08-30 paper/paper.Rmd
spell check

@@
 We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
-Our work may be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to researchers thinking about how to develop forecasts that perform similary to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
+Our work may be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to researchers thinking about how to develop forecasts that perform similarly to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
 

### 2022-08-30 paper/paper.Rmd
add changes from @nikosbosse

@@
 The majority of the difference in performance was linked to a few forecasts from a period of high incidence.
-We note that despite performing worse than the hub ensemble the majority of the time our surrogate model was better calibrated, and less liable to underpredict.
+We note that despite performing worse than the hub ensemble the majority of the time our surrogate model showed better calibration, and was less liable to underpredict.
 The central part of the surrogate models forecast distribution performed more similarly to ensemble forecasts distribution whilst the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecast performance.
@@
 We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
-Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers developing forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
+Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers aiming to develop forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
 bibliography: library.bib
@@
 These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
-However, ensemble forecasts have a range of potential downsides with the most obvious being their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project, and that forecasts from these ensembles can be difficult to link to the underlying model assumption.
-In climate forecasting emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
-However, within epidemiological modelling, we often want to use our models for applications other than forecasting and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
+However, ensemble forecasts have a range of potential downsides.
+The most obvious is their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
+Forecasts from these ensembles can be hard to interpret as it is difficult to link them to the underlying model assumption.
+In climate forecasting, emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
+However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
 This means that non-parametric emulation is less suitable for use in epidemiological forecasting though for epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@REFLSHTMTBEMULATOR].
@@
 
-The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and ECDC forecast hub [@ECDC] in 2021.
-All of these collaborations, which the authors (particularly SF and JB have been involved with managing),  ensemble contributions from multiple independent teams using a similar ensembling approach (most recently an unweighted median ensemble [@REF]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting [@REF].
+The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and the ECDC forecast hub [@ECDC] in 2021.
+All of these collaborations, which the authors (particularly SF and JB) have been involved with managing,  ensembled contributions from multiple independent teams using a similar approach (most recently an unweighted median ensemble [@REF]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting [@REF].
 However, they have struggled to provide feedback to those contributing forecasts that would allow them to improve their forecast approaches, have found it difficult to improve the quality of their ensemble forecasts [@REF], and have struggled to maintain incentives for forecasters to continue to contribute which may impact the long-term quality of the ensembles they produce.
@@
 
-Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with aim of producing ensemble-like performance without sacrificing interpretability.
+Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with the aim of producing ensemble-like performance without sacrificing interpretability.
 The first simplification we make is to model only weekly data, rather than using daily data and then aggregating, this means that the impact of daily reporting artefacts is mitigated.
-It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model this leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
+It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model. 
+This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
 The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
@@
 This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence. 
-The only observation for which we do not make an adaption is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
+The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
 This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model (where variance is linked to the mean)  is a sensible choice though this is an area of current research.
@@
 We briefly outline some of the summary statistics it produces and their epidemiological relevance. 
-We then evaluate its real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying its empirical coverage.
+We then evaluate its real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
 We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed. 
@@
   r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
-  r_t &= \left(\mathcal{H}(r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
+  r_t &= \left(\mathcal{H}[a](r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
   \epsilon_t &= \mathcal{H}(t > 0) \beta \epsilon_{t-1} + \eta_t
@@
 
-Where $\mathcal{H}$ is the Heaviside step function and is defined such that it attains the value of 1 if the argument is true and the value of 0 otherwise.
+Where $\mathcal{H}$ is the Heaviside step function and is defined such that it attains the value of 1 if the argument is true and the value of 0 otherwise, and $\epsilon_t$ and $\eta_t$ are error terms.
 The following priors are used,
@@
 Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15.
-For real-time usage convergence was not assessed but during model development, the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergent transitions and exceedance of the maximum tree depth [@cmdstanr].
+For real-time usage, convergence was not assessed, but during model development, the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergent transitions and exceedance of the maximum tree depth [@cmdstanr].
 During development, posterior predictions were also visually compared to observed data.
@@
 Performance was most comparable in Slovakia but in the United Kingdom and Germany in particular the surrogate model performed visibly worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
-For both the ensemble and the surrogate performance visibly decreased as the forecast horizon increased though the relative difference in performance between models was relatively stable across horizons.
+For both the ensemble and the surrogate, performance visibly decreased as the forecast horizon increased though the relative difference in performance between models was relatively stable across horizons.
 In general, in the study period, the ensemble appeared to be liable to underpredict and the surrogate model appeared more likely to over-predict. 
@@
 Looking across the distribution of relative scores by horizon we see little evidence that relative forecast quality decreased with time, though the tail of the distribution appears slightly heavier at the four-week time horizon (Figure \@ref(fig:assemble-eval) a).
-`r relative_wis_thresholds[, better_than]` of surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecat, and `r relative_wis_thresholds[, more_than_100_percent]` scored performed worse than 100% more than the comparable ensemble forecast.
+`r relative_wis_thresholds[, better_than]` of surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecast, and `r relative_wis_thresholds[, more_than_100_percent]` scored performed worse than 100% more than the comparable ensemble forecast.
 
@@
 
-Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by larger intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) c).
+Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by differences in the outer intervals [b][c][d]and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) c).
 This is true across forecast horizons but the magnitude of the different increases.
@@
 # References
+
+[a]Isn't the indicator function with 1(.) just more common and potentially easier to read?
+[b]I would be very surprised if that were true. The reason is that the outer intervals get tiny weights in the WIS and performance usually is more strongly influenced by the inner intervals.
+[c]This hangs on if the output from score is already weighted or not which I was not clear on?
+[d]Which was my understanding of the implementation (weighting happens in score and summarise_score is just an averaging?)
+
+https://epiforecasts.io/scoringutils/reference/score_quantile.html
+
+Perhaps there is an implementation nuance? I agree I was surprised by the magnitude of this difference.

### 2022-08-30 paper/paper.Rmd
add more comments from @nikosbosse

@@
 
-Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly precise at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
+Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly precise at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) c).
 In contrast, the surrogate ensemble forecast was generally equally well calibrated at shorter horizons with a tendency to produce conservative forecasts.
@@
 
-Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by differences in the outer intervals [b][c][d]and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) c).
+Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by differences in the outer intervals [b][c][d][e]and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
 This is true across forecast horizons but the magnitude of the different increases.
@@
 
-Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty.
-In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction.
-Evaluating the relative performance of the surrogate model compared to the ECDC forecast hub ensemble we found that the mean performance was substantially worse but that this was largely driven by a relatively small subset of forecasts.
+Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty[f][g].
+In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction[h].
+Evaluating the relative performance of the surrogate model compared to the ECDC forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of forecasts.
 The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
@@
 Perhaps there is an implementation nuance? I agree I was surprised by the magnitude of this difference.
+[e]Aah I was just looking at Figure 3c here (I think it needs to be 3b).
+Oooh ok I think I now know what you mean.
+I would think that Figure 3b) is misleading.
+For every interval x, this is the score from the surrogate over the score for the ensemble for that interval x, right?
+Two reasons it's misleading:
+1) I think the values that you are dividing just become very small and therefore what you get is unstable, because the denominator gets smaller and smaller
+2) the outer values don't influence the average a lot, because you first add the values for the individual intervals up (to get one WIS for ensemble and surrogate) and then you divide these values.
+The plot suggests that you divide the values for all the quantiles and then average to get your relative score
+[f]WIS decompositions would be nice to look at. Although I guess can be left to future Sam (or never Sam)
+[g]Suggest add to slack as a suggestion for discussion there (close once done please).
+[h]As Johannes mentioned a while ago: It is a priori unclear what the "correct" distribution of overprediction, underprediction and dispersion is.
+So I think it is technically not correct to talk about a bias towards over / underprediction.
+Adding here as a comment because it might be relevant throughout.
+We could look at the bias value from scoringutils.

### 2022-08-31 paper/paper.Rmd
comments from @bisaloo

@@
 Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available.
-Typically, the ensemble forecast produced each week comprises at least 5 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
+Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
 Most submitted forecasts use available daily data, along with potentially other leading predictors, to produce daily forecasts that are then aggregated into weekly forecasts.
@@
 Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
-Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the ECDC forecasting hub. 
+Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the ECDC forecasting hub.[a][b] 
 In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
@@
 
-In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions and summarising its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action.
+In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions and summarising its implementation, with a focus on minimal resource use and reproducibility, as a GitHub Actions workflow.
 We briefly outline some of the summary statistics it produces and their epidemiological relevance. 
@@
   r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
-  r_t &= \left(\mathcal{H}[a](r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
+  r_t &= \left(\mathcal{H}[c](r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
   \epsilon_t &= \mathcal{H}(t > 0) \beta \epsilon_{t-1} + \eta_t
@@
 
-Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by differences in the outer intervals [b][c][d][e]and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
+Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by differences in the outer intervals [d][e][f][g][h]and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
 This is true across forecast horizons but the magnitude of the different increases.
@@
 
-Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty[f][g].
-In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction[h].
+Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty.
+In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction[i].
 Evaluating the relative performance of the surrogate model compared to the ECDC forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of forecasts.
@@
 Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases.
-That being said no forecast hub has switched to these alternative ensemble designs for their operational forecast implying they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.
+That being said no forecast hub has switched to these alternative ensemble designs for their operational forecast suggesting they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.
 Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
@@
 
-[a]Isn't the indicator function with 1(.) just more common and potentially easier to read?
-[b]I would be very surprised if that were true. The reason is that the outer intervals get tiny weights in the WIS and performance usually is more strongly influenced by the inner intervals.
-[c]This hangs on if the output from score is already weighted or not which I was not clear on?
-[d]Which was my understanding of the implementation (weighting happens in score and summarise_score is just an averaging?)
+[a]For this to be really informative, you would need to specify the CPU clock speed. There are huge disparities between the most recent 16-core machines and older ones.
+[b]Agree it could/should list the Azure type and clock speed but disagree it brings no information as the comparison is to a 2 core CI process available in 2022 (i.e when the paper is written).
+
+I think perhaps adding this detail could be a bit overwhelming for an introduction but worth adding to try it out
+[c]Isn't the indicator function with 1(.) just more common and potentially easier to read?
+[d]I would be very surprised if that were true. The reason is that the outer intervals get tiny weights in the WIS and performance usually is more strongly influenced by the inner intervals.
+[e]This hangs on if the output from score is already weighted or not which I was not clear on?
+[f]Which was my understanding of the implementation (weighting happens in score and summarise_score is just an averaging?)
 
@@
 Perhaps there is an implementation nuance? I agree I was surprised by the magnitude of this difference.
-[e]Aah I was just looking at Figure 3c here (I think it needs to be 3b).
+[g]Aah I was just looking at Figure 3c here (I think it needs to be 3b).
 Oooh ok I think I now know what you mean.
@@
 The plot suggests that you divide the values for all the quantiles and then average to get your relative score
-[f]WIS decompositions would be nice to look at. Although I guess can be left to future Sam (or never Sam)
-[g]Suggest add to slack as a suggestion for discussion there (close once done please).
-[h]As Johannes mentioned a while ago: It is a priori unclear what the "correct" distribution of overprediction, underprediction and dispersion is.
+[h]That sounds like a wording/interpretation issue vs a change the plot issue?
+
+Do you have suggestions for that wording or if suggesting the plot is changed what it should be changed to?
+[i]As Johannes mentioned a while ago: It is a priori unclear what the "correct" distribution of overprediction, underprediction and dispersion is.
 So I think it is technically not correct to talk about a bias towards over / underprediction.

### 2022-09-20 paper/paper.Rmd
update with comments from @sbfnk

--- 2022-08-30 paper/paper.Rmd
@@
 In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models.
-Here we use observations of the behaviour of the ECDC forecast ensemble, combined with our own forecasting experience,  to specify a series of model assumptions.
-We then develop a forecast model using these assumptions and compare its performance to the ECDC forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon.
+  Here we use observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience, to specify a series of model assumptions modelling incident reported cases of COVID-19.
+We then develop a forecast model using these assumptions and compare its performance to the European forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon.
   We found that this surrogate model performed visibly similarly to a subset of example locations though with increased uncertainty.
@@
 We note that despite performing worse than the hub ensemble the majority of the time our surrogate model showed better calibration, and was less liable to underpredict.
-The central part of the surrogate models forecast distribution performed more similarly to ensemble forecasts distribution whilst the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecast performance.
-We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
+The central part of the surrogate model’s forecast distribution performed more similarly to the ensemble forecast’s distribution whilst the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecast performance.
+  We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble.
+However more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
 We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
@@
 
-# Introduction {-}
+# Introduction {-}[a][b]
 
-Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade with significant success [@REF].
+Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@REF].
 These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
-However, ensemble forecasts have a range of potential downsides.
-The most obvious is their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
-Forecasts from these ensembles can be hard to interpret as it is difficult to link them to the underlying model assumption.
+
+At the same time, ensemble forecasts have a range of potential downsides.
+The most significant is their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
+Forecasts from these ensembles can be hard to interpret as it is difficult to link them to the underlying model assumptions.
+Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute, and difficulty improving the quality of the ensemble [@REF].
+Each of these may impact the long-term quality of the resulting ensemble forecasts.
+
+Here, we focus on addressing the issues of resource cost, and interpretability.
 In climate forecasting, emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
 However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
-This means that non-parametric emulation is less suitable for use in epidemiological forecasting though for epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@REFLSHTMTBEMULATOR].
-In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, limited computational resources, and an easily generalised implementation.
+This means that non-parametric emulation is less suitable for use in epidemiological forecasting (while noting that epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@REFLSHTMTBEMULATOR]).
+In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, and limited computational resources with an easily generalised implementation.
+
+In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, limited computational resources with an easily generalised implementation.
 
@@
 Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available.
-Typically, the ensemble forecast produced each week comprises at least 5 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
+Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
 Most submitted forecasts use available daily data, along with potentially other leading predictors, to produce daily forecasts that are then aggregated into weekly forecasts.
-Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts.
+Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts [@REF].
 In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
-Real-time evaluation has shown that each of these classes of models may perform well though in general semi-mechanistic models, and models that include human insights have outperformed other approaches during certain periods [@REF].
-We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@REF], and human insight-based forecasts [@REF] with our forecasts performing relatively well compared to the majority of those submitted but rarely outperforming significantly.
+Real-time evaluation has shown that each of these classes of models may perform well, although in general, semi-mechanistic models and models that include human insights have outperformed other approaches during certain periods [@REF].
+We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@REF], and human insight-based forecasts [@REF].
+
+Our forecasts have performed relatively well compared to the majority of other submitted forecasts, but rarely outperforming significantly.
 The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@REF].
-Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
+Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
 Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the ECDC forecasting hub. 
@@
 
-Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with the aim of producing ensemble-like performance without sacrificing interpretability.
-The first simplification we make is to model only weekly data, rather than using daily data and then aggregating, this means that the impact of daily reporting artefacts is mitigated.
+Based on the above insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with the aim of producing ensemble-like performance without sacrificing interpretability.
+The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
+This mitigates the impact of daily reporting artefacts.
 It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
 This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
+
 The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
@@
 As discussed in [@gostic] this leads to varying amounts of bias depending on the epidemic phase.
+
 The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@EpiNow2].
@@
 This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence.
+
 The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
@@
 
-In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions and summarising its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action.
+In this study, we evaluate an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions. 
+We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble. 
+First, we define the model and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub Actions workflow.  
 We briefly outline some of the summary statistics it produces and their epidemiological relevance.
-We then evaluate its real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
+We then evaluate its real-time performance in comparison to the European forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
 We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed.  
 Finally, we summarise our findings, discuss their implications, and highlight areas for potential study.
-We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models
+We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
 
@@
 
-  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
-  We included all locations covered by the ECDC forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
+We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
+We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
 Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@REF].
@@
 As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
-  This follows the common practice of the forecasting hub projects in treating the latest available data as the forecast target regardless of the number of potential data revisions.
+This follows the practice of the European forecasting hub project in treating the latest available data as the forecast target regardless of the number of potential data revisions [@REF].
 Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.
 
-  The ECDC forecast hub requests forecasts for one to four-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
+The European forecast hub requests forecasts for one to four-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
 No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
@@
 Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week).
-  The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team).
+The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team and timely submission).
 An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
-  Submitted forecasts and target observations are available from the ECDC forecast hub GitHub repository [@REF]. We provide code in the repository of this study to streamline access.
+Submitted forecasts and target observations are available from the European forecast hub GitHub repository [@REF]. We provide code in the repository of this study to streamline access.
 
@@
   r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
-  r_t &= \left(\mathcal{H}[a](r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
+  r_t &= \left(\mathcal{H}(r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
   \epsilon_t &= \mathcal{H}(t > 0) \beta \epsilon_{t-1} + \eta_t
@@
 
-## Forecast evaluation {-}
+## Forecast evaluation[c][d] {-}
 
@@
 We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
-Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the ECDC forecast hub  [@REF].
+Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub  [@REF].
 We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).
@@
 To maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
-We explore relative performance by forecast horizon, by month and horizon, and by location and horizon.
+We explore relative performance by forecast horizon, by month and horizon, and by location and horizo[e]n.
 
-In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% credible intervals and by quantile.
+In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile.[f]
 Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon.
@@
 The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
-`cmstanr` surfaces several settings that trade off between sampling speed and the robustness of the approach.
+`cmdstanr` surfaces several settings that trade off between sampling speed and the robustness of the approach.
 Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15.
@@
 
-To download and manipulate forecasts from the ECDC forecasting hub [@EuroHub] we use  the `data.table` (`1.14.2`) [@data.table] and `gh` (`1.3.0`) [@gh] R packages.
+To download and manipulate forecasts from the European forecasting hub [@EuroHub] we use  the `data.table` (`1.14.2`) [@data.table] and `gh` (`1.3.0`) [@gh] R packages.
 We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, visualise forecasts and summary measures, and summarise forecasts.
@@
 
-In our study period, incidence rates across ECDC nations and in the UK were primarily driven by the spread of novel subvariants of concern related to the Omicron variant and changes in population susceptibility.
-Many nations, such as the UK saw large BA.1 waves in January that then resulted in declining incidence rates through February.
+In our study period, incidence rates across European nations and in the UK were primarily driven by the spread of novel subvariants of concern related to the Omicron variant and changes in population susceptibility.
+Many nations, such as the UK, saw large BA.1 waves in January that then resulted in declining incidence rates through February (Figure \@ref(fig:vis-forecasts)).
 From late February, through to the end of May, most nations then saw another wave, typically with lower reported incidence rates, driven by BA.2.
@@
 Unlike earlier periods in the pandemic, our study period did not see the use of new non-pharmaceutical interventions (NPIs) in response to increasing COVID-19 incidence in most locations.
-In addition, in general, ascertainment rates likely reduced over time in most locations due to reductions in routine testing, and reductions in test availability.
+In addition, ascertainment rates likely reduced over time in most locations due to reductions in routine testing, and reductions in test availability.
 Whilst both the reduced use of NPIs and testing generally occurred across nations our study period also marked an increase in the heterogeneity of the response to the COVID-19 pandemic with nations changing policy at different times and to different degrees.
@@
 
-We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK making `r locations` unique locations).
-In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the ECDC baseline model).
+We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
+In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the European baseline model).
 Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
 Of the remaining models submitted `r length(local_models)` were submitted in only one location.
-Single location models were clustered in a few locations, particularly Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the ECDC forecasting hub project).
+Single location models were clustered in a few locations, particularly Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project).
 Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
@@
 In our example set of locations, the absolute performance of the ensemble and the surrogate model was similar in all locations at short forecast horizons though this varied by location.
-Performance was most comparable in Slovakia but in the United Kingdom and Germany in particular the surrogate model performed visibly worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
+Performance was most similar between the two in Slovakia but in the United Kingdom and Germany in particular the surrogate model performed visibly worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
 For both the ensemble and the surrogate, performance visibly decreased as the forecast horizon increased though the relative difference in performance between models was relatively stable across horizons.
@@
 Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
-In comparison to the ensemble model the surrogate model appeared less likely to place weight on infeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in February in Slovakia and Poland).
+In comparison to the ensemble model the surrogate model appeared less likely to place weight on unfeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in February in Slovakia and Poland).
 Italy and Germany, which had the highest number of submitted models of the example locations had little clear difference in performance between the ensemble and surrogate models compared to other example locations.
@@
 Incidence rates are shown on a log scale.
-30\\%, 60\\%, and 90\\% credible intervals are shown.
+30\\%, 60\\%, and 90\\% prediction intervals are shown.
 The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 18}
@@
 
-Evaluating the ensemble and surrogate models using the weighted interval score across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
-Indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by approximately 100% on average with the surrogate model performing worst at the one and four-week horizon.
-However, much of this outperformance was driven by a small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
+Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by approximately 100% on average with the surrogate model performing worst at the one and four-week horizon.
+Much of this outperformance was driven by a small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
 If we instead consider median relative performance (note not a proper scoring rule) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
-Looking across the distribution of relative scores by horizon we see little evidence that relative forecast quality decreased with time, though the tail of the distribution appears slightly heavier at the four-week time horizon (Figure \@ref(fig:assemble-eval) a).
-`r relative_wis_thresholds[, better_than]` of surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecast, and `r relative_wis_thresholds[, more_than_100_percent]` scored performed worse than 100% more than the comparable ensemble forecast.
+We see little evidence that relative forecast quality decreased with increasing forecast horizon, though the tail of the distribution appears slightly heavier at the four-week time horizon (Figure \@ref(fig:assemble-eval) a).
+`r relative_wis_thresholds[, better_than]` of individual surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecast, and `r relative_wis_thresholds[, more_than_100_percent]` had a more than 100% worse WIS than the comparable ensemble forecast.
 
-The surrogate model's relative performance varied over time with substantially worse performance from January to March compared to later in the year across all forecast horizons (Figure \@ref(fig:assemble-eval) b).
+The surrogate model's relative performance varied over time with substantially worse performance from January to March compared to later in the year across all forecast horizons based on changes in the relative score distribution and its summary statistics (Figure \@ref(fig:assemble-eval) b).
 The majority of the difference in performance appeared to be driven by a thicker right tail with this being a particular feature of forecasts at longer horizons.
-Notably, forecast performance in March had a bimodal distribution at the four-week horizon with a substantial fraction of surrogate forecasts outperforming the ensemble.
-This variation in performance may have been linked to the BA.2 wave which peaked in most locations during this period.
+Forecast performance in March had a bimodal distribution at the four-week horizon with a substantial fraction of surrogate forecasts outperforming the ensemble.
+This variation in performance may have been linked to the BA.2 wave which peaked in most locations during this period if the surrogate model was more likely to overpredict peak incidence than the ensemble forecast.
 
-There was also substantial variation across forecast locations with the surrogate performing well in some locations at some forecast horizons, for example, the four-week horizon in the United Kingdom, and badly in others, for example, the four-week forecast in Switzerland (Figure \@ref(fig:assemble-eval) c).
+There was also substantial variation across forecast locations with the surrogate performing relatively well in some locations at some forecast horizons, for example, the four-week horizon in the United Kingdom, and badly in others, for example, the four-week forecast in Switzerland (Figure \@ref(fig:assemble-eval) c).
 In general, across locations, as observed overall, relative forecast performance was comparable across horizons though with a heavier right tail at longer horizons.
@@
 
-### Forecast calibration {-}
+### Forecast calibration[g][h] {-}
 
-Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% credible intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger credible intervals.
-When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon and then became progressively less well calibrated as the forecast horizon increased with the worst calibration at the 90% credible interval (Figure \@ref(fig:plot-coverage) a).
-In comparison, the surrogate forecast was less well calibrated at the one-week forecast horizon with a tendency to have a larger empirical coverage than required at longer horizons the surrogate forecast transitioned to being well calibrated and then had lower coverage than optimal though to a much lower degree than the ensemble forecast.
-This was not the case for the 90% credible interval where the surrogate model stably covered slightly more than the expected interval indicating forecasts were somewhat overly uncertain for this interval.
+Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger prediction intervals (Figure \@ref(fig:plot-coverage) a).
+When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon and then became progressively less well calibrated as the forecast horizon increased with the worst calibration at the 90% prediction interval (Figure \@ref(fig:plot-coverage) a).
+In comparison, the surrogate forecast was less well calibrated than the ensemble forecast at the one-week forecast horizon with a tendency to have a larger empirical coverage than required  (Figure \@ref(fig:plot-coverage) a).
+At longer horizons and narrower prediction intervals the surrogate forecast transitioned to being well calibrated and then had lower coverage than optimal though greater coverage than the ensemble forecast.
+This was not the case for the 90% prediction interval where the surrogate model stably covered slightly more than the expected interval indicating forecasts were overly uncertain for this interval.
 
-Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly precise at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
-In contrast, the surrogate ensemble forecast was generally equally well calibrated at shorter horizons with a tendency to produce conservative forecasts.
+Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly certain at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) c).
+In contrast, the surrogate ensemble forecast was generally equally well calibrated at shorter horizons with a tendency to underpredict.
 At longer horizons, however, quantiles below the median were liable to have poor coverage suggesting a bias towards over-predicting.
 
-Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by differences in the outer intervals [b][c][d]and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) c).
-This is true across forecast horizons but the magnitude of the different increases.
-This implies that the surrogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecasts we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
+Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by differences in the outer intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
+This is true across forecast horizons but the magnitude of the difference increases.
+This implies that the surrogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecast population weighted incidence we see that the median for the ensemble forecast[i] was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
 However, when we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
@@
 
-```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate and ensemble forecasts at the 90\\%, 60\\%, and 30\\% credible intervals stratified by forecast horizon.
-Ideally, a well-calibrated forecast should have empirical coverage for a given credible interval that equals the range of that interval (i.e 90\\%).
+```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate and ensemble forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon.
+Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the range of that interval (i.e 90\\%).
 b.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast.
@@
 
-In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test-positive reported COVID-19 cases in ECDC nations and the UK.
-We first defined a set of assumptions for how the surrogate model should behave based on our observations of the ECDC forecast hub ensemble, our experience submitting forecasts to various forecast Hubs, and our observations of the performance of other models.
-We restricted ourselves to a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources.
+In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the European forecast hub multi-team ensemble for forecasting test-positive reported COVID-19 cases in European nations.
+We first defined a set of assumptions for how the surrogate model should behave based on our observations of the European forecast hub ensemble, our experience submitting forecasts to various forecast hubs, and our observations of the performance of other models.
+We aimed for a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources.
 We further provide a fully reproducible workflow for running and evaluating this model using GitHub actions facilitating others to do the same.
 
-Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty.
-In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction.
-Evaluating the relative performance of the surrogate model compared to the ECDC forecast hub ensemble we found that the mean performance was substantially worse but that this was largely driven by a relatively small subset of forecasts.
-The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
-Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to increasing incidence rates across forecast locations linked to BA.2.
-Unlike our earlier forecasting efforts, the effect of outlier forecasts was much reduced though the surrogate model was less likely to outperform the ensemble than our earlier forecasting model methodology [@refNikosinGermany].
-In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
-The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger credible intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to overcover at large intervals.
+Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble on the log scale though with greater uncertainty.[j][k]
+In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction[l].
+Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of forecasts.
+The median relative performance was only marginally[m] worse across forecast horizons with the distribution of relative scores having a heavy right tail.
+Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
+In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods used for real-time forecasting [@REFPREVIOUSFORECASTING] which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
+The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger prediction intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to be overly uncertain at large intervals.
 Stratifying the relative weighted interval score by interval we found that larger intervals (i.e the tails of the forecast distribution) were responsible for the majority of the difference in performance.
-Indeed, the median point performance of the forecasts was comparable though the surrogate model performed substantially worse on average due to its skewed performance distribution.
+Indeed, the median point performance of the surrogate and ensemble models was comparable though the surrogate model performed substantially worse on average due to its skewed performance distribution.
 
@@
 
-Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the ECDC forecast hub).
+Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the European forecast hub).
 This means that we can be confident we have captured how our forecast methodology behaves in real-world usage, at least within our study period. 
@@
 Importantly in this study, we focussed on replicating the forecast hub ensembles' observed behaviour rather than attempting to define an optimal forecast for forecast consumers.
-It is likely that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
-However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature, then our approach makes sense and may be optimal for producing a "good" forecast.
-Our focus on replicating the performance of the hub ensemble is also useful as it highlights some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
-It also highlights some of the differences between our surrogate forecast model and the ensemble that may lead to new insights, such as the generally poor coverage of the ensemble that could not be explained by the assumptions we used in developing our surrogate methodology.
+It is possible that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
+However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@REF], then our approach makes sense and may be optimal for producing a "good" forecast.
+Our focus on replicating the performance of the hub ensemble is also useful as the surrogate model may highlight some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
+It also highlights some of the differences between our surrogate forecast model and the ensemble that may lead to new insights in understanding the mechanisms leading to the ensembles behaviour, such as the generally poor coverage of the ensemble that could not be explained by the assumptions we used in developing our surrogate methodology.
 Whilst we normalised reported cases to be population-adjusted incidence rates our results are conditional on the use of the weighted interval score as an evaluation metric.
@@
 This bias could explain the relatively poor performance of the surrogate model, compared to the ensemble, despite the surrogate model being comparably well-calibrated.
-We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick closely to the methodology used by the ECDC forecast hub as these choices inform the development of submitted models and so are key to our findings.
+We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick relatively closely to the methodology used by the European forecast hub, aside from the use of population weighting in order to facilitate comparison between forecast locations, as these choices inform the development of submitted models and so are key to our findings.
 
@@
 Few studies focus on delivering computationally feasible forecasting models in a reproducible framework backed by an openly accessible compute platform. 
-However, the CDC [@REF], ECDC [@REF], and Germany/Poland [@REF] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles.
-In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study) and comparatively little attention has been paid to coverage.
-The CDC forecast hub identified the top 5 performing models and noted the structural assumptions they made [@REF].
+However, the US [@REF], European [@REF], and Germany/Poland [@REF] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles.
+In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study).
+The poor calibration of the forecast ensembles produced by median Hub ensembles as been noted repeatedly [REFEVAN; REFJOHANNES; REFKATH] but little progress has been made understanding the causes or suggesting alternatives forecasts for policy makers to use.
+The US forecast hub identified the top 5 performing models and noted the structural assumptions they made [@REF].
 However, they did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
-Similarly, the Germany and Poland, and ECDC forecasting hubs were able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
+Similarly, the Germany and Poland forecasting hubs was able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
 All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@REF; @REF; @REF].
 In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions after having concluded it was a favourable choice [@REF]. 
-The CDC forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling [@REF].
-Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases.
-That being said no forecast hub has switched to these alternative ensemble designs for their operational forecast implying they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.
+The CDC, and European, forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling [@REF].
+Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the CDC forecasting hub and did not outperform in the case of the European forecasting hub.
+No forecast hub has switched to these alternative ensemble designs for their operational forecast suggesting they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.
 Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
@@
 If new forecast ensemble methods are adopted as best practice by forecast hubs then follow-up work attempting to create surrogate forecast models should also use these approaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example, the tendency to be poorly calibrated.
-Lastly, here we have only explored a surrogate for an ensemble for a single disease and a limited set of locations meaning our findings are difficult to generalise.
-Follow-up work should explore whether this behaviour holds across diseases and locations.
+Lastly, here we have only explored a surrogate for an ensemble for a single disease, a limited set of locations, and a single target (incident cases), meaning our findings are difficult to generalise.
+Follow-up work should explore whether this behaviour holds across diseases, locations, and epidemiological targets where the behaviour of ensembles is notably different [@REF].
 However, this work is limited by the lack of large-scale forecast ensembling projects despite them showing obvious promise to improve the forecasts available to stakeholders.
@@
 SA conceived the study, developed the initial set of assumptions for the surrogate model, implemented the model into code, designed and conducted the forecast evaluation, and wrote the first draft of the manuscript.
-All other authors provided feedback on the manuscript and contributed to revisions.
+All other authors provided feedback on the manuscript and analyses and contributed to revisions.
 SF provided funding.
-HG reviewed the code and reproducibility of the analysis.
+HG reviewed the code and reproducibility of the analyses.
 
@@
 
-SA and SF were funded by a Wellcome senior fellowship to SF (), KS was funded by, and JB was funded by.
+SA and SF were funded by a Wellcome senior fellowship to SF (), KS and HG were funded by an ECDC grant to SF, and JB was funded by.
 
@@
 
-# References
+# References {-}
 
-[a]Isn't the indicator function with 1(.) just more common and potentially easier to read?
-[b]I would be very surprised if that were true. The reason is that the outer intervals get tiny weights in the WIS and performance usually is more strongly influenced by the inner intervals.
-[c]This hangs on if the output from score is already weighted or not which I was not clear on?
-[d]Which was my understanding of the implementation (weighting happens in score and summarise_score is just an averaging?)
+[a]Rejected my suggested changes moving around text - instead, put the suggestions here:
 
-https://epiforecasts.io/scoringutils/reference/score_quantile.html
+https://docs.google.com/document/d/1GbAsu1ZZsnm6Wl0WQA0g6LbpRucrLs4gRjpg97DUAA8/edit
+[b]Ah, only just saw this - will port my suggestions over.
+[c]shouldn't we exclude the surrogate model from the ensemble before comparing performance? Or else mention in the discussion that they aren't completely independent.
+[d]I think mention in the discussion as I would like to avoid recomputing.
+[e]From SF: Quantify the number of very poorly performing forecasts for the surrogate vs ensemble model.
+[f]From SF: what do we consider good / relatively good / poor performance in terms of calibration?
+[g]There are quite a few qualitative statements here (see individual comments below) - would be good to quantify with numbers, or otherwise state explicitly in the methods how we evaluate calibration.
+[h]Note I've closed out the other related comments in the following section just to be able to see the wood for the trees.
 
-Perhaps there is an implementation nuance? I agree I was surprised by the magnitude of this difference.
+Personally, I would rather look at a figure but if you feel it would help the reader happy to put the numbers in the text.
+[i]would be good to break this down a bit more, too, at least by horizon
+[j]I'm not sure we need this statement as it comes down largely to how you plot things - e.g. Fig. 1b would look quite different on a natural scale.
+[k]added log scale. I think we want to make some statement about what we see when we visualise the forecasts but it may not be this one.
+[l]As Johannes mentioned a while ago: It is a priori unclear what the "correct" distribution of overprediction, underprediction and dispersion is.
+So I think it is technically not correct to talk about a bias towards over / underprediction.
+Adding here as a comment because it might be relevant throughout.
+We could look at the bias value from scoringutils.
+[m]See above - a bit hard to assess the scale of this and where we delineat substantial vs. marginal

### 2022-09-21 paper/paper.Rmd

@@
 ---
-title: Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble
+title: Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble
 preprint: true
@@
 
-The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and the ECDC forecast hub [@ECDC] in 2021.
+The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and the European forecast hub [@ECDC] in 2021.
 All of these collaborations, which the authors (particularly SF and JB) have been involved with managing,  ensembled contributions from multiple independent teams using a similar approach (most recently an unweighted median ensemble [@REF]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting [@REF].
@@
 Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
-Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the ECDC forecasting hub. 
+Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the European forecasting hub. 
 In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
@@
 
-# Load ECDC forecasts. See data-raw/get-hub-forecasts.R for munging
+# Load European forecasts. See data-raw/get-hub-forecasts.R for munging
 # Merge with truth and rescale to incidence rate per 100,000

### 2022-09-21 paper/paper.Rmd
Close #4 and add detail on GitHub actions compute

@@
 Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
-Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the European forecasting hub. 
+Our previous methodology also requires significant computational resources running for an hour on a Azure D v5-series 16-core machine when producing forecasts for the European forecasting hub [@AzurePricing].
 In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
@@
 
-To ensure the reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with a built Docker image [@Boettiger:2015dw] (via GitHub Actions) is provided in the code repository.
-Weekly forecasts are also made using `renv` and GitHub actions to ensure they require limited compute and that our implementation is independent of local resources facilitating democratised access.
+To ensure the reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with a built Docker image [@Boettiger:2015dw] (via GitHub Actions [@GitHubActionsAbout]) is provided in the code repository.
+Weekly forecasts were made using `renv` and based GitHub Actions free tier as available in 2022 to ensure they require limited compute and that our implementation is independent of local resources facilitating democratised access.
+The free GitHub Actions runner we used for all forecasts was Ubuntu `20.04` based with 2 cores (x86_64), 7 GB of RAM, and 14 GB of SSD space.
 The code for this analysis can be found here: https://github.com/epiforecasts/simplified-forecaster-evaluation
@@
 If new forecast ensemble methods are adopted as best practice by forecast hubs then follow-up work attempting to create surrogate forecast models should also use these approaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example, the tendency to be poorly calibrated.
+In September 2022, GitHub announced support for hosted GitHub Action runners with additional compute power [@GitHubActionsLargeRunners].
+Whilst a paid feature this may allow more compute attensive models, with fewer potental performance trade-offs, to be easily democratised though only if funds are available to support the hosting costs.
+One potential research area is to explore forecasting methods that can be used with a range of compute resources though this would require extensive evaluation and documentation to make it clear to users what the trade-offs between compute usage and forecast performance are.
 Lastly, here we have only explored a surrogate for an ensemble for a single disease, a limited set of locations, and a single target (incident cases), meaning our findings are difficult to generalise.

### 2022-09-21 paper/paper.Rmd
close #5 by adding references

@@
 
-Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@REF].
+Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Sherratt2022-ue].
 These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
@@
 Forecasts from these ensembles can be hard to interpret as it is difficult to link them to the underlying model assumptions.
-Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute, and difficulty improving the quality of the ensemble [@REF].
+Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
 Each of these may impact the long-term quality of the resulting ensemble forecasts.
@@
 This means that non-parametric emulation is less suitable for use in epidemiological forecasting (while noting that epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@REFLSHTMTBEMULATOR]).
-In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, and limited computational resources with an easily generalised implementation.
 
@@
 
-The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and the European forecast hub [@ECDC] in 2021.
-All of these collaborations, which the authors (particularly SF and JB) have been involved with managing,  ensembled contributions from multiple independent teams using a similar approach (most recently an unweighted median ensemble [@REF]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting [@REF].
-However, they have struggled to provide feedback to those contributing forecasts that would allow them to improve their forecast approaches, have found it difficult to improve the quality of their ensemble forecasts [@REF], and have struggled to maintain incentives for forecasters to continue to contribute which may impact the long-term quality of the ensembles they produce.
+The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and the European forecast hub [@Sherratt2022-ue] in 2021.
+All of these collaborations, which the authors (particularly SF and JB) have been involved with managing,  ensembled contributions from multiple independent teams using a similar approach (most recently an unweighted median ensemble [@Sherratt2022-ue]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
+However, they have struggled to provide feedback to those contributing forecasts that would allow them to improve their forecast approaches, have found it difficult to improve the quality of their ensemble forecasts [@Sherratt2022-ue], and have struggled to maintain incentives for forecasters to continue to contribute which may impact the long-term quality of the ensembles they produce.
 
@@
 Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts [@REF].
-In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
+In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
 Real-time evaluation has shown that each of these classes of models may perform well, although in general, semi-mechanistic models and models that include human insights have outperformed other approaches during certain periods [@REF].
-We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@REF], and human insight-based forecasts [@REF].
+We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@Bosse2022-hm], and human insight-based forecasts [@Bosse2022-hm].
 
@@
 
-We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
+We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
 We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
@@
 As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
-This follows the practice of the European forecasting hub project in treating the latest available data as the forecast target regardless of the number of potential data revisions [@REF].
+This follows the practice of the European forecasting hub project in treating the latest available data as the forecast target regardless of the number of potential data revisions [@Sherratt2022-ue].
 Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.
@@
 An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
-Submitted forecasts and target observations are available from the European forecast hub GitHub repository [@REF]. We provide code in the repository of this study to streamline access.
+Submitted forecasts and target observations are available from the European forecast hub GitHub repository [@EuroHubGitHub]. We provide code in the repository of this study to streamline access.
 
@@
 We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
-Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub  [@REF].
+Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub  [@Sherratt2022-ue].
 We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).
@@
 It is possible that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
-However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@REF], then our approach makes sense and may be optimal for producing a "good" forecast.
+However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@REF; [@Sherratt2022-ue]], then our approach makes sense and may be optimal for producing a "good" forecast.
 Our focus on replicating the performance of the hub ensemble is also useful as the surrogate model may highlight some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
@@
 This bias could explain the relatively poor performance of the surrogate model, compared to the ensemble, despite the surrogate model being comparably well-calibrated.
-We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick relatively closely to the methodology used by the European forecast hub, aside from the use of population weighting in order to facilitate comparison between forecast locations, as these choices inform the development of submitted models and so are key to our findings.
+We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick relatively closely to the methodology used by the European forecast hub [@Sherratt2022-ue], aside from the use of population weighting in order to facilitate comparison between forecast locations, as these choices inform the development of submitted models and so are key to our findings.
 
@@
 Few studies focus on delivering computationally feasible forecasting models in a reproducible framework backed by an openly accessible compute platform. 
-However, the US [@REF], European [@REF], and Germany/Poland [@REF] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles.
+However, the US [@REF], European [@Sherratt2022-ue], and Germany/Poland [@REF] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles.
 In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study).
-The poor calibration of the forecast ensembles produced by median Hub ensembles as been noted repeatedly [REFEVAN; REFJOHANNES; REFKATH] but little progress has been made understanding the causes or suggesting alternatives forecasts for policy makers to use.
+The poor calibration of the forecast ensembles produced by median Hub ensembles as been noted repeatedly [REFEVAN; REFJOHANNES; @Sherratt2022-ue] but little progress has been made understanding the causes or suggesting alternatives forecasts for policy makers to use.
 The US forecast hub identified the top 5 performing models and noted the structural assumptions they made [@REF].
@@
 Similarly, the Germany and Poland forecasting hubs was able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
-All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@REF; @REF; @REF].
-In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions after having concluded it was a favourable choice [@REF]. 
-The CDC, and European, forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling [@REF].
+All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@REF; @REF; @Sherratt2022-ue].
+In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions. 
+The CDC, and European, forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling [@REF; @Sherratt2022-ue].
 Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the CDC forecasting hub and did not outperform in the case of the European forecasting hub.
@@
 Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
-Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not the model-based forecasts [@refNikosinGermany].
+Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not the model-based forecasts [@Bosse2022-hm].
 

### 2022-09-21 paper/paper.Rmd
flesh out references

@@
 
-The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and the European forecast hub [@Sherratt2022-ue] in 2021.
+The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@Bracher2021-wk] in late 2020, and the European forecast hub [@Sherratt2022-ue] in 2021.
 All of these collaborations, which the authors (particularly SF and JB) have been involved with managing,  ensembled contributions from multiple independent teams using a similar approach (most recently an unweighted median ensemble [@Sherratt2022-ue]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
@@
 Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available.
-Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
+Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
 Most submitted forecasts use available daily data, along with potentially other leading predictors, to produce daily forecasts that are then aggregated into weekly forecasts.
-Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts [@REF].
+Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts [@REF; @Bosse2022-hm].
 In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
-Real-time evaluation has shown that each of these classes of models may perform well, although in general, semi-mechanistic models and models that include human insights have outperformed other approaches during certain periods [@REF].
-We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@Bosse2022-hm], and human insight-based forecasts [@Bosse2022-hm].
+Real-time evaluation has shown that each of these classes of models may perform well, although in general, semi-mechanistic models and models that include human insights have outperformed other approaches during certain periods [@Bosse2022-hm].
+We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts [@Bosse2022-hm].
 
-Our forecasts have performed relatively well compared to the majority of other submitted forecasts, but rarely outperforming significantly.
-The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@REF].
+Our forecasts have performed relatively well compared to the majority of other submitted forecasts, but rarely outperforming significantly (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm]).
+The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
 Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
@@
 In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
-We found that our human insight-based forecasts significantly outperformed our model-based forecasts and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@REF].
+We found that our human insight-based forecasts significantly outperformed our model-based forecasts and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].
 Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts.
@@
 
-Based on the above insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with the aim of producing ensemble-like performance without sacrificing interpretability.
+Based on the above insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
 The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
@@
 However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
-As discussed in [@gostic] this leads to varying amounts of bias depending on the epidemic phase.
+As discussed in [@Gostic2020-vw] this leads to varying amounts of bias depending on the epidemic phase.
 
-The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@EpiNow2].
+The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
 This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
@@
 We briefly outline some of the summary statistics it produces and their epidemiological relevance.
-We then evaluate its real-time performance in comparison to the European forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
+We then evaluate its real-time performance in comparison to the European forecast hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
 We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed.  
@@
 Of the remaining models submitted `r length(local_models)` were submitted in only one location.
-Single location models were clustered in a few locations, particularly Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project).
+Single location models were clustered in a few locations, particularly Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
 Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
@@
 Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
-In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods used for real-time forecasting [@REFPREVIOUSFORECASTING] which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
+In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods used for real-time forecasting [@Bosse2022-hm; @EuroHub; @EpiNow2] which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
 The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger prediction intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to be overly uncertain at large intervals.
@@
 
-Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the European forecast hub).
+Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the European forecast hub [@Sherratt2022-ue]).
 This means that we can be confident we have captured how our forecast methodology behaves in real-world usage, at least within our study period. 
@@
 It is possible that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
-However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@REF; [@Sherratt2022-ue]], then our approach makes sense and may be optimal for producing a "good" forecast.
+However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@REF; @Sherratt2022-ue], then our approach makes sense and may be optimal for producing a "good" forecast.
 Our focus on replicating the performance of the hub ensemble is also useful as the surrogate model may highlight some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
@@
 Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
-Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not the model-based forecasts [@Bosse2022-hm].
+Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not in our the model-based forecasts [@Bosse2022-hm; @Bracher2021-wk].
 

### 2022-09-21 paper/paper.Rmd
add US hub reference and name change

@@
 
-Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Sherratt2022-ue].
+Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
 These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
@@
 
-The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@Bracher2021-wk] in late 2020, and the European forecast hub [@Sherratt2022-ue] in 2021.
+The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the US forecasting hub [@Cramer2022-sp] in early 2020, the Germany/Poland forecast hub [@Bracher2021-wk] in late 2020, and the European forecast hub [@Sherratt2022-ue] in 2021.
 All of these collaborations, which the authors (particularly SF and JB) have been involved with managing,  ensembled contributions from multiple independent teams using a similar approach (most recently an unweighted median ensemble [@Sherratt2022-ue]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
@@
 We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, visualise forecasts and summary measures, and summarise forecasts.
-Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils], and the `scoringRules` R package (1.0.1) [@REF].
+Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils], and the `scoringRules` R package (1.0.1) [@scoringRules].
 
@@
 It is possible that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
-However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@REF; @Sherratt2022-ue], then our approach makes sense and may be optimal for producing a "good" forecast.
+However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue], then our approach makes sense and may be optimal for producing a "good" forecast.
 Our focus on replicating the performance of the hub ensemble is also useful as the surrogate model may highlight some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
@@
 Few studies focus on delivering computationally feasible forecasting models in a reproducible framework backed by an openly accessible compute platform. 
-However, the US [@REF], European [@Sherratt2022-ue], and Germany/Poland [@REF] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles.
+However, the US [@Cramer2022-sp], European [@Sherratt2022-ue], and Germany/Poland [@Bracher2021-wk] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles.
 In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study).
-The poor calibration of the forecast ensembles produced by median Hub ensembles as been noted repeatedly [REFEVAN; REFJOHANNES; @Sherratt2022-ue] but little progress has been made understanding the causes or suggesting alternatives forecasts for policy makers to use.
-The US forecast hub identified the top 5 performing models and noted the structural assumptions they made [@REF].
+The poor calibration of the forecast ensembles produced by median Hub ensembles as been noted repeatedly [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue] but little progress has been made understanding the causes or suggesting alternatives forecasts for policy makers to use.
+The US forecast hub identified the top 5 performing models and noted the structural assumptions they made [@Cramer2022-sp].
 However, they did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
 Similarly, the Germany and Poland forecasting hubs was able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
-All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@REF; @REF; @Sherratt2022-ue].
+All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
 In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions. 
-The CDC, and European, forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling [@REF; @Sherratt2022-ue].
-Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the CDC forecasting hub and did not outperform in the case of the European forecasting hub.
+The US [@Cramer2022-sp], and European [@Sherratt2022-ue]], forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling.
+Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the UUS forecasting hub and did not outperform in the case of the European forecasting hub.
 No forecast hub has switched to these alternative ensemble designs for their operational forecast suggesting they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.

### 2022-09-21 paper/paper.Rmd
reesolve [c] from @sbfnk

@@
 
-## Forecast evaluation[c][d] {-}
+## Forecast evaluation[d] {-}
 
@@
 However, as our study has been conducted with a focus on reproducibility and openness our findings can be replicated or extended by others regardless of compute availability (due to our use of GitHub actions as a compute platform which is freely available to researchers).
+An additional downside to this approach is that the hub ensemble includes forecasts from our surrogate model, increasing the similarity between the two approaches.
+This is difficult to avoid without retrospectively re-calculating the ensemble using the same approach as taken by the hub which would reduce the independence of the hub ensemble as a source of truth to compare our forecasts against.
+Given the number of forecasts submitted in most locations, and the Europeans forecast hubs practice of not calculating an ensemble when fewer than 3 independent forecasts were available, the bias on our results caused by this limitation should be relatively small.
 Importantly in this study, we focussed on replicating the forecast hub ensembles' observed behaviour rather than attempting to define an optimal forecast for forecast consumers.
@@
 [b]Ah, only just saw this - will port my suggestions over.
-[c]shouldn't we exclude the surrogate model from the ensemble before comparing performance? Or else mention in the discussion that they aren't completely independent.
 [d]I think mention in the discussion as I would like to avoid recomputing.

### 2022-09-21 paper/paper.Rmd
simplify use of the Heavside step function closing #9

@@
   r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
-  r_t &= \left(\mathcal{H}(r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
+  r_t &= \left(\mathcal{H}(r_{t-1}) \xi_{+} + (1 - \mathcal{H}(r_{t-1})) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
   \epsilon_t &= \mathcal{H}(t > 0) \beta \epsilon_{t-1} + \eta_t
@@
 
-Where $\mathcal{H}$ is the Heaviside step function and is defined such that it attains the value of 1 if the argument is true and the value of 0 otherwise, and $\epsilon_t$ and $\eta_t$ are error terms.
+Where $\mathcal{H}(r_{t-1})$ is the Heaviside step function and is defined such that it attains the value of 1 if the argument is greater than 0 and 0 otherwise, and $\epsilon_t$ and $\eta_t$ are error terms.
 The following priors are used,

### 2022-09-21 paper/paper.Rmd
promote remaining comments to issues

@@
 
-# Introduction {-}[a][b]
+# Introduction {-}
 
@@
 To maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
-We explore relative performance by forecast horizon, by month and horizon, and by location and horizo[e]n.
+We explore relative performance by forecast horizon, by month and horizon, and by location and horizon.
 
-In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile.[f]
+In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile.
 Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon.
@@
 
-### Forecast calibration[g][h] {-}
+### Forecast calibration {-}
 
@@
 This is true across forecast horizons but the magnitude of the difference increases.
-This implies that the surrogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecast population weighted incidence we see that the median for the ensemble forecast[i] was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
+This implies that the surrogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecast population weighted incidence we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
 However, when we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
@@
 
-Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble on the log scale though with greater uncertainty.[j][k]
-In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction[l].
+Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble on the log scale though with greater uncertainty.
+In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction.
 Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of forecasts.
-The median relative performance was only marginally[m] worse across forecast horizons with the distribution of relative scores having a heavy right tail.
+The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
 Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
@@
 # References {-}
-
-[a]Rejected my suggested changes moving around text - instead, put the suggestions here:
-
-https://docs.google.com/document/d/1GbAsu1ZZsnm6Wl0WQA0g6LbpRucrLs4gRjpg97DUAA8/edit
-[b]Ah, only just saw this - will port my suggestions over.
-[d]I think mention in the discussion as I would like to avoid recomputing.
-[e]From SF: Quantify the number of very poorly performing forecasts for the surrogate vs ensemble model.
-[f]From SF: what do we consider good / relatively good / poor performance in terms of calibration?
-[g]There are quite a few qualitative statements here (see individual comments below) - would be good to quantify with numbers, or otherwise state explicitly in the methods how we evaluate calibration.
-[h]Note I've closed out the other related comments in the following section just to be able to see the wood for the trees.
-
-Personally, I would rather look at a figure but if you feel it would help the reader happy to put the numbers in the text.
-[i]would be good to break this down a bit more, too, at least by horizon
-[j]I'm not sure we need this statement as it comes down largely to how you plot things - e.g. Fig. 1b would look quite different on a natural scale.
-[k]added log scale. I think we want to make some statement about what we see when we visualise the forecasts but it may not be this one.
-[l]As Johannes mentioned a while ago: It is a priori unclear what the "correct" distribution of overprediction, underprediction and dispersion is.
-So I think it is technically not correct to talk about a bias towards over / underprediction.
-Adding here as a comment because it might be relevant throughout.
-We could look at the bias value from scoringutils.
-[m]See above - a bit hard to assess the scale of this and where we delineat substantial vs. marginal

### 2022-09-21 paper/paper.Rmd
moore references

@@
 No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
-Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@https://covid19forecasthub.eu/community.html].
+Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@EuroHub].
 Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week).
@@
   r_t &= \left(\mathcal{H}(r_{t-1}) \xi_{+} + (1 - \mathcal{H}(r_{t-1})) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
-  \epsilon_t &= \mathcal{H}(t > 0) \beta \epsilon_{t-1} + \eta_t
+  \epsilon_t &= \mathcal{H}(t) \beta \epsilon_{t-1} + \eta_t
 \end{align*}
 
-Where $\mathcal{H}(r_{t-1})$ is the Heaviside step function and is defined such that it attains the value of 1 if the argument is greater than 0 and 0 otherwise, and $\epsilon_t$ and $\eta_t$ are error terms.
+Where $\mathcal{H}(x)$ is the Heaviside step function and is defined such that it attains the value of 1 if the argument is greater than 0 and 0 otherwise, and $\epsilon_t$ and $\eta_t$ are error terms.
 The following priors are used,
@@
 
-## Forecast evaluation[d] {-}
+## Forecast evaluation {-}
 

### 2022-09-22 paper/paper.Rmd
update use of data to be locked at 1st Sept 2022 #1

@@
 We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
-Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@REF].
-We used the latest available observed data as of the 20th of August 2022.
+Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@JHU].
+We used the latest available observed data as of the 1st of September 2022 (commit `f6922c3e4bdcb055abcbba8e73472afacac4cf40` from [@EuroHubGitHub]).
+Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.
 As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
-This follows the practice of the European forecasting hub project in treating the latest available data as the forecast target regardless of the number of potential data revisions [@Sherratt2022-ue].
-Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.
+To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomolous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomolous data [@Sherratt2022-ue].
+We defined anomolous data based on the implementation used by the [@EuroHub] where a datapoint is considered anomolous if a future revision alters it by more than 5%.
 

### 2022-09-22 data-raw/get-population.R

@@ -3,8 +3,9 @@ library(data.table)
 library(here)
 
 # Download population data from the hub
+# as available on the 1st of September 2022
 population <- fread(
-  "https://raw.githubusercontent.com/covid19-forecast-hub-europe/covid19-forecast-hub-europe/main/data-locations/locations_eu.csv" # nolint
+  "https://raw.githubusercontent.com/covid19-forecast-hub-europe/covid19-forecast-hub-europe/f6922c3e4bdcb055abcbba8e73472afacac4cf40/data-locations/locations_eu.csv" # nolint
 )
 
 fwrite(population, here("data", "population.csv"))

### 2022-09-22 data-raw/get-truth.R

@@ -2,9 +2,9 @@
 library(data.table)
 library(here)
 
-# Download JHU data from the hub
+# Download JHU data from the hub - as available on the 1st of September 2022
 jhu <- fread(
-  "https://raw.githubusercontent.com/covid19-forecast-hub-europe/covid19-forecast-hub-europe/main/data-truth/JHU/truth_JHU-Incident%20Cases.csv" # nolint
+  "https://raw.githubusercontent.com/covid19-forecast-hub-europe/covid19-forecast-hub-europe/f6922c3e4bdcb055abcbba8e73472afacac4cf40/data-truth/JHU/truth_JHU-Incident%20Cases.csv" # nolint
 )
 
 # Format date

### 2022-09-22 data-raw/score-forecasts.R

@@ -25,6 +25,7 @@ anomalies <- fread(here("data", "anomalies.csv")) |>
   DT(, location_name := NULL)
 
 # Filter out anomalies from evaluation set for truth data
+# Based on: https://github.com/covid19-forecast-hub-europe/covid19-forecast-hub-europe/blob/39823425e9ea5d66c3dc0e7a55fa7ba5433d7df2/code/evaluation/load_and_score_models.r#L29 # nolint
 clean_forecasts_w_truth <- forecasts_with_truth |>
   DT(!anomalies, on = c("location", "target_end_date")) |>
   DT(, previous_end_date := forecast_date - 2) |>

### 2022-09-22 data-raw/get-hub-metadata.R
add anomalies to metadata

@@ -30,10 +30,44 @@ hub_forecasts <- hub_forecasts[grepl("inc case", target)]
 hub_forecasts <- hub_forecasts |>
   DT(!model %in% "EuroCOVIDhub-ensemble")
   
-# Streamline to metadata
-metadata <- hub_forecasts |>
+# Load anomalies
+anomalies <- fread(here("data", "anomalies.csv")) |>
+  DT(, location_name := NULL) |>
+  DT(, anomaly := TRUE)
+
+# Tag anomalies from evaluation set for truth data
+# Based on: https://github.com/covid19-forecast-hub-europe/covid19-forecast-hub-europe/blob/39823425e9ea5d66c3dc0e7a55fa7ba5433d7df2/code/evaluation/load_and_score_models.r#L29 # nolint
+metadata <- anomalies |>
+  copy() |>
+  DT(hub_forecasts, on = c("location", "target_end_date")) |>
+  DT(, previous_end_date := forecast_date - 2)
+
+metadata <- anomalies |>
+  DT(metadata,
+     on = c("target_end_date" = "previous_end_date", "location")
+  ) |>
+  DT(, c("target_end_date", "anomaly") := NULL) |>
+  setnames(
+    c("i.target_end_date", "i.anomaly"), c("target_end_date", "anomaly")
+  ) |>
+  DT(is.na(anomaly), anomaly := FALSE) |>
+  DT(, anomaly := as.logical(anomaly))
+
+# Streamline to unqiue forecast metadata
+# + count anomalies
+metadata <- metadata |>
+  DT(
+    target %in% c("1 wk ahead inc case", "2 wk ahead inc case",
+                  "3 wk ahead inc case", "4 wk ahead inc case")
+  ) |>
+  DT(, .(model, location, forecast_date, anomaly, target)) |>
+  unique() |>
+  DT(,
+   `:=`(anomaly = any(anomaly), n_anomaly = sum(anomaly)),
+   by = c("location", "forecast_date", "model")
+  ) |>
   DT(target %in% "1 wk ahead inc case") |>
-  DT(, .(model, location, forecast_date, target_end_date)) |>
+  DT(, .(model, location, forecast_date, anomaly, n_anomaly)) |>
   unique()
 
 # Save metadata

### 2022-09-22 data-raw/update.sh

@@ -4,9 +4,6 @@
 # Download hub forecasts of interest
 Rscript data-raw/get-hub-forecasts.R
 
-# Make hub metadata for study period
-Rscript data-raw/get-hub-metadata.R
-
 # Download truth data
 Rscript data-raw/get-truth.R
 
@@ -15,6 +12,9 @@ Rscript data-raw/get-population.R
 
 # Get anomalies data
 Rscript data-raw/get-anomalies.R
+
+# Make hub metadata for study period
+Rscript data-raw/get-hub-metadata.R
 
 # Calculate per forecast and quantile scores
 Rscript data-raw/score-forecasts.R

### 2022-09-22 paper/paper.Rmd
add anomalies to metadata
add discussion of anomalies, limitations, and further work

@@
 ```{r metadata}
+# Excluded forecast
+exclude_forecasts <- metadata |>
+  DT(, sum(n_anomaly))
+
+included_forecasts <- metadata |>
+  DT(, sum(n))
+
+per_excluded <- round(exclude_forecasts / included_forecasts * 100, 1)
+
+# Forecast dates + locations with anomalies
+exclusions <- metadata |>
+  DT(, .(forecast_date, location, anomaly)) |>
+  unique() |>
+  DT(, .(total = .N, exclude = sum(anomaly), include = .N - sum(anomaly))) |>
+  DT(, per := round(exclude / total * 100, 1))
+
 # Number of locations
@@
 forecast_dates <- metadata |>
-  DT(, uniqueN(target_end_date))
+  DT(, uniqueN(forecast_date))
 
@@
 
+`r nrow(anomalies)` observations, stratified by week and location, were defined to be anomalous within the study period by the European forecast hub [@EuroHub].
+Data anomalies were not randomly distributed with some locations being particlarly prone to data revisions including Lithuania (with `r anomalies[location == "LT", .N]` weeks with data anomalies), and Portugal (with `r anomalies[location == "PT", .N]` weeks with data anomalies).
+Anomalies  were also not evenly distributed over time with a higher proportion occurring earlier in the study period (potentially due to our choice to extract data from the 1st of September which effectively truncated anomalies).
+`r per_excluded`% of forecasts were excluded across all horizons due to anomalies in the observed data.
+Aggregated across horizons `r exclusions$per`% of forecasts included at least one week with anomalous data.
+
 ## Forecast evaluation {-}
@@
 Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not in our the model-based forecasts [@Bosse2022-hm; @Bracher2021-wk].
+Finally, our results are potentially sensitive to the definition used to define anonamous observations (which generally related to retrospective data revisions).
+Here we follow the practice of the European forecast hub [@EuroHub] of excluding forecasts for weeks with a data revision of more than 5% and forecasts made based on data that is subsequently revised by more than 5%.
 
@@
 One potential research area is to explore forecasting methods that can be used with a range of compute resources though this would require extensive evaluation and documentation to make it clear to users what the trade-offs between compute usage and forecast performance are.
+More work is needed to understand the best practice treatment of data revisions when evaluating forecasts and the potential sources for bias these may cause.
 Lastly, here we have only explored a surrogate for an ensemble for a single disease, a limited set of locations, and a single target (incident cases), meaning our findings are difficult to generalise.

### 2022-09-22 paper/paper.Rmd
fix bug caused by changes to metadata code

@@
 models_per_forecast <- metadata |>
-  summarise_forecasts_by(by = "target_end_date")
+  summarise_forecasts_by(by = "forecast_date")
 
@@
     summarise_forecasts_by(
-      var = "location", by = c("model", "target_end_date")
+      var = "location", by = c("model", "forecast_date")
     ) |>
@@
   summarise_forecasts_by(
-    var = "location", by = c("model", "target_end_date")
+    var = "location", by = c("model", "forecast_date")
   ) |>
@@
 models_per_loc_date <- metadata |>
-  summarise_forecasts_by(var = "model", by = c("target_end_date", "location"))
+  summarise_forecasts_by(var = "model", by = c("forecast_date", "location"))
 

### 2022-09-22 paper/paper.Rmd
fix bug caused by changes to metadata code

@@
 forecast_dates_by_models <- metadata |>
-  summarise_forecasts_by(var = "target_end_date", by = "model") |>
+  summarise_forecasts_by(var = "forecast_date", by = "model") |>
   DT(order(n)) |>

### 2022-09-23 paper/paper.Rmd
add references for emulation

@@
 Here, we focus on addressing the issues of resource cost, and interpretability.
-In climate forecasting, emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
+In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
 However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
-This means that non-parametric emulation is less suitable for use in epidemiological forecasting (while noting that epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@REFLSHTMTBEMULATOR]).
+This means that non-parametric emulation is less suitable for use in epidemiological forecasting (while noting that epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@Iskauskas2022-fz]).
 

### 2022-09-23 paper/paper.Rmd
fix anomaly bug from #1

@@
 We then develop a forecast model using these assumptions and compare its performance to the European forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon.
-  We found that this surrogate model performed visibly similarly to a subset of example locations though with increased uncertainty.
-Unsurprisingly, using the weighted interval score, we concluded that average performance was substantially worse but that median performance was more comparable.
+  We found that this surrogate model performed visibly relatively similarly to a subset of example locations though with increased uncertainty.
+Unsurprisingly, using the weighted interval score, we concluded that average performance was substantially worse relative to the ensemble forecast but that median performance was more comparable.
 The majority of the difference in performance was linked to a few forecasts from a period of high incidence.
@@
 Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
-These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
+These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts and being trusted by policy-makers.
 
@@
 Here, we focus on addressing the issues of resource cost, and interpretability.
-In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
+In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
 However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
@@
 forecast_dates <- metadata |>
-  DT(, uniqueN(forecast_date))
+  DT(, unique(forecast_date))
 
@@
 We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
-In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the European baseline model).
+In total `r nrow(metadata)` forecasts were made across all locations, with unique `r forecast_dates` forecast dates and `r models` independent forecast models (including the European hub baseline model).
 Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.

### 2022-09-23 data-raw/get-hub-metadata.R

@@ -2,6 +2,7 @@
 library(data.table)
 library(here)
 library(gh)
+library(lubridate)
 
 # Load functions
 source(here("R", "get-hub-forecasts.R"))
@@ -40,7 +41,11 @@ anomalies <- fread(here("data", "anomalies.csv")) |>
 metadata <- anomalies |>
   copy() |>
   DT(hub_forecasts, on = c("location", "target_end_date")) |>
-  DT(, previous_end_date := forecast_date - 2)
+  DT(,
+    previous_end_date := ceiling_date(
+      forecast_date - 4, week_start = 6, unit = "week"
+    )
+  )
 
 metadata <- anomalies |>
   DT(metadata,
@@ -60,14 +65,19 @@ metadata <- metadata |>
     target %in% c("1 wk ahead inc case", "2 wk ahead inc case",
                   "3 wk ahead inc case", "4 wk ahead inc case")
   ) |>
-  DT(, .(model, location, forecast_date, anomaly, target)) |>
+  DT(, .(model, location, forecast_date, target_end_date, anomaly, target)) |>
   unique() |>
   DT(,
-   `:=`(anomaly = any(anomaly), n = .N, n_anomaly = sum(anomaly)),
+   `:=`(
+      anomaly = any(anomaly), n = .N, n_anomaly = sum(anomaly),
+      target_end_date = min(target_end_date)
+    ),
    by = c("location", "forecast_date", "model")
   ) |>
   DT(target %in% "1 wk ahead inc case") |>
-  DT(, .(model, location, forecast_date, anomaly, n_anomaly, n)) |>
+  DT(,
+    .(model, location, forecast_date, target_end_date, anomaly, n_anomaly, n)
+  ) |>
   unique()
 
 # Save metadata

### 2022-09-23 paper/paper.Rmd
fix paper issues related to anomaly bug identified #1

@@
 exclusions <- metadata |>
-  DT(, .(forecast_date, location, anomaly)) |>
+  DT(, .(target_end_date, location, anomaly)) |>
   unique() |>
@@
 forecast_dates <- metadata |>
-  DT(, unique(forecast_date))
+  DT(, uniqueN(target_end_date))
 
@@
     summarise_forecasts_by(
-      var = "location", by = c("model", "forecast_date")
+      var = "location", by = c("model", "target_end_date")
     ) |>
@@
   summarise_forecasts_by(
-    var = "location", by = c("model", "forecast_date")
+    var = "location", by = c("model", "target_end_date")
   ) |>
@@
 models_per_loc_date <- metadata |>
-  summarise_forecasts_by(var = "model", by = c("forecast_date", "location"))
+  summarise_forecasts_by(var = "model", by = c("target_end_date", "location"))
 
@@
 forecast_dates_by_models <- metadata |>
-  summarise_forecasts_by(var = "forecast_date", by = "model") |>
+  summarise_forecasts_by(var = "target_end_date", by = "model") |>
   DT(order(n)) |>

### 2022-09-23 paper/paper.Rmd
close #1 by reviewing paper + other improvements

@@
 Unsurprisingly, using the weighted interval score, we concluded that average performance was substantially worse relative to the ensemble forecast but that median performance was more comparable.
-The majority of the difference in performance was linked to a few forecasts from a period of high incidence.
+The majority of the difference in performance was linked to a subset of forecasts with worse performance early in 2022.
 We note that despite performing worse than the hub ensemble the majority of the time our surrogate model showed better calibration, and was less liable to underpredict.
-The central part of the surrogate model’s forecast distribution performed more similarly to the ensemble forecast’s distribution whilst the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecast performance.
+The central part of the surrogate model’s forecast distribution performed more similarly to the ensemble forecast’s distribution but maintained a right skewed performance distribution.
   We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble.
-However more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
-We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
+However, more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
+We also conclude that our findings may be largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
 Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers aiming to develop forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
@@
 truth <- fread(here("data", "truth.csv"))
+
+# Load observations defined as anomalies
+anomalies <- fread(here("data", "anomalies.csv"))
 
@@
 coverage <- calc_coverage(scores)
+
+overall_coverage <- coverage |>
+ DT(, .(coverage = round(mean(coverage) * 100, 1)), by = c("model", "range")) |>
+ dcast(range ~ model, value.var = "coverage") |>
+ DT(, target := c(30, 60, 90)) |>
+ DT(, rel_ensemble := round(Ensemble / target * 100, 1) - 100) |>
+ DT(, rel_surrogate := round(Surrogate / target * 100, 1) - 100)
+
+overall_coverage_by_horizon <- coverage |>
+ DT(,
+    .(coverage = round(mean(coverage) * 100, 1)),
+    by = c("model", "range", "horizon")
+  ) |>
+ dcast(range + horizon ~ model, value.var = "coverage")
 ```
@@
 We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
-In total `r nrow(metadata)` forecasts were made across all locations, with unique `r forecast_dates` forecast dates and `r models` independent forecast models (including the European hub baseline model).
+In total `r nrow(metadata)` forecasts were made across all locations, with `r forecast_dates` unique forecast dates and `r models` independent forecast models (including the European hub baseline model).
 Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
@@
 `r nrow(anomalies)` observations, stratified by week and location, were defined to be anomalous within the study period by the European forecast hub [@EuroHub]. 
+Forecasts for these observations were excluded as were forecasts for forecast weeks where they were the latest available data.
 Data anomalies were not randomly distributed with some locations being particlarly prone to data revisions including Lithuania (with `r anomalies[location == "LT", .N]` weeks with data anomalies), and Portugal (with `r anomalies[location == "PT", .N]` weeks with data anomalies).
@@
 
-In our example set of locations, the absolute performance of the ensemble and the surrogate model was similar in all locations at short forecast horizons though this varied by location.
-Performance was most similar between the two in Slovakia but in the United Kingdom and Germany in particular the surrogate model performed visibly worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
+In our example set of locations, the absolute performance of the ensemble and the surrogate model was visually similar on the log scale in all locations at short forecast horizons though this varied by location.
+Performance was most similar in Slovakia but in the United Kingdom and Germany in particular the surrogate model performed visibly worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
 For both the ensemble and the surrogate, performance visibly decreased as the forecast horizon increased though the relative difference in performance between models was relatively stable across horizons.
@@
 
-Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by approximately 100% on average with the surrogate model performing worst at the one and four-week horizon.
-Much of this outperformance was driven by a small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
-If we instead consider median relative performance (note not a proper scoring rule) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
-We see little evidence that relative forecast quality decreased with increasing forecast horizon, though the tail of the distribution appears slightly heavier at the four-week time horizon (Figure \@ref(fig:assemble-eval) a).
+Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by at least 25% and that the relative performance of the surrogate model degraded as the forecast horizon increased.
+Much of this outperformance, especially at longer forecast horizons, was driven by a relatively small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a).
+If we instead consider median relative performance (note this is not a proper scoring rule) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
+This would suggest that an increasingly skewed score distribution as the forecast horizon increased is responsible for the increase in the mean relative score (Figure \@ref(fig:assemble-eval) a).
 `r relative_wis_thresholds[, better_than]` of individual surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecast, and `r relative_wis_thresholds[, more_than_100_percent]` had a more than 100% worse WIS than the comparable ensemble forecast.
@@
 The majority of the difference in performance appeared to be driven by a thicker right tail with this being a particular feature of forecasts at longer horizons.
-Forecast performance in March had a bimodal distribution at the four-week horizon with a substantial fraction of surrogate forecasts outperforming the ensemble.
+Forecast performance in March had a bimodal distribution at the four-week horizon with a substantial fraction of surrogate forecasts outperforming the ensemble and a substantial fraction substantially underperforming.
 This variation in performance may have been linked to the BA.2 wave which peaked in most locations during this period if the surrogate model was more likely to overpredict peak incidence than the ensemble forecast.
@@
 There was also substantial variation across forecast locations with the surrogate performing relatively well in some locations at some forecast horizons, for example, the four-week horizon in the United Kingdom, and badly in others, for example, the four-week forecast in Switzerland (Figure \@ref(fig:assemble-eval) c).
-In general, across locations, as observed overall, relative forecast performance was comparable across horizons though with a heavier right tail at longer horizons.
+In general, across locations, as observed overall, relative forecast performance degraded across horizons with a heavier right tail at longer horizons.
 Some locations showed less of this behaviour, for example, Spain, and in some, it was very dominant, for example, Switzerland.
@@
 
-Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger prediction intervals (Figure \@ref(fig:plot-coverage) a).
-When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon and then became progressively less well calibrated as the forecast horizon increased with the worst calibration at the 90% prediction interval (Figure \@ref(fig:plot-coverage) a).
+Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction interval, though with a tendency to over cover, with empirical coverage of `r overall_coverage[range %in% "30% interval", Surrogate]`%, `r overall_coverage[range %in% "60% interval", Surrogate]`%, `r overall_coverage[range %in% "90% interval", Surrogate]`% respectively.
+The ensemble model was less well calibrated, with a tendency to under cover with empirical coverage of `r overall_coverage[range %in% "30% interval", Ensemble]`%, `r overall_coverage[range %in% "60% interval", Ensemble]`%, `r overall_coverage[range %in% "90% interval", Ensemble]`% respectively (Figure \@ref(fig:plot-coverage) a).
+When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon, and then became progressively less well calibrated as the forecast horizon increased (Figure \@ref(fig:plot-coverage) a).
 In comparison, the surrogate forecast was less well calibrated than the ensemble forecast at the one-week forecast horizon with a tendency to have a larger empirical coverage than required  (Figure \@ref(fig:plot-coverage) a).
-At longer horizons and narrower prediction intervals the surrogate forecast transitioned to being well calibrated and then had lower coverage than optimal though greater coverage than the ensemble forecast.
-This was not the case for the 90% prediction interval where the surrogate model stably covered slightly more than the expected interval indicating forecasts were overly uncertain for this interval.
+At longer horizons and narrower prediction intervals the surrogate forecast became better calibrated though with a tendency to under cover.
+This was not the case for the 90% prediction interval where the surrogate model covered more than the expected interval, for all horizons, indicating forecasts were overly uncertain for this interval regardless of horizon.
 
 Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly certain at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) c).
-In contrast, the surrogate ensemble forecast was generally equally well calibrated at shorter horizons with a tendency to underpredict.
-At longer horizons, however, quantiles below the median were liable to have poor coverage suggesting a bias towards over-predicting.
+This behaviour became more prominant as forecast horizon increased.
+In contrast, the surrogat forecast was generally equally well calibrated at across horizons with a tendency to over cover for intervals above the median.
+At longer horizons, however, quantiles below the median became liable to have under coverage.
 
-Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by differences in the outer intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
+Breaking down the relative weighted interval score by forecast interval we observe that the surrogate modeel produces forecasts that differ most from the ensemble in the outer intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
 This is true across forecast horizons but the magnitude of the difference increases.
-This implies that the surrogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecast population weighted incidence we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
-However, when we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
-`r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) as noted elsewhere.
-This indicates a right skew in performance for the surrogate model point forecasts not observed for the ensemble forecast which matches that observed when evaluating the full predictive distributions of both forecasts.
+If we consider the absolute error of the median forecast population weighted incidence we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
+When we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
+`r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) which is comparable on the relative scale to the difference in performance observed using WIS.
+This implies a similarly skewed distribution for the relaive point forecasts from the surrogate model point forecasts compared to the ensemble forecast which again  matches that observed when evaluating the full predictive distributions of both forecasts.
 
@@
 In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction.
-Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of forecasts.
-The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
+Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of vey poorly performing forecasts.
+Median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
 Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
-In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods used for real-time forecasting [@Bosse2022-hm; @EuroHub; @EpiNow2] which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
-The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger prediction intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to be overly uncertain at large intervals.
-Stratifying the relative weighted interval score by interval we found that larger intervals (i.e the tails of the forecast distribution) were responsible for the majority of the difference in performance.
-Indeed, the median point performance of the surrogate and ensemble models was comparable though the surrogate model performed substantially worse on average due to its skewed performance distribution.
+In general, the relative performance of the surrogate model degraded as forecast horizons increas with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
+The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
+Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though relative performance of the surrogate model generally improved.
 
@@
 In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study).
-The poor calibration of the forecast ensembles produced by median Hub ensembles as been noted repeatedly [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue] but little progress has been made understanding the causes or suggesting alternatives forecasts for policy makers to use.
-The US forecast hub identified the top 5 performing models and noted the structural assumptions they made [@Cramer2022-sp].
-However, they did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
+The poor calibration of the forecast ensembles produced by median Hub ensembles has been noted repeatedly [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue] but little progress has been made understanding the causes or suggesting alternatives forecasts for policy makers to use.
+Progress understanding which structural model features lead to better infectious disease forecasts has been limited.
+The US forecast hub identified the top 5 performing models and noted the structural assumptions they made but couldn't directly link assumptions with performance [@Cramer2022-sp].
+They also did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
 Similarly, the Germany and Poland forecasting hubs was able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
@@
 We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
-We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
-Our work may be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to researchers thinking about how to develop forecasts that perform similarly to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
+We also conclude that our findings may belargely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
+Our work is useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to researchers thinking about how to develop forecasts that perform similarly to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
 

### 2022-09-26 paper/paper.Rmd
partial edits to reflect suggestions from #7

@@
 
-Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
-These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts and being trusted by policy-makers.
+Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Reich2022-jo].
+These approaches have been inspired by the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providng forecasts that are trusted by stakholders.
+The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic.
+Over 2020 and 2021, teams established COVID-19 Forecasting Hubs covering the US [@@Cramer2022-sp], Germany and Poland [@Bracher2021-wk], and Europe [@Sherratt2022-ue] (the latter two including authors of this study).
+These projects were motivated partially to meet an increased need from stakeholders for reliable forecast models, but also by the perceived low quality of many forecasts otherwise being produced and publicised.
+
+Whilst there is robust and consistent evidence that ensemble forecasts provide reliable and performant forecacasts across domains [@Reich2022-jo] they also have a range of downsides.
+The most significant is their considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
+Forecasts from these ensembles can also be hard to interpret as it is difficult to link them to the underlying model assumptions.
+Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
+Each of these may impact the long-term quality of the resulting ensemble forecasts and have implications for end-users.
+
+Here, we draw insights from ensemble forecasts produced and endorsed by the various COVID-19 forecast hubs, as well as our own forecasting work, to propose and evaluate a "surrogate" forecast model.
+This seeks to emulate the observed behaviour of recent COVID-19 ensembles forecasting reported cases, using a minimal set of easily communicated assumptions, and limited computational resources with an easily generalised implementation.
+The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
+In addition to this, we seek to provide a simple, and robust, forecasting system that can be easily resused by others, perhaps where a large scale forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
+
+In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
+However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
+This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz] utility, it may be less suitable for use in replicating epidemiological ensemble forecasts.
+
+In each of the COVID-19 Forecast Hubs, multiple independent teams contribute forecasts of reported cases, along with potentially other metrics, for a given set of locations, which the Hub collates and uses to produce an ensemble forecast.
+Each Hub uses a similar approach focusing on forecasting by epidemiological week (i.e.
+Sunday to Saturday) over a time horizon of a few weeks horizon.
+In general, 4 main classes of forecast models are submitted.
+This includes statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations such as individual-based models [@REF], semi-mechanistic approaches that blend statistical and mechanistic approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
+Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce daily forecasts.
+Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation.
+
+We have contributed a range of forecasts to these collaborations involving statistical and semi-mechanistic models [@REF], and human insight-based forecasts [@REF].
+With our semi-mechanistic forecast modelling, we have focussed on trying to model the underlying delays from infection through to symptom onset, and test positivity using non-exponential probability distributions whilst also modelling the complexity of daily, within the week, reporting periodicity [@REF].
+With these models, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data.
+Our model-based Bayesian forecast methodology requires significant computational resources, running for approximately aan hour on a 16-core machine when producing forecasts for the European forecasting hub.
+In contrast, our human insight models implicitly included unobserved interventions or other changes in the epidemiology and reporting process.
+The human insight models were…[@REF].
+Resources for the human insight models required an initial investment in building interactive forecasting infrastructure, followed by ongoing weekly participation from multiple independent forecasters.
+
+Our forecast submissions have not clearly over- or under- performed other forecasts submitted to the forecasting hubs.
+A clear weakness of our previous work has been a lack of robustness to large outliers in reporting and to changes in reporting patterns which has driven extremely poor performance at times.
+We found that our human insight-based forecasts significantly outperformed our model-based forecasts on average and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@REF].
+
+Considering multi-model ensembles, the respective Hubs produce an ensemble forecast each week comprising at least 3 independent forecasts [@REF].
+To create the ensemble, as of September 2022, the US and European Hubs both used an unweighted median ensemble [@REF].
+Evaluations of each of the COVID-19 Forecast Hubs have shown that these ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting [@REF].
+Unlike our forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts.
+Similarly to our human insight forecasts, the ensembles also demonstrated some ability to forecast future changes in trends that were not present in the observed data.
+This could indicate the inclusion of either human insight models, or assumptions about future interventions.
+In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence.
+On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
+Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under-, versus over-, predicting.
+There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models.
+
+Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
+The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
+This mitigates the impact of daily reporting artefacts.
+It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
+This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
+
+The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
+This reduces computational requirements and removes the need for external information on the delay from infection to report which can potentially be misspecified leading to potentially reduced forecast performance.
+However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
+As discussed in [@Gostic2020-vw] this leads to varying amounts of bias depending on the epidemic phase.
+
+The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
+This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the complexity, and computational overhead of the model.
+To model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
+This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence.
+
+The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
+This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model (where variance is linked to the mean) is a sensible choice.
+We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].
+
+In this study, we evaluate an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions.
+We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble.
+First, we define the model and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub Actions workflow.
+We briefly outline some of the summary statistics it produces and their epidemiological relevance.
+We then evaluate its real-time performance in comparison to the European forecast hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
+We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed.
+Finally, we summarise our findings, discuss their implications, and highlight areas where more work is needed.
+We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
+
+## Old
 

### 2022-09-26 paper/paper.Rmd
close #7 by adding in changes and editting

--- 2022-09-23 paper/paper.Rmd
@@
 
-Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
-These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts and being trusted by policy-makers.
+Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Reich2022-jo].
+These approaches have been inspired by the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providng forecasts that are trusted by stakholders.
 
-At the same time, ensemble forecasts have a range of potential downsides.
-The most significant is their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
-Forecasts from these ensembles can be hard to interpret as it is difficult to link them to the underlying model assumptions.
+Whilst there is robust and consistent evidence that ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
+The most significant is their considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
+Forecasts from these ensembles can also be hard to interpret as it is difficult to link them to the underlying model assumptions.
 Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
-Each of these may impact the long-term quality of the resulting ensemble forecasts.
+Each of these may impact the long-term quality of the resulting ensemble forecasts and have implications for end-users.
 
-Here, we focus on addressing the issues of resource cost, and interpretability.
+The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic due to a pressing need for reliable forecasts and a perception that many publicly available forecasts were low quality.
+Over 2020 and 2021, teams established COVID-19 Forecasting Hubs covering the US [@Cramer2022-sp], Germany and Poland [@Bracher2021-wk], and Europe [@Sherratt2022-ue] (the latter two including authors of this study).
+All of these collaborations, which the author have been involved with managing, ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
+However, less progress has been made mitigating the downsides of these projects and improving access to the high quality and robust forecasts they seek to generate.
+
+In this work, we draw insights from ensemble forecasts produced and endorsed by these forecast hubs, as well as our own forecasting work, to propose and evaluate a "surrogate" forecast model.
+This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justified assumptions, and limited computational resources with an easily generalised implementation.
+The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
+In addition to this, we seek to provide a simple, and robust, forecasting system that can be easily reused by others, perhaps where a large scale forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
+
 In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
-However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
-This means that non-parametric emulation is less suitable for use in epidemiological forecasting (while noting that epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@Iskauskas2022-fz]).
+However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
+This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz] utility, it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.
 
-In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, limited computational resources with an easily generalised implementation.
+In order to understand the behaviour of the forecast hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
+These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their own forecast models and then use them to produce weekly forecast that they then submit to the central hub team.
+Each hub targets a range of metrics including test positive reported cases, repoorted deaths, and hospitalisations, has a specific geographic focus, and asks for weekly forecasts (using epidemiological weeks i.e.
+Sunday to Saturday) over a time horizon of a few weeks.
+Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
+Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
+Here, we focus on reported cases and primarily on the European forecast hub but our observations hold, in our view, across COVID-19 forecast hubs and to a less degree targets.
+In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
+Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].
 
-The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the US forecasting hub [@Cramer2022-sp] in early 2020, the Germany/Poland forecast hub [@Bracher2021-wk] in late 2020, and the European forecast hub [@Sherratt2022-ue] in 2021.
-All of these collaborations, which the authors (particularly SF and JB) have been involved with managing,  ensembled contributions from multiple independent teams using a similar approach (most recently an unweighted median ensemble [@Sherratt2022-ue]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
-However, they have struggled to provide feedback to those contributing forecasts that would allow them to improve their forecast approaches, have found it difficult to improve the quality of their ensemble forecasts [@Sherratt2022-ue], and have struggled to maintain incentives for forecasters to continue to contribute which may impact the long-term quality of the ensembles they produce.
-
-These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a one to four-week time horizon.
-Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available.
-Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
-Most submitted forecasts use available daily data, along with potentially other leading predictors, to produce daily forecasts that are then aggregated into weekly forecasts.
-Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts [@REF; @Bosse2022-hm].
-In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
-Real-time evaluation has shown that each of these classes of models may perform well, although in general, semi-mechanistic models and models that include human insights have outperformed other approaches during certain periods [@Bosse2022-hm].
-We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts [@Bosse2022-hm].
-
-Our forecasts have performed relatively well compared to the majority of other submitted forecasts, but rarely outperforming significantly (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm]).
+We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts.
+Our forecast submissions have not clearly over- or under- performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
 The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
-Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
+Based on our observations our forecasts have generally captured the current trend relatively well, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
 Our previous methodology also requires significant computational resources running for an hour on a Azure D v5-series 16-core machine when producing forecasts for the European forecasting hub [@AzurePricing].
 In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
-We found that our human insight-based forecasts significantly outperformed our model-based forecasts and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].
+We found that our human insight-based forecasts significantly outperformed our model-based forecasts on average and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].
+
 Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts.
@@
 Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting.
-There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models.
+There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models
 
-Based on the above insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
+Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
 The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
@@
 The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
-This reduces computational requirements and removes the need for external information on the delay from infection to report which can potentially be misspecified leading to potentially reduced forecast performance.
+This reduces removes the need for, potentially misspecified, external information on the delay from infection to report, and reduces computational requirements due to a reduction in model complexity. 
 However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
@@
 The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
-This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model (where variance is linked to the mean)  is a sensible choice though this is an area of current research.
-We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations.
+This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model, where variance is linked to the mean, is a sensible choic. 
+We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].
 
@@
 We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed.
-Finally, we summarise our findings, discuss their implications, and highlight areas for potential study.
-We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
+Finally, we summarise our findings, discuss their implications, and highlight areas where more work is needed.
+We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
 

### 2022-09-27 paper/paper.Rmd
make forecast visualisation more naunced #14

@@
 
-```{r vis-forecasts, fig.cap = "a.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom.
-b.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks).
-Incidence rates are shown on a log scale.
+```{r vis-forecasts, fig.cap = a.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one and four weeks).
 30\\%, 60\\%, and 90\\% prediction intervals are shown.
-The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 18}
+The black line and points are the notified cases as of the date of data extraction rather than those available at the time.
+b.) A replicate of a.) but with incidence rates on the log scale.
+c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "95%", fig.height = 12, fig.width = 18}
 p <- (
-  plot_wis(wis, locs) +
-   theme(strip.text.x = element_blank()) |
-  plot_forecasts(forecasts, locs, ranges) +
-    guides(fill = guide_none(), fill_ramp = guide_none(), col = guide_none())
+  plot_forecast_custom(forecasts, log = FALSE) |
+  plot_forecast_custom(forecasts, log = TRUE) +
+  labs(y = "") |
+  plot_wis(wis, locs)
 ) +
-  plot_layout(widths = c(1, 4), guides = "collect") +
+  plot_layout(widths = c(2, 2, 1), guides = "collect") +
   plot_annotation(tag_levels = "a") &

### 2022-09-27 paper/paper.Rmd
fix caption

@@
 
-```{r vis-forecasts, fig.cap = a.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one and four weeks).
+```{r vis-forecasts, fig.cap = "a.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one and four weeks).
 30\\%, 60\\%, and 90\\% prediction intervals are shown.

### 2022-09-27 data-raw/score-forecasts.R
add forecasts ready for scoring

@@ -41,6 +41,9 @@ clean_forecasts_w_truth <- forecasts_with_truth |>
 # Check forecasts to be scored
 check_forecasts(clean_forecasts_w_truth)
 
+# Save forecasts to be scored
+fwrite(clean_forecasts_w_truth, here("data", "forecasts_ready_for_scoring.csv"))
+
 # Score forecasts
 scores <- clean_forecasts_w_truth |>
   score()

### 2022-09-27 paper/paper.Rmd
add forecasts ready for scoring
move absolute error discussion and add additional across horizon ae figure #17

@@
 library(ggridges)
+library(ggh4x)
 library(patchwork)
@@
   rename_models()
+
+# Load forecasts processed ready for scoring
+forecasts_ready_for_scoring <- fread(
+  here("data", "forecasts_ready_for_scoring.csv")
+)
 
@@
   calc_relative_score(cols = c("horizon")) |>
-  DT(, ris := round(relative_interval_score, 2))
+  DT(, ris := round(relative_interval_score, 2)) |>
+  DT(, interval_score := NULL)
+setnames(
+  relative_wis_by_horizon, c("relative_interval_score"),
+  c("Weighted interval score")
+)
+
+relative_ae_by_horizon <- scores |>
+  summarise_scores(by = c("horizon", "model")) |>
+  DT(, interval_score := ae_median) |>
+  calc_relative_score(cols = c("horizon")) |>
+  DT(, ris := round(relative_interval_score, 2)) |>
+  DT(, interval_score := NULL)
+setnames(
+  relative_ae_by_horizon, c("relative_interval_score", "ris"),
+  c("Absolute error of the median", "raem")
+)
+
+relative_summary <- relative_wis_by_horizon |>
+  merge(
+    relative_ae_by_horizon, by = c("horizon", "model")
+  ) |>
+  melt(
+    measure.vars = c("Weighted interval score", "Absolute error of the median")
+  )
 
@@
 
-Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by at least 25% and that the relative performance of the surrogate model degraded as the forecast horizon increased.
+Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by at least 25% and that the relative performance of the surrogate model degraded as the forecast horizon increased  (Figure \@ref(fig:assemble-eval) c).
 Much of this outperformance, especially at longer forecast horizons, was driven by a relatively small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a).
@@
 Some locations showed less of this behaviour, for example, Spain, and in some, it was very dominant, for example, Switzerland.
+
+If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
+`r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) which is comparable on the relative scale to the difference in performance observed using WIS.
+As when using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
+`r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`).
 
@@
   ) |
-  plot_rwis_location
+  (
+    plot_relative_summary(relative_summary) +
+    plot_rwis_location +
+    plot_layout(heights = c(1, 7), guides = "collect")
+  )
 ) +
@@
 This is true across forecast horizons but the magnitude of the difference increases.
-If we consider the absolute error of the median forecast population weighted incidence we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
-When we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
-`r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) which is comparable on the relative scale to the difference in performance observed using WIS.
-This implies a similarly skewed distribution for the relaive point forecasts from the surrogate model point forecasts compared to the ensemble forecast which again  matches that observed when evaluating the full predictive distributions of both forecasts.
 

### 2022-09-27 R/evaluation-plots.R

@@ -100,3 +100,19 @@ plot_relative_wis <- function(relative_wis, alpha = 0.8,
     theme_scoringutils() +
     labs(x = "Relative weighted interval score")
 }
+
+plot_relative_summary <- function(relative_summary) {
+  relative_summary |>
+    ggplot() +
+    aes(x = horizon, y = value, col = variable) +
+    geom_hline(yintercept = 1, linetype = 2) +
+    geom_point() +
+    geom_line() +
+    scale_color_brewer(palette = "Dark2") +
+    theme_scoringutils() +
+    theme(legend.position = "bottom") +
+    labs(
+      y = "Relative score", x = "Forecast horizon (weeks)", col = "Metric"
+    ) +
+    lims(y = c(0, NA))
+}

### 2022-09-27 paper/paper.Rmd
improve handling of forecast visualisation and close #14

@@
   Here we use observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience, to specify a series of model assumptions modelling incident reported cases of COVID-19.
-We then develop a forecast model using these assumptions and compare its performance to the European forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon.
+We then develop a forecast model using these assumptions and compare its performance to the European forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon using the weighted interval score.
   We found that this surrogate model performed visibly relatively similarly to a subset of example locations though with increased uncertainty.
-Unsurprisingly, using the weighted interval score, we concluded that average performance was substantially worse relative to the ensemble forecast but that median performance was more comparable.
+Unsurprisingly, we concluded that average performance was substantially worse relative to the ensemble forecast but that median performance was more comparable.
 The majority of the difference in performance was linked to a subset of forecasts with worse performance early in 2022.
@@
 
-We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
+We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS), and the absolute error (AE) of the median forecast.
 Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub [@Sherratt2022-ue].
@@
 
-To understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
+To understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS and rAE) by dividing the WIS/AE for the surrogate model by the WIS/AE of the ensemble model for all locations and forecast horizons.
 To maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
@@
 In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile.
-Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon.
+Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon, to assess the relative difference in performance across the predictive distribution.
 
@@
 
-In our example set of locations, the absolute performance of the ensemble and the surrogate model was visually similar on the log scale in all locations at short forecast horizons though this varied by location.
-Performance was most similar in Slovakia but in the United Kingdom and Germany in particular the surrogate model performed visibly worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
-For both the ensemble and the surrogate, performance visibly decreased as the forecast horizon increased though the relative difference in performance between models was relatively stable across horizons.
-In general, in the study period, the ensemble appeared to be liable to underpredict and the surrogate model appeared more likely to over-predict. 
+In our example set of locations, the absolute performance of the ensemble and the surrogate model was visually similar on the log scale in all locations at short forecast horizons though this varied by location (Figure \@ref(fig:vis-forecasts) b).
+On the natural scale the difference in performance was more marked, specially for periods of high incidence and at longer horizons (Figure \@ref(fig:vis-forecasts) a).
+Performance was not homogeneous across our set of example locations with the surrogate model performing similarly to the ensemble in Slovakia whilst in the United Kingdom and Germany the surrogate model performed visibly substantially worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
+For both the ensemble and the surrogate, performance visibly decreased as the forecast horizon increased with this being particularly noticable for the surrogate model during periods of high incidence.
+In general, in the study period, the ensemble appeared to be more liable to underpredict and the surrogate model appeared more likely to over-predict. 
 Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
 In comparison to the ensemble model the surrogate model appeared less likely to place weight on unfeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in February in Slovakia and Poland).
-Italy and Germany, which had the highest number of submitted models of the example locations had little clear difference in performance between the ensemble and surrogate models compared to other example locations.
 
@@
 Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble on the log scale though with greater uncertainty.
+Visual differences were more marked on the natural scale with the surrogate model forecasting spuriously high peak incidence.
 In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction.
@@
 The findings from our study may also be useful for informing this improved set of assumptions. 
+A particular focus should be understanding why our surrogate mdoel was liable to overestimate peak incidence and what simply additional assumptions may be used to mitigate this.
 In addition, the model we derived based on our assumptions was likely not optimal both in terms of compute time and accuracy at reproducing ensemble-like behaviour.

### 2022-09-27 paper/paper.Rmd
expannd on AE discussion to close #17

@@
 
+If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast again outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
+`r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`).
+As when using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (`r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` vs.
+`r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`).
+Across forecast horizons the same pattern of outperformance holds.
+However, the difference in relative performance was less than when the full probability distribution was accounted for with this becoming more marked as the forecast horizon increased (Figure \@ref(fig:assemble-eval) c).
+
 The surrogate model's relative performance varied over time with substantially worse performance from January to March compared to later in the year across all forecast horizons based on changes in the relative score distribution and its summary statistics (Figure \@ref(fig:assemble-eval) b).
@@
 Some locations showed less of this behaviour, for example, Spain, and in some, it was very dominant, for example, Switzerland.
-
-If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
-`r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) which is comparable on the relative scale to the difference in performance observed using WIS.
-As when using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
-`r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`).
 

### 2022-09-27 paper/paper.Rmd
add bias plot for #15

@@
 ```{r relative-interval-score}
+summary_performance <- scores |>
+  summarise_scores(by = c("model", "horizon"))
+
 relative_interval_score <- scores |>
@@
 b.) The density of the relative score by month for a given forecast horizon stratified by the one and four-week forecast horizon.
-c.) The density of the relative score by forecast location stratified by the one and four-week forecast horizon.
+c.) The average relative weighted interval score and absolute error for the surrogate model compared to the ensemble forecast by forecast horizon.
+d.) The density of the relative score by forecast location stratified by the one and four-week forecast horizon.
 The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast.
@@
 
-Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly certain at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) c).
+Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly certain at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
 This behaviour became more prominant as forecast horizon increased.
@@
 
-Breaking down the relative weighted interval score by forecast interval we observe that the surrogate modeel produces forecasts that differ most from the ensemble in the outer intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
+Breaking down the relative weighted interval score by forecast interval we observe that the surrogate modeel produces forecasts that differ most from the ensemble in the outer intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) c).
 This is true across forecast horizons but the magnitude of the difference increases.
@@
 Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the range of that interval (i.e 90\\%).
-b.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast.
-c.) Empirical coverage by quantile for both the surrogate and ensemble forecasts.
+b.) Empirical coverage by quantile for both the surrogate and ensemble forecasts.
 A well-calibrated forecast should have empirical quantiles that match the theoretical ones.
-The green area of this figure corresponds to conservative forecasts."}
+The green area of this figure corresponds to conservativ.
+forecasts.
+c.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast.
+d.) Bias of the ensemble and surrogate forecasts stratified by horizon."}
 p <- (
@@
     (
-      plot_coverage_range(coverage, ranges) /
-      plot_rel_score_by_quantile(relative_interval_score)
+      plot_coverage_range(coverage, ranges)
+    )
+  ) | (
+    plot_coverage_quantiles(scores) + (
+      plot_rel_score_by_quantile(relative_interval_score) +
+      plot_bias(summary_performance)
     ) +
-    plot_layout(heights = c(3, 1), guides = "collect")
-  ) |
-  plot_coverage_quantiles(scores)
+    plot_layout(heights = c(2, 1), guides = "collect")
+  )
 ) +

### 2022-09-27 paper/paper.Rmd
add bias methods detail + reference

@@
 
-In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile.
+In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile [@scoringutils].
+We also calculate the bias [@Funk2019-vi] of both forecasting approaches, stratified by forecast horizon.
+This metric aims to capture the tendency for a forecast to under- or over-predict.
+It captures the average proportion of the mass of the forecast distribution that is above or below the true value (and so can range from -1 to 1) with an unbiased forecast having an average bias value of 0.
 Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon, to assess the relative difference in performance across the predictive distribution.

### 2022-09-30 paper/paper.Rmd
review introduction

@@
 
-Whilst there is robust and consistent evidence that ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
+Whilst there is robust and consistent evidence that multi-team ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
 The most significant is their considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
@@
 Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
-Each of these may impact the long-term quality of the resulting ensemble forecasts and have implications for end-users.
+Each of these issues may impact the long-term quality of the resulting ensemble forecasts and have implications for end-users.
 
@@
 Over 2020 and 2021, teams established COVID-19 Forecasting Hubs covering the US [@Cramer2022-sp], Germany and Poland [@Bracher2021-wk], and Europe [@Sherratt2022-ue] (the latter two including authors of this study).
-All of these collaborations, which the author have been involved with managing, ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
-However, less progress has been made mitigating the downsides of these projects and improving access to the high quality and robust forecasts they seek to generate.
+All of these collaborations ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
+However, less progress has been made mitigating the downsides of these projects and improving access to the high quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.
 
 In this work, we draw insights from ensemble forecasts produced and endorsed by these forecast hubs, as well as our own forecasting work, to propose and evaluate a "surrogate" forecast model.
-This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justified assumptions, and limited computational resources with an easily generalised implementation.
+This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
 The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
-In addition to this, we seek to provide a simple, and robust, forecasting system that can be easily reused by others, perhaps where a large scale forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
+In addition to this, we seek to provide a simple, and robust, forecasting system that can be easily reused by others, perhaps where a forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
 
-In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
+In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
 However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
-This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz] utility, it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.
+This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @harles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.
 
@@
 Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
-Here, we focus on reported cases and primarily on the European forecast hub but our observations hold, in our view, across COVID-19 forecast hubs and to a less degree targets.
+Here, we focus on reported cases and primarily on the European forecast hub but our observations hold, in our view, across COVID-19 forecast hubs and to a lesser degree targets.
 In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
@@
 
-We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts.
+We have contributed a range of forecasts to these forecast collaborations, generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts.
 Our forecast submissions have not clearly over- or under- performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
@@
 Based on our observations our forecasts have generally captured the current trend relatively well, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
-Our previous methodology also requires significant computational resources running for an hour on a Azure D v5-series 16-core machine when producing forecasts for the European forecasting hub [@AzurePricing].
+Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing].
 In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
@@
 
-Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
+Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
 The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
@@
 The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
-This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model, where variance is linked to the mean, is a sensible choic. 
+This choice is motivated by our belief that the underlying transmission process is an exponential discrete one and therefore a count error model with a log link function, where variance is linked to the mean, is a sensible choice.
 We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].
@@
   (
+    (
       plot_relative_summary(relative_summary) +
+      plot_layout(guides = "keep")
+    )
     plot_rwis_location +

### 2022-09-30 paper/paper.Rmd
review paper

@@
 We then develop a forecast model using these assumptions and compare its performance to the European forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon using the weighted interval score.
-  We found that this surrogate model performed visibly relatively similarly to a subset of example locations though with increased uncertainty.
-Unsurprisingly, we concluded that average performance was substantially worse relative to the ensemble forecast but that median performance was more comparable.
-The majority of the difference in performance was linked to a subset of forecasts with worse performance early in 2022.
+  We found that this surrogate model performed visibly somewhat similarly to a subset of example locations though with increased uncertainty and poor performance at forecasting peak incidence.
+Unsurprisingly, we concluded that average performance was substantially worse relative to the ensemble forecast but that median performance was more comparable though still worse.
+The majority of the difference in performance was linked to a subset of forecasts with the majority of these being from early 2022.
 We note that despite performing worse than the hub ensemble the majority of the time our surrogate model showed better calibration, and was less liable to underpredict.
@@
 
-In this study, we evaluate an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions.
+In this study, we evaluate an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles based on a set of clear assumptions.
 We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble.
@@
 To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomolous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomolous data [@Sherratt2022-ue].
-We defined anomolous data based on the implementation used by the [@EuroHub] where a datapoint is considered anomolous if a future revision alters it by more than 5%.
+We defined anomolous data based on the implementation used by [@EuroHub] where a datapoint is considered anomolous if a future revision alters it by more than 5%.
 
@@
 
-We model the expectation ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(P)) process by epidemiological week ($t$).
+We model the expectation ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(p)) process by epidemiological week ($t$) on the log scale.
 The model is initialised by assuming that the initially reported cases are representative with a small amount of error (2.5%).
@@
 On the natural scale the difference in performance was more marked, specially for periods of high incidence and at longer horizons (Figure \@ref(fig:vis-forecasts) a).
-Performance was not homogeneous across our set of example locations with the surrogate model performing similarly to the ensemble in Slovakia whilst in the United Kingdom and Germany the surrogate model performed visibly substantially worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
-For both the ensemble and the surrogate, performance visibly decreased as the forecast horizon increased with this being particularly noticable for the surrogate model during periods of high incidence.
-In general, in the study period, the ensemble appeared to be more liable to underpredict and the surrogate model appeared more likely to over-predict. 
+Performance was not homogeneous across our set of example locations with the surrogate model performing similarly to the ensemble in Slovakia whilst in the United Kingdom and Germany the surrogate model performed substantially worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
+For both the ensemble and the surrogate, performance decreased as the forecast horizon increased with this being particularly noticable for the surrogate model during periods of high incidence.
+In general, in the study period the ensemble appeared to be better able to forecast peak incidence. 
 Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
@@
 This behaviour became more prominant as forecast horizon increased.
-In contrast, the surrogat forecast was generally equally well calibrated at across horizons with a tendency to over cover for intervals above the median.
-At longer horizons, however, quantiles below the median became liable to have under coverage.
+In contrast, the surrogate forecast was generally equally well calibrated across horizons with a tendency to over cover for intervals above the median.
+At longer horizons, however, quantiles below the median became liable to under cover.
 
@@
 This is true across forecast horizons but the magnitude of the difference increases.
+
+Calculating the bias of the forecasts from each model we see that the (Figure \@ref(fig:plot-coverage) d) ensemble forecast is initially biased towards undprediction but this bias reduces as the forecast horizon increases.
+In comparison, the surrogate forecast model is biased towards over prection for all forecast horizons with the magnitude of this bias appearing to increase linearly with forecast horizon.
 
@@
 In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the European forecast hub multi-team ensemble for forecasting test-positive reported COVID-19 cases in European nations.
-We first defined a set of assumptions for how the surrogate model should behave based on our observations of the European forecast hub ensemble, our experience submitting forecasts to various forecast hubs, and our observations of the performance of other models.
+We first defined a set of assumptions for how the surrogate model should behave based on our observations of the European forecast hub ensemble, our experience submitting forecasts to various forecast hubs.
 We aimed for a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources.
@@
 Visual differences were more marked on the natural scale with the surrogate model forecasting spuriously high peak incidence.
-In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction.
-Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of vey poorly performing forecasts.
+In a subset of example locations, we observed some variation in performance across locations, that the ensemble better captured peak incidence, and that the surrogate model appeared biased toward overprediction.
+Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse, and that relative performance decreased with forecast horizon.
 Median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
+The difference in mean and median relative performance suggested a skewed distribution in scores, which we confirmed visually. 
+This means that a relatively small fraction of forecasts were responsible for a substantially portion for the difference in performance.
+Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though relative performance of the surrogate model generally improved.
 Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
 In general, the relative performance of the surrogate model degraded as forecast horizons increas with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
-The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
-Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though relative performance of the surrogate model generally improved.
+The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and larger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
+The ensemble forecast was biased towards under prediction at short to medium forecast horizons but unbiased at longer horizons. 
+In comparison the surrogate model was biased towards over prediction and this biased increased linearly with forecast horizon.
 
@@
 In September 2022, GitHub announced support for hosted GitHub Action runners with additional compute power [@GitHubActionsLargeRunners].
-Whilst a paid feature this may allow more compute attensive models, with fewer potental performance trade-offs, to be easily democratised though only if funds are available to support the hosting costs.
+Whilst a paid feature this may allow more compute intensive models, with fewer potental performance trade-offs, to be easily democratised though only if funds are available to support the hosting costs.
 One potential research area is to explore forecasting methods that can be used with a range of compute resources though this would require extensive evaluation and documentation to make it clear to users what the trade-offs between compute usage and forecast performance are.
-More work is needed to understand the best practice treatment of data revisions when evaluating forecasts and the potential sources for bias these may cause.
+More work is needed to understand the best practice treatment of data revisions when evaluating forecasts and the potential bias these may cause.
 Lastly, here we have only explored a surrogate for an ensemble for a single disease, a limited set of locations, and a single target (incident cases), meaning our findings are difficult to generalise.
@@
 
-SA and SF were funded by a Wellcome senior fellowship to SF (), KS and HG were funded by an ECDC grant to SF, and JB was funded by.
+SA and SF were funded by a Wellcome senior fellowship to SF (210758/Z/18/Z), KS and HG were funded by an ECDC grant to SF, and JB was funded by.
 
@@
 We thank the forecast hub team for publishing all data in an accessible format.
+We thank the Epiforecast group for a productive discussion of an early version of this analysis.
 We thank Molly for being a good Labrador.

### 2022-09-30 paper/paper.Rmd
fix plot

@@
       plot_layout(guides = "keep")
-    )
+    ) +
     plot_rwis_location +

### 2022-09-30 paper/paper.Rmd
add KS as a code reviewer
spell check

@@
 Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Reich2022-jo].
-These approaches have been inspired by the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providng forecasts that are trusted by stakholders.
+These approaches have been inspired by the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providng forecasts that are trusted by stakeholders.
 
@@
 All of these collaborations ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
-However, less progress has been made mitigating the downsides of these projects and improving access to the high quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.
+However, less progress has been made in mitigating the downsides of these projects and improving access to the high-quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.
 
-In this work, we draw insights from ensemble forecasts produced and endorsed by these forecast hubs, as well as our own forecasting work, to propose and evaluate a "surrogate" forecast model.
+In this work, we draw insights from ensemble forecasts produced and endorsed by these forecast hubs, as well as our forecasting work, to propose and evaluate a "surrogate" forecast model.
 This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
@@
 
-In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
-However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
+In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysics [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
+However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models, in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
 This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @harles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.
 
-In order to understand the behaviour of the forecast hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
-These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their own forecast models and then use them to produce weekly forecast that they then submit to the central hub team.
-Each hub targets a range of metrics including test positive reported cases, repoorted deaths, and hospitalisations, has a specific geographic focus, and asks for weekly forecasts (using epidemiological weeks i.e.
+To understand the behaviour of the forecast hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
+These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team.
+Each hub targets a range of metrics including test-positive reported cases, reported deaths, and hospitalisations has a specific geographic focus and asks for weekly forecasts (using epidemiological weeks i.e.
 Sunday to Saturday) over a time horizon of a few weeks.
@@
 
-We have contributed a range of forecasts to these forecast collaborations, generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts.
-Our forecast submissions have not clearly over- or under- performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
+We have contributed a range of forecasts to these forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
+Our forecast submissions have not over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
 The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
-Based on our observations our forecasts have generally captured the current trend relatively well, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
+Based on our observations our forecasts have generally captured the current trend relatively well but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
 Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing].
@@
 On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
-Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting.
+Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under vs over predicting.
 There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models
 
-Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
+Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability.
 The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
@@
 As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
-To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomolous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomolous data [@Sherratt2022-ue].
-We defined anomolous data based on the implementation used by [@EuroHub] where a datapoint is considered anomolous if a future revision alters it by more than 5%.
+To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomalous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomalous data [@Sherratt2022-ue].
+We defined anomalous data based on the implementation used by [@EuroHub] where a datapoint is considered anomalous if a future revision alters it by more than 5%.
 
@@
 We also calculate the bias [@Funk2019-vi] of both forecasting approaches, stratified by forecast horizon.
-This metric aims to capture the tendency for a forecast to under- or over-predict.
+This metric aims to capture the tendency for a forecast to under or over-predict.
 It captures the average proportion of the mass of the forecast distribution that is above or below the true value (and so can range from -1 to 1) with an unbiased forecast having an average bias value of 0.
@@
 The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
-`cmdstanr` surfaces several settings that trade off between sampling speed and the robustness of the approach.
+`cmdstanr` surfaces several settings that trade-off between sampling speed and the robustness of the approach.
 Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15.
@@
 To ensure the reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with a built Docker image [@Boettiger:2015dw] (via GitHub Actions [@GitHubActionsAbout]) is provided in the code repository.
-Weekly forecasts were made using `renv` and based GitHub Actions free tier as available in 2022 to ensure they require limited compute and that our implementation is independent of local resources facilitating democratised access.
+Weekly forecasts were made using `renv` and based on GitHub Actions free tier as available in 2022 to ensure they require limited compute and that our implementation is independent of local resources facilitating democratised access.
 The free GitHub Actions runner we used for all forecasts was Ubuntu `20.04` based with 2 cores (x86_64), 7 GB of RAM, and 14 GB of SSD space.
@@
 
-We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
+We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, the European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
 In total `r nrow(metadata)` forecasts were made across all locations, with `r forecast_dates` unique forecast dates and `r models` independent forecast models (including the European hub baseline model).
@@
 Of the remaining models submitted `r length(local_models)` were submitted in only one location.
-Single location models were clustered in a few locations, particularly Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
+Single location models were clustered in a few locations, particularly in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
 Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
@@
 Forecasts for these observations were excluded as were forecasts for forecast weeks where they were the latest available data.
-Data anomalies were not randomly distributed with some locations being particlarly prone to data revisions including Lithuania (with `r anomalies[location == "LT", .N]` weeks with data anomalies), and Portugal (with `r anomalies[location == "PT", .N]` weeks with data anomalies).
+Data anomalies were not randomly distributed with some locations being particularly prone to data revisions including Lithuania (with `r anomalies[location == "LT", .N]` weeks with data anomalies), and Portugal (with `r anomalies[location == "PT", .N]` weeks with data anomalies).
 Anomalies were also not evenly distributed over time with a higher proportion occurring earlier in the study period (potentially due to our choice to extract data from the 1st of September which effectively truncated anomalies).
@@
 In our example set of locations, the absolute performance of the ensemble and the surrogate model was visually similar on the log scale in all locations at short forecast horizons though this varied by location (Figure \@ref(fig:vis-forecasts) b).
-On the natural scale the difference in performance was more marked, specially for periods of high incidence and at longer horizons (Figure \@ref(fig:vis-forecasts) a).
+On the natural scale the difference in performance was more marked, especially for periods of high incidence and at longer horizons (Figure \@ref(fig:vis-forecasts) a).
 Performance was not homogeneous across our set of example locations with the surrogate model performing similarly to the ensemble in Slovakia whilst in the United Kingdom and Germany the surrogate model performed substantially worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
-For both the ensemble and the surrogate, performance decreased as the forecast horizon increased with this being particularly noticable for the surrogate model during periods of high incidence.
-In general, in the study period the ensemble appeared to be better able to forecast peak incidence. 
+For both the ensemble and the surrogate, performance decreased as the forecast horizon increased with this being particularly noticeable for the surrogate model during periods of high incidence.
+In general, in the study period, the ensemble appeared to be better able to forecast peak incidence. 
 Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
@@
 
-Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by at least 25% and that the relative performance of the surrogate model degraded as the forecast horizon increased  (Figure \@ref(fig:assemble-eval) c).
+Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one-week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two-week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three-week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four-week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by at least 25% and that the relative performance of the surrogate model degraded as the forecast horizon increased  (Figure \@ref(fig:assemble-eval) c).
 Much of this outperformance, especially at longer forecast horizons, was driven by a relatively small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a).
@@
 `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`).
-As when using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (`r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` vs.
+When using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (`r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` vs.
 `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`).
@@
 In comparison, the surrogate forecast was less well calibrated than the ensemble forecast at the one-week forecast horizon with a tendency to have a larger empirical coverage than required  (Figure \@ref(fig:plot-coverage) a).
-At longer horizons and narrower prediction intervals the surrogate forecast became better calibrated though with a tendency to under cover.
-This was not the case for the 90% prediction interval where the surrogate model covered more than the expected interval, for all horizons, indicating forecasts were overly uncertain for this interval regardless of horizon.
+At longer horizons and narrower prediction intervals, the surrogate forecast became better calibrated though with a tendency to under cover.
+This was not the case for the 90% prediction interval where the surrogate model covered more than the expected interval, for all horizons, indicating forecasts were overly uncertain for this interval regardless of the horizon.
 
 Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly certain at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
-This behaviour became more prominant as forecast horizon increased.
+This behaviour became more prominent as the forecast horizon increased.
 In contrast, the surrogate forecast was generally equally well calibrated across horizons with a tendency to over cover for intervals above the median.
@@
 
-Breaking down the relative weighted interval score by forecast interval we observe that the surrogate modeel produces forecasts that differ most from the ensemble in the outer intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) c).
+Breaking down the relative weighted interval score by forecast interval we observe that the surrogate model produces forecasts that differ most from the ensemble in the outer intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) c).
 This is true across forecast horizons but the magnitude of the difference increases.
 
-Calculating the bias of the forecasts from each model we see that the (Figure \@ref(fig:plot-coverage) d) ensemble forecast is initially biased towards undprediction but this bias reduces as the forecast horizon increases.
-In comparison, the surrogate forecast model is biased towards over prection for all forecast horizons with the magnitude of this bias appearing to increase linearly with forecast horizon.
+Calculating the bias of the forecasts from each model we see that the (Figure \@ref(fig:plot-coverage) d) ensemble forecast is initially biased towards underprediction but this bias reduces as the forecast horizon increases.
+In comparison, the surrogate forecast model is biased towards overprediction for all forecast horizons with the magnitude of this bias appearing to increase linearly with the forecast horizon.
 
@@
 In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the European forecast hub multi-team ensemble for forecasting test-positive reported COVID-19 cases in European nations.
-We first defined a set of assumptions for how the surrogate model should behave based on our observations of the European forecast hub ensemble, our experience submitting forecasts to various forecast hubs.
+We first defined a set of assumptions for how the surrogate model should behave based on our observations of the European forecast hub ensemble, and our experience submitting forecasts to various forecast hubs.
 We aimed for a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources.
@@
 Visual differences were more marked on the natural scale with the surrogate model forecasting spuriously high peak incidence.
-In a subset of example locations, we observed some variation in performance across locations, that the ensemble better captured peak incidence, and that the surrogate model appeared biased toward overprediction.
-Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse, and that relative performance decreased with forecast horizon.
-Median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
+In a subset of example locations, we observed some variation in performance across locations, that the ensemble better-captured peak incidence, and that the surrogate model appeared biased toward overprediction.
+Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse and that relative performance decreased with the forecast horizon.
+The median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
 The difference in mean and median relative performance suggested a skewed distribution in scores, which we confirmed visually. 
-This means that a relatively small fraction of forecasts were responsible for a substantially portion for the difference in performance.
-Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though relative performance of the surrogate model generally improved.
+This means that a relatively small fraction of forecasts were responsible for a substantial portion of the difference in performance.
+Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though the relative performance of the surrogate model generally improved.
 Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
-In general, the relative performance of the surrogate model degraded as forecast horizons increas with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
+In general, the relative performance of the surrogate model degraded as forecast horizons increased with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
 The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and larger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
 The ensemble forecast was biased towards under prediction at short to medium forecast horizons but unbiased at longer horizons. 
-In comparison the surrogate model was biased towards over prediction and this biased increased linearly with forecast horizon.
+In comparison, the surrogate model was biased towards over prediction and this bias increased linearly with the forecast horizon.
 
@@
 This means that we can be confident we have captured how our forecast methodology behaves in real-world usage, at least within our study period. 
-The downside of this approach is that it was not possible to update the surrogate model over time in response to initial evaluation or to explore other parameterisations that might be more successful of which there are likely several.
+The downside of this approach is that it was not possible to update the surrogate model over time in response to the initial evaluation or to explore other parameterisations that might be more successful of which there are likely several.
 However, as our study has been conducted with a focus on reproducibility and openness our findings can be replicated or extended by others regardless of compute availability (due to our use of GitHub actions as a compute platform which is freely available to researchers).
@@
 This is difficult to avoid without retrospectively re-calculating the ensemble using the same approach as taken by the hub which would reduce the independence of the hub ensemble as a source of truth to compare our forecasts against.
-Given the number of forecasts submitted in most locations, and the Europeans forecast hubs practice of not calculating an ensemble when fewer than 3 independent forecasts were available, the bias on our results caused by this limitation should be relatively small.
+Given the number of forecasts submitted in most locations and the European forecast hubs' practice of not calculating an ensemble when fewer than 3 independent forecasts were available, the bias in our results caused by this limitation should be relatively small.
 Importantly in this study, we focussed on replicating the forecast hub ensembles' observed behaviour rather than attempting to define an optimal forecast for forecast consumers.
@@
 Our focus on replicating the performance of the hub ensemble is also useful as the surrogate model may highlight some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
-It also highlights some of the differences between our surrogate forecast model and the ensemble that may lead to new insights in understanding the mechanisms leading to the ensembles behaviour, such as the generally poor coverage of the ensemble that could not be explained by the assumptions we used in developing our surrogate methodology.
+It also highlights some of the differences between our surrogate forecast model and the ensemble that may lead to new insights into the mechanisms leading to the ensemble's behaviour, such as the generally poor coverage of the ensemble that could not be explained by the assumptions we used in developing our surrogate methodology.
 Whilst we normalised reported cases to be population-adjusted incidence rates our results are conditional on the use of the weighted interval score as an evaluation metric.
@@
 This bias could explain the relatively poor performance of the surrogate model, compared to the ensemble, despite the surrogate model being comparably well-calibrated.
-We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick relatively closely to the methodology used by the European forecast hub [@Sherratt2022-ue], aside from the use of population weighting in order to facilitate comparison between forecast locations, as these choices inform the development of submitted models and so are key to our findings.
+We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick relatively closely to the methodology used by the European forecast hub [@Sherratt2022-ue], aside from the use of population weighting to facilitate comparison between forecast locations, as these choices inform the development of submitted models and so are key to our findings.
 
@@
 In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study).
-The poor calibration of the forecast ensembles produced by median Hub ensembles has been noted repeatedly [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue] but little progress has been made understanding the causes or suggesting alternatives forecasts for policy makers to use.
-Progress understanding which structural model features lead to better infectious disease forecasts has been limited.
+The poor calibration of the forecast ensembles produced by median Hub ensembles has been noted repeatedly [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue] but little progress has been made in understanding the causes or suggesting alternatives forecasts for policymakers to use.
+Progress in understanding which structural model features lead to better infectious disease forecasts has been limited.
 The US forecast hub identified the top 5 performing models and noted the structural assumptions they made but couldn't directly link assumptions with performance [@Cramer2022-sp].
 They also did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
-Similarly, the Germany and Poland forecasting hubs was able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
+Similarly, the Germany and Poland forecasting hubs were able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
 All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
@@
 Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
-Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not in our the model-based forecasts [@Bosse2022-hm; @Bracher2021-wk].
-Finally, our results are potentially sensitive to the definition used to define anonamous observations (which generally related to retrospective data revisions).
+Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not in our model-based forecasts [@Bosse2022-hm; @Bracher2021-wk].
+Finally, our results are potentially sensitive to the definition used to define anomalous observations (generally related to retrospective data revisions).
 Here we follow the practice of the European forecast hub [@EuroHub] of excluding forecasts for weeks with a data revision of more than 5% and forecasts made based on data that is subsequently revised by more than 5%.
@@
 The findings from our study may also be useful for informing this improved set of assumptions. 
-A particular focus should be understanding why our surrogate mdoel was liable to overestimate peak incidence and what simply additional assumptions may be used to mitigate this.
+A particular focus should be on understanding why our surrogate model was liable to overestimate peak incidence and what simply additional assumptions may be used to mitigate this.
 In addition, the model we derived based on our assumptions was likely not optimal both in terms of compute time and accuracy at reproducing ensemble-like behaviour.
@@
 In September 2022, GitHub announced support for hosted GitHub Action runners with additional compute power [@GitHubActionsLargeRunners].
-Whilst a paid feature this may allow more compute intensive models, with fewer potental performance trade-offs, to be easily democratised though only if funds are available to support the hosting costs.
-One potential research area is to explore forecasting methods that can be used with a range of compute resources though this would require extensive evaluation and documentation to make it clear to users what the trade-offs between compute usage and forecast performance are.
+Whilst a paid feature this may allow more compute-intensive models, with fewer potential performance trade-offs, to be easily democratised though only if funds are available to support the hosting costs.
+One potential research area is to explore forecasting methods that can be used with a range of computing resources though this would require extensive evaluation and documentation to make it clear to users what the trade-offs between compute usage and forecast performance are.
 More work is needed to understand the best practice treatment of data revisions when evaluating forecasts and the potential bias these may cause.
@@
 We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
-We also conclude that our findings may belargely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
+We also conclude that our findings may be largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
 Our work is useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to researchers thinking about how to develop forecasts that perform similarly to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
@@
 SF provided funding.
-HG reviewed the code and reproducibility of the analyses.
+HG and KS reviewed the code and reproducibility of the analyses.
 

### 2022-10-03 paper/paper.Rmd
update scoringutils and refactor intro paragraph

@@
 Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Reich2022-jo].
-These approaches have been inspired by the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providng forecasts that are trusted by stakeholders.
+These approaches have been inspired by the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providing forecasts that are trusted by stakeholders.
 
 Whilst there is robust and consistent evidence that multi-team ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
-The most significant is their considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
-Forecasts from these ensembles can also be hard to interpret as it is difficult to link them to the underlying model assumptions.
+The most significant is the difficulty in interpreting the underlying mechanisms for the forecasts they produce. 
+This impacts users' trust, how easily ensemble performance can be improved, and how easily contributor forecasts can be improved.
+Forecasts from these ensembles also require considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
 Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
 Each of these issues may impact the long-term quality of the resulting ensemble forecasts and have implications for end-users. 
+There has been limited critical feedback on forecasting ensembling projects for infectious disease epidemiology with little evaluation of the effort required to produce them compared to their benefits for improving forecast performance.
+What evaluation there has been has mainly focussed on secondary data analysis.
 
@@
 The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
-In addition to this, we seek to provide a simple, and robust, forecasting system that can be easily reused by others, perhaps where a forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
+In addition to this, we seek to provide a simple, and robust, forecasting system that others can easily reuse, perhaps where a forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
 
@@
 The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
-This reduces removes the need for, potentially misspecified, external information on the delay from infection to report, and reduces computational requirements due to a reduction in model complexity. 
+This removes the need for, potentially misspecified, external information on the delay from infection to report, and reduces computational requirements due to a reduction in model complexity. 
 However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
@@
 
-The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
+The final simplification is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
 This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
@@
 In our study period, incidence rates across European nations and in the UK were primarily driven by the spread of novel subvariants of concern related to the Omicron variant and changes in population susceptibility.
-Many nations, such as the UK, saw large BA.1 waves in January that then resulted in declining incidence rates through February (Figure \@ref(fig:vis-forecasts)).
+Many countries, such as the UK, saw large BA.1 waves in January that then resulted in declining incidence rates through February (Figure \@ref(fig:vis-forecasts)).
 From late February, through to the end of May, most nations then saw another wave, typically with lower reported incidence rates, driven by BA.2.

### 2022-10-03 paper/paper.Rmd
partial implementation of comments from @jbracher

@@
 However, more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
-We also conclude that our findings may be largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
+We also conclude that our findings may be largely driven by the choice of evaluation measure used by the forecast hub.
+If the assumptions underlying this measure do not reflect forecast users' expectations then the forecast hub submissions may also not reflect these expectations.
 Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers aiming to develop forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
@@
 Whilst there is robust and consistent evidence that multi-team ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
-The most significant is the difficulty in interpreting the underlying mechanisms for the forecasts they produce. 
+The most significant is the difficulty in interpreting them.
+This relates both to the underlying mechanisms for the forecasts they produce and to understanding if and when their behaviour can is desirable.
 This impacts users' trust, how easily ensemble performance can be improved, and how easily contributor forecasts can be improved.
 Forecasts from these ensembles also require considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
-Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
+Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute and adjust their models to changing conditions [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
 Each of these issues may impact the long-term quality of the resulting ensemble forecasts and have implications for end-users. 
@@
 The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic due to a pressing need for reliable forecasts and a perception that many publicly available forecasts were low quality.
-Over 2020 and 2021, teams established COVID-19 Forecasting Hubs covering the US [@Cramer2022-sp], Germany and Poland [@Bracher2021-wk], and Europe [@Sherratt2022-ue] (the latter two including authors of this study).
+Over 2020 and 2021, teams established COVID-19 Forecasting Hubs covering the US [@Cramer2022-sp], Germany and Poland [@Bracher2021-wk], and Europe [@Sherratt2022-ue] (all three including authors of this study).
 All of these collaborations ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
+Both the US and European forecast hubs were supported and received funding from, public health agencies (the Center for Disease Control (CDC) and European Center for Disease Prevention and Control (ECDC) respectively) with their forecasts being adopted into the official output from these agencies.
 However, less progress has been made in mitigating the downsides of these projects and improving access to the high-quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.
@@
 However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models, in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
-This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @harles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.
+This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @charles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.
 
@@
 Here, we focus on reported cases and primarily on the European forecast hub but our observations hold, in our view, across COVID-19 forecast hubs and to a lesser degree targets.
-In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
+We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator, and they are generally the most challenging to predict.
+In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-ft; @Adam2020-if], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
 Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].
@@
 We have contributed a range of forecasts to these forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
-Our forecast submissions have not over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
+Our forecast submissions have not systematically over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
 The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
@@
 In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
-We found that our human insight-based forecasts significantly outperformed our model-based forecasts on average and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].
+We found that our human insight-based forecasts outperformed our model-based forecasts on average. 
+This was particularly the case when forecasting cases and at longer forecast horizons.
+We hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].
 
@@
 They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions.
-In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence.
+In general, in comparison to our submitted forecasts, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence.
 On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
@@
 
-## Data {-}
+## Setting of the European COVID-19 Forecast Hub {-}
 
@@
 
-We model the expectation ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(p)) process by epidemiological week ($t$) on the log scale.
+We model the expectation ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(1)) process by epidemiological week ($t$) on the log scale.
 The model is initialised by assuming that the initially reported cases are representative with a small amount of error (2.5%).
@@
 
-Where $r_t$ can be interpreted as the growth rate.
+where $r_t$ can be interpreted as the weekly growth rate.
 $r_t$ is then modelled as a piecewise constant differenced AR(1) process modified such that the dependence of $r_{t-1}$ is multiplied by a decay factor ($\xi_{+,-}$) that varies dynamically according to the sign of $r_{t-1}$.
@@
   r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
-  r_t &= \left(\mathcal{H}(r_{t-1}) \xi_{+} + (1 - \mathcal{H}(r_{t-1})) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
-  \epsilon_t &= \mathcal{H}(t) \beta \epsilon_{t-1} + \eta_t
+  r_t &= \left(\mathbf{1}_{r_{t-1} > 0} \xi_{+} + \mathbf{1}_{r_{t-1} \leq 0}\xi_{-}\right)r_{t-1} + \epsilon_t  \\
+  \epsilon_t &= \mathbf{1}_{t > 0} \beta \epsilon_{t-1} + \eta_t
 \end{align*}
 
-Where $\mathcal{H}(x)$ is the Heaviside step function and is defined such that it attains the value of 1 if the argument is greater than 0 and 0 otherwise, and $\epsilon_t$ and $\eta_t$ are error terms.
+where $\epsilon_t$ and $\eta_t$ are error terms. 
 The following priors are used,
@@
 
-### Summary statistics {-}
-
-As well as posterior predictions and forecasts for notifications the model also returns several epidemiological summary statistics which may be useful for drawing inferences about underlying transmission dynamics.
-These are the log scale weekly growth rate ($g^{o, \delta}_t$), and the instantaneous effective reproduction number ($R^{o, \delta}_t$).
-These are calculated as follows:
-
-\begin{align*}
-  g^{o, \delta}_t &= T_s r^{o, \delta}_t \\
-  R^{o, \delta}_t &= e^{\left(T_s r^{o, \delta}_t\right)} \\
-\end{align*}
-
-$T_s = 7$ is a parameter that defines the weekly timespan over which the summary metrics apply dependent on the time step of the data.
-
 ## Forecast evaluation {-}
@@
 
-We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS), and the absolute error (AE) of the median forecast.
+We evaluate forecasts for all locations and horizons quantitatively using the absolute error (AE) of the median forecast and the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021].
+The WIS is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
 Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub [@Sherratt2022-ue].
@@
 ```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate and ensemble forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon.
-Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the range of that interval (i.e 90\\%).
+Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the nominal level of the interval (i.e., 30\\%, 60\\% and 90\\%, respectively).
 b.) Empirical coverage by quantile for both the surrogate and ensemble forecasts.
 A well-calibrated forecast should have empirical quantiles that match the theoretical ones.
-The green area of this figure corresponds to conservativ.
-forecasts.
+The green area of this figure corresponds to conservative forecasts.
 c.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast.
@@
 
-## Limitations {-}
+## Strengths and Weaknesses {-}
 
 Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the European forecast hub [@Sherratt2022-ue]).
-This means that we can be confident we have captured how our forecast methodology behaves in real-world usage, at least within our study period. 
+This means that we are not subject to hindsight bias. 
 The downside of this approach is that it was not possible to update the surrogate model over time in response to the initial evaluation or to explore other parameterisations that might be more successful of which there are likely several.
@@
 It is possible that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
-However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue], then our approach makes sense and may be optimal for producing a "good" forecast.
+However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue], then our approach may make sense as a way of producing a "good" forecast.
 Our focus on replicating the performance of the hub ensemble is also useful as the surrogate model may highlight some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
@@
 In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions. 
-The US [@Cramer2022-sp], and European [@Sherratt2022-ue]], forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling.
-Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the UUS forecasting hub and did not outperform in the case of the European forecasting hub.
-No forecast hub has switched to these alternative ensemble designs for their operational forecast suggesting they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.
+The US [@Ray2022-zq], and European [@Sherratt2022-ue]], forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling.
+Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the US forecasting hub and did not outperform in the case of the European forecasting hub.
+No forecast hub has switched to these alternative ensemble designs for their operational forecast of reported cases, though the US hub has switched to a trained ensemble for death forecasts.
+This suggests that the hub teams do not think the evidence base is strong enough for trained ensembles to be used by forecast consumers for reported cases and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.
 Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
@@
 We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
-We also conclude that our findings may be largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements.
+We also conclude that our findings may be largely driven by the choice of evaluation measure used by the forecast hub.
+While this measure has desirable mathematical properties and is routinely used in a similar form e.g., in weather forecasting, it is subject to debate whether it appropriately reflects forecast users' requirements and perceptions as to what makes a good forecast.
 Our work is useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to researchers thinking about how to develop forecasts that perform similarly to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
@@
 
-JB and SF, and KS have received funding to run forecast hub projects.
+SF, JB, JS, BB, and HG have coordinated Forecast Hub platforms. SF and KS received funding from the European Center for Disease Prevention and Control to this end.
 
@@
 All other authors provided feedback on the manuscript and analyses and contributed to revisions.
-SF provided funding.
 HG and KS reviewed the code and reproducibility of the analyses.
@@
 
-SA and SF were funded by a Wellcome senior fellowship to SF (210758/Z/18/Z), KS and HG were funded by an ECDC grant to SF, and JB was funded by.
+SA and SF were funded by a Wellcome senior fellowship to SF (210758/Z/18/Z), KS and HG were funded by an ECDC grant to SF.
+JB acknowledges support from the Helmholtz Foundation via the SIM-610 CARD Information and Data Science Pilot Project.
 

### 2022-10-04 paper/paper.Rmd
change axis text size

@@
     legend.position = "bottom", legend.margin = margin(),
-    legend.justification = "centre"
+    legend.justification = "centre", axis.title = element_text(size = 12)
   )

### 2022-10-04 paper/paper.Rmd
more changes from @jbracher

--- 2022-10-03 paper/paper.Rmd
@@
   Here we use observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience, to specify a series of model assumptions modelling incident reported cases of COVID-19.
-We then develop a forecast model using these assumptions and compare its performance to the European forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon using the weighted interval score.
+We then develop a forecast model using these assumptions and compare its performance to the European Forecast Hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon using the weighted interval score.
   We found that this surrogate model performed visibly somewhat similarly to a subset of example locations though with increased uncertainty and poor performance at forecasting peak incidence.
@@
 However, more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
-We also conclude that our findings may be largely driven by the choice of evaluation measure used by the forecast hub.
-If the assumptions underlying this measure do not reflect forecast users' expectations then the forecast hub submissions may also not reflect these expectations.
+We also conclude that our findings may be largely driven by the choice of evaluation measure used by the Forecast Hub.
+If the assumptions underlying this measure do not reflect forecast users' expectations then the Forecast Hub submissions may also not reflect these expectations.
 Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers aiming to develop forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
@@
 All of these collaborations ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
-Both the US and European forecast hubs were supported and received funding from, public health agencies (the Center for Disease Control (CDC) and European Center for Disease Prevention and Control (ECDC) respectively) with their forecasts being adopted into the official output from these agencies.
+Both the US and European Forecast Hubs were supported and received funding from public health agencies (the Center for Disease Control (CDC) and European Center for Disease Prevention and Control (ECDC) respectively) with their forecasts being adopted into the official output from these agencies.
 However, less progress has been made in mitigating the downsides of these projects and improving access to the high-quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.
 
-In this work, we draw insights from ensemble forecasts produced and endorsed by these forecast hubs, as well as our forecasting work, to propose and evaluate a "surrogate" forecast model.
+In this work, we draw insights from ensemble forecasts produced and endorsed by these Forecast Hubs, as well as our forecasting work, to propose and evaluate a "surrogate" forecast model.
 This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
 The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
-In addition to this, we seek to provide a simple, and robust, forecasting system that others can easily reuse, perhaps where a forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
+In addition to this, we seek to provide a simple, and robust, forecasting system that others can easily reuse, perhaps where a Forecast Hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
 
@@
 
-To understand the behaviour of the forecast hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
+To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
 These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team.
@@
 Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
-Here, we focus on reported cases and primarily on the European forecast hub but our observations hold, in our view, across COVID-19 forecast hubs and to a lesser degree targets.
+Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
 We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator, and they are generally the most challenging to predict.
@@
 They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions.
-In general, in comparison to our submitted forecasts, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence.
+In comparison to our submitted forecasts, the ensemble forecasts were less reactive to changes in trend such as from stable or reducing case incidence to increasing incidence.
 On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
@@
 We briefly outline some of the summary statistics it produces and their epidemiological relevance.
-We then evaluate its real-time performance in comparison to the European forecast hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
+We then evaluate its real-time performance in comparison to the European Forecast Hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
 We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed.
 Finally, we summarise our findings, discuss their implications, and highlight areas where more work is needed.
-We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
+We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 Forecast Hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
 
@@
 
-The European forecast hub requests forecasts for one to four-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
+The European Forecast Hub requests forecasts for one to four-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
 No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
-Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@EuroHub].
+Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@EuroHub; @Bracher2022-ug; @Cramer2022-sp].
 Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week).
@@
 An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
-Submitted forecasts and target observations are available from the European forecast hub GitHub repository [@EuroHubGitHub]. We provide code in the repository of this study to streamline access.
+Submitted forecasts and target observations are available from the European Forecast Hub GitHub repository [@EuroHubGitHub]. We provide code in the repository of this study to streamline access.
 
@@
 
-We model the expectation ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(1)) process by epidemiological week ($t$) on the log scale.
+We model the expectation ($\lambda_t$) of reported cases ($C_t$) given past reported cases as an order 1 autoregressive (AR(1)) process by epidemiological week ($t$) on the log scale.
 The model is initialised by assuming that the initially reported cases are representative with a small amount of error (2.5%).
@@
 The WIS is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
-Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub [@Sherratt2022-ue].
+Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European Forecast Hub [@Sherratt2022-ue].
 We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).
@@
 
-`r nrow(anomalies)` observations, stratified by week and location, were defined to be anomalous within the study period by the European forecast hub [@EuroHub]. 
+`r nrow(anomalies)` observations, stratified by week and location, were defined to be anomalous within the study period by the European Forecast Hub [@EuroHub]. 
 Forecasts for these observations were excluded as were forecasts for forecast weeks where they were the latest available data.
@@
     legend.position = "bottom", legend.margin = margin(),
-    legend.justification = "centre"
+    legend.justification = "centre", axis.title = element_text(size = 15)
   )
@@
 
-In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the European forecast hub multi-team ensemble for forecasting test-positive reported COVID-19 cases in European nations.
-We first defined a set of assumptions for how the surrogate model should behave based on our observations of the European forecast hub ensemble, and our experience submitting forecasts to various forecast hubs.
+In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the European Forecast Hub multi-team ensemble for forecasting test-positive reported COVID-19 cases in European nations.
+We first defined a set of assumptions for how the surrogate model should behave based on our observations of the European Forecast Hub ensemble, and our experience submitting forecasts to various Forecast Hubs.
 We aimed for a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources.
@@
 
-Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble on the log scale though with greater uncertainty.
+Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the Forecast Hub ensemble on the log scale though with greater uncertainty.
 Visual differences were more marked on the natural scale with the surrogate model forecasting spuriously high peak incidence.
 In a subset of example locations, we observed some variation in performance across locations, that the ensemble better-captured peak incidence, and that the surrogate model appeared biased toward overprediction.
-Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse and that relative performance decreased with the forecast horizon.
+Evaluating the relative performance of the surrogate model compared to the European Forecast Hub ensemble we found that the mean performance was substantially worse and that relative performance decreased with the forecast horizon.
 The median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
@@
 In general, the relative performance of the surrogate model degraded as forecast horizons increased with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
-The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and larger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
+The Forecast Hub ensemble was poorly calibrated, particularly at longer forecast horizons and larger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
 The ensemble forecast was biased towards under prediction at short to medium forecast horizons but unbiased at longer horizons. 
@@
 
-Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the European forecast hub [@Sherratt2022-ue]).
+Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the European Forecast Hub [@Sherratt2022-ue]).
 This means that we are not subject to hindsight bias. 
@@
 This is difficult to avoid without retrospectively re-calculating the ensemble using the same approach as taken by the hub which would reduce the independence of the hub ensemble as a source of truth to compare our forecasts against.
-Given the number of forecasts submitted in most locations and the European forecast hubs' practice of not calculating an ensemble when fewer than 3 independent forecasts were available, the bias in our results caused by this limitation should be relatively small.
-Importantly in this study, we focussed on replicating the forecast hub ensembles' observed behaviour rather than attempting to define an optimal forecast for forecast consumers.
+Given the number of forecasts submitted in most locations and the European Forecast Hubs' practice of not calculating an ensemble when fewer than 3 independent forecasts were available, the bias in our results caused by this limitation should be relatively small.
+Importantly in this study, we focussed on replicating the Forecast Hub ensembles' observed behaviour rather than attempting to define an optimal forecast for forecast consumers.
 It is possible that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
-However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue], then our approach may make sense as a way of producing a "good" forecast.
+However, if we start from the view that the Forecast Hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue], then our approach may make sense as a way of producing a "good" forecast.
 Our focus on replicating the performance of the hub ensemble is also useful as the surrogate model may highlight some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
@@
 This bias could explain the relatively poor performance of the surrogate model, compared to the ensemble, despite the surrogate model being comparably well-calibrated.
-We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick relatively closely to the methodology used by the European forecast hub [@Sherratt2022-ue], aside from the use of population weighting to facilitate comparison between forecast locations, as these choices inform the development of submitted models and so are key to our findings.
+We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick relatively closely to the methodology used by the European Forecast Hub [@Sherratt2022-ue], aside from the use of population weighting to facilitate comparison between forecast locations, as these choices inform the development of submitted models and so are key to our findings.
 
@@
 Progress in understanding which structural model features lead to better infectious disease forecasts has been limited.
-The US forecast hub identified the top 5 performing models and noted the structural assumptions they made but couldn't directly link assumptions with performance [@Cramer2022-sp].
+The US Forecast Hub identified the top 5 performing models and noted the structural assumptions they made but couldn't directly link assumptions with performance [@Cramer2022-sp].
 They also did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
 Similarly, the Germany and Poland forecasting hubs were able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
-All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
+All comparable Forecast Hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
 In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions. 
@@
 Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the US forecasting hub and did not outperform in the case of the European forecasting hub.
-No forecast hub has switched to these alternative ensemble designs for their operational forecast of reported cases, though the US hub has switched to a trained ensemble for death forecasts.
+No Forecast Hub has switched to these alternative ensemble designs for their operational forecast of reported cases, though the US hub has switched to a trained ensemble for death forecasts.
 This suggests that the hub teams do not think the evidence base is strong enough for trained ensembles to be used by forecast consumers for reported cases and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.
-Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
+Other studies have been published evaluating single forecast models in comparison to ensemble performance from the Forecast Hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
 Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not in our model-based forecasts [@Bosse2022-hm; @Bracher2021-wk].
 Finally, our results are potentially sensitive to the definition used to define anomalous observations (generally related to retrospective data revisions).
-Here we follow the practice of the European forecast hub [@EuroHub] of excluding forecasts for weeks with a data revision of more than 5% and forecasts made based on data that is subsequently revised by more than 5%.
+Here we follow the practice of the European Forecast Hub [@EuroHub] of excluding forecasts for weeks with a data revision of more than 5% and forecasts made based on data that is subsequently revised by more than 5%.
 
@@
 As we have hypothesised that the use of absolute scoring measures is inappropriate and leads to performance characteristics that are unlikely to be favoured by forecast stakeholders more work should be done in this area.
-If new forecast ensemble methods are adopted as best practice by forecast hubs then follow-up work attempting to create surrogate forecast models should also use these approaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example, the tendency to be poorly calibrated.
+If new forecast ensemble methods are adopted as best practice by Forecast Hubs then follow-up work attempting to create surrogate forecast models should also use these approaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example, the tendency to be poorly calibrated.
 In September 2022, GitHub announced support for hosted GitHub Action runners with additional compute power [@GitHubActionsLargeRunners].
@@
 We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
-We also conclude that our findings may be largely driven by the choice of evaluation measure used by the forecast hub.
+We also conclude that our findings may be largely driven by the choice of evaluation measure used by the Forecast Hub.
 While this measure has desirable mathematical properties and is routinely used in a similar form e.g., in weather forecasting, it is subject to debate whether it appropriately reflects forecast users' requirements and perceptions as to what makes a good forecast.
@@
 We thank the ECDC for supporting the forecasting hub, and all forecasters who submitted forecasts for making this study possible.
-We thank the forecast hub team for publishing all data in an accessible format.
+We thank the Forecast Hub team for publishing all data in an accessible format.
 We thank the Epiforecast group for a productive discussion of an early version of this analysis.

### 2022-10-04 paper/paper.Rmd
more changes from @jbracher

@@
 \end{align*}
-
-where $r_t$ can be interpreted as the weekly growth rate.
+where the mean and variance of the negative binomial are defined as \[ \mathbb{E}[n] = \mu \ \ \
+\text{ and } \ \ \ \text{Var}[n] = \mu + \frac{\mu^2}{\phi}. \]
+Here $r_t$ can be interpreted as the weekly growth rate.
 $r_t$ is then modelled as a piecewise constant differenced AR(1) process modified such that the dependence of $r_{t-1}$ is multiplied by a decay factor ($\xi_{+,-}$) that varies dynamically according to the sign of $r_{t-1}$.
@@
 In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile [@scoringutils].
-We also calculate the bias [@Funk2019-vi] of both forecasting approaches, stratified by forecast horizon.
+We also calculate the bias (see [@scoringutils] and [@Funk2019-vi] for a more detailed definition) of both forecasting approaches, stratified by forecast horizon.
 This metric aims to capture the tendency for a forecast to under or over-predict.
@@
 
-## Data summary {-}
+##  Summary of the European COVID-19 Forecast Hub Setting {-}
 
@@
     legend.position = "bottom", legend.margin = margin(),
-    legend.justification = "centre", axis.title = element_text(size = 15)
+    legend.justification = "centre", axis.title = element_text(size = 18),
+    legend.text = element_text(size = 18)
   )
@@
 
-If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast again outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs.
-`r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`).
-When using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (`r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` vs.
-`r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`).
+If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast again outperformed the surrogate forecast (rAE for the surrogate compared to the ensemble `r overall_ae_median[model == "Surrogate", Mean] / overall_ae_median[model == "Ensemble", Mean] |> round(digits = 2)`).
+When using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (rAE `r overall_ae_median[model == "Surrogate", Median] / overall_ae_median[model == "Ensemble", Median] |> round(digits = 2)`).
 Across forecast horizons the same pattern of outperformance holds.
@@
 
-Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly certain at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
+Stratifying calibration by quantile and forecast horizon the ensemble forecast was conservative at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
 This behaviour became more prominent as the forecast horizon increased.

### 2022-10-07 paper/paper.Rmd
reorg introduction and axis tics from @sbfnk

@@
 abstract: > 
-  Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts, especially during the COVID-19 pandemic.
-However, these forecasts require a coordinated team to produce and are effectively black boxes.
+  Multi-model and multi-team ensemble forecasts have become widely used in order to generate reliable short-term predictions of infectious disease spread.
+Notably, various public health agencies have used them to leverage academic disease modelling during the COVID-19 pandemic.
+However, ensemble forecasts are essentially black boxes and require extensive effort from numerous participating groups as well as a coordination team.
 In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models.
-  Here we use observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience, to specify a series of model assumptions modelling incident reported cases of COVID-19.
-We then develop a forecast model using these assumptions and compare its performance to the European Forecast Hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon using the weighted interval score.
-  We found that this surrogate model performed visibly somewhat similarly to a subset of example locations though with increased uncertainty and poor performance at forecasting peak incidence.
-Unsurprisingly, we concluded that average performance was substantially worse relative to the ensemble forecast but that median performance was more comparable though still worse.
-The majority of the difference in performance was linked to a subset of forecasts with the majority of these being from early 2022.
-We note that despite performing worse than the hub ensemble the majority of the time our surrogate model showed better calibration, and was less liable to underpredict.
-The central part of the surrogate model’s forecast distribution performed more similarly to the ensemble forecast’s distribution but maintained a right skewed performance distribution.
-  We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble.
-However, more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements.
-We also conclude that our findings may be largely driven by the choice of evaluation measure used by the Forecast Hub.
-If the assumptions underlying this measure do not reflect forecast users' expectations then the Forecast Hub submissions may also not reflect these expectations.
-Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers aiming to develop forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
+Here we use observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience to identify a set of properties present in current ensemble forecasts.
+We then develop a parsimonious forecast model with the aim of mirroring these properties.
+We assess forecasts generated from this model in real time over a six-month period (15th of January 2022 to the 19th of July 2022) and for multiple European countries.
+We focus on forecasts of cases one to four weeks ahead and compare them to those by the European forecast hub ensemble.
+We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence.
+As a result, it falls short of the ensemble in terms of average performance.
+The performance differences, however, seem to be driven mainly by a subset of time points, and in terms of forecast calibration the proposed model actually outperforms the ensemble.
+We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble, but more work is needed to understand the implicit epidemiological model that it represents.
+We moreover note that our findings may be driven by the choice of evaluation measure, which while widely used in the different forecast hubs, may not perfectly reflect forecast users' expectations or requirements.
+Our work will be of use to consumers of ensemble forecasts who wish to better understand their inherent properties, and to researchers aiming to generate reliable forecasts in the absence of the resources required for multi-model ensemble approaches.
 bibliography: library.bib
@@
 This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @charles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.
-
-To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
-These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team.
-Each hub targets a range of metrics including test-positive reported cases, reported deaths, and hospitalisations has a specific geographic focus and asks for weekly forecasts (using epidemiological weeks i.e.
-Sunday to Saturday) over a time horizon of a few weeks.
-Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
-Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
-Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
-We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator, and they are generally the most challenging to predict.
-In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-ft; @Adam2020-if], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
-Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].
-
-We have contributed a range of forecasts to these forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
-Our forecast submissions have not systematically over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
-The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
-Based on our observations our forecasts have generally captured the current trend relatively well but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
-Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing].
-In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
-We found that our human insight-based forecasts outperformed our model-based forecasts on average. 
-This was particularly the case when forecasting cases and at longer forecast horizons.
-We hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].
-
-Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts.
-They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions.
-In comparison to our submitted forecasts, the ensemble forecasts were less reactive to changes in trend such as from stable or reducing case incidence to increasing incidence.
-On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
-Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under vs over predicting.
-There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models
-
-Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability.
-The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
-This mitigates the impact of daily reporting artefacts.
-It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
-This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
-
-The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
-This removes the need for, potentially misspecified, external information on the delay from infection to report, and reduces computational requirements due to a reduction in model complexity. 
-However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
-As discussed in [@Gostic2020-vw] this leads to varying amounts of bias depending on the epidemic phase.
-
-The final simplification is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
-This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
-To model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
-This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence.
-
-The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
-This choice is motivated by our belief that the underlying transmission process is an exponential discrete one and therefore a count error model with a log link function, where variance is linked to the mean, is a sensible choice.
-We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].
 
@@
 ## Model {-}
+
+### Motivation {-}
+
+To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
+These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team.
+Each hub targets a range of metrics including test-positive reported cases, reported deaths, and hospitalisations has a specific geographic focus and asks for weekly forecasts (using epidemiological weeks i.e.
+Sunday to Saturday) over a time horizon of a few weeks.
+Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
+Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
+Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
+We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator, and they are generally the most challenging to predict.
+In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-ft; @Adam2020-if], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
+Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].
+
+We have contributed a range of forecasts to these forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
+Our forecast submissions have not systematically over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
+The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
+Based on our observations our forecasts have generally captured the current trend relatively well but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
+Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing].
+In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
+We found that our human insight-based forecasts outperformed our model-based forecasts on average. 
+This was particularly the case when forecasting cases and at longer forecast horizons.
+We hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].
+
+Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts.
+They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions.
+In comparison to our submitted forecasts, the ensemble forecasts were less reactive to changes in trend such as from stable or reducing case incidence to increasing incidence.
+On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
+Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under vs over predicting.
+There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models
+
+Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability.
+The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
+This mitigates the impact of daily reporting artefacts.
+It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
+This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
+
+The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
+This removes the need for, potentially misspecified, external information on the delay from infection to report, and reduces computational requirements due to a reduction in model complexity. 
+However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
+As discussed in [@Gostic2020-vw] this leads to varying amounts of bias depending on the epidemic phase.
+
+The final simplification is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
+This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
+To model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
+This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence.
+
+The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
+This choice is motivated by our belief that the underlying transmission process is an exponential discrete one and therefore a count error model with a log link function, where variance is linked to the mean, is a sensible choice.
+We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].
 
@@
     legend.position = "bottom", legend.margin = margin(),
-    legend.justification = "centre", axis.title = element_text(size = 18),
-    legend.text = element_text(size = 18)
+    legend.justification = "centre", axis.title = element_text(size = 16)
   )

### 2022-10-07 R/evaluation-plots.R

@@ -50,7 +50,8 @@ plot_forecast_custom <- function(forecasts, log) {
     guides(
       fill = guide_none(), fill_ramp = guide_none(), col = guide_none()
     ) +
-    theme(strip.text.y = element_blank())
+    theme(strip.text.y = element_blank()) +
+    scale_x_date(date_breaks = "1 month", date_labels = "%b")
 }
 
 plot_wis <- function(wis, locs) {
@@ -70,6 +71,7 @@ plot_wis <- function(wis, locs) {
     scale_y_continuous(
       labels = ~ scales::comma(.x, accuracy = 1), trans = "log"
     ) +
+    scale_x_date(date_breaks = "1 month", date_labels = "%b") +
     facet_grid(rows = vars(location_name), scales = "free_y") +
     labs(
       x = "Date", y = "Weighted interval score",

### 2022-10-07 paper/paper.Rmd
add model colours to figure legends

@@
 
-```{r vis-forecasts, fig.cap = "a.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one and four weeks).
+```{r vis-forecasts, fig.cap = "a.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom, by forecast horizon (one and four weeks) for the surrogate model (orange) and forecast ensemble (green).
 30\\%, 60\\%, and 90\\% prediction intervals are shown.
@@
 Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one-week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two-week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three-week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four-week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by at least 25% and that the relative performance of the surrogate model degraded as the forecast horizon increased  (Figure \@ref(fig:assemble-eval) c).
-Much of this outperformance, especially at longer forecast horizons, was driven by a relatively small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a).
-If we instead consider median relative performance (note this is not a proper scoring rule) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
+Much of this outperformance, especially at longer forecast horizons, was driven by a small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
+If we instead consider median relative performance (note this is not a proper scoring rule and should not be used to choose between models) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
 This would suggest that an increasingly skewed score distribution as the forecast horizon increased is responsible for the increase in the mean relative score (Figure \@ref(fig:assemble-eval) a).
@@
 
-```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate and ensemble forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon.
+```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate (orange) and ensemble (green) forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon.
 Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the nominal level of the interval (i.e., 30\\%, 60\\% and 90\\%, respectively).

### 2022-10-07 paper/paper.Rmd
clean up mean and variance definition

@@
 However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models, in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
-This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @charles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.
+This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @Charles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.
 
@@
 We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator, and they are generally the most challenging to predict.
-In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-ft; @Adam2020-if], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
+In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-ft; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
 Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].
@@
 \end{align*}
-where the mean and variance of the negative binomial are defined as \[ \mathbb{E}[n] = \mu \ \ \
-\text{ and } \ \ \ \text{Var}[n] = \mu + \frac{\mu^2}{\phi}. \]
+where the mean and variance of the negative binomial are defined as \[ \mathbb{E}[n] = \lambda_t \ \ \
+\text{ and } \ \ \ \text{Var}[n] = \lambda_t + \frac{\lambda_t^2}{\phi}. \]
 Here $r_t$ can be interpreted as the weekly growth rate.

### 2022-10-07 paper/paper.Rmd
chase up remaining references

@@
 We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator, and they are generally the most challenging to predict.
-In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-ft; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
+In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models, mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@Srivastava2020-qv; @Li2021-cj], semi-mechanistic approaches that blend both of these approaches [@Castro2021-lj; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-jn; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Karlen2020-pt; @Bosse2022-hm].
 Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].
@@
 Lastly, here we have only explored a surrogate for an ensemble for a single disease, a limited set of locations, and a single target (incident cases), meaning our findings are difficult to generalise.
-Follow-up work should explore whether this behaviour holds across diseases, locations, and epidemiological targets where the behaviour of ensembles is notably different [@REF].
+Follow-up work should explore whether this behaviour holds across diseases, locations, and epidemiological targets where the behaviour of ensembles is notably different.
 However, this work is limited by the lack of large-scale forecast ensembling projects despite them showing obvious promise to improve the forecasts available to stakeholders.

### 2022-10-07 paper/paper.Rmd
switch template

@@
   bookdown::pdf_book:
-    base_format: rticles::peerj_article
+    base_format: rticles::arxiv_article
     extra_dependencies: ["float"]
@@
 
+```{r}
+observations <- data.table(
+  No. = ,
+  Observation = c(
+    "Robust to daily reporting artefacts",
+
+  )
+)
+```
 To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].

### 2022-10-07 paper/paper.Rmd
add table

--- 2022-10-07 paper/paper.Rmd
@@
 
+```{r}
+observations <- data.table(
+  No. = 1:5,
+  Observation = c(
+    "Robust to daily reporting artefacts",
+    "Some ability to forecast future trend changes",
+    "Less reactive to apparent observed changes in trend",
+    "Sharper forecasts",
+    "A tendency towards under prediction"
+  )
+)
+
+assumptions <- data.table(
+  Assumption = c(
+    "Reported cases can be modelled using weekly data and a generative process discretised by week", # nolint
+    "Reported cases can be modelled as if they represented infections",
+    "The growth rate of infections can be represented as an auto-regressive process with an order of 1 week", # nolint
+    "Unobserved interventions and more general changes in transmission towards a stable state can be represented using a multiplicative decay parameter" # nolint
+  ),
+  Observation = c("1, and 2", "-", "3 and 4", "2, and 5")
+)
+knitr::kables(
+  list(knitr::kable(observations), knitr::kable(assumptions)),
+  caption = "(\\#tab:observations) Observations on the relative performance of the Forecast Hub ensemble compared to our forecast submissions and assumptions/simplifications based on these observations." # nolint
+)
+```
+
 To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
@@
 
-Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability.
+Based on these insights (summarised in Table \@ref(tab:observations)), here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability.
 The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
@@
 
-The final simplification is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
+The final simplification is to model the growth rate as a differenced auto-regressive process with an order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
 This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
@@
 We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].
+Our derived set of assumptions and simplifications is summarised in Table \@ref(tab:observations).
 

### 2022-10-10 paper/paper.Rmd
move table
more changes + comments from @sbfnk

--- 2022-10-07 paper/paper.Rmd
@@
 Notably, various public health agencies have used them to leverage academic disease modelling during the COVID-19 pandemic.
-However, ensemble forecasts are essentially black boxes and require extensive effort from numerous participating groups as well as a coordination team.
+However, ensemble forecasts are difficult to interpret and require extensive effort from numerous participating groups as well as a coordination team.
 In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models.
-Here we use observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience to identify a set of properties present in current ensemble forecasts.
-We then develop a parsimonious forecast model with the aim of mirroring these properties.
-We assess forecasts generated from this model in real time over a six-month period (15th of January 2022 to the 19th of July 2022) and for multiple European countries.
-We focus on forecasts of cases one to four weeks ahead and compare them to those by the European forecast hub ensemble.
-We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence.
-As a result, it falls short of the ensemble in terms of average performance.
-The performance differences, however, seem to be driven mainly by a subset of time points, and in terms of forecast calibration the proposed model actually outperforms the ensemble.
+Here we used observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience to identify a set of properties present in current ensemble forecasts.
+We then developed a parsimonious forecast model with the aim of mirroring these properties.
+We assess forecasts generated from this model in real time over a six-month period (the 15th of January 2022 to the 19th of July 2022) and for multiple European countries.
+We focused on forecasts of cases one to four weeks ahead and compared them to those by the European forecast hub ensemble.
+We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence as measured by the Weighted Interval Score.
+The performance differences, however, seem to be partially driven by a subset of time points, and the proposed model appears better calibrated than the ensemble.
 We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble, but more work is needed to understand the implicit epidemiological model that it represents.
-We moreover note that our findings may be driven by the choice of evaluation measure, which while widely used in the different forecast hubs, may not perfectly reflect forecast users' expectations or requirements.
-Our work will be of use to consumers of ensemble forecasts who wish to better understand their inherent properties, and to researchers aiming to generate reliable forecasts in the absence of the resources required for multi-model ensemble approaches.
 bibliography: library.bib
@@
 Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Reich2022-jo].
-These approaches have been inspired by the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providing forecasts that are trusted by stakeholders.
-
-Whilst there is robust and consistent evidence that multi-team ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
-The most significant is the difficulty in interpreting them.
-This relates both to the underlying mechanisms for the forecasts they produce and to understanding if and when their behaviour can is desirable.
-This impacts users' trust, how easily ensemble performance can be improved, and how easily contributor forecasts can be improved.
-Forecasts from these ensembles also require considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
-Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute and adjust their models to changing conditions [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
-Each of these issues may impact the long-term quality of the resulting ensemble forecasts and have implications for end-users. 
-There has been limited critical feedback on forecasting ensembling projects for infectious disease epidemiology with little evaluation of the effort required to produce them compared to their benefits for improving forecast performance.
-What evaluation there has been has mainly focussed on secondary data analysis.
+The experience of other domains has inspired these approaches, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providing forecasts that stakeholders trust.
 
@@
 All of these collaborations ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
-Both the US and European Forecast Hubs were supported and received funding from public health agencies (the Center for Disease Control (CDC) and European Center for Disease Prevention and Control (ECDC) respectively) with their forecasts being adopted into the official output from these agencies.
-However, less progress has been made in mitigating the downsides of these projects and improving access to the high-quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.
+Both the US and European Forecast Hubs were supported and received funding from public health agencies (the Center for Disease Control (CDC) and European Center for Disease Prevention and Control (ECDC) respectively) with their forecasts being adopted into the official output from these agencies
 
-In this work, we draw insights from ensemble forecasts produced and endorsed by these Forecast Hubs, as well as our forecasting work, to propose and evaluate a "surrogate" forecast model.
-This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
+Whilst there is robust and consistent evidence that multi-team ensemble forecasts provide reliable and performant forecasts across domains [@Reich2022-jo] they also have a range of downsides.
+The most significant is the difficulty in interpreting them.
+This relates both to the underlying mechanisms for the forecasts they produce and to understanding if and when their behaviour is desirable.
+This impacts users' trust, how easily ensemble performance can be improved, and how easily contributor forecasts can be improved.
+Forecasts from these ensembles also require considerable resource cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
+Additional challenges with maintaining multi-team collaborations can include providing detailed feedback to those contributing forecasts that would allow them to improve their forecast approaches, providing incentives for forecasters to continue to contribute and adjust their models to changing conditions, and difficulty improving the quality of the ensemble by learning from past predictive performance [@Sherratt2022-ue].
+Each of these issues may impact the long-term quality of the resulting forecasts and have implications for end-users. 
+Little progress has so far been made in mitigating these downsides or in improving access to the high-quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.
+There has also been limited critical feedback on the structure of forecasting ensembling projects for infectious disease epidemiology and little evaluation of the effort required to produce them relative to their benefits for improving forecast performance.
+
+In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysics [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
+These approaches generally take the same inputs as the models they seek to emulate and then are trained based on the output from those models.
+In the context of epidemioloigcal models non-parametric emulation has been used to allow the rapid exploration of the parameter space of complex models that would otherwise be resource prohibative [@Iskauskas2022-fz; @Charles2022-ds].
+Replicating forecasts from resource intensive models may also be a promising avenue of future research. 
+However, as an approach it may be less useful for resolving some of the issues of multi-team and multi-model forecasts as it does not provide interpretability, key for stakeholder take-up, it is not clear how the methods commonly used perform out of sample, and emulating a quantile based forecast is an unadressed methodological challenge as far as we are aware.
+
+In this work, we draw insights from ensemble forecasts produced and endorsed by the COVID-19 Forecast Hubs, as well as our forecasting work, to propose and evaluate a "surrogate" forecast model.
+This seeks to reproduce ensemble performance by mimicking its behaviour based on a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
 The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
@@
 
-In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysics [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
-However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models, in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
-This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @Charles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.
-
-In this study, we evaluate an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles based on a set of clear assumptions.
+In order to achieve these aims, we evaluate an initial attempt at developing a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles based on a set of clear assumptions.
 We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble.
-First, we define the model and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub Actions workflow.
-We briefly outline some of the summary statistics it produces and their epidemiological relevance.
+In this work, we first define the model and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub Actions workflow [@GitHubActionsAbout].  
 We then evaluate its real-time performance in comparison to the European Forecast Hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
-We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed.
+We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where it performs less well.
 Finally, we summarise our findings, discuss their implications, and highlight areas where more work is needed.
@@
 
-We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
+To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
+These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team.
+Each hub targets a range of metrics including test-positive reported cases, reported deaths, and hospitalisations has a specific geographic focus and asks for weekly forecasts (using epidemiological weeks i.e.
+Sunday to Saturday) over a time horizon of a few weeks.
+Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
+Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
+We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator for other metrics such as hospitalisations [@Meakin2022-hw], and they are generally the most challenging to predict [@Sherratt2022-ue].
+In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models, mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@Srivastava2020-qv; @Li2021-cj], semi-mechanistic approaches that blend both of these approaches [@Castro2021-lj; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-jn; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Karlen2020-pt; @Bosse2022-hm].
+Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].
+
+We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth` and defined in the following section).
 We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
@@
 We used the latest available observed data as of the 1st of September 2022 (commit `f6922c3e4bdcb055abcbba8e73472afacac4cf40` from [@EuroHubGitHub]).
-Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.
+Incidence was aggregated by epidemiological week (i.e. Sunday through Saturday).
 As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
@@
 No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
-Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@EuroHub; @Bracher2022-ug; @Cramer2022-sp].
 Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week).
-The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team and timely submission).
+The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team and timely submission)  [@Sherratt2022-ue].
 An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
@@
 
-## Model {-}
+## Observations based on previous forecasts {-}
 
-### Motivation {-}
-
-To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
-These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team.
-Each hub targets a range of metrics including test-positive reported cases, reported deaths, and hospitalisations has a specific geographic focus and asks for weekly forecasts (using epidemiological weeks i.e.
-Sunday to Saturday) over a time horizon of a few weeks.
-Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
-Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
-Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
-We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator, and they are generally the most challenging to predict.
-In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models, mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@Srivastava2020-qv; @Li2021-cj], semi-mechanistic approaches that blend both of these approaches [@Castro2021-lj; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-jn; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Karlen2020-pt; @Bosse2022-hm].
-Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].
-
-We have contributed a range of forecasts to these forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
+We have contributed a range of forecasts to COVID-19 forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
 Our forecast submissions have not systematically over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
@@
 Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing].
+This resource usage is likely beyond the capacity of many interested in having access to state of the art short-term forecasts of infectious disease.
 In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
@@
 Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under vs over predicting.
-There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models
 
-Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability.
+```{r observations}
+observations <- data.table(
+  No. = 1:5,
+  Observation = c(
+    "Robust to daily reporting artefacts",
+    "Some ability to forecast future trend changes",
+    "Less reactive to apparent observed changes in trend",
+    "Sharper forecasts",
+    "A tendency towards under prediction"
+  )
+)
+
+knitr::kable(observations),
+  caption = "(\\#tab:observations) Observations on the relative performance of the Forecast Hub ensemble compared to our forecast submissions." # nolint
+)
+```
+
+## Model {-}
+
+### Assumptions and simplifications {-}
+
+Based on these observations (summarised in Table \@ref(tab:observations)), here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability and with a lower cost.
 The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
@@
 This removes the need for, potentially misspecified, external information on the delay from infection to report, and reduces computational requirements due to a reduction in model complexity. 
-However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
+However, this sacrifices some of the interpretability of the forecast model as any transmission statistics we now calculate will be based on reported cases and not latent infections.
 As discussed in [@Gostic2020-vw] this leads to varying amounts of bias depending on the epidemic phase.
 
-The final simplification is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
+The final simplification is to model the growth rate as a differenced auto-regressive process with an order 1 rather than using a gaussian process-based method as we have done in other work [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
 This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
@@
 We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].
+Our set of assumptions and simplifications are summarised in Table \@ref(tab:observations).
+
+```{r assumptions}
+assumptions <- data.table(
+  Assumption = c(
+    "Reported cases can be modelled using weekly data and a generative process discretised by week", # nolint
+    "Reported cases can be modelled as if they represented infections",
+    "The growth rate of infections can be represented as an auto-regressive process with an order of 1 week", # nolint
+    "Unobserved interventions and more general changes in transmission towards a stable state can be represented using a multiplicative decay parameter" # nolint
+  ),
+  Observation = c("1, and 2", "-", "3 and 4", "2, and 5")
+)
+knitr::kable(assumptions,
+  caption = "(\\#tab:assumptions) Assumptions/simplifications based on observations of the relative performance of Forecast Hub ensembles compared to our forecast submissions." # nolint
+)
+```
 
@@
 \end{align*}
-where the mean and variance of the negative binomial are defined as \[ \mathbb{E}[n] = \lambda_t \ \ \
-\text{ and } \ \ \ \text{Var}[n] = \lambda_t + \frac{\lambda_t^2}{\phi}. \]
+where the mean and variance of the negative binomial are defined as \[ \mathbb{E}[C_t] = \lambda_t \ \ \
+\text{ and } \ \ \ \text{Var}[C_t] = \lambda_t + \frac{\lambda_t^2}{\phi}. \]
 Here $r_t$ can be interpreted as the weekly growth rate.
@@
   \beta &\sim \text{Normal}\left(0, 0.25 \right) \\
-  \eta_t &\sim \text{Half-Normal}\left(0, \sigma \right) \\
-  \sigma &\sim \text{Normal}\left(0, 0.2 \right) \\
+  \eta_t &\sim \text{Normal}\left(0, \sigma \right) \\
+  \sigma &\sim \text{Half-Normal}\left(0, 0.2 \right) \\
   \frac{1}{\sqrt{\phi}} &\sim \text{Half-Normal}(0, 1) 
@@
 
-To standardise forecasts across forecast locations we first normalised both weekly notified test positive cases and forecast test positive cases by the population in the forecast region scaled to be an incidence rate per 10,000 people.
+To standardise the magnitude of observations and forecasts across forecast locations we first normalised both weekly notified test positive cases and forecast test positive cases by the population in the forecast region scaled to be an incidence rate per 10,000 people.
 We then evaluated a subset of forecasts from the following countries Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom visually by forecast horizon (1-4 weeks) for both the ensemble and surrogate model.
@@
 The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`)  [@forecast.vocs].
-We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`.
+We note that our use of an indicator function introduces a discontinuity to the posterior making it less suited for use with `stan`.
 Other model formulations without this feature would be more efficient and robust.
@@
 
-## Forecast evaluation {-}
-
-### Visualisation of forecasts by horizon {-}
-
-In our example set of locations, the absolute performance of the ensemble and the surrogate model was visually similar on the log scale in all locations at short forecast horizons though this varied by location (Figure \@ref(fig:vis-forecasts) b).
-On the natural scale the difference in performance was more marked, especially for periods of high incidence and at longer horizons (Figure \@ref(fig:vis-forecasts) a).
-Performance was not homogeneous across our set of example locations with the surrogate model performing similarly to the ensemble in Slovakia whilst in the United Kingdom and Germany the surrogate model performed substantially worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
-For both the ensemble and the surrogate, performance decreased as the forecast horizon increased with this being particularly noticeable for the surrogate model during periods of high incidence.
-In general, in the study period, the ensemble appeared to be better able to forecast peak incidence. 
-Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
-In comparison to the ensemble model the surrogate model appeared less likely to place weight on unfeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in February in Slovakia and Poland).
-
 ```{r vis-forecasts, fig.cap = "a.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom, by forecast horizon (one and four weeks) for the surrogate model (orange) and forecast ensemble (green).
@@
 b.) A replicate of a.) but with incidence rates on the log scale.
-c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "95%", fig.height = 12, fig.width = 18}
+c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "95%", fig.height = 8, fig.width = 12}
 p <- (
@@
 
+## Forecast evaluation {-}
+
+### Visualisation of forecasts by horizon {-}
+
+In our example set of locations, the absolute performance of the ensemble and the surrogate model was visually similar on the log scale in all locations at short forecast horizons though this varied by location (Figure \@ref(fig:vis-forecasts) b).
+On the natural scale the difference in performance was more marked, especially for periods of peak incidence and at longer horizons (Figure \@ref(fig:vis-forecasts) a).
+Performance was not homogeneous across our set of example locations with the surrogate model performing similarly to the ensemble in Slovakia whilst in the United Kingdom and Germany the surrogate model performed substantially worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
+For both the ensemble and the surrogate, performance decreased as the forecast horizon increased with this being particularly noticeable for the surrogate model during periods of peak incidence.
+In general, in the study period, the ensemble appeared to be better able to forecast peak incidence. 
+Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
+In comparison to the ensemble model the surrogate model appeared less likely to place weight on unfeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in February in Slovakia and Poland).
+
 ### Relative forecast evaluation {-}
@@
 
-If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast again outperformed the surrogate forecast (rAE for the surrogate compared to the ensemble `r overall_ae_median[model == "Surrogate", Mean] / overall_ae_median[model == "Ensemble", Mean] |> round(digits = 2)`).
-When using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (rAE `r overall_ae_median[model == "Surrogate", Median] / overall_ae_median[model == "Ensemble", Median] |> round(digits = 2)`).
+If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast again outperformed the surrogate forecast (rAE for the surrogate compared to the ensemble `r round(overall_ae_median[model == "Surrogate", Mean] / overall_ae_median[model == "Ensemble", Mean], digits = 2)`).
+If we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (rAE `r round(overall_ae_median[model == "Surrogate", Median] / overall_ae_median[model == "Ensemble", Median], digits = 2)`).
 Across forecast horizons the same pattern of outperformance holds.
@@
 
-```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale.
+```{r assemble-eval, fig.height = 10, fig.width = 10, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale.
 a.) The density of the relative score by horizon.
@@
 
-Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction interval, though with a tendency to over cover, with empirical coverage of `r overall_coverage[range %in% "30% interval", Surrogate]`%, `r overall_coverage[range %in% "60% interval", Surrogate]`%, `r overall_coverage[range %in% "90% interval", Surrogate]`% respectively.
+Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction interval, though with a tendency to slightly over cover, with empirical coverage of `r overall_coverage[range %in% "30% interval", Surrogate]`%, `r overall_coverage[range %in% "60% interval", Surrogate]`%, `r overall_coverage[range %in% "90% interval", Surrogate]`% respectively.
 The ensemble model was less well calibrated, with a tendency to under cover with empirical coverage of `r overall_coverage[range %in% "30% interval", Ensemble]`%, `r overall_coverage[range %in% "60% interval", Ensemble]`%, `r overall_coverage[range %in% "90% interval", Ensemble]`% respectively (Figure \@ref(fig:plot-coverage) a).
@@
 
-```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate (orange) and ensemble (green) forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon.
+```{r plot-coverage, fig.height = 8, fig.width = 8, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate (orange) and ensemble (green) forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon.
 Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the nominal level of the interval (i.e., 30\\%, 60\\% and 90\\%, respectively).
@@
 It is possible that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
-However, if we start from the view that the Forecast Hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue], then our approach may make sense as a way of producing a "good" forecast.
+However, if we start from the view that the Forecast Hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue], then our approach may make sense as a way of producing a "good" forecast without sacrificing interpretability.
+
+Developing forecast methodologies with limited resources is critical as whilst improving predictive performance is a key goal of short-term forecasting it is also important that forecast models be accessible as this makes it easier to iteratively improve them, and makes them more flexible when used in real-time setting [@Osthus2022-yw].
+An example of the lack of flexibility of the Forecast Hub ensemble, and its constituent models, is the ensembles response to upswings linked to variant dynamics with the growth of one variant being temporally hidden by the decline of another.
+Rather than forecasting this ahead of time the Forecast ensembles generally only reacted to changes in the observed data indicating that variant information was not being used by forecasters.
+Unlike the Forecast Hub ensemble the surrogate model can be, and indeed has been [@Abbott2021-ly], be extended to capture this. 
+Other examples where additional transient information is available to forecasters can be readily thought of implying this is a general advantage of simpler methods.
+
 Our focus on replicating the performance of the hub ensemble is also useful as the surrogate model may highlight some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
@@
 In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions. 
-The US [@Ray2022-zq], and European [@Sherratt2022-ue]], forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling.
+The US [@Ray2022-zq], and European [@Sherratt2022-ue], forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling.
 Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the US forecasting hub and did not outperform in the case of the European forecasting hub.
@@
 The findings from our study may also be useful for informing this improved set of assumptions. 
-A particular focus should be on understanding why our surrogate model was liable to overestimate peak incidence and what simply additional assumptions may be used to mitigate this.
+A particular focus should be on understanding why our surrogate model was liable to overestimate peak incidence and what simple additional assumptions may be used to mitigate this.
 In addition, the model we derived based on our assumptions was likely not optimal both in terms of compute time and accuracy at reproducing ensemble-like behaviour.
@@
 As we identified that the tails of our predictive distributions were responsible for a large proportion of the difference in performance compared to the forecast ensemble it may be the case that post-processing of forecasts from our surrogate model would enhance their similarity to the forecast ensemble.
-This seems likely to improve out-of-sample performance but does not help with the secondary aim of understanding the implicit assumptions driving the performance of multi-model, multi-team infectious disease forecast ensembles.
+This seems likely to improve out-of-sample performance but does not help with understanding the implicit assumptions driving the performance of multi-model, multi-team infectious disease forecast ensembles.
 As we have hypothesised that the use of absolute scoring measures is inappropriate and leads to performance characteristics that are unlikely to be favoured by forecast stakeholders more work should be done in this area.
@@
 
-SA and SF were funded by a Wellcome senior fellowship to SF (210758/Z/18/Z), KS and HG were funded by an ECDC grant to SF.
+SA,SF, KS and HG were funded by a Wellcome senior fellowship to SF (210758/Z/18/Z), KS and HG were further funded by an ECDC grant to SF.
 JB acknowledges support from the Helmholtz Foundation via the SIM-610 CARD Information and Data Science Pilot Project.

### 2022-10-10 paper/paper.Rmd
tidy up references
clean up comman in table

@@
 
-knitr::kable(observations),
+knitr::kable(observations
   caption = "(\\#tab:observations) Observations on the relative performance of the Forecast Hub ensemble compared to our forecast submissions." # nolint

### 2022-10-10 paper/paper.Rmd
catch another missing comma

@@
 
-knitr::kable(observations
+knitr::kable(observations,
   caption = "(\\#tab:observations) Observations on the relative performance of the Forecast Hub ensemble compared to our forecast submissions." # nolint

### 2022-10-10 paper/paper.Rmd
tidy up and rescale

--- 2022-10-10 paper/paper.Rmd
@@
 library(scales)
+library(knitr)
+library(kableExtra)
 library(here)
@@
 
-knitr::kable(observations
-  caption = "(\\#tab:observations) Observations on the relative performance of the Forecast Hub ensemble compared to our forecast submissions." # nolint
-)
+observations |>
+  kable(
+  caption = "(\\#tab:observations) Observations on the relative performance of the Forecast Hub ensemble compared to our forecast submissions.", # nolint
+  booktabs = TRUE
+  ) |>
+  kable_styling()
 ```
@@
 )
-knitr::kable(assumptions,
-  caption = "(\\#tab:assumptions) Assumptions/simplifications based on observations of the relative performance of Forecast Hub ensembles compared to our forecast submissions." # nolint
-)
+
+assumptions |>
+  kable(
+  caption = "(\\#tab:assumptions) Assumptions/simplifications based on observations of the relative performance of Forecast Hub ensembles compared to our forecast submissions.", # nolint
+  booktabs = TRUE
+  ) |>
+  kable_styling()
 ```
@@
 b.) A replicate of a.) but with incidence rates on the log scale.
-c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "95%", fig.height = 8, fig.width = 12}
+c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "95%", fig.height = 10, fig.width = 10}
 p <- (
@@
     legend.position = "bottom", legend.margin = margin(),
-    legend.justification = "centre", axis.title = element_text(size = 16)
+    legend.justification = "centre"
   )
@@
 
-```{r assemble-eval, fig.height = 10, fig.width = 10, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale.
+```{r assemble-eval, fig.height = 8, fig.width = 8, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale.
 a.) The density of the relative score by horizon.

### 2022-10-10 paper/paper.Rmd
make legends vertical

@@
     legend.position = "bottom", legend.margin = margin(),
-    legend.justification = "centre"
+    legend.justification = "centre", legend.justification = "vertical"
   )

### 2022-10-10 paper/paper.Rmd
figure tweaks

--- 2022-10-10 paper/paper.Rmd
@@
   ) |>
-  kable_styling()
+  kable_styling(full_width = FALSE)
 ```
@@
 b.) A replicate of a.) but with incidence rates on the log scale.
-c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "95%", fig.height = 10, fig.width = 10}
+c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "95%", fig.height = 8, fig.width = 10}
 p <- (
@@
     legend.position = "bottom", legend.margin = margin(),
-    legend.justification = "centre"
+    legend.justification = "centre", legend.direction = "vertical"
   )

### 2022-10-10 paper/paper.Rmd
more plot fiddling:

@@
 On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
-Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under vs over predicting.
+Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under vs over predicting. Our observations are summarised in Table \@ref(tab:observations).
 
@@
 
-Based on these observations (summarised in Table \@ref(tab:observations)), here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability and with a lower cost.
+Based on our observations of forecast performance (summarised in Table \@ref(tab:observations)), here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability and with a lower cost to produce.
 The first simplification we make is to model only weekly data, rather than using daily data and then aggregating.
@@
 We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].
-Our set of assumptions and simplifications are summarised in Table \@ref(tab:observations).
+Our set of assumptions and simplifications are summarised in Table \@ref(tab:assumptions).
 
@@
     "The growth rate of infections can be represented as an auto-regressive process with an order of 1 week", # nolint
-    "Unobserved interventions and more general changes in transmission towards a stable state can be represented using a multiplicative decay parameter" # nolint
+    "Unobserved interventions and more general changes in transmission towards a stable state can be represented \n using a multiplicative decay parameter" # nolint
   ),
@@
 b.) A replicate of a.) but with incidence rates on the log scale.
-c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "95%", fig.height = 8, fig.width = 10}
+c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "98%", fig.height = 8, fig.width = 12}
 p <- (
@@
 
-```{r assemble-eval, fig.height = 8, fig.width = 8, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale.
+```{r assemble-eval, fig.height = 8, fig.width = 8, out.width = "98%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale.
 a.) The density of the relative score by horizon.
@@
     legend.position = "bottom", legend.margin = margin(),
-    legend.justification = "centre", legend.direction = "vertical"
+    legend.justification = "centre", legend.box = "vertical"
   )
@@
 
-```{r plot-coverage, fig.height = 8, fig.width = 8, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate (orange) and ensemble (green) forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon.
+```{r plot-coverage, fig.height = 8, fig.width = 8, out.width = "98%", fig.cap = "a.) Empirical coverage of the surrogate (orange) and ensemble (green) forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon.
 Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the nominal level of the interval (i.e., 30\\%, 60\\% and 90\\%, respectively).
@@
     (
-      plot_coverage_range(coverage, ranges)
+      plot_coverage_range(coverage, ranges) +
+       guides(
+        fill = guide_none(), point_color = guide_none(), point = guide_none()
+       )
     )

### 2022-10-10 paper/paper.Rmd
tweak figure labels

@@
   melt(
-    measure.vars = c("Weighted interval score", "Absolute error of the median")
+    measure.vars = c("Weighted interval score", "Absolute error")
   )

### 2022-10-11 paper/paper.Rmd
final comments from @kathsherratt

@@
 All of these collaborations ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
-Both the US and European Forecast Hubs were supported and received funding from public health agencies (the Center for Disease Control (CDC) and European Center for Disease Prevention and Control (ECDC) respectively) with their forecasts being adopted into the official output from these agencies
+Both the US and European Forecast Hubs were supported and received funding from public health agencies (the Center for Disease Control (CDC) and European Center for Disease Prevention and Control (ECDC) respectively) with their forecasts being adopted into the official output from these agencies.
 
@@
 These approaches generally take the same inputs as the models they seek to emulate and then are trained based on the output from those models.
-In the context of epidemioloigcal models non-parametric emulation has been used to allow the rapid exploration of the parameter space of complex models that would otherwise be resource prohibative [@Iskauskas2022-fz; @Charles2022-ds].
+In the context of epidemiological models, non-parametric emulation has been used to allow the rapid exploration of the parameter space of complex models that would otherwise be resource prohibitive [@Iskauskas2022-fz; @Charles2022-ds].
 Replicating forecasts from resource intensive models may also be a promising avenue of future research. 
-However, as an approach it may be less useful for resolving some of the issues of multi-team and multi-model forecasts as it does not provide interpretability, key for stakeholder take-up, it is not clear how the methods commonly used perform out of sample, and emulating a quantile based forecast is an unadressed methodological challenge as far as we are aware.
+However, as an approach it may be less useful for resolving some of the issues of multi-team and multi-model forecasts as it does not provide interpretability, key for stakeholder take-up.
+Additionally, it is not clear how the methods commonly used perform out of sample, and emulating a quantile based forecast is an unaddressed methodological challenge as far as we are aware.
 
@@
 This seeks to reproduce ensemble performance by mimicking its behaviour based on a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
-The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
-In addition to this, we seek to provide a simple, and robust, forecasting system that others can easily reuse, perhaps where a Forecast Hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
+The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecasts widely considered the gold standard for COVID-19 forecasting.
+In addition to this, we seek to provide a simple and robust forecasting system that others can easily reuse, perhaps where a Forecast Hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
 
@@
 We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble.
-In this work, we first define the model and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub Actions workflow [@GitHubActionsAbout].  
+In this work, we first define the model and summarise its implementation, with a focus on minimal resource use and reproducibility as a GitHub Actions workflow [@GitHubActionsAbout].  
 We then evaluate its real-time performance in comparison to the European Forecast Hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
@@
 Finally, we summarise our findings, discuss their implications, and highlight areas where more work is needed.
-We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 Forecast Hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
+We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 Forecast Hub ensembles, provide a sensible, low-resource, surrogate modelwhere large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
 
@@
 These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team.
-Each hub targets a range of metrics including test-positive reported cases, reported deaths, and hospitalisations has a specific geographic focus and asks for weekly forecasts (using epidemiological weeks i.e.
+Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus; and asks for weekly forecasts (using epidemiological weeks i.e.
 Sunday to Saturday) over a time horizon of a few weeks.
@@
 
-The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
+The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models.
+Instead we make use of a negative binomial observation model allowing the inclusion of overdispersion.
 This choice is motivated by our belief that the underlying transmission process is an exponential discrete one and therefore a count error model with a log link function, where variance is linked to the mean, is a sensible choice.
-We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].
+We suggest that part of the reason the hub ensembles exhibit such sharpness is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations [@bracherEvaluatingEpidemicForecasts2021].
 Our set of assumptions and simplifications are summarised in Table \@ref(tab:assumptions).
@@
 
-To standardise the magnitude of observations and forecasts across forecast locations we first normalised both weekly notified test positive cases and forecast test positive cases by the population in the forecast region scaled to be an incidence rate per 10,000 people.
-We then evaluated a subset of forecasts from the following countries Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom visually by forecast horizon (1-4 weeks) for both the ensemble and surrogate model.
+We standardised the magnitude of observations and forecasts across forecast locations by scaling both weekly notified test positive cases and forecast test positive cases by the population in the forecast region to an incidence rate per 10,000 people.
+We then visually evaluated forecasts from a subset of locations by forecast horizon (1 and 4 weeks) on both the natural and log scales.
+The countries in this subset were Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom.
 These countries were selected to include forecasts based on different numbers and types of submitted forecast models, to be at least partially representative of the full sample of forecast locations, and to include nations for which the authors had a good understanding of local data and transmission dynamics in the study period.
@@
 Many countries, such as the UK, saw large BA.1 waves in January that then resulted in declining incidence rates through February (Figure \@ref(fig:vis-forecasts)).
-From late February, through to the end of May, most nations then saw another wave, typically with lower reported incidence rates, driven by BA.2.
-This wave was typically characterised by a lower peak than the BA.1 wave with a more gradual decrease in incidence.
-The end of our study period was characterised by the gradual take-over of the BA.4/BA.5 subvariants that again had a lower peak and lower absolute growth rates.
+From late February through to the end of May, most nations saw another wave driven by BA.2.
+This wave typically saw lower reported incidence rates, and was characterised by a lower peak than the BA.1 wave with a more gradual decrease in incidence.
+The end of our study period was dominated by the gradual take-over of the BA.4/BA.5 subvariants that again had a lower peak and lower absolute growth rates.
 Unlike earlier periods in the pandemic, our study period did not see the use of new non-pharmaceutical interventions (NPIs) in response to increasing COVID-19 incidence in most locations.
@@
 Across forecast horizons the same pattern of outperformance holds.
-However, the difference in relative performance was less than when the full probability distribution was accounted for with this becoming more marked as the forecast horizon increased (Figure \@ref(fig:assemble-eval) c).
+However, the difference in relative performance was less than when the full probability distribution was accounted for, with this becoming more marked as the forecast horizon increased (Figure \@ref(fig:assemble-eval) c).
 
@@
 Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the European Forecast Hub [@Sherratt2022-ue]).
-This means that we are not subject to hindsight bias. 
+This means that our results are not subject to hindsight bias. 
 The downside of this approach is that it was not possible to update the surrogate model over time in response to the initial evaluation or to explore other parameterisations that might be more successful of which there are likely several.
@@
 Developing forecast methodologies with limited resources is critical as whilst improving predictive performance is a key goal of short-term forecasting it is also important that forecast models be accessible as this makes it easier to iteratively improve them, and makes them more flexible when used in real-time setting [@Osthus2022-yw].
-An example of the lack of flexibility of the Forecast Hub ensemble, and its constituent models, is the ensembles response to upswings linked to variant dynamics with the growth of one variant being temporally hidden by the decline of another.
-Rather than forecasting this ahead of time the Forecast ensembles generally only reacted to changes in the observed data indicating that variant information was not being used by forecasters.
-Unlike the Forecast Hub ensemble the surrogate model can be, and indeed has been [@Abbott2021-ly], be extended to capture this. 
+An example of the lack of flexibility of the Forecast Hub ensemble, and its constituent models, is the ensembles response to upswings linked to variant dynamics, with the growth of one variant being temporally hidden by the decline of another.
+Rather than forecasting this ahead of time the Forecast ensembles generally only reacted to changes in the observed data indicating that variant information was not being used by most forecasters.
+Unlike the Forecast Hub ensemble the surrogate model can be, and indeed has been [@Abbott2021-ly], easily extended to capture this. 
 Other examples where additional transient information is available to forecasters can be readily thought of implying this is a general advantage of simpler methods.
@@
 However, the US [@Cramer2022-sp], European [@Sherratt2022-ue], and Germany/Poland [@Bracher2021-wk] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles.
-In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study).
-The poor calibration of the forecast ensembles produced by median Hub ensembles has been noted repeatedly [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue] but little progress has been made in understanding the causes or suggesting alternatives forecasts for policymakers to use.
+In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score, as in this study).
+
+The poor calibration of the forecast ensembles produced by median Hub ensembles has been noted repeatedly [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue] but little progress has been made in understanding the causes or suggesting alternatives.
 Progress in understanding which structural model features lead to better infectious disease forecasts has been limited.
-The US Forecast Hub identified the top 5 performing models and noted the structural assumptions they made but couldn't directly link assumptions with performance [@Cramer2022-sp].
-They also did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
+The US Forecast Hub identified the top 5 performing models and noted the structural assumptions they made, but couldn't directly link assumptions with performance [@Cramer2022-sp].
+They also did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions (as done in this study to motivate the surrogate model), or explore the performance of a forecasting model designed with these assumptions in mind.
 Similarly, the Germany and Poland forecasting hubs were able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
-All comparable Forecast Hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
+All comparable Forecast Hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness - though this was rarely fully defined - and should be the output used by forecast consumers [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
 In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions. 
@@
 This suggests that the hub teams do not think the evidence base is strong enough for trained ensembles to be used by forecast consumers for reported cases and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study.
-Other studies have been published evaluating single forecast models in comparison to ensemble performance from the Forecast Hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
+
+Other studies have been published evaluating single forecast models in comparison to ensemble performance from the Forecast Hub.
+In general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric.
 Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not in our model-based forecasts [@Bosse2022-hm; @Bracher2021-wk].

### 2022-10-11 paper/paper.Rmd
fix bug in labels

@@
   relative_ae_by_horizon, c("relative_interval_score", "ris"),
-  c("Absolute error of the median", "raem")
+  c("Absolute error", "raem")
 )

### 2022-10-11 paper/paper.Rmd
add citation links
also render as html
final comments from @nikosbosse

@@
 We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence as measured by the Weighted Interval Score.
-The performance differences, however, seem to be partially driven by a subset of time points, and the proposed model appears better calibrated than the ensemble.
+The performance differences, however, seem to be partially driven by a subset of time points, and the proposed model appears better probabilistically calibrated than the ensemble.
 We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble, but more work is needed to understand the implicit epidemiological model that it represents.
 bibliography: library.bib
+link-citations: true
 output:
@@
     extra_dependencies: ["float"]
+    keep_tex: true
 ---
@@
   ) |>
-  kable_styling(full_width = FALSE)
+  kable_styling(full_width = TRUE)
 ```

### 2022-10-11 paper/paper.Rmd
read through for typos

@@
 abstract: > 
-  Multi-model and multi-team ensemble forecasts have become widely used in order to generate reliable short-term predictions of infectious disease spread.
+  Multi-model and multi-team ensemble forecasts have become widely used to generate reliable short-term predictions of infectious disease spread.
 Notably, various public health agencies have used them to leverage academic disease modelling during the COVID-19 pandemic.
@@
 Here we used observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience to identify a set of properties present in current ensemble forecasts.
-We then developed a parsimonious forecast model with the aim of mirroring these properties.
-We assess forecasts generated from this model in real time over a six-month period (the 15th of January 2022 to the 19th of July 2022) and for multiple European countries.
+We then developed a parsimonious forecast model intending to mirror these properties.
+We assess forecasts generated from this model in real time over six months (the 15th of January 2022 to the 19th of July 2022) and for multiple European countries.
 We focused on forecasts of cases one to four weeks ahead and compared them to those by the European forecast hub ensemble.
@@
 Each of these issues may impact the long-term quality of the resulting forecasts and have implications for end-users. 
-Little progress has so far been made in mitigating these downsides or in improving access to the high-quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.
+Little progress has so far been made in mitigating these downsides or in improving access to the high-quality and robust forecasts they seek to generate for geographies without coverage or for other infectious diseases.
 There has also been limited critical feedback on the structure of forecasting ensembling projects for infectious disease epidemiology and little evaluation of the effort required to produce them relative to their benefits for improving forecast performance.
@@
 These approaches generally take the same inputs as the models they seek to emulate and then are trained based on the output from those models.
-In the context of epidemiological models, non-parametric emulation has been used to allow the rapid exploration of the parameter space of complex models that would otherwise be resource prohibitive [@Iskauskas2022-fz; @Charles2022-ds].
-Replicating forecasts from resource intensive models may also be a promising avenue of future research. 
-However, as an approach it may be less useful for resolving some of the issues of multi-team and multi-model forecasts as it does not provide interpretability, key for stakeholder take-up.
-Additionally, it is not clear how the methods commonly used perform out of sample, and emulating a quantile based forecast is an unaddressed methodological challenge as far as we are aware.
+In the context of epidemiological models, non-parametric emulation has been used to allow the rapid exploration of the parameter space of complex models that would otherwise be resource-prohibitive [@Iskauskas2022-fz; @Charles2022-ds].
+Replicating forecasts from resource-intensive models may also be a promising avenue of future research. 
+However, as an approach, it may be less useful for resolving some of the issues of multi-team and multi-model forecasts as it does not provide interpretability, key for stakeholder take-up.
+Additionally, it is not clear how the methods commonly used perform out of sample, and emulating a quantile-based forecast is an unaddressed methodological challenge as far as we are aware.
 
@@
 
-In order to achieve these aims, we evaluate an initial attempt at developing a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles based on a set of clear assumptions.
+To achieve these aims, we evaluate an initial attempt at developing a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles based on a set of clear assumptions.
 We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble.
@@
 Finally, we summarise our findings, discuss their implications, and highlight areas where more work is needed.
-We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 Forecast Hub ensembles, provide a sensible, low-resource, surrogate modelwhere large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
+We aim for this work to highlight some of the potential implicit assumptions of current COVID-19 Forecast Hub ensembles, provide a sensible, low-resource, surrogate model where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.
 
@@
 These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team.
-Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus; and asks for weekly forecasts (using epidemiological weeks i.e.
+Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus, and asks for weekly forecasts (using epidemiological weeks i.e.
 Sunday to Saturday) over a time horizon of a few weeks.
@@
 To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomalous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomalous data [@Sherratt2022-ue].
-We defined anomalous data based on the implementation used by [@EuroHub] where a datapoint is considered anomalous if a future revision alters it by more than 5%.
+We defined anomalous data based on the implementation used by [@EuroHub] where a data point is considered anomalous if a future revision alters it by more than 5%.
 
@@
 Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing].
-This resource usage is likely beyond the capacity of many interested in having access to state of the art short-term forecasts of infectious disease.
+This resource usage is likely beyond the capacity of many interested in having access to state-of-the-art short-term forecasts of infectious diseases.
 In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
@@
 The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models.
-Instead we make use of a negative binomial observation model allowing the inclusion of overdispersion.
+Instead, we make use of a negative binomial observation model allowing the inclusion of overdispersion.
 This choice is motivated by our belief that the underlying transmission process is an exponential discrete one and therefore a count error model with a log link function, where variance is linked to the mean, is a sensible choice.
@@
 In our study period, incidence rates across European nations and in the UK were primarily driven by the spread of novel subvariants of concern related to the Omicron variant and changes in population susceptibility.
-Many countries, such as the UK, saw large BA.1 waves in January that then resulted in declining incidence rates through February (Figure \@ref(fig:vis-forecasts)).
+Many countries, such as the UK, saw large BA.1 waves in January, resulting in declining incidence rates through February (Figure \@ref(fig:vis-forecasts)).
 From late February through to the end of May, most nations saw another wave driven by BA.2.
@@
 Unlike earlier periods in the pandemic, our study period did not see the use of new non-pharmaceutical interventions (NPIs) in response to increasing COVID-19 incidence in most locations.
-In addition, ascertainment rates likely reduced over time in most locations due to reductions in routine testing, and reductions in test availability.
+In addition, ascertainment rates likely reduced over time in most locations due to reductions in routine testing and test availability.
 Whilst both the reduced use of NPIs and testing generally occurred across nations our study period also marked an increase in the heterogeneity of the response to the COVID-19 pandemic with nations changing policy at different times and to different degrees.
@@
 Of the remaining models submitted `r length(local_models)` were submitted in only one location.
-Single location models were clustered in a few locations, particularly in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
+Single-location models were clustered in a few locations, particularly in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
 Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
@@
 
-```{r vis-forecasts, fig.cap = "a.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom, by forecast horizon (one and four weeks) for the surrogate model (orange) and forecast ensemble (green).
+```{r vis-forecasts, fig.cap = "a.) Forecasts of notified test-positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom, by forecast horizon (one and four weeks) for the surrogate model (orange) and forecast ensemble (green).
 30\\%, 60\\%, and 90\\% prediction intervals are shown.
@@
 c.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast.
-d.) Bias of the ensemble and surrogate forecasts stratified by horizon."}
+d.) The bias of the ensemble and surrogate forecasts stratified by horizon."}
 p <- (
@@
 The Forecast Hub ensemble was poorly calibrated, particularly at longer forecast horizons and larger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
-The ensemble forecast was biased towards under prediction at short to medium forecast horizons but unbiased at longer horizons. 
+The ensemble forecast was biased towards under-prediction at short to medium forecast horizons but unbiased at longer horizons. 
 In comparison, the surrogate model was biased towards over prediction and this bias increased linearly with the forecast horizon.
@@
 Given the number of forecasts submitted in most locations and the European Forecast Hubs' practice of not calculating an ensemble when fewer than 3 independent forecasts were available, the bias in our results caused by this limitation should be relatively small.
-Importantly in this study, we focussed on replicating the Forecast Hub ensembles' observed behaviour rather than attempting to define an optimal forecast for forecast consumers.
+Notably in this study, we focussed on replicating the Forecast Hub ensembles' observed behaviour rather than attempting to define an optimal forecast for forecast consumers.
 It is possible that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
@@
 
-Developing forecast methodologies with limited resources is critical as whilst improving predictive performance is a key goal of short-term forecasting it is also important that forecast models be accessible as this makes it easier to iteratively improve them, and makes them more flexible when used in real-time setting [@Osthus2022-yw].
+Developing forecast methodologies with limited resources is critical as whilst improving predictive performance is a key goal of short-term forecasting it is also important that forecast models be accessible as this makes it easier to iteratively improve them, and makes them more flexible when used in real-time settings [@Osthus2022-yw].
 An example of the lack of flexibility of the Forecast Hub ensemble, and its constituent models, is the ensembles response to upswings linked to variant dynamics, with the growth of one variant being temporally hidden by the decline of another.

### 2022-10-11 paper/paper.Rmd
final comments from @jbracher

@@
   - code: 2
-    address: Chair of Statistics and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
+    address: Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
 abstract: > 
@@
 We focused on forecasts of cases one to four weeks ahead and compared them to those by the European forecast hub ensemble.
-We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence as measured by the Weighted Interval Score.
-The performance differences, however, seem to be partially driven by a subset of time points, and the proposed model appears better probabilistically calibrated than the ensemble.
+We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence (as measured by the Weighted Interval Score).
+The performance differences, however, seem to be partially due to a subset of time points, and the proposed model appears better probabilistically calibrated than the ensemble.
 We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble, but more work is needed to understand the implicit epidemiological model that it represents.
@@
 The experience of other domains has inspired these approaches, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providing forecasts that stakeholders trust.
-
 The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic due to a pressing need for reliable forecasts and a perception that many publicly available forecasts were low quality.
@@
 All of these collaborations ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
-Both the US and European Forecast Hubs were supported and received funding from public health agencies (the Center for Disease Control (CDC) and European Center for Disease Prevention and Control (ECDC) respectively) with their forecasts being adopted into the official output from these agencies.
+Both the US and European Forecast Hubs were supported and received funding from public health agencies (the Center for Disease Control, CDC, and European Center for Disease Prevention and Control, ECDC, respectively) with their forecasts used in official communications by these agencies.
 
@@
 The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecasts widely considered the gold standard for COVID-19 forecasting.
-In addition to this, we seek to provide a simple and robust forecasting system that others can easily reuse, perhaps where a Forecast Hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.
+In addition to this, we seek to provide a simple and robust forecasting system that others can easily reuse, perhaps where a Forecast Hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and improved forecast performance.
 
@@
 These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team.
-Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus, and asks for weekly forecasts (using epidemiological weeks i.e.
+Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus, and asks for weekly forecasts (using MMWR epidemiological weeks i.e.
 Sunday to Saturday) over a time horizon of a few weeks.
@@
 Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week).
-The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team and timely submission) [@Sherratt2022-ue].
+The ensemble forecast was constructed by taking the median of all forecasts for each predictive quantile without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team and timely submission) [@Sherratt2022-ue].
 An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
@@
 \end{align*}
-where the mean and variance of the negative binomial are defined as \[ \mathbb{E}[C_t] = \lambda_t \ \ \
-\text{ and } \ \ \ \text{Var}[C_t] = \lambda_t + \frac{\lambda_t^2}{\phi}. \]
+where the mean and variance of the negative binomial are given by \[ \mathbb{E}[C_{t} \mid \lambda_{t}] = \lambda_t \ \ \
+\text{ and } \ \ \ \text{Var}[C_{t} \mid \lambda_{t}] = \lambda_t + \frac{\lambda_t^2}{\phi}. \]
 Here $r_t$ can be interpreted as the weekly growth rate.
@@
 
-We standardised the magnitude of observations and forecasts across forecast locations by scaling both weekly notified test positive cases and forecast test positive cases by the population in the forecast region to an incidence rate per 10,000 people.
+We standardised the magnitude of observations and forecasts across forecast locations, in order to facilitate comparison, by scaling both weekly notified test positive cases and forecast test positive cases by the population in the forecast region to an incidence rate per 10,000 people.
+This differs from the approach typically taken by the Forecast Hubs where no population standardisation is used [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
 We then visually evaluated forecasts from a subset of locations by forecast horizon (1 and 4 weeks) on both the natural and log scales.
@@
       plot_relative_summary(relative_summary) +
-      plot_layout(guides = "keep")
+      plot_rwis_location 
     ) +
-    plot_rwis_location +
     plot_layout(heights = c(1, 7), guides = "collect")
@@
 ) +
-  plot_layout(guides = "collect", widths = c(1, 1)) +
+  plot_layout(guides = "collect") +
   plot_annotation(tag_levels = "a") &
@@
 It also highlights some of the differences between our surrogate forecast model and the ensemble that may lead to new insights into the mechanisms leading to the ensemble's behaviour, such as the generally poor coverage of the ensemble that could not be explained by the assumptions we used in developing our surrogate methodology.
-Whilst we normalised reported cases to be population-adjusted incidence rates our results are conditional on the use of the weighted interval score as an evaluation metric.
-As this proper scoring rule measures absolute performance this means that forecasts during periods of higher incidence are given more weight than forecasts from periods of low incidence.
+Whilst we normalised reported cases to be population-adjusted incidence rates, and so can more easily compare across locations than using the approach commonly implemented by the Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue], our results are still conditional on the use of the weighted interval score as an evaluation metric.
+As this proper scoring rule scales with the order of magnitude of the predicted quantity this means that forecasts during periods of higher incidence are given more weight than forecasts from periods of low incidence.
 It also means that overprediction is penalised more than underprediction as incidence rates are bounded at zero but relatively weakly bounded by populations at the upper bound (as incidence rates are typically only a small fraction of the overall population).
@@
 
-Whilst we derived our surrogate model from a range of assumptions based on observing ensemble forecasts behaviour and the behaviour and structure of submitted models this was by no means as structured or complete a process as it could have been.
+Whilst we derived our surrogate model from a range of assumptions based on observing ensemble forecasts behaviour and the behaviour and structure of submitted models avenues for future improvement remain in terms of improving the approach used to elicit these observations.
 In follow-up work, a more rigorous approach to this could be taken to further refine this set of assumptions, in particular using the input of a wider pool of researchers. 
@@
 Follow-up work should explore whether this behaviour holds across diseases, locations, and epidemiological targets where the behaviour of ensembles is notably different.
-However, this work is limited by the lack of large-scale forecast ensembling projects despite them showing obvious promise to improve the forecasts available to stakeholders.
+However, this is limited to infectious diseases with similar large-scale forecast ensembling projects.
+These projects remain relatively rare despite them showing obvious promise to improve the forecasts available to stakeholders.
 

### 2022-10-12 paper/paper.Rmd
update mmwy reference
add final comments from @sbfnk and @seabbs

@@
 title: Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble
-preprint: true
+params:
+ preprint: TRUE
+preprint: `r params$preprint`
 author: 
   - name: Sam Abbott
-    affiliation: 1
+    affiliation: 1, 2
     corresponding: true
@@
   - name: Katharine Sherratt
-    affiliation: 1
+    affiliation: 1, 2
   - name: Nikos Bosse
-    affiliation: 1
+    affiliation: 1, 2
   - name: Hugo Gruson
-    affiliation: 1
+    affiliation: 1, 2 
   - name:  Johannes Bracher
-    affiliation: 2
+    affiliation: 3
   - name: Sebastian Funk
-    affiliation: 1
+    affiliation: 1, 2
 affiliation:
   - code: 1
-    address: London School of Hygiene and Tropical Medicine, London, UK
+    address: The Centre for Mathematical Modelling of Infectious Diseases, London School of Hygiene & Tropical Medicine, London, UK
   - code: 2
+    address: Department of Infectious Disease Epidemiology, London School of Hygiene & Tropical Medicine, London, UK
+  - code: 3
     address: Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
@@
     keep_tex: true
+   bookdown::html_document2:
+    keep_toc: true        
 ---
@@
 In the context of epidemiological models, non-parametric emulation has been used to allow the rapid exploration of the parameter space of complex models that would otherwise be resource-prohibitive [@Iskauskas2022-fz; @Charles2022-ds].
-Replicating forecasts from resource-intensive models may also be a promising avenue of future research. 
-However, as an approach, it may be less useful for resolving some of the issues of multi-team and multi-model forecasts as it does not provide interpretability, key for stakeholder take-up.
-Additionally, it is not clear how the methods commonly used perform out of sample, and emulating a quantile-based forecast is an unaddressed methodological challenge as far as we are aware.
+These methods may be less useful for resolving some of the issues of multi-team and multi-model forecasts as they do not provide interpretability, key for stakeholder take-up.
+Additionally, it is not clear how these methods perform out of sample, or how they would be applied to a quantile-based forecast.
 
@@
 The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecasts widely considered the gold standard for COVID-19 forecasting.
-In addition to this, we seek to provide a simple and robust forecasting system that others can easily reuse, perhaps where a Forecast Hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and improved forecast performance.
+Our secondary aim is to provide the basis for a robust forecasting system that others can easily reuse both in operational contexts and as a platform for future research.
 
@@
 Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus, and asks for weekly forecasts (using MMWR epidemiological weeks i.e.
-Sunday to Saturday) over a time horizon of a few weeks.
+Sunday to Saturday [@MMWRweek]) over a time horizon of a few weeks.
 Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
@@
 On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
-Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under vs over predicting. Our observations are summarised in Table \@ref(tab:observations).
+Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under- vs overpredicting. Our observations are summarised in Table \@ref(tab:observations).
 
@@
     "Sharper forecasts",
-    "A tendency towards under prediction"
+    "A tendency towards underprediction"
   )
@@
 Similarly the prior for $\beta$ has been chosen to be weakly informative that there is weak auto-correlation in differenced growth rates.
-$\sigma$ has also been made weakly informative under the assumption that the potential change in growth rates in a single timestep should be relatively small.
+$\sigma$ has also been made weakly informative under the assumption that the potential change in growth rates in a single time-step should be relatively small.
 
@@
 The code for the forecasting model defined above along with the infrastructure required to forecast using GitHub Actions can be found here: https://github.com/seabbs/ecdc-weekly-growth-forecasts
-Versions archived on Zenodo are available [@REF] and [@REF].
+Versions archived on Zenodo are available [@AnalysisCode
+] and [@ModelCode].
 
@@
 
-Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction interval, though with a tendency to slightly over cover, with empirical coverage of `r overall_coverage[range %in% "30% interval", Surrogate]`%, `r overall_coverage[range %in% "60% interval", Surrogate]`%, `r overall_coverage[range %in% "90% interval", Surrogate]`% respectively.
-The ensemble model was less well calibrated, with a tendency to under cover with empirical coverage of `r overall_coverage[range %in% "30% interval", Ensemble]`%, `r overall_coverage[range %in% "60% interval", Ensemble]`%, `r overall_coverage[range %in% "90% interval", Ensemble]`% respectively (Figure \@ref(fig:plot-coverage) a).
+Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction interval, though with a tendency to be slightly underconfident, with empirical coverage of `r overall_coverage[range %in% "30% interval", Surrogate]`%, `r overall_coverage[range %in% "60% interval", Surrogate]`%, `r overall_coverage[range %in% "90% interval", Surrogate]`% respectively.
+The ensemble model was less well calibrated, with a tendency to be overconfident with empirical coverage of `r overall_coverage[range %in% "30% interval", Ensemble]`%, `r overall_coverage[range %in% "60% interval", Ensemble]`%, `r overall_coverage[range %in% "90% interval", Ensemble]`% respectively (Figure \@ref(fig:plot-coverage) a).
 When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon, and then became progressively less well calibrated as the forecast horizon increased (Figure \@ref(fig:plot-coverage) a).
 In comparison, the surrogate forecast was less well calibrated than the ensemble forecast at the one-week forecast horizon with a tendency to have a larger empirical coverage than required  (Figure \@ref(fig:plot-coverage) a).
-At longer horizons and narrower prediction intervals, the surrogate forecast became better calibrated though with a tendency to under cover.
+At longer horizons and narrower prediction intervals, the surrogate forecast became better calibrated though with a tendency to be overconfident.
 This was not the case for the 90% prediction interval where the surrogate model covered more than the expected interval, for all horizons, indicating forecasts were overly uncertain for this interval regardless of the horizon.
@@
 This behaviour became more prominent as the forecast horizon increased.
-In contrast, the surrogate forecast was generally equally well calibrated across horizons with a tendency to over cover for intervals above the median.
-At longer horizons, however, quantiles below the median became liable to under cover.
+In contrast, the surrogate forecast was generally equally well calibrated across horizons with a tendency to be under confident for intervals above the median.
+At longer horizons, however, quantiles below the median were over confident.
 
@@
 The ensemble forecast was biased towards under-prediction at short to medium forecast horizons but unbiased at longer horizons. 
-In comparison, the surrogate model was biased towards over prediction and this bias increased linearly with the forecast horizon.
+In comparison, the surrogate model was biased towards overprediction and this bias increased linearly with the forecast horizon.
 
@@
 
-## Conclusions {-}
+# Conclusions {-}
 
@@
 Our work is useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to researchers thinking about how to develop forecasts that perform similarly to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
+
+# Acknowledgments {-}
+
+We thank the ECDC for supporting the forecasting hub, and all forecasters who submitted forecasts for making this study possible.
+We thank the Forecast Hub team for publishing all data in an accessible format.
+We thank the Epiforecast group for a productive discussion of an early version of this analysis.
+We thank Molly for being a good Labrador.
 
@@
 
+And are archived here:
+
+https://doi.org/10.5281/zenodo.7189308, https://doi.org/10.5281/zenodo.7189620
+
 ## Funding {-}
@@
 
-# Acknowledgments {-}
-
-We thank the ECDC for supporting the forecasting hub, and all forecasters who submitted forecasts for making this study possible.
-We thank the Forecast Hub team for publishing all data in an accessible format.
-We thank the Epiforecast group for a productive discussion of an early version of this analysis.
-We thank Molly for being a good Labrador.
-
 # References {-}

### 2022-10-12 paper/paper.Rmd
make sure ranges are passed as expected

@@
     "A tendency towards underprediction"
+    "Modelling the reporting process appears to have little impact"
   )
@@
   ),
-  Observation = c("1, and 2", "-", "3 and 4", "2, and 5")
+  Observation = c("1, and 2", "6", "3 and 4", "2, and 5")
 )
@@
 p <- (
-  plot_forecast_custom(forecasts, log = FALSE) |
-  plot_forecast_custom(forecasts, log = TRUE) +
+  plot_forecast_custom(forecasts, log = FALSE, ranges = ranges) |
+  plot_forecast_custom(forecasts, log = TRUE, ranges = ranges) +
   labs(y = "") |

### 2022-10-12 R/evaluation-plots.R

@@ -43,10 +43,10 @@ plot_forecasts <- function(forecasts, locs, ranges, log = FALSE) {
   return(plot)
 }
 
-plot_forecast_custom <- function(forecasts, log) {
+plot_forecast_custom <- function(forecasts, log, ranges) {
   forecasts |>
     DT(horizon == 1 | horizon == 4) |>
-    plot_forecasts(locs, ranges, log = log) +
+    plot_forecasts(locs, ranges = ranges, log = log) +
     guides(
       fill = guide_none(), fill_ramp = guide_none(), col = guide_none()
     ) +

### 2022-10-12 paper/paper.Rmd
manually specify table column width

@@
     "Sharper forecasts",
-    "A tendency towards underprediction"
+    "A tendency towards underprediction",
     "Modelling the reporting process appears to have little impact"
@@
   ) |>
-  kable_styling(full_width = TRUE)
+  kable_styling(full_width = FALSE) |>
+  column_spec(1, width = "3in")
 ```

### 2022-10-12 paper/paper.Rmd
fix params
fix figure legend
change title

@@
 ---
-title: Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble
+title: Evaluating an epidemiologically motivated surrogate model of a multi-model ensemble
 params:

### 2022-10-12 paper/paper.Rmd
try and fix parameterised rmarkdown

@@
 title: Evaluating an epidemiologically motivated surrogate model of a multi-model ensemble
+output:
+  bookdown::pdf_book:
+    base_format: rticles::peerj_article
+    extra_dependencies: ["float"]
+    keep_tex: true
+  bookdown::html_document2:
+    keep_toc: true   
 params:
   preprint: TRUE  
-preprint: `r params$preprint`
+preprint: params$preprint
 author: 
@@
 link-citations: true   
-output:
-  bookdown::pdf_book:
-    base_format: rticles::peerj_article
-    extra_dependencies: ["float"]
-    keep_tex: true
-   bookdown::html_document2:
-    keep_toc: true        
 ---

### 2022-10-12 paper/paper.Rmd
drop markdown parameterisation for now

@@
   preprint: TRUE  
-preprint: params$preprint
+preprint: TRUE
 author: 

### 2022-10-12 paper/paper.Rmd
correct paper numbering

@@
 observations <- data.table(
-  No. = 1:5,
+  No. = 1:6,
   Observation = c(

### 2022-11-16 paper/paper.Rmd
reviewer 1 introducition and abstract minor comments

@@
 params:
-  preprint: TRUE  
-preprint: TRUE
+  preprint: FALSE  
+preprint: FALSE
 author: 
@@
 However, ensemble forecasts are difficult to interpret and require extensive effort from numerous participating groups as well as a coordination team.
-In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models.
+In other fields, resource requirements have been reduced by training simplified models that reproduce some of the observed behaviour of more complex models.
 Here we used observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience to identify a set of properties present in current ensemble forecasts.
-We then developed a parsimonious forecast model intending to mirror these properties.
+We then developed a parsimonious surrogate forecast model intending to mirror these properties.
 We assess forecasts generated from this model in real time over six months (the 15th of January 2022 to the 19th of July 2022) and for multiple European countries.
@@
 We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence (as measured by the Weighted Interval Score).
-The performance differences, however, seem to be partially due to a subset of time points, and the proposed model appears better probabilistically calibrated than the ensemble.
-We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble, but more work is needed to understand the implicit epidemiological model that it represents.
+However, the proposed model appears better probabilistically calibrated than the ensemble.
+We conclude that our surrogate forecast model may have captured some of the dynamics of the hub ensemble, but more work is needed to understand the implicit epidemiological model that it represents.
 bibliography: library.bib
@@
 In this work, we draw insights from ensemble forecasts produced and endorsed by the COVID-19 Forecast Hubs, as well as our forecasting work, to propose and evaluate a "surrogate" forecast model.
-This seeks to reproduce ensemble performance by mimicking its behaviour based on a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
+This surrogate model seeks to reproduce ensemble performance by mimicking its behaviour based on a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
 The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecasts widely considered the gold standard for COVID-19 forecasting.
@@
 In this work, we first define the model and summarise its implementation, with a focus on minimal resource use and reproducibility as a GitHub Actions workflow [@GitHubActionsAbout].  
-We then evaluate its real-time performance in comparison to the European Forecast Hub ensemble by visualising forecasts, using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
-We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where it performs less well.
+We then evaluate the surrogate model's real-time performance in comparison to the European Forecast Hub ensemble by visualising forecasts using the weighted interval score [@bracherEvaluatingEpidemicForecasts2021], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
+We highlight settings where this model performs well as a surrogate for the ensemble forecast and areas where it performs less well.
 Finally, we summarise our findings, discuss their implications, and highlight areas where more work is needed.

### 2023-01-09 paper/paper.Rmd
add additional paper changes]

@@
 Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
-We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator for other metrics such as hospitalisations [@Meakin2022-hw], and they are generally the most challenging to predict [@Sherratt2022-ue].
+We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator for other metrics such as hospitalisations [@Meakin2022-hw], and they are generally the most challenging to predict as they do not have a clear leading indicator [@Sherratt2022-ue].
 In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models, mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@Srivastava2020-qv; @Li2021-cj], semi-mechanistic approaches that blend both of these approaches [@Castro2021-lj; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-jn; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Karlen2020-pt; @Bosse2022-hm].
@@
 
-We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth` and defined in the following section).
-We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
-Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@JHU].
+We extracted forecasts and data on notified weekly COVID-19 cases from the European Forecasting Hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth` and defined in the following section).
+We included all locations covered by the European Forecasting Hub, which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
+Data on notified weekly cases were originally sourced from the Johns Hopkins University (JHU) curated data repository [@JHU].
 We used the latest available observed data as of the 1st of September 2022 (commit `f6922c3e4bdcb055abcbba8e73472afacac4cf40` from [@EuroHubGitHub]).
@@
 As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
-To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomalous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomalous data [@Sherratt2022-ue].
+To account for this we followed the practice of the European Forecasting Hub project in excluding forecasts made using anomalous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomalous data [@Sherratt2022-ue].
 We defined anomalous data based on the implementation used by [@EuroHub] where a data point is considered anomalous if a future revision alters it by more than 5%.
@@
 Based on our observations our forecasts have generally captured the current trend relatively well but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns.
-Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing].
+Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European Forecasting Hub [@AzurePricing].
 This resource usage is likely beyond the capacity of many interested in having access to state-of-the-art short-term forecasts of infectious diseases.
@@
 
-To download and manipulate forecasts from the European forecasting hub [@EuroHub] we use the `data.table` (`1.14.2`) [@data.table] and `gh` (`1.3.0`) [@gh] R packages.
+To download and manipulate forecasts from the European Forecasting Hub [@EuroHub] we use the `data.table` (`1.14.2`) [@data.table] and `gh` (`1.3.0`) [@gh] R packages.
 We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, visualise forecasts and summary measures, and summarise forecasts.
@@
 
-We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, the European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
+We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European Forecasting Hub (nations of the European Union, the European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
 In total `r nrow(metadata)` forecasts were made across all locations, with `r forecast_dates` unique forecast dates and `r models` independent forecast models (including the European hub baseline model).
@@
 Of the remaining models submitted `r length(local_models)` were submitted in only one location.
-Single-location models were clustered in a few locations, particularly in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
+Single-location models were clustered in a few locations, particularly in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European Forecasting Hub project [@Sherratt2022-ue]).
 Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
@@
 The US [@Ray2022-zq], and European [@Sherratt2022-ue], forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling.
-Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the US forecasting hub and did not outperform in the case of the European forecasting hub.
+Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the US forecasting hub and did not outperform in the case of the European Forecasting Hub.
 No Forecast Hub has switched to these alternative ensemble designs for their operational forecast of reported cases, though the US hub has switched to a trained ensemble for death forecasts.
